### 2017-03-09 manuscript.Rmd
first commit with references

---
title: "On g~0~"
author: ''
date: ''
output:
  word_document:
    reference_docx: manuscriptstyle.docx
csl: global-change-biology.csl
bibliography: references.bib
---

\
Remko A. Duursma^1^, Chris Blackman^1^, Belinda Medlyn^1^

\
^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia

\
**Corresponding author**: 
Remko Duursma

\
E: remkoduursma@gmail.com

\
T: +61(0)45701806

# Abstract

# Introduction

1. Importance of stomatal conductance. Prevalence of the Ball-Berry model, and g0. What is g0? Point of this paper.

2. Nighttime conductance not necessarily same as minimum conductance

@lombardozzi2017 puts g0 into CLM, notes in introduction that 'conductance during the night (and at other times when A = 0 ) is thus represented using g0'.

@barnard2013 makes the same assumption

@dekauwe2015 makes this point as well, their Fig. A2 (can we add that also?)

@barbour2007 considerable stomatal control during the night; mirrors daytime

@howard2007 nighttime conductance lower in drought. Nighttime 2-5x higher than cuticular.

We are not reviewing nighttime conductance here, that has been done extensively by others [@caird2007]

2b. Variation in g0

@misson2004 g0 decreases with pre-dawn water potential

3. How do we estimate g0 from gas exchange?

Poorly, that's how! Problem of correlation between g1 and g0; reason why compilations like Lin et al. set g0 to zero.
Problem that g0 is frequently negative (e.g. @heroult2013)

4. Components of minimum conductance

- Stomata simply do not close all the way, might stay open regardless of photosynthesis, or are broken.

5. Cuticular conductance

- Kerstiens
- H2O not CO2? [@boyer1997]
- gmin as a plastic trait. Temperature response generally unresolved [@kerstiens2006]
- @riederer2001 correlation between temperature and gcut (non-stomatal cuticles)
- @burghardt2003 gmin review
- @schreiber1996 cuticle review

@waldencoleman2013 Very interesting! gdark is different from gmin. Whole-plant WUE correlated with gdark (but not gmin). Lots of references on gdark and gmin and their variation. Closed system to measure gdark+gmin; gmin on detached leaves.

@santrucek Direct measurements showing much higher 'cuticular' conductance of the leaf side with stomata, though stomata were closed. In agreement with @kerstiens1996 who showed 10x higher gmin for stomatal side vs. non-stomatal cuticle.

@messinger2006 Conclude that Ci response of gs is the same in the dark or when DCMU was applied (which disrupts electron transport). However, Fig5A-B indicate that gs was much higher in the dark than when DCMU was applied (although the ‘response’ was similar). This indicates gs response to low PAR (darkness) is more than just responding to what photosynthesis does.

@riederer2001 Review on cuticular conductance. gcut is not correlated with cuticle thickness. No reference to stomata. Cuticles are even less permeable than polymers of equivalent thickness used in packaging.

@fanourakis2013 gmin (desiccation gmin) much higher in high RH grown plants, at any value of leaf water content (Table 1). This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss (Hypothesis 2).

@burghardt2003 gmin was the same as cuticular permeance (Table 3) (Gmin is lowest gs in desiccating leaves, which includes cuticle and closed stomata). This demonstrates that stomata were able to completely close, and not lose significant amounts of water (with the exception of Hedera helix).

@leuning1995 discusses g0

@caird2007 For many species, gnight is not stable throughout the night period. Endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [lots of refs]

@korner1994 table of minimum conductance. Also points out (without reference) that g0 in droughted leaves is certainly lower than g0 in dark or very high CO2 leaves.

@egea2011 Assumes g0 is constant, does not vary with water stress, and that it represents a ‘cuticular conductance’.

## Figures

1. Ball-Berry relation, showing g0 on intercept. Second panel correlation ellipse g0 ~ g1 for one study. Third panel g0 vs. R2 for the Lin dataset.

2. DeKauwe2015 figure A2 redone (minimum photosynthesis, minimum conductance)

3. Kerstiens data: differences between methods, connection to climate data (speciesmap)

4. Blackman data: cuticular conductance is a plastic trait

# References {.unnumbered}

### 2017-03-10 manuscript.Rmd
messy analyses of lin, miner

@@
 
+@manzoni assumes minimum conductance is cuticular only, because CO2 does not diffuse across it but water does.
+Seems to work in his modelling context but appears to be in disagreement with the observation that gmin >> gcut.
+
+Various measures of minimum conductance
+
+- gcut (cuticle without stomata)
+- gmin (minimum conductance in situ, stomata closed)
+- g0 (minimum conductance in the Ball-Berry-type model)
+- gdark (nighttime conductance)
+
+gdark > gmin > gcut
+g0?
+
 ## Figures

### 2017-03-10 R/analysis_lin2015.R

@@ -1,0 +1,82 @@
+
+
+fits <- lmList(Cond ~ BBopti | fitgroup, data=lin2015a, na.action=na.omit)
+p <- as.data.frame(coef(fits)) %>% 
+  rownames_to_column(var = "fitgroup") %>%  
+  rename(g0 = `(Intercept)`, g1 = BBopti) %>% 
+  mutate(R2 = sapply(fits, function(x)summary(x)$adj.r.squared),
+         rmse = sapply(fits, function(x)summary(x)$sigma),
+         cor = sapply(fits, function(x)summary(x, correlation=T)$correlation[2,1]),
+         g0_lci = sapply(fits, function(x)confint(x)[1,1]),
+         g0_uci = sapply(fits, function(x)confint(x)[1,2])) %>% 
+  left_join(summarize(lin2015a, 
+                      sd_bbopti = sd(BBopti, na.rm=TRUE),
+                      cv_bbopti = sd_bbopti/mean(BBopti, na.rm=TRUE),
+                      sd_cond = sd(Cond, na.rm=TRUE),
+                      Qrange = max(PARin) - min(PARin),
+                      Drange = max(VPD) - min(VPD),
+                      Trange = max(Tleaf) - min(Tleaf)), 
+            by="fitgroup")
+
+fits2 <- nlsList(Cond ~ g0 + 1.6*(1 + g1/sqrt(VPD))*(Photo / CO2S) | fitgroup, 
+                 start=list(g0=0, g1=3),
+                 data=lin2015a, na.action=na.omit)
+p2 <- as.data.frame(coef(fits2)) %>% 
+  rename(g0f = g0, g1f = g1) %>%
+  mutate(rmsef = sapply(fits2, function(x)summary(x)$sigma))
+p <- cbind(p, p2)
+
+fits3 <- nlsList(Cond ~ 1.6*(1 + g1/sqrt(VPD))*(Photo / CO2S) | fitgroup, 
+                 start=list(g1=3),
+                 data=lin2015a, na.action=na.omit)
+p3 <- as.data.frame(coef(fits3)) %>% 
+  rename(g1n = g1) %>%
+  mutate(rmsen = sapply(fits3, function(x)summary(x)$sigma))
+p <- cbind(p, p3)
+
+
+
+hist(p$g0, breaks=30)
+
+
+
+
+par(mfrow=c(1,2))
[…]

### 2017-03-10 R/load.R

@@ -1,0 +1,25 @@
+
+pacman::p_load(dplyr, tidyr, nlme, nlshelper, tibble)
+
+
+
+if(!dir.exists("download"))dir.create("download")
+
+linfn <- "download/lin2015.csv"
+if(!file.exists(linfn)){
+  download.file("https://ndownloader.figshare.com/files/1886204", linfn)
+}
+
+lin2015 <- read.csv(linfn, stringsAsFactors = FALSE) %>%
+  mutate(fitgroup = paste(Datacontrib, Species, sep="_"),
+         Ci = CO2S - Photo/(Cond/1.6),
+         CiCa = Ci / CO2S,
+         BBopti = Photo/(CO2S*sqrt(VPD)))
+
+lin2015a <- group_by(lin2015, fitgroup) %>%
+  filter(n() > 15,
+         PARin < 2500,
+         PARin >= 0,
+         Datacontrib != "Derek Eamus",
+         fitgroup != "Jon Bennie_Zornia")
+

### 2017-03-10 R/martinpaul.R

@@ -1,0 +1,37 @@
+
+source("R/load.R")
+
+library(car)
+
+x <- subset(lin2015, fitgroup == "Nicolas Martin-StPaul_Quercus ilex_StPaul_Puechabon")
+fit <- lm(Cond ~ BBopti, data=x)
+
+el <- confidenceEllipse(fit, draw=FALSE)
+b <- bootCase(fit)
+
+
+windows(8,4)
+par(mfrow=c(1,2), mar=c(4,4,1,1), mgp=c(2.5,0.5,0), tcl=0.2, las=1,
+    cex.lab=1.1, cex.axis=0.9)
+with(x, plot(BBopti, Cond, xlim=c(0, max(BBopti)), ylim=c(0, 0.15),
+             panel.first=add_regres_line(fit),
+             xlab=expression(A/(C[a]*sqrt(D))),
+             ylab=expression(g[s]~(mol~m^-2~s^-1)),
+             pch=16, cex=0.8, col="dimgrey"))
+box()
+
+plot(el, type='l', lty=3, 
+     xlab=expression(g[0]), ylab=expression(g[1]),
+     ylim=c(3,5))
+points(b, pch=16, cex=0.3, col="dimgrey")
+points(coef(fit)[1],coef(fit)[2], pch=19)
+
+
+
+
+confint(lm(Cond ~ BBopti, data=x))
+
+
+
+
+

### 2017-03-10 R/miner.R

@@ -1,0 +1,80 @@
+
+
+url <- "http://onlinelibrary.wiley.com/doi/10.1111/pce.12871/full"
+
+library(rvest)
+library(stringr)
+
+tab <- read_html(url, encoding="latin1")
+
+tab <- html_nodes(tab, "table")
+
+table1 <- html_table(tab[1])
+
+df <- as.data.frame(table1[[1]][-1,]) 
+names(df) <- c("species","growthcond","model","m","g0","method","R2","ref")
+
+df$m <- gsub("\\(.+\\)","", df$m)
+
+df$m <- sapply(strsplit(df$m, "±"), "[", 1)
+
+df$m <- gsub(" a", "", df$m)
+df$m <- gsub("a", "", df$m)
+df$m <- gsub("\U00002009", "-", df$m)
+df$m <- gsub("≈", "", df$m)
+df$m <- gsub(" to ", "-", df$m)
+
+df$m <- str_trim(df$m)
+
+df$m <- gsub("[^\\.0-9]", "--", df$m)
+df$m <- gsub("^--", "", df$m)
+df$m <- gsub("--$", "", df$m)
+
+l <- strsplit(df$m, "--", fixed=TRUE)
+
+df$m <- sapply(l, function(x){
+  x <- as.numeric(x)
+  if(length(x) == 1)
+    return(x)
+  else
+    return((x[2] + x[1])/2)
+})
+df$m[which(df$m > 100)] <- df$m[which(df$m > 100)]/10
+
+df <- subset(df, model == "BB")
+
+g0 <- df$g0
+
+g0 <- gsub("\\(.+\\)","", g0)
+
+g0 <- gsub(" a", "", g0)
+g0 <- gsub("a", "", g0)
+g0 <- str_trim(g0)
+g0 <- gsub("[^\\.0-9]", "--", g0)
+l <- strsplit(g0, "\U00002009")
+g0 <- sapply(l, "[", 1)
+
+
+ii <- c(1,5,107,179,181) 
+g0l <- strsplit(g0[ii], "[^\\.0-9]")
+
+g0l <- sapply(g0l, function(x){
+  x <- as.numeric(x)
+  if(length(x) == 1)
+    return(x)
+  else
+    return((x[3] + x[1])/2)
+})
+g0[ii] <- g0l
+
+# the ones that don't convert; we don't want those!
+df$g0 <- as.numeric(g0)
[…]

### 2017-03-14 R/analysis_lin2015.R
figures tidies; some new references; gmin analysis kerstiens etc

@@ -1,82 +1,0 @@
-
-
-fits <- lmList(Cond ~ BBopti | fitgroup, data=lin2015a, na.action=na.omit)
-p <- as.data.frame(coef(fits)) %>% 
-  rownames_to_column(var = "fitgroup") %>%  
-  rename(g0 = `(Intercept)`, g1 = BBopti) %>% 
-  mutate(R2 = sapply(fits, function(x)summary(x)$adj.r.squared),
-         rmse = sapply(fits, function(x)summary(x)$sigma),
-         cor = sapply(fits, function(x)summary(x, correlation=T)$correlation[2,1]),
-         g0_lci = sapply(fits, function(x)confint(x)[1,1]),
-         g0_uci = sapply(fits, function(x)confint(x)[1,2])) %>% 
-  left_join(summarize(lin2015a, 
-                      sd_bbopti = sd(BBopti, na.rm=TRUE),
-                      cv_bbopti = sd_bbopti/mean(BBopti, na.rm=TRUE),
-                      sd_cond = sd(Cond, na.rm=TRUE),
-                      Qrange = max(PARin) - min(PARin),
-                      Drange = max(VPD) - min(VPD),
-                      Trange = max(Tleaf) - min(Tleaf)), 
-            by="fitgroup")
-
-fits2 <- nlsList(Cond ~ g0 + 1.6*(1 + g1/sqrt(VPD))*(Photo / CO2S) | fitgroup, 
-                 start=list(g0=0, g1=3),
-                 data=lin2015a, na.action=na.omit)
-p2 <- as.data.frame(coef(fits2)) %>% 
-  rename(g0f = g0, g1f = g1) %>%
-  mutate(rmsef = sapply(fits2, function(x)summary(x)$sigma))
-p <- cbind(p, p2)
-
-fits3 <- nlsList(Cond ~ 1.6*(1 + g1/sqrt(VPD))*(Photo / CO2S) | fitgroup, 
-                 start=list(g1=3),
-                 data=lin2015a, na.action=na.omit)
-p3 <- as.data.frame(coef(fits3)) %>% 
-  rename(g1n = g1) %>%
-  mutate(rmsen = sapply(fits3, function(x)summary(x)$sigma))
-p <- cbind(p, p3)
-
-
-
-hist(p$g0, breaks=30)
-
-
-
-
-par(mfrow=c(1,2))
[…]

### 2017-03-14 R/figures.R

@@ -1,0 +1,106 @@
+
+
+figure1 <- function(gdfr){
+  
+  par(mar=c(3,5,1,0.5), cex.lab=1.2)
+  fit <- lm(log10(gmin) ~ method-1, data=gdfr)
+  cis <- confint(fit)
+  quans <- sapply(split(gdfr, gdfr$method), function(x)quantile(log10(x$gmin), probs=c(0.05, 0.95)))
+  gmins <- with(gdfr, tapply(log10(gmin), method, mean, na.rm=TRUE))
+  
+  plot(1:5, gmins, pch=19, cex=1.2, axes=FALSE, xlab="",
+       panel.first=segments(x0=1:5, x1=1:5, y0=quans[1,], y1=quans[2,], col="grey"),
+       ylim=c(0,2),
+       xlim=c(0.5,5.5),
+       ylab=expression(Conductance~~(mmol~m^-2~s^-1)))
+  axis(1, at=1:5, labels=c(expression(g["cut,isol"]), 
+                           expression(g["min,seal"]), 
+                           expression(g[min]), 
+                           expression(g[dark]),
+                           expression(g[0])))
+  magaxis(side=2, unlog=2)
+  box()
+  arrows(x0=1:5, x1=1:5, y0=cis[,1], y1=cis[,2], angle=90, length=0.1, code=3)
+  
+  
+}
+
+
+figure2 <- function(lin2015){
+  
+  x <- subset(lin2015, fitgroup == "Nicolas Martin-StPaul_Quercus ilex_StPaul_Puechabon")
+  fit <- lm(Cond ~ BBopti, data=x)
+  
+  el <- confidenceEllipse(fit, draw=FALSE)
+  set.seed(123)
+  b <- bootCase(fit)
+  
+  par(mfrow=c(1,2), mar=c(4,4,1,1), mgp=c(2.5,0.5,0), tcl=0.2, las=1,
+      cex.lab=1.1, cex.axis=0.9)
+  with(x, plot(BBopti, Cond, ylim=c(0, 0.15),
+               xlim=c(0,0.04),
+               panel.first=add_regres_line(fit),
+               xlab=expression(A/(C[a]*sqrt(D))),
+               ylab=expression(g[s]~(mol~m^-2~s^-1)),
+               pch=16, cex=0.8, col="dimgrey"))
+  box()
+  
+  plot(el, type='l', lty=3, 
[…]

### 2017-03-14 R/functions.R

@@ -1,0 +1,52 @@
+to.pdf <- function(expr, filename, ..., verbose=TRUE) {
+  if(!file.exists(dirname(filename)))
+    dir.create(dirname(filename), recursive=TRUE)
+  if ( verbose )
+    cat(sprintf("Creating %s\n", filename))
+  pdf(filename, ...)
+  on.exit(dev.off())
+  eval.parent(substitute(expr))
+}
+
+
+
+fits_lin2015 <- function(lin2015a){
+  
+  fits <- lmList(Cond ~ BBopti | fitgroup, data=lin2015a, na.action=na.omit)
+  p <- as.data.frame(coef(fits)) %>% 
+    rownames_to_column(var = "fitgroup") %>%  
+    rename(g0 = `(Intercept)`, g1 = BBopti) %>% 
+    mutate(R2 = sapply(fits, function(x)summary(x)$adj.r.squared),
+           rmse = sapply(fits, function(x)summary(x)$sigma),
+           cor = sapply(fits, function(x)summary(x, correlation=T)$correlation[2,1]),
+           g0_lci = sapply(fits, function(x)confint(x)[1,1]),
+           g0_uci = sapply(fits, function(x)confint(x)[1,2])) %>% 
+    left_join(summarize(lin2015a, 
+                        sd_bbopti = sd(BBopti, na.rm=TRUE),
+                        cv_bbopti = sd_bbopti/mean(BBopti, na.rm=TRUE),
+                        sd_cond = sd(Cond, na.rm=TRUE),
+                        Qrange = max(PARin) - min(PARin),
+                        Drange = max(VPD) - min(VPD),
+                        Trange = max(Tleaf) - min(Tleaf)), 
+              by="fitgroup")
+  
+  fits2 <- nlsList(Cond ~ g0 + 1.6*(1 + g1/sqrt(VPD))*(Photo / CO2S) | fitgroup, 
+                   start=list(g0=0, g1=3),
+                   data=lin2015a, na.action=na.omit)
+  p2 <- as.data.frame(coef(fits2)) %>% 
+    rename(g0f = g0, g1f = g1) %>%
+    mutate(rmsef = sapply(fits2, function(x)summary(x)$sigma))
+  p <- cbind(p, p2)
+  
+  fits3 <- nlsList(Cond ~ 1.6*(1 + g1/sqrt(VPD))*(Photo / CO2S) | fitgroup, 
+                   start=list(g1=3),
[…]

### 2017-03-14 R/load.R

@@ -1,10 +1,14 @@
 
-pacman::p_load(dplyr, tidyr, nlme, nlshelper, tibble)
+pacman::p_load(car, dplyr, tidyr, nlme, nlshelper, tibble, magicaxis)
 
 
 
 if(!dir.exists("download"))dir.create("download")
+if(!dir.exists("output"))dir.create("output")
+source("R/functions.R")
+source("R/figures.R")
 
+# Lin et al. 2015
 linfn <- "download/lin2015.csv"
 if(!file.exists(linfn)){
   download.file("https://ndownloader.figshare.com/files/1886204", linfn)
@@ -20,6 +24,41 @@ lin2015a <- group_by(lin2015, fitgroup) %>%
   filter(n() > 15,
          PARin < 2500,
          PARin >= 0,
-         Datacontrib != "Derek Eamus",
+         Datacontrib != "Derek Eamus",  # sorry derek you are an outlier
          fitgroup != "Jon Bennie_Zornia")
+
+# Lin 2015 fits
+lin2015coef <- fits_lin2015(lin2015a)
+
+
+# Kerstiens 1996
+kerst <- read.csv("data/kerstiens1996_table1.csv", stringsAsFactors = FALSE) %>%
+  filter(gmin > 0,
+         method %in% c("A","C","D")) %>%
+  mutate(method = dplyr::recode(method, A="gcut_isol", C="gcut_seal", D="gmin"),
+         gmin = 10^5 * 10^-3 * gmin / 40)   # convert to mmol m-2 s-1 assume 40 mol m-3
+
+# Miner et al. 2016
+miner <- read.csv("data/Miner_table1.csv")
+
+
+# Lombardozzi et al 2017
+lombar <- read.csv("data/lombardozzi_gnight.csv") %>%
+  rename(gmin = gnight) %>%
+  filter(gmin > 0) %>%
+  mutate(method="gnight")
+
+
+g0s <- data.frame(gmin=1000 * c(lin2015coef$g0, miner$g0),
+                  method="g0", stringsAsFactors = FALSE) %>%
+  filter(!is.na(gmin),
+         gmin > 0)
+
+gdfr <- bind_rows(kerst, lombar, g0s) %>%
+  mutate(method = factor(method, levels=c("gcut_isol","gcut_seal","gmin","gnight", "g0")))
+
+
+
+
+
 

### 2017-03-14 R/lombardozzi.R

@@ -1,0 +1,25 @@
+fn <- "C:\\Users\\30022860\\Google Drive\\Work\\Projects\\WUE Project\\Various\\g0\\Supplementary Table 1 v4_FinalForRevision_MZ.pdf"
+
+#devtools::install_github("ropensci/tabulizerjars", quick=T)
+#devtools::install_github("ropensci/tabulizer", quick=T)
+
+
+library(tabulizer)
+out <- extract_tables(fn)
+
+# list, each page is a node
+
+tab <- do.call(c, lapply(out, function(x)x[,3]))
+
+tab <- tab[tab != ""]
+
+gnight <- as.numeric(tab)
+gnight <- gnight[!is.na(gnight)]
+
+write.csv(data.frame(gnight=gnight), "data/lombardozzi_gnight.csv", row.names=FALSE)
+
+hist(log10(gnight), axes=FALSE)
+axis(2)
+library(magicaxis)
+magaxis(side=1, unlog=1)
+

### 2017-03-14 R/make_figures.R

@@ -1,0 +1,5 @@
+
+to.pdf(figure1(gdfr), "output/figure1.pdf", height=5, width=4)
+to.pdf(figure2(lin2015), "output/figure2.pdf", height=4, width=8)
+to.pdf(figure3(lin2015coef, miner), "output/figure3.pdf", height=4, width=8)
+to.pdf(figure4(lin2015), "output/figure4.pdf", height=5, width=5)

### 2017-03-14 R/martinpaul.R

@@ -1,37 +1,0 @@
-
-source("R/load.R")
-
-library(car)
-
-x <- subset(lin2015, fitgroup == "Nicolas Martin-StPaul_Quercus ilex_StPaul_Puechabon")
-fit <- lm(Cond ~ BBopti, data=x)
-
-el <- confidenceEllipse(fit, draw=FALSE)
-b <- bootCase(fit)
-
-
-windows(8,4)
-par(mfrow=c(1,2), mar=c(4,4,1,1), mgp=c(2.5,0.5,0), tcl=0.2, las=1,
-    cex.lab=1.1, cex.axis=0.9)
-with(x, plot(BBopti, Cond, xlim=c(0, max(BBopti)), ylim=c(0, 0.15),
-             panel.first=add_regres_line(fit),
-             xlab=expression(A/(C[a]*sqrt(D))),
-             ylab=expression(g[s]~(mol~m^-2~s^-1)),
-             pch=16, cex=0.8, col="dimgrey"))
-box()
-
-plot(el, type='l', lty=3, 
-     xlab=expression(g[0]), ylab=expression(g[1]),
-     ylim=c(3,5))
-points(b, pch=16, cex=0.3, col="dimgrey")
-points(coef(fit)[1],coef(fit)[2], pch=19)
-
-
-
-
-confint(lm(Cond ~ BBopti, data=x))
-
-
-
-
-

### 2017-03-14 R/miner.R

@@ -72,6 +72,8 @@ df$g0 <- as.numeric(g0)
 df$R2 <- as.numeric(df$R2)
 
 
+write.csv(df, "data/Miner_table1.csv", row.names=FALSE)
+
 with(df, plot(R2, g0))
 l <- loess(g0 ~ R2, data=df)
 plot_loess(l, add=TRUE)

### 2017-03-14 manuscript.Rmd
figures tidies; some new references; gmin analysis kerstiens etc
added simulate figure with and without g0

@@
 
-@barnard2013 makes the same assumption
+@barnard2013 makes the same assumption, without any discussion of the possibility that daytime g0 might be different from nighttime conductance
 
@@
 - @schreiber1996 cuticle review
+
+[@medlyn_reconciling_2011] does not develop theory for g0, instead adds it as a minimum conductancce because that's how it is!
+
+When fitting a Ball-Berry type model, negative values are quite common.
+A typical approach is to avoid this by setting some value based on guesswork [e.g.
+@leuning_modelling_1990], which we assume is quite common when inspecting the many rounded values in the review by
+
+The first Ball-Berry model did not have a g0
 
@@
 
+[@leuning_critical-appraisal_1995]
+
 Various measures of minimum conductance
@@
 g0?
+
+@rescodedios2016 early night gs much (much) lower than pre-dawn gs, in E. camaldulensis
+@rescodedios2013 nighttime fluctuation gs independent of environment
 

### 2017-03-14 R/figures.R

@@ -104,3 +104,52 @@ figure4 <- function(lin2015){
 }
 
 
+figure5 <- function(){
+  
+  par(mfrow=c(2,2), mar=c(4,4,1,1), mgp=c(2.5,0.5,0), tcl=0.2, las=1,
+      cex.axis=0.9, cex.lab=1.1)
+  
+  g1 <- 4
+  g0 <- 0.03
+  
+  lty0 <- 1
+  lty1 <- 5
+  col0 <- "black"
+  col1 <- "black"
+  
+  r0 <- Photosyn(PPFD=seq(20,1000, length=101),
+                 g1=g1, g0=0, Vcmax=70, Jmax=140,
+                 VPD=2, Tleaf=20, Ca=400)
+  r1 <- Photosyn(PPFD=seq(20,1000, length=101),
+                 g1=g1, g0=g0, Vcmax=70, Jmax=140,
+                 VPD=2, Tleaf=20, Ca=400)
+  
+  with(r0, plot(PPFD, ALEAF/GS, type='l', ylim=c(0,80), col=col0, lty=lty0))
+  with(r1, lines(PPFD, ALEAF/GS, col=col1, lty=lty1))
+  
+  with(r0, plot(PPFD, Ci, type='l',
+                col=col0, lty=lty0,
+                ylim=c(250,400)))
+  with(r1, lines(PPFD, Ci, col=col1, lty=lty1))
+  
+  tleafs <- seq(10,45,length=101)
+  vpds <- 0.000605 * tleafs^2.39
+  r0 <- Photosyn(PPFD=1500,
+                 g1=g1, g0=0, Vcmax=70, Jmax=140,
+                 VPD=vpds, Tleaf=tleafs, Ca=400)
+  r1 <- Photosyn(PPFD=1500,
+                 g1=g1, g0=g0, Vcmax=70, Jmax=140,
+                 VPD=vpds, Tleaf=tleafs, Ca=400)
+  
+  with(r0, plot(Tleaf, ELEAF, type='l', 
+                col=col0, lty=lty0,
+                ylim=c(0,10)))
+  with(r1, lines(Tleaf, ELEAF, col=col1, lty=lty1))
+  
+  with(r0, plot(VPD, Ci, type='l', 
+                col=col0, lty=lty0,
+                ylim=c(0,400)))
+  with(r1, lines(VPD, Ci, col=col1, lty=lty1))
+  
+}
+

### 2017-03-14 R/make_figures.R

@@ -3,3 +3,4 @@ to.pdf(figure1(gdfr), "output/figure1.pdf", height=5, width=4)
 to.pdf(figure2(lin2015), "output/figure2.pdf", height=4, width=8)
 to.pdf(figure3(lin2015coef, miner), "output/figure3.pdf", height=4, width=8)
 to.pdf(figure4(lin2015), "output/figure4.pdf", height=5, width=5)
+to.pdf(figure5(), "output/figure5.pdf", height=6, width=6)

### 2017-03-15 manuscript.Rmd
Added WTC4 figure; manuscript organized a bit with headers; figures inserted; html output option

@@
 output:
+  html_document:
+    theme: flatly
   word_document:
@@
 
+```{r echo=FALSE}
+includefigs <- TRUE
+library(png)
+library(grid)
+
+plotfig <- function(...)grid.raster(readPNG(...))
+```
+
 # Abstract
@@
 
-1. Importance of stomatal conductance. Prevalence of the Ball-Berry model, and g0. What is g0? Point of this paper.
+Importance of stomatal conductance. Prevalence of the Ball-Berry model, and g0. What is g0? Point of this paper.
 
-2. Nighttime conductance not necessarily same as minimum conductance
+# Do we need g~0~ in simulations of water and carbon fluxes?
+
+The first Ball-Berry model did not have a g0
+
+[@medlyn_reconciling_2011] does not develop theory for g0, instead adds it as a minimum conductancce because that's how it is!
+
+g~0~ modifies not just conductance in zero light, but also low light and high VPD/ high T.
+It is commonly observed and makes sense that Ci converges to Ca at zero light, but the coupled leaf gas exchange model does not predict that unless a g0 is included [@leuning1995].
+
+similarly at high T. See also Kala. Figure 1.
 
@@
 
-@barbour2007 considerable stomatal control during the night; mirrors daytime
+@messinger2006 Conclude that Ci response of gs is the same in the dark or when DCMU was applied (which disrupts electron transport).
+However, Fig5A-B indicate that gs was much higher in the dark than when DCMU was applied (although the ‘response’ was similar).
+This indicates gs response to low PAR (darkness) is more than just responding to what photosynthesis does.
 
-@howard2007 nighttime conductance lower in drought. Nighttime 2-5x higher than cuticular.
+# g~0~ is poorly estimated from regression
 
-We are not reviewing nighttime conductance here, that has been done extensively by others [@caird2007]
-
-2b. Variation in g0
-
-@misson2004 g0 decreases with pre-dawn water potential
-
-3. How do we estimate g0 from gas exchange?
-
-Poorly, that's how! Problem of correlation between g1 and g0; reason why compilations like Lin et al. set g0 to zero.
-Problem that g0 is frequently negative (e.g. @heroult2013)
-
-4. Components of minimum conductance
-
-- Stomata simply do not close all the way, might stay open regardless of photosynthesis, or are broken.
-
-5. Cuticular conductance
-
-- Kerstiens
-- H2O not CO2? [@boyer1997]
-- gmin as a plastic trait. Temperature response generally unresolved [@kerstiens2006]
-- @riederer2001 correlation between temperature and gcut (non-stomatal cuticles)
-- @burghardt2003 gmin review
-- @schreiber1996 cuticle review
-
-[@medlyn_reconciling_2011] does not develop theory for g0, instead adds it as a minimum conductancce because that's how it is!
+Barnard and Bauerle said so too but gave no details.
+Even in very good datasets, g0 and g1 are strongly correlated (Figure 2).
+Across datasets, g0 increases when the model fits poorly (Figure 3), and the width of the confidence interval increases when less variation in a ball-berry coefficient.
+Often g0 is even negative [@heroult2013; @leuning1990], so it is often set to some arbitrary value
 
@@
 
-The first Ball-Berry model did not have a g0
+# Components of g0
+
+- gcut (cuticle without stomata)
+- gmin (minimum conductance in situ, stomata closed)
+- g0 (minimum conductance in the Ball-Berry-type model)
+- gdark (nighttime conductance)
+
+gdark > gmin > gcut
+g0?
+
+@egea2011 Assumes g0 is constant, does not vary with water stress, and that it represents a ‘cuticular conductance’.
 
@@
 
+## nighttime
+Nighttime conductance not necessarily same as minimum conductance
+We are not reviewing nighttime conductance here, that has been done extensively by others [@caird2007]
+
+gnight not constant 
+@barbour2007 considerable stomatal control during the night; mirrors daytime
+@howard2007 nighttime conductance lower in drought. Nighttime 2-5x higher than cuticular.
+
+@misson2004 g0 decreases with pre-dawn water potential
+
+Stomata simply do not close all the way, might stay open regardless of photosynthesis, or are broken.
+
+@rescodedios2016 early night gs much (much) lower than pre-dawn gs, in E. camaldulensis
+@rescodedios2013 nighttime fluctuation gs independent of environment
+
+@caird2007 For many species, gnight is not stable throughout the night period.
+Endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [lots of refs]
+
+@korner1994 table of minimum conductance. Also points out (without reference) that g0 in droughted leaves is certainly lower than g0 in dark or very high CO2 leaves.
+
+## minimum conductance
+
+true cuticular
+MLD method (Figure 6) gives higher values but this is probably more useful for models (since these are whole leaves)
+
 @santrucek Direct measurements showing much higher 'cuticular' conductance of the leaf side with stomata, though stomata were closed.
@@
 non-stomatal cuticle.
-
-@messinger2006 Conclude that Ci response of gs is the same in the dark or when DCMU was applied (which disrupts electron transport).
-However, Fig5A-B indicate that gs was much higher in the dark than when DCMU was applied (although the ‘response’ was similar).
-This indicates gs response to low PAR (darkness) is more than just responding to what photosynthesis does.
 
@@
 
-@leuning1995 discusses g0
+conductance: cuticle + closed stomata + cracks, broken stomata
 
-@caird2007 For many species, gnight is not stable throughout the night period.
-Endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [lots of refs]
-
-@korner1994 table of minimum conductance. Also points out (without reference) that g0 in droughted leaves is certainly lower than g0 in dark or very high CO2 leaves.
-
-@egea2011 Assumes g0 is constant, does not vary with water stress, and that it represents a ‘cuticular conductance’.
+- Kerstiens
+@boyer1997 shows that water but not CO~2~ diffuses across non-stomatal cuticles.
 
@@
 
-[@leuning_critical-appraisal_1995]
+- gmin as a plastic trait. Temperature response generally unresolved [@kerstiens2006]
+- @riederer2001 correlation between temperature and gcut (non-stomatal cuticles)
+- @burghardt2003 gmin review
+- @schreiber1996 cuticle review
 
-Various measures of minimum conductance
-
-- gcut (cuticle without stomata)
-- gmin (minimum conductance in situ, stomata closed)
-- g0 (minimum conductance in the Ball-Berry-type model)
-- gdark (nighttime conductance)
-
-gdark > gmin > gcut
-g0?
-
-@rescodedios2016 early night gs much (much) lower than pre-dawn gs, in E. camaldulensis
-@rescodedios2013 nighttime fluctuation gs independent of environment
+# Future directions
 
@@
 
-1. Ball-Berry relation, showing g0 on intercept. Second panel correlation ellipse g0 ~ g1 for one study. Third panel g0 vs. R2 for the Lin dataset.
+```{r fig.height=6,echo=FALSE, eval=includefigs}
+plotfig("output/figure1.png")
+```
 
-2. DeKauwe2015 figure A2 redone (minimum photosynthesis, minimum conductance)
+**Figure 1**.
+Simulations with the plantecophys R package, with and without a g~0~.
+For the bottom two panels, VPD and T~air~ were covaried based on an empirical relationship, reflecting the fact that they covary in field conditions.
+For all simulations it is assumed that T~leaf~ is equal to T~air~.
+Parameters were, etc.
 
-3. Kerstiens data: differences between methods, connection to climate data (speciesmap)
+```{r fig.height=4,echo=FALSE, eval=includefigs}
+plotfig("output/figure2.png")
+```
 
-4. Blackman data: cuticular conductance is a plastic trait
+**Figure 2**.
+Example linear regression of stomatal conductance (g~s~) against a stomatal index, here the linearized version of the model of Medlyn et al.
+(2011).
+Example data is taken from Lin et al.
+(2015), chosen is a dataset with a very good fit as an example.
+Solid line is regression line, shaded area a 95% confidence interval.
+Right panel, visualized correlation between estimated slope (g~1~) and intercept (g~0~) of the regression.
+Dotted ellipse is a bivariate 95% confidence interval (using confidenceEllipse from the car R package), symbols 1000 bootstrap samples of the coefficients (using bootCase in the car R package).
+
+```{r fig.height=4,echo=FALSE, eval=includefigs}
+plotfig("output/figure3.png")
+```
+
+**Figure 3**.
+
+```{r fig.width=4,echo=FALSE, eval=includefigs}
+plotfig("output/figure4.png")
+```
+
+**Figure 4**.
+
+```{r fig.height=5, fig.width=5, echo=FALSE, eval=includefigs}
+plotfig("output/figure5.png")
+```
+
+**Figure 5**.
+
+```{r fig.height=4,echo=FALSE, eval=includefigs}
+plotfig("output/figure6.png")
+```
+
+**Figure 6**.
 

### 2017-03-15 R/figures.R

@@ -1,6 +1,6 @@
 
 
-figure1 <- function(gdfr){
+figure_gmin_review <- function(gdfr){
   
   par(mar=c(3,5,1,0.5), cex.lab=1.2)
   fit <- lm(log10(gmin) ~ method-1, data=gdfr)
@@ -26,14 +26,20 @@ figure1 <- function(gdfr){
 }
 
 
-figure2 <- function(lin2015){
+figure_g0g1_cor <- function(lin2015){
   
   x <- subset(lin2015, fitgroup == "Nicolas Martin-StPaul_Quercus ilex_StPaul_Puechabon")
+  
+  # for bootCase to work
+  assign("x",x,envir=.GlobalEnv)
+  
   fit <- lm(Cond ~ BBopti, data=x)
   
   el <- confidenceEllipse(fit, draw=FALSE)
   set.seed(123)
+  
   b <- bootCase(fit)
+  rm(x, envir=.GlobalEnv)
   
   par(mfrow=c(1,2), mar=c(4,4,1,1), mgp=c(2.5,0.5,0), tcl=0.2, las=1,
       cex.lab=1.1, cex.axis=0.9)
@@ -57,7 +63,7 @@ figure2 <- function(lin2015){
 
 
 
-figure3 <- function(lin2015coef, miner){
+figure_R2g0 <- function(lin2015coef, miner){
   
   par(mfrow=c(1,2), mar=c(5,5,1,1), cex.axis=0.9)
   with(lin2015coef, {
@@ -83,7 +89,7 @@ figure3 <- function(lin2015coef, miner){
 }
 
 
-figure4 <- function(lin2015){
+figure_amings <- function(lin2015){
   
   par(mar=c(5,5,1,1), cex.axis=0.9, cex.lab=1.1)
   minags <- group_by(lin2015, fitgroup) %>%
@@ -104,7 +110,7 @@ figure4 <- function(lin2015){
 }
 
 
-figure5 <- function(){
+figure_sim <- function(){
   
   par(mfrow=c(2,2), mar=c(4,4,1,1), mgp=c(2.5,0.5,0), tcl=0.2, las=1,
       cex.axis=0.9, cex.lab=1.1)
@@ -124,10 +130,18 @@ figure5 <- function(){
                  g1=g1, g0=g0, Vcmax=70, Jmax=140,
                  VPD=2, Tleaf=20, Ca=400)
   
-  with(r0, plot(PPFD, ALEAF/GS, type='l', ylim=c(0,80), col=col0, lty=lty0))
[…]

### 2017-03-15 R/functions.R

@@ -50,3 +50,34 @@ fits_lin2015 <- function(lin2015a){
 }
 
 
+# Simple function for placing labels on a figure.
+plotlabel <- function(txt, where, inset=0.08, inset.x=inset, inset.y=inset,...){
+  u <- par()$usr
+  if(grepl("left",where))x <- u[1] + inset.x*(u[2]-u[1])
+  if(grepl("right",where))x <- u[2] - inset.x*(u[2]-u[1])
+  if(grepl("bottom",where))y <- u[3] + inset.y*(u[4]-u[3])
+  if(grepl("top",where))y <- u[4] - inset.y*(u[4]-u[3])
+  
+  text(x,y,txt,font=2,...)
+}
+
+
+convert_all_pdf <- function(path, ...){
+  o <- getwd()
+  on.exit(setwd(o))
+  setwd(path)
+  pdfs <- dir(pattern="[.]pdf")
+  for(i in seq_along(pdfs)){
+    convert_pdf_png(pdfs[i])
+    message(i)
+  }
+}
+
+convert_pdf_png <- function(filename, fnout=NULL, overwrite=TRUE, res=600, resize=100){
+  if(is.null(fnout))fnout <- gsub("[.]pdf",".png",filename)
+  if(!file.exists(fnout) | overwrite){
+    cmd <- sprintf("convert -density %s -resize %s%% -colorspace CMYK %s %s",res, resize, filename,fnout) 
+    shell(cmd)
+  }
+}
+

### 2017-03-15 R/load.R

@@ -1,5 +1,5 @@
 
-pacman::p_load(car, dplyr, tidyr, nlme, nlshelper, tibble, magicaxis)
+pacman::p_load(car, dplyr, tidyr, nlme, nlshelper, tibble, magicaxis, plantecophys)
 
 
 
@@ -58,6 +58,12 @@ gdfr <- bind_rows(kerst, lombar, g0s) %>%
   mutate(method = factor(method, levels=c("gcut_isol","gcut_seal","gmin","gnight", "g0")))
 
 
+# Blackman, WTC4
+wtc4gmin <- read.csv("data/wtc4_gmin_detached.csv") %>%
+  rename(gmin = gmin_mmol_m2_s) %>%
+  mutate(ch_temp_fac = as.factor(ch_temp))
+
+wtc4gdark <- read.csv("data/wtc4_gnight.csv")
 
 
 

### 2017-03-15 R/make_figures.R

@@ -1,6 +1,13 @@
 
-to.pdf(figure1(gdfr), "output/figure1.pdf", height=5, width=4)
-to.pdf(figure2(lin2015), "output/figure2.pdf", height=4, width=8)
-to.pdf(figure3(lin2015coef, miner), "output/figure3.pdf", height=4, width=8)
-to.pdf(figure4(lin2015), "output/figure4.pdf", height=5, width=5)
-to.pdf(figure5(), "output/figure5.pdf", height=6, width=6)
+to.pdf(figure_sim(), "output/figure1.pdf", height=6, width=6)
+to.pdf(figure_g0g1_cor(lin2015), "output/figure2.pdf", height=4, width=8)
+to.pdf(figure_R2g0(lin2015coef, miner), "output/figure3.pdf", height=4, width=8)
+
+to.pdf(figure_gmin_review(gdfr), "output/figure4.pdf", height=5, width=4)
+to.pdf(figure_amings(lin2015), "output/figure5.pdf", height=5, width=5)
+
+to.pdf(figure_wtc4_gmin(wtc4gmin, wtc4gdark), "output/figure6.pdf", height=4, width=8)
+
+
+convert_all_pdf("output")
+

### 2017-03-15 manuscript.Rmd
kerstiens has species; load TRY category dataset

@@
 Across datasets, g0 increases when the model fits poorly (Figure 3), and the width of the confidence interval increases when less variation in a ball-berry coefficient.
-Often g0 is even negative [@heroult2013; @leuning1990], so it is often set to some arbitrary value
+Often g0 is even negative [@heroult2013; @leuning_modelling_1990], so it is often set to some arbitrary value
 
@@
 
+Hygen 1951 early work showing MLD method
+
 true cuticular
@@
 
-@santrucek Direct measurements showing much higher 'cuticular' conductance of the leaf side with stomata, though stomata were closed.
+santrucek Direct measurements showing much higher 'cuticular' conductance of the leaf side with stomata, though stomata were closed.
 In agreement with @kerstiens1996 who showed 10x higher gmin for stomatal side vs.

### 2017-03-15 R/load.R

@@ -1,5 +1,5 @@
 
-pacman::p_load(car, dplyr, tidyr, nlme, nlshelper, tibble, magicaxis, plantecophys)
+pacman::p_load(car, dplyr, tidyr, nlme, nlshelper, tibble, magicaxis, plantecophys, readxl)
 
 
 
@@ -7,6 +7,21 @@ if(!dir.exists("download"))dir.create("download")
 if(!dir.exists("output"))dir.create("output")
 source("R/functions.R")
 source("R/figures.R")
+
+
+# You must find TRY categorical traits yourself and place it in /data
+# I cannot share it here (registration required)
+if(!file.exists("data/trydb.rds")){
+  tryfile <- "data/TRY_Categorical_Traits_Lookup_Table_2012_03_17_TestRelease.xlsx"
+  if(!file.exists(tryfile))stop("Place the TRY look up table in data/.")
+  trydb <- read_excel(tryfile)
+  trydb <- trydb[,c("AccSpeciesName","PhylogeneticGroup","PlantGrowthForm","LeafType","LeafPhenology")]
+  saveRDS(trydb, "data/trydb.rds")
+} else {
+  trydb <- readRDS("data/trydb.rds")
+}
+
+
 
 # Lin et al. 2015
 linfn <- "download/lin2015.csv"
@@ -58,12 +73,16 @@ gdfr <- bind_rows(kerst, lombar, g0s) %>%
   mutate(method = factor(method, levels=c("gcut_isol","gcut_seal","gmin","gnight", "g0")))
 
 
+kerst2 <- inner_join(kerst, trydb, by=c("species" = "AccSpeciesName"))
+
+
 # Blackman, WTC4
 wtc4gmin <- read.csv("data/wtc4_gmin_detached.csv") %>%
   rename(gmin = gmin_mmol_m2_s) %>%
   mutate(ch_temp_fac = as.factor(ch_temp))
 
 wtc4gdark <- read.csv("data/wtc4_gnight.csv")
+
 
 
 

### 2017-03-16 manuscript.Rmd
figure split kerst by group; add riederer data; methods

@@
 Importance of stomatal conductance. Prevalence of the Ball-Berry model, and g0. What is g0? Point of this paper.
+
+# Methods
+
+## Data
+
+Kerstiens 1996.
+Permeances were converted to molar fluxes assuming 41 mol m-3 (but see Kerstiens for a note on this - it is apparently not appropriate).
+Data from Riederer2001 not included because overlap with Kerstiens; but reported values very similar.
+
+```{r, echo=FALSE, eval=FALSE}
+with(subset(kerst, method == "gcut_isol"), summary(gmin))
+with(subset(rieder, gmin < 100000), summary(gmin))
+```
+
+Compilation of night-time conductance from lombardozzi.
+
+Minimum conductance from Blackman (WTC4). See Drake for WTC4 methods.
 
@@
 Closed system to measure gdark+gmin; gmin on detached leaves.
+
+## Cuticular
+
+Not really a conductance, since water does not diffuse as a gas through the cuticle.
+Instead it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle.
+The main barrier to diffusion is actually a thin layer at the leaf surface, and hence gcut does not depend on the thickness of the cuticle (Rieder Schreiber 1996).
 

### 2017-03-16 R/figures.R

@@ -25,6 +25,70 @@ figure_gmin_review <- function(gdfr){
   
 }
 
+figure_gmin_review_2 <- function(gdfr, minags){
+  par(mar=c(3,5,1,0.5), cex.lab=1.2)
+  fit <- lm(log10(gmin) ~ method-1, data=gdfr)
+  cis <- confint(fit)
+  quans <- sapply(split(gdfr, gdfr$method), function(x)quantile(log10(x$gmin), probs=c(0.05, 0.95)))
+  gmins <- with(gdfr, tapply(log10(gmin), method, mean, na.rm=TRUE))
+  
+  plot(1:5, 10^gmins, pch=19, cex=1.2, xlab="",
+       axes=FALSE,
+       panel.first=segments(x0=1:5, x1=1:5, y0=10^quans[1,], y1=10^quans[2,], col="grey"),
+       ylim=c(0,50),
+       xlim=c(0.5,6.5),
+       ylab=expression(Conductance~~(mmol~m^-2~s^-1)))
+  axis(1, at=1:6, labels=c(expression(g["cut,isol"]),
+                           expression(g["min,seal"]),
+                           expression(g[min]),
+                           expression(g[dark]),
+                           expression(g[0]),
+                           expression(g["s,low A"])))
+  axis(2)
+  box()
+  arrows(x0=1:5, x1=1:5, y0=10^cis[,1], y1=10^cis[,2], angle=90, length=0.1, code=3)
+  
+  quang <- quantile(minags$gsmin, probs=c(0.05, 0.95))
+  points(6, 1000*mean(minags$gsmin), pch=19, cex=1.2,
+         panel.first=segments(x0=6, x1=6, y0=1000*quang[1], y1=1000*quang[2], col="grey"))
+  cig <- 1000*(mean(minags$gsmin) + c(-2,2)*sd(minags$gsmin)/sqrt(nrow(minags)))
+  arrows(x0=6, x1=6, y0=cig[1], y1=cig[2], angle=90, length=0.1, code=3)
+}
+
+
+figure_gmin_review_3 <- function(){
+  
+  library(multcomp)
+  
+  plot_kerst2 <- function(yvar = "PhylogeneticGroup", meth = "gcut_isol", data=kerst2, ...){
+    
+    data <- data[data$method == meth,]
+    data$yvar <- as.factor(data[,yvar])
+    
[…]

### 2017-03-16 R/load.R

@@ -48,12 +48,17 @@ lin2015a <- group_by(lin2015, fitgroup) %>%
 lin2015coef <- fits_lin2015(lin2015a)
 
 
+# Riederer and Schreiber 2001
+rieder <- read.csv("data/riederer2001_table1.csv") %>%
+  mutate(gmin = 10^5 * 10^-3 * gmin / 41)
+
 # Kerstiens 1996
 kerst <- read.csv("data/kerstiens1996_table1.csv", stringsAsFactors = FALSE) %>%
   filter(gmin > 0,
          method %in% c("A","C","D")) %>%
   mutate(method = dplyr::recode(method, A="gcut_isol", C="gcut_seal", D="gmin"),
-         gmin = 10^5 * 10^-3 * gmin / 40)   # convert to mmol m-2 s-1 assume 40 mol m-3
+         gmin = 2 * 10^5 * 10^-3 * gmin / 41)   # convert to mmol m-2 s-1 assume 41 mol m-3
+                # 2 because gmin was expressed as per total leaf area.
 
 # Miner et al. 2016
 miner <- read.csv("data/Miner_table1.csv")
@@ -101,6 +106,15 @@ wtc4gmin <- read.csv("data/wtc4_gmin_detached.csv") %>%
 wtc4gdark <- read.csv("data/wtc4_gnight.csv")
 
 
+# From Lin2015, data where A < threshold.
+minags <- group_by(lin2015, fitgroup) %>%
+  summarize(
+    Amin = min(Photo, na.rm=TRUE),
+    gsmin = min(Cond, na.rm=TRUE),
+    Qrange = max(PARin) - min(PARin),
+    Pathway = unique(Pathway)[1]
+  ) %>%
+  filter(Amin < 2)
 
 
 

### 2017-03-16 R/make_figures.R

@@ -3,10 +3,10 @@ to.pdf(figure_sim(), "output/figure1.pdf", height=6, width=6)
 to.pdf(figure_g0g1_cor(lin2015), "output/figure2.pdf", height=4, width=8)
 to.pdf(figure_R2g0(lin2015coef, miner), "output/figure3.pdf", height=4, width=8)
 
-to.pdf(figure_gmin_review(gdfr), "output/figure4.pdf", height=5, width=4)
-to.pdf(figure_amings(lin2015), "output/figure5.pdf", height=5, width=5)
+to.pdf(figure_gmin_review_2(gdfr, minags), "output/figure4.pdf", height=5, width=5)
+#to.pdf(figure_amings_lin2015), "output/figure5.pdf", height=5, width=5)
 
-to.pdf(figure_wtc4_gmin(wtc4gmin, wtc4gdark), "output/figure6.pdf", height=4, width=8)
+to.pdf(figure_wtc4_gmin(wtc4gmin, wtc4gdark), "output/figure5.pdf", height=4, width=8)
 
 
 convert_all_pdf("output")

### 2017-03-20 manuscript.Rmd
some notes

@@
 
+Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015]
+
 # Components of g0
@@
 g0?
+
+@schuster2016 thermal tolerance and gmin relationship across species.
 
@@
 Instead it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle.
-The main barrier to diffusion is actually a thin layer at the leaf surface, and hence gcut does not depend on the thickness of the cuticle (Rieder Schreiber 1996).
+The main barrier to diffusion is actually a thin layer at the leaf surface, and hence gcut does not depend on the thickness of the cuticle  [@riederer2001].
+
+gcut is probably too low in the way it is measured (as a permeance), due to low water content of the cuticle in those conditions [@kerstiens1996]
+
+gseal on average not different from gcut - but for individual studies they frequently differ a lot indicating variable role of stomatal contribution (santrucek)
 
@@
 - @riederer2001 correlation between temperature and gcut (non-stomatal cuticles)
+- @schuster2016 correlation temperature, gcut AND gmin for Nerium oleander (page 59).
+T range much wider than Blackman though...
+But Schuster gmin not different from gcut at most temperatures, which is a bit strange too.
+page 90 few gmin values from drying curves that we can add.
+
 - @burghardt2003 gmin review

### 2017-03-21 manuscript.Rmd
figure captions

--- 2017-03-16 manuscript.Rmd
@@
 **Figure 1**.
-Simulations with the plantecophys R package, with and without a g~0~.
-For the bottom two panels, VPD and T~air~ were covaried based on an empirical relationship, reflecting the fact that they covary in field conditions.
+Simulations  of intrinsic water-use efficiency (A~n~/g~s~), intercellular CO~2~ concentration (C~i~) and leaf transpiration rate (E~L~) with three values of g~0~.
+For the top two panels, all environmental drivers were held constant except the photosynthetic photon flux density (PPFD), demonstrating that the inclusion of g~0~ is necessary to simulate the decrease in A~n~/g~s~ at low PPFD.
+For the bottom two panels, VPD and T~air~ were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions.
 For all simulations it is assumed that T~leaf~ is equal to T~air~.
@@
 **Figure 2**.
-Example linear regression of stomatal conductance (g~s~) against a stomatal index, here the linearized version of the model of Medlyn et al.
+Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al.
 (2011).
-Example data is taken from Lin et al.
-(2015), chosen is a dataset with a very good fit as an example.
-Solid line is regression line, shaded area a 95% confidence interval.
-Right panel, visualized correlation between estimated slope (g~1~) and intercept (g~0~) of the regression.
-Dotted ellipse is a bivariate 95% confidence interval (using confidenceEllipse from the car R package), symbols 1000 bootstrap samples of the coefficients (using bootCase in the car R package).
+This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al.
+(2015).
+The solid line is regression line, and the shaded area a 95% confidence interval for the mean.
+The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression.
+The dotted ellipse is a bivariate 95% confidence interval for slope and intercept.
+The symbols represent 1000 bootstrap samples of the coefficients.
 
@@
 **Figure 3**.
+Correlation between estimated g~0~ and the R^2^ of the regression.
+In the left panel, we fitted the linearized form of the Medlyn et al.
+(2011) model to each of the datasets in the Lin et al.
+(2015) database.
+Vertical lines are 95% confidence intervals.
+In the right panel, a very similar association between g~0~ and R^2^ is presented by the review of Miner et al.
+(2017), using the Ball-Berry model.
 
@@
 **Figure 4**.
+Comparison of various estimates of minimum conductance.
+Error bars are standard errors, grey vertical lines 90% quantiles indicating wide range of variation for any variable.
+g~cut,isol~ : cuticual permeance measured on isolated cuticles that lack stomata (data from Kerstiens 1996), g~min,seal~ : minimum conductance based on leaf dry-down curves where the side of the leaf that bears stomata is sealed (data from Kerstiens 1996), g~min~ : minimum conductance based on leaf-dry-down curves (data from Kerstiens 1996, and more added), g~dark : stomatal conductance at night, measured in field conditions (data from Lombardozzi et al.
+2017), g~0~ fitted g~0~ in the Medlyn et al.
+(2011) model with the Lin et al.
+(2015) database, g~s,low A~ : for the Lin et al.
+(2015) database, minimum g~s~ when net assimilation rate was less than 2 $mu$ mol m^-2^ s^-1^, calculated once for each dataset.
 
@@
 **Figure 5**.
+Analysis of g~min~ from Figure 4 by plant growth form and phylogenetic group.
+Different letters denote significant differences at $alpha$ = 0.05.
+The groups Trees are split by angiosperms and gymnosperms.
 

### 2017-03-21 R/figures.R

@@ -102,9 +102,9 @@ figure_gmin_review_3 <- function(kerst2){
 
 
 
-figure_g0g1_cor <- function(lin2015){
+figure_g0g1_cor <- function(lin2015, group, legend=FALSE){
   
-  x <- subset(lin2015, fitgroup == "Nicolas Martin-StPaul_Quercus ilex_StPaul_Puechabon")
+  x <- subset(lin2015, fitgroup == group)
   
   # for bootCase to work
   assign("x",x,envir=.GlobalEnv)
@@ -119,8 +119,11 @@ figure_g0g1_cor <- function(lin2015){
   
   par(mfrow=c(1,2), mar=c(4,4,1,1), mgp=c(2.5,0.5,0), tcl=0.2, las=1,
       cex.lab=1.1, cex.axis=0.9)
-  with(x, plot(BBopti, Cond, ylim=c(0, 0.15),
+  with(x, plot(BBopti, Cond, 
                xlim=c(0,0.04),
+               ylim=c(0,0.15),
+               # ylim=c(0, max(Cond)),
+               # xlim=c(0,max(BBopti)),
                panel.first=add_regres_line(fit),
                xlab=expression(A/(C[a]*sqrt(D))),
                ylab=expression(g[s]~(mol~m^-2~s^-1)),
@@ -130,10 +133,14 @@ figure_g0g1_cor <- function(lin2015){
   plot(el, type='l', lty=3, 
        xlab=expression(g[0]~~(mol~m^-2~s^-1)), 
        ylab=expression(g[1]~~(kPa^-0.5)),
+       xlim=c(0,0.025),
        ylim=c(3,5))
+       # xlim=c(min(b[,1]), max(b[,1])),
+       # ylim=c(min(b[,2]), max(b[,2])))
   points(b, pch=16, cex=0.3, col="dimgrey")
   points(coef(fit)[1],coef(fit)[2], pch=19)
   
+  if(legend)legend("topright", group, bty='n', cex=0.6)
 }
 
 
@@ -186,18 +193,24 @@ figure_sim <- function(){
       cex.axis=0.9, cex.lab=1.1)
   
   g1 <- 4
-  g0 <- 0.03
+  g0_1 <- 0.01
+  g0_2 <- 0.03
   
   lty0 <- 1
   lty1 <- 5
+  lty2 <- 4
   col0 <- "black"
   col1 <- "black"
+  col2 <- "black"
   
   r0 <- Photosyn(PPFD=seq(20,1000, length=101),
[…]

### 2017-03-21 R/make_figures.R

@@ -1,16 +1,30 @@
 
 to.pdf(figure_sim(), "output/figure1.pdf", height=6, width=6)
-to.pdf(figure_g0g1_cor(lin2015), "output/figure2.pdf", height=4, width=8)
+to.pdf(figure_g0g1_cor(lin2015, "Nicolas Martin-StPaul_Quercus ilex_StPaul_Puechabon"), 
+       "output/figure2.pdf", height=4, width=8)
+
 to.pdf(figure_R2g0(lin2015coef, miner), "output/figure3.pdf", height=4, width=8)
 
-to.pdf(figure_gmin_review_3(kerst2), "output/figure4.pdf", height=4, width=3.5)
+to.pdf(figure_gmin_review_2(gdfr, minags), "output/figure4.pdf", height=4, width=4.5)
+to.pdf(figure_gmin_review_3(kerst2), "output/figure5.pdf", height=4, width=3.5)
 #to.pdf(figure_amings_lin2015), "output/figure5.pdf", height=5, width=5)
 
 #to.pdf(figure_wtc4_gmin(wtc4gmin, wtc4gdark), "output/figure5.pdf", height=4, width=8)
-to.pdf(figure_wtc4_gmin_2(wtc4gmin), "output/figure5.pdf", height=4, width=3.5)
+to.pdf(figure_wtc4_gmin_2(wtc4gmin), "output/figure6.pdf", height=4, width=3.5)
 
-to.pdf(figure_hakea_gmin(lopw), "output/figure6.pdf", height=5, width=6)
+to.pdf(figure_hakea_gmin(lopw), "output/figure7.pdf", height=5, width=6)
 
 
 #convert_all_pdf("output")
+
+
+
+
+
+# pdf("allfits.pdf")
+# for(g in lin2015coef$fitgroup){
+#   figure_g0g1_cor(lin2015, g)
+# }
+# dev.off()
+
 

### 2017-03-21 manuscript.Rmd
Hakea figure updated

@@
 \
-Remko A. Duursma^1^, Chris Blackman^1^, Belinda Medlyn^1^
+Remko A. Duursma^1^, Chris Blackman^1^, Rosana Lopéz^1^, Belinda Medlyn^1^
 
@@
 ```{r echo=FALSE}
-includefigs <- TRUE
+includefigs <- FALSE
 library(png)
@@
 plotfig <- function(...)grid.raster(readPNG(...))
+
+source("R/load.R")
+
+pval <- function(...)formatPval(..., includeEquality=TRUE)
 ```
@@
 
-Importance of stomatal conductance. Prevalence of the Ball-Berry model, and g0. What is g0? Point of this paper.
+Importance of stomatal conductance. Prevalence of the Ball-Berry model. Parameters are g1 and g0. 
+What is g0? Point of this paper.
+
+We argue that
+
+- we need g0 to get responses at low light correct, and probably at high T
+- g0 is poorly estimated as the intercept, with the usual regression technique
+- can we estimate g0 independently?
+Some assume that g0 = gdark but that's rubbish.
+It is also much larger than cuticular conductance, but seems roughly consistent with gmin (minimum conductance from leaf dry-down curves).
+- there is inter-specific variation in gmin that is related to growth form 
+- gmin is a plastic trait as it responds to growth at elevated temperature, and growth in dry soil
+- models should consider night-time conductance and daily minimum conductance as distinctly separate processes.
 
@@
 
-```{r, echo=FALSE, eval=FALSE}
-with(subset(kerst, method == "gcut_isol"), summary(gmin))
-with(subset(rieder, gmin < 100000), summary(gmin))
-```
-
 Compilation of night-time conductance from lombardozzi.
@@
 Minimum conductance from Blackman (WTC4). See Drake for WTC4 methods.
+gmin Hakea, same methods (measurement T?)
 
@@
 
-[@medlyn_reconciling_2011] does not develop theory for g0, instead adds it as a minimum conductancce because that's how it is!
+[@medlyn_reconciling_2011] does not develop theory for g0, instead adds it as a minimum conductance because that's how it is!
 
@@
 
+ **Cite Figure 1**
+
 # g~0~ is poorly estimated from regression
@@
 
-# Components of g0
+**Cite Figure 2 and 3**
+
+# Definitions of minimum conductance
 
@@
 g0?
+
+**Cite Figure 4**
 
@@
 
-## Cuticular
+## Cuticular permeance
 
@@
 
-## nighttime
+## Night-time conductance
+
 Nighttime conductance not necessarily same as minimum conductance
@@
 
-## minimum conductance
+## Minimum conductance from leaf-dry down curves
 
@@
 - @schreiber1996 cuticle review
+
+** cite figures 5,6 and 7**
 
@@
 Error bars are standard errors, grey vertical lines 90% quantiles indicating wide range of variation for any variable.
-g~cut,isol~ : cuticual permeance measured on isolated cuticles that lack stomata (data from Kerstiens 1996), g~min,seal~ : minimum conductance based on leaf dry-down curves where the side of the leaf that bears stomata is sealed (data from Kerstiens 1996), g~min~ : minimum conductance based on leaf-dry-down curves (data from Kerstiens 1996, and more added), g~dark : stomatal conductance at night, measured in field conditions (data from Lombardozzi et al.
-2017), g~0~ fitted g~0~ in the Medlyn et al.
+g~cut,isol~ : cuticual permeance measured on isolated cuticles that lack stomata (data from Kerstiens 1996), g~min,seal~ : minimum conductance based on leaf dry-down curves where the side of the leaf that bears stomata is sealed (data from Kerstiens 1996), g~min~ : minimum conductance based on leaf-dry-down curves (data from Kerstiens 1996, with a few extra data points - to be added), g~dark : stomatal conductance in the dark at night, measured in field conditions (data from Lombardozzi et al.
+2017), g~0~ : fitted g~0~ in the Medlyn et al.
 (2011) model with the Lin et al.
@@
 Different letters denote significant differences at $alpha$ = 0.05.
-The groups Trees are split by angiosperms and gymnosperms.
+The groups 'Angiosperm' and 'Gymnosperm' include trees only.
 
@@
 **Figure 6**.
+Minimum conductance measured from leaf dry-down curves on *Eucalyptus parramattensis* grown in whole-tree chambers.
+Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C).
+Measurements of g~min~ were carried out at five different temperatures on replicate leaves.
+There is no effect of measurement T on g~min~, but leaves of trees grown in elevated temperature showed consistently lower g~min~.
+
+```{r fig.height=4,echo=FALSE, eval=includefigs}
+plotfig("output/figure7.png")
+```
+
+**Figure 7**.
+Minimum conductance measured from leaf dry-down curves on a variety *Hakea* species, a genus native to Australia.
+Plants were grown in containers in a grow house, and supplied with ample water or underwent long-term mild drought stress.
+Bars are labelled by leaf form : broadleaf (B) or needle-like (N).
+g~min~ was higher for species with needle-like leaves, both for well-watered plants (Wilcoxon test, p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`) and droughted plants (p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`).
 

### 2017-03-21 R/figures.R

@@ -378,7 +378,8 @@ figure_hakea_gmin <- function(lopw){
   bar_cols <- c("lightgrey","black")
   par(mar=c(7,4,1,1), las=3, yaxs="i",
       mgp=c(2.5, 0.5,0), tcl=0.2, cex.axis=0.9, cex.lab=1.1)
-  barplot(t(as.matrix(lopw2[,2:3])), beside=T,
+  m <- t(as.matrix(lopw2[,c("gmin.w","gmin.d")]))
+  b <- barplot(m, beside=T,
           ylim=c(0,1.4),
           col=bar_cols,
           names.arg=abbrev_hak(lopw2$species),
@@ -386,7 +387,10 @@ figure_hakea_gmin <- function(lopw){
   box()
   legend("topleft", c("Well-watered","Drought stress"), fill=bar_cols,
          bty='n', cex=0.9)
-  
+  y <- m[1,]
+  x <- colMeans(b)
+  ch <- toupper(substr(as.character(lopw2$leaf.form),1,1))
+  text(x,y,ch, pos=3, font=3, cex=0.7)
 }
 
 

### 2017-03-21 R/load.R

@@ -1,7 +1,8 @@
 
 pacman::p_load(Hmisc, car, dplyr, tidyr, nlme, nlshelper, 
                forcats, tibble, magicaxis, 
-               plantecophys, readxl, multcomp)
+               plantecophys, readxl, multcomp,
+               reporttools)
 
 
 
@@ -129,6 +130,26 @@ minags <- group_by(lin2015, fitgroup) %>%
 lopez <- read.csv("data/lopez_gmin_hakea.csv") %>% group_by(species, treatment) %>%
   summarize(gmin = mean(gmin))
 
+# some meta data for this dataset (leaf form, MAP)
+lopmet <- read.csv("data/lopez_gmin_hakea_meta.csv")
+
+# Wide format
 lopw <- reshape(as.data.frame(lopez), direction="wide", idvar="species", timevar="treatment")
+
+lopw <- dplyr::select(lopmet, species, leaf.form) %>% distinct() %>%
+  right_join(lopw, by="species")
+
+# Version 2; by population merged with lopmet (because MAP varies by population)
+lopez2 <- read.csv("data/lopez_gmin_hakea.csv") %>% group_by(species, population, treatment) %>%
+  summarize(gmin = mean(gmin)) %>% 
+  inner_join(lopmet, by=c("species","population"))
+
+
+
+
+
+
+
+
 
 

### 2017-03-21 R/make_figures.R

@@ -15,7 +15,7 @@ to.pdf(figure_wtc4_gmin_2(wtc4gmin), "output/figure6.pdf", height=4, width=3.5)
 to.pdf(figure_hakea_gmin(lopw), "output/figure7.pdf", height=5, width=6)
 
 
-#convert_all_pdf("output")
+convert_all_pdf("output")
 
 
 

### 2017-03-22 manuscript.Rmd
updated Lopez data

@@
 ```{r echo=FALSE}
-includefigs <- FALSE
+includefigs <- TRUE
 library(png)
@@
 Bars are labelled by leaf form : broadleaf (B) or needle-like (N).
-g~min~ was higher for species with needle-like leaves, both for well-watered plants (Wilcoxon test, p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`) and droughted plants (p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`).
+For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`).
 

### 2017-03-22 R/figures.R

@@ -380,7 +380,7 @@ figure_hakea_gmin <- function(lopw){
       mgp=c(2.5, 0.5,0), tcl=0.2, cex.axis=0.9, cex.lab=1.1)
   m <- t(as.matrix(lopw2[,c("gmin.w","gmin.d")]))
   b <- barplot(m, beside=T,
-          ylim=c(0,1.4),
+          ylim=c(0,10),
           col=bar_cols,
           names.arg=abbrev_hak(lopw2$species),
           ylab=expression(g[min]~~(mmol~m^-2~s^-1)))

### 2017-03-23 manuscript.Rmd
Added data to review; more references and some methods written

@@
 
-Importance of stomatal conductance. Prevalence of the Ball-Berry model. Parameters are g1 and g0. 
-What is g0? Point of this paper.
+Prevalence of the Ball-Berry class of stomatal conductance models characterized by dependence on assimilation rate, CO~2~ and VPD or RH.
+
+\
+(1)
+$$ g_s = g_0 + g_1 \frac{A_n}{C_a}f(D)$$
+
+where f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), g~1~ the slope parameter, and g~0~ an intercept.
+The Medlyn model is actually non-linear but can be expressed in an approximate linear form of Eq.
+1.
+
+The g~0~ parameter is the intercept, i.e. g~s~ when A is zero. Depending on the form of f(D), it can also represent g~s~ when D is very large.
+
+The objective of this paper is to review sources of variation and common misconceptions on the g~0~ parameter.
 
@@
 
-Kerstiens 1996.
-Permeances were converted to molar fluxes assuming 41 mol m-3 (but see Kerstiens for a note on this - it is apparently not appropriate).
-Data from Riederer2001 not included because overlap with Kerstiens; but reported values very similar.
+We compare estimates of g~cut~, g~min~, and g~dark~ from the literature.
+Measurements of g~cut~ and g~min~ were taken from the compilation by @kerstiens1996, with additional g~min~ measurements from @schuster2016, @burghardt2003, @cavenderbares2007, and @phillips2010 (for the latter study, raw data were obtained and g~min~ calculated).
+It is customary for g~cut~ and g~min~ to be expressed per unit total leaf surface area, however, since we aim a comparison with g~s~ and g~dark~, all values were converted to per unit projected surface area.
+Units were converted to molar fluxes assuming 41 mol m^-3^ (cf.
+@kerstiens1996).
+For g~dark~, we used a recent compilation of measurements from @lombardozzi2017.
 
-Compilation of night-time conductance from lombardozzi.
+For the combined dataset, we looked up plant growth form and phylogenetic group (angiosperm or gymnosperm) from the TRY categorical traits database [@kattge2011].
 
-Minimum conductance from Blackman (WTC4). See Drake for WTC4 methods.
-gmin Hakea, same methods (measurement T?)
+We present two case studies demonstrating the acclimation of g~min~ to environmental drivers.
+Both studies use identical methods.
+The first dataset was collected on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature.
+Leaves were sampled from XXX, n=XXX.
+The second case study was on *Hakea* plants.
+Details here (measurement T?).
+Methods for gmin from drying curves here.
+
+## Model
+
+\
+(2)
+$$ A_n = \frac{g_s}{1.6}(C_a - C_i) $$
+
+\
+where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~).
+
+\
+(3)
+$$ A_n = \frac{a \cdot (C-i - \Gamma^{*})}{C_i + b} - R_d$$
+
+where a = V~cmax~ or ~J/4~ (where J goes to Jmax at high light) depending on the limitation to photosynthesis, and b also varies.
+Dependence on temperature enters through V~cmax~, J~max~, R~d~, $\Gamma^{*}$ [@medlyn2002].
 
@@
 
-[@medlyn_reconciling_2011] does not develop theory for g0, instead adds it as a minimum conductance because that's how it is!
+[@collatz1992] shows that including g0 affects response of gs to RH in the coupled leaf gas exchange model.
+And "It seems possible that other conditions such as severe drought stress may affect the minimum value of  g  as well, though we currently have no evidence to support this."
+
+[@medlyn_reconciling_2011] does not develop theory for g~0~, instead adds it as a minimum conductance because that's how it is!
 
@@
 Instead it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle.
-The main barrier to diffusion is actually a thin layer at the leaf surface, and hence gcut does not depend on the thickness of the cuticle  [@riederer2001].
+The main barrier to diffusion is actually a thin layer at the leaf surface, and hence gcut does not depend on the thickness of the cuticle  [@riederer2001; @anfodillo2002].
+@onoda2012 found that thicker cuticles are more resistant to tearing (force to tear and tensile strength both approx proportional to cuticle thickness)
 
@@
 @korner1994 table of minimum conductance. Also points out (without reference) that g0 in droughted leaves is certainly lower than g0 in dark or very high CO2 leaves.
+
+@howard2007 gdark much higher than gmin (Helianthus annuus). In general this is the case [@caird2007]
 
@@
 
-@manzoni assumes minimum conductance is cuticular only, because CO2 does not diffuse across it but water does.
+@manzoni2011 assumes minimum conductance is cuticular only, because CO2 does not diffuse across it but water does.
 Seems to work in his modelling context but appears to be in disagreement with the observation that gmin >> gcut.
@@
 - @schreiber1996 cuticle review
+
+[@anfodillo2002] gmin high altitude.
+Four-fold increase gmin from February (winter) to August in Picea abies, probably due to incomplete stomatal closure not changes in cuticle properties.
+Cite common finding that gmin increases with altitude (top left page 485).
+
+[@herrick1991] gmin increased with altitude in Picea rubens.
+
+[@jordan2007] large increase gmin in the herb Agastachys odorata with leaf age
+
+Sack et al. 2010 protocol for measurement
+
+[@cavenderbares2007] gdark similar to gmin in oaks (unusual!). Lower gmin in droughted oaks.
 
@@
 (2015) database, minimum g~s~ when net assimilation rate was less than 2 $mu$ mol m^-2^ s^-1^, calculated once for each dataset.
+Minimum conductance and cuticular conductance are commonly expressed per all surface area but here projected was used because that's more common for g0, gdark.
 
@@
 # References {.unnumbered}
+
+Sack, L., Scoffoni, C., PrometheusWiki contributors, 2010.
+"Minimum epidermal conductance (gmin, a.k.a.
+cuticular conductance)," PrometheusWiki, http://www.publish.csiro.au/prometheuswiki/tiki-pagehistory.php?page=Minimum epidermal conductance (gmin, a.k.a.
+cuticular conductance)&preview=7 (accessed March 23, 2017).

### 2017-03-23 R/load.R

@@ -57,9 +57,7 @@ rieder <- read.csv("data/riederer2001_table1.csv") %>%
 kerst <- read.csv("data/kerstiens1996_table1.csv", stringsAsFactors = FALSE) %>%
   filter(gmin > 0,
          method %in% c("A","C","D")) %>%
-  mutate(method = dplyr::recode(method, A="gcut_isol", C="gcut_seal", D="gmin"),
-         gmin = 2 * 10^5 * 10^-3 * gmin / 41)   # convert to mmol m-2 s-1 assume 41 mol m-3
-                # 2 because gmin was expressed as per total leaf area.
+  mutate(method = dplyr::recode(method, A="gcut_isol", C="gcut_seal", D="gmin"))
 # Notes:
 # A - astomatous cuticle, removed from leaf (permeance)
 # B - detached leaves, stomatal side sealed
@@ -68,6 +66,18 @@ kerst <- read.csv("data/kerstiens1996_table1.csv", stringsAsFactors = FALSE) %>%
 # E1 - gnight / gdark
 # E2 - minimum conductance measured during the day (vague!)
 # E3 - presumed stomatal closure (v high CO2, ABA, drought, VPD) (vague!)
+
+# Add some more data to kerst. Manually entered from various sources.
+gminrev <- read.csv("data/gmin_review_literature.csv",stringsAsFactors = FALSE) %>% 
+  mutate(method = "gmin") %>%
+  dplyr::select(gmin, method, species)
+
+kerst <- bind_rows(kerst, gminrev) %>% 
+  mutate(gmin = 2 * 10^5 * 10^-3 * gmin / 41)   
+# convert to mmol m-2 s-1 assume 41 mol m-3
+# 2 because gmin was expressed as per total leaf area.
+
+
 
 # Miner et al. 2016
 miner <- read.csv("data/Miner_table1.csv")

### 2017-03-23 R/figures.R
Added data to review; more references and some methods written

@@ -58,7 +58,7 @@ figure_gmin_review_2 <- function(gdfr, minags){
 
 figure_gmin_review_3 <- function(kerst2){
   
-  plot_kerst2 <- function(yvar = "PhylogeneticGroup", meth = "gcut_isol", data=kerst2, ...){
+  plot_kerst2 <- function(yvar = "PhylogeneticGroup", meth = "gmin", data=kerst2, ...){
     
     data <- data[data$method == meth,]
     data$yvar <- as.factor(data[,yvar])
@@ -75,7 +75,7 @@ figure_gmin_review_3 <- function(kerst2){
          xlim=c(0.7, n+0.3), ...)
     axis(1, at=1:n, labels=capitalize(levels(data$yvar)))
     arrows(x0=1:n, x1=1:n, y0=ci[,1], y1=ci[,2], angle=90, length=0.05, code=3)
-    text(x=1:n, y=ci[,2], lets, pos=3)
+    text(x=1:n, y=ci[,2], lets, pos=3, cex=0.8)
     axis(2)
     box()
   }

### 2017-03-27 manuscript.Rmd
more references and new outline

@@
 ```{r echo=FALSE}
-includefigs <- TRUE
+includefigs <- F
 library(png)
@@
 ```
+
+# new outline
+
+1. Introduction (short): questions about g0.
+2. Need for inclusion of g0 in models
+3. Comparison of values for various possible definitions
+4. Estimation of g~0~ by fitting 
+5. Nighttime conductance (v. short)
+6. Cuticular conductance (short)
+7. Minimum conductance (longer)
+8. Recommendations and future work
 
@@
 - models should consider night-time conductance and daily minimum conductance as distinctly separate processes.
+
+g~0~ is often defined as g~s~ when A approaches zero when PPFD approaches zero [@leuning1995] but it is clear from Eq.
+1 that g~0~ is also attained when A approaches zero due to other environmental drivers, in particular temperature, and drought stress (which may either reduce photosynthetic capacity or g~1~, or both, see Zhou).
+It is not clear nor obvious whether these various minima should equal each other.
+
+One estimate of g~0~ (low PPFD) comes from night-time measurements of conductance, which have been known for a long time to be substantially larger than zero.
+It is however not clear whether daytime g~0~ at low PPFD, for darkened leaves, is comparable in magnitude to g~0~ in the dark at night.
+Evidence from experiments where environmental conditions are held constant suggests that g~s~ varies tremendously throughout the day independent of these drivers, and has therefore been attributed to an internal clock (Resco).
 
@@
 
+Implemented in the *plantecophys* R package [@duursma2015]
+
 # Do we need g~0~ in simulations of water and carbon fluxes?
@@
 The first Ball-Berry model did not have a g0
+
+@leuning1995 says "g0 > 0, otherwise ci/ca would be constant for all values of Q.
+The rate of convergence of ci to ca at low A (low Q) depends on the value of g0; when g0 is large, ci deviates from the asymptotic value at higher values of A than when g0 is small".
 
@@
 
+"We have assumed in the following equations that when stomata are closed the leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." [@jarvis1976]
+
+The difficulty of this can be seen from the fact that a) many studies set g0 to something rather than fitting it (e.g.
+@leuning1995 uses 0.01 for all); presumably because unreliable estimates were obtained, b) negative fitted values are common [@leuning1995; @herault2013; @miner2016].
+
 Barnard and Bauerle said so too but gave no details.
@@
 @leuning_modelling_1990], which we assume is quite common when inspecting the many rounded values in the review by
+
+g~0~ fitted is higher than minimum g~s~ from datasets.
+If we restric the fittings to those were g~0~ had a relatively narrow width of the confidence interval (< 0.03 mol m-2 s-1), then the average was `r with(subset(lin2015coef, g0_uci - g0_lci < 0.03), mean(g0))` compared to `r mean(lin2015coef$g0)` for all.
 

### 2017-03-27 manuscript.Rmd
fixed broken refs

@@
 
-[@medlyn_reconciling_2011] does not develop theory for g~0~, instead adds it as a minimum conductance because that's how it is!
+[@medlyn2011] does not develop theory for g~0~, instead adds it as a minimum conductance because that's how it is!
 
@@
 The difficulty of this can be seen from the fact that a) many studies set g0 to something rather than fitting it (e.g.
-@leuning1995 uses 0.01 for all); presumably because unreliable estimates were obtained, b) negative fitted values are common [@leuning1995; @herault2013; @miner2016].
+@leuning1995 uses 0.01 for all); presumably because unreliable estimates were obtained, b) negative fitted values are common [@leuning1995; @heroult2013; @miner2016].
 
@@
 Across datasets, g0 increases when the model fits poorly (Figure 3), and the width of the confidence interval increases when less variation in a ball-berry coefficient.
-Often g0 is even negative [@heroult2013; @leuning_modelling_1990], so it is often set to some arbitrary value
+Often g0 is even negative [@heroult2013; @leuning1990], so it is often set to some arbitrary value
 
@@
 A typical approach is to avoid this by setting some value based on guesswork [e.g.
-@leuning_modelling_1990], which we assume is quite common when inspecting the many rounded values in the review by
+@leuning1990], which we assume is quite common when inspecting the many rounded values in the review by
 

### 2017-03-27 manuscript.Rmd
fixed references again

@@
 - can we estimate g0 independently?
-Some assume that g0 = gdark but that's rubbish.
+Some assume that g0 = gdark (also @miner2017 !) but that's rubbish.
 It is also much larger than cuticular conductance, but seems roughly consistent with gmin (minimum conductance from leaf dry-down curves).
@@
 (3)
-$$ A_n = \frac{a \cdot (C-i - \Gamma^{*})}{C_i + b} - R_d$$
+$$ A_n = \frac{a \cdot (C_i - \Gamma^{*})}{C_i + b} - R_d$$
 
@@
 The difficulty of this can be seen from the fact that a) many studies set g0 to something rather than fitting it (e.g.
-@leuning1995 uses 0.01 for all); presumably because unreliable estimates were obtained, b) negative fitted values are common [@leuning1995; @heroult2013; @miner2016].
+@leuning1995 uses 0.01 for all); presumably because unreliable estimates were obtained, b) negative fitted values are common [@leuning1995; @heroult2013; @miner2017].
 
@@
 Vertical lines are 95% confidence intervals.
-In the right panel, a very similar association between g~0~ and R^2^ is presented by the review of Miner et al.
-(2017), using the Ball-Berry model.
+In the right panel, a very similar association between g~0~ and R^2^ is presented by the review of @miner2017, using the Ball-Berry model.
 

### 2017-03-29 manuscript.Rmd
more references; new data added for brodribb, phillips, review; review data better organized and figures updated

--- 2017-03-27 manuscript.Rmd
@@
 ```{r echo=FALSE}
-includefigs <- F
+includefigs <- TRUE
 library(png)
@@
 ```
-
-# new outline
-
-1. Introduction (short): questions about g0.
-2. Need for inclusion of g0 in models
-3. Comparison of values for various possible definitions
-4. Estimation of g~0~ by fitting 
-5. Nighttime conductance (v. short)
-6. Cuticular conductance (short)
-7. Minimum conductance (longer)
-8. Recommendations and future work
 
@@
 - models should consider night-time conductance and daily minimum conductance as distinctly separate processes.
-
-g~0~ is often defined as g~s~ when A approaches zero when PPFD approaches zero [@leuning1995] but it is clear from Eq.
-1 that g~0~ is also attained when A approaches zero due to other environmental drivers, in particular temperature, and drought stress (which may either reduce photosynthetic capacity or g~1~, or both, see Zhou).
-It is not clear nor obvious whether these various minima should equal each other.
-
-One estimate of g~0~ (low PPFD) comes from night-time measurements of conductance, which have been known for a long time to be substantially larger than zero.
-It is however not clear whether daytime g~0~ at low PPFD, for darkened leaves, is comparable in magnitude to g~0~ in the dark at night.
-Evidence from experiments where environmental conditions are held constant suggests that g~s~ varies tremendously throughout the day independent of these drivers, and has therefore been attributed to an internal clock (Resco).
 
@@
 
-Implemented in the *plantecophys* R package [@duursma2015]
-
 # Do we need g~0~ in simulations of water and carbon fluxes?
@@
 The first Ball-Berry model did not have a g0
-
-@leuning1995 says "g0 > 0, otherwise ci/ca would be constant for all values of Q.
-The rate of convergence of ci to ca at low A (low Q) depends on the value of g0; when g0 is large, ci deviates from the asymptotic value at higher values of A than when g0 is small".
 
@@
 [@medlyn_reconciling_2011] does not develop theory for g~0~, instead adds it as a minimum conductance because that's how it is!
+
+[@hari1986; @hari1999] does not mention it (their Eq 6 assumes g0 = 0)
 
@@
 
+We need a non-zero minimum conductance when simulating water fluxes in very dry soil [@blackman2016; @anderegg2012]
+Also @gleason2014
+
 # g~0~ is poorly estimated from regression
-
-"We have assumed in the following equations that when stomata are closed the leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." [@jarvis1976]
-
-The difficulty of this can be seen from the fact that a) many studies set g0 to something rather than fitting it (e.g.
-@leuning1995 uses 0.01 for all); presumably because unreliable estimates were obtained, b) negative fitted values are common [@leuning1995; @herault2013; @miner2016].
 
@@
 @leuning_modelling_1990], which we assume is quite common when inspecting the many rounded values in the review by
-
-g~0~ fitted is higher than minimum g~s~ from datasets.
-If we restric the fittings to those were g~0~ had a relatively narrow width of the confidence interval (< 0.03 mol m-2 s-1), then the average was `r with(subset(lin2015coef, g0_uci - g0_lci < 0.03), mean(g0))` compared to `r mean(lin2015coef$g0)` for all.
 
@@
 Closed system to measure gdark+gmin; gmin on detached leaves.
+
+@muchow1989 correlation stomatal density and gmin suggesting leaky stomata
+
+@brodribb2014 59-94% of transpiration during leaf dry down was from stomatal side of the leaf in seven conifer species, suggesting large component of gmin is leaky stomata.
+
+@brodribb2014 gmin across conifer species was correlated with the rainfall of the driest quarter, again suggesting gmin is a plastic trait
 
@@
 [@cavenderbares2007] gdark similar to gmin in oaks (unusual!). Lower gmin in droughted oaks.
+
+[@duarte2016] fitted g0 with BB and BBOpti.
+Lower g0 in heat wave treatment.
+Also lower gdark in heat wave treatment.
+But g1 or m not affected by heat waves (again example where a gmin is a responsive parameter!)
 

### 2017-03-29 R/figures.R

@@ -25,16 +25,16 @@ figure_gmin_review <- function(gdfr){
   
 }
 
-figure_gmin_review_2 <- function(gdfr, minags){
+figure_gmin_review_2 <- function(gdfr){
   par(mar=c(3,5,1,0.5), cex.lab=1.2)
   fit <- lm(log10(gmin) ~ method-1, data=gdfr)
   cis <- confint(fit)
   quans <- sapply(split(gdfr, gdfr$method), function(x)quantile(log10(x$gmin), probs=c(0.05, 0.95)))
   gmins <- with(gdfr, tapply(log10(gmin), method, mean, na.rm=TRUE))
   
-  plot(1:5, 10^gmins, pch=19, cex=1.2, xlab="",
+  plot(1:6, 10^gmins, pch=19, cex=1.2, xlab="",
        axes=FALSE,
-       panel.first=segments(x0=1:5, x1=1:5, y0=10^quans[1,], y1=10^quans[2,], col="grey"),
+       #panel.first=segments(x0=1:5, x1=1:5, y0=10^quans[1,], y1=10^quans[2,], col="grey"),
        ylim=c(0,50),
        xlim=c(0.5,6.5),
        ylab=expression(Conductance~~(mmol~m^-2~s^-1)))
@@ -46,19 +46,15 @@ figure_gmin_review_2 <- function(gdfr, minags){
                            expression(g["s,low A"])))
   axis(2)
   box()
-  arrows(x0=1:5, x1=1:5, y0=10^cis[,1], y1=10^cis[,2], angle=90, length=0.1, code=3)
-  
-  quang <- quantile(minags$gsmin, probs=c(0.05, 0.95))
-  points(6, 1000*mean(minags$gsmin), pch=19, cex=1.2,
-         panel.first=segments(x0=6, x1=6, y0=1000*quang[1], y1=1000*quang[2], col="grey"))
-  cig <- 1000*(mean(minags$gsmin) + c(-2,2)*sd(minags$gsmin)/sqrt(nrow(minags)))
-  arrows(x0=6, x1=6, y0=cig[1], y1=cig[2], angle=90, length=0.1, code=3)
+  arrows(x0=1:6, x1=1:6, y0=10^cis[,1], y1=10^cis[,2], angle=90, length=0.1, code=3)
 }
 
 
-figure_gmin_review_3 <- function(kerst2){
+figure_gmin_review_3 <- function(dat){
   
[…]

### 2017-03-29 R/load.R

@@ -17,12 +17,17 @@ source("R/figures.R")
 if(!file.exists("data/trydb.rds")){
   tryfile <- "data/TRY_Categorical_Traits_Lookup_Table_2012_03_17_TestRelease.xlsx"
   if(!file.exists(tryfile))stop("Place the TRY look up table in data/.")
-  trydb <- read_excel(tryfile)
-  trydb <- trydb[,c("AccSpeciesName","PhylogeneticGroup","PlantGrowthForm","LeafType","LeafPhenology")]
+  trydb <- read_excel(tryfile) %>% 
+    rename(species = AccSpeciesName)
+  trydb <- trydb[,c("species","PhylogeneticGroup","PlantGrowthForm","LeafType","LeafPhenology")]
   saveRDS(trydb, "data/trydb.rds")
 } else {
   trydb <- readRDS("data/trydb.rds")
 }
+
+# Additional growth form / phylogenetic group from googling missing species
+gf_add <- read.csv("data/growthform_additional.csv", stringsAsFactors = FALSE)
+trydb <- bind_rows(trydb, gf_add)
 
 
 
@@ -49,15 +54,23 @@ lin2015a <- group_by(lin2015, fitgroup) %>%
 lin2015coef <- fits_lin2015(lin2015a)
 
 
+
+
 # Riederer and Schreiber 2001
-rieder <- read.csv("data/riederer2001_table1.csv") %>%
-  mutate(gmin = 10^5 * 10^-3 * gmin / 41)
+rieder <- read.csv("data/riederer2001_table1.csv", stringsAsFactors = FALSE) %>%
+  filter(gmin < 10000) %>%
+  mutate(gmin = 2 * 10^5 * 10^-3 * gmin / 41,
+         source = "Riederer and Schreiber 2001",
+         method = "gcut_isol")
 
 # Kerstiens 1996
 kerst <- read.csv("data/kerstiens1996_table1.csv", stringsAsFactors = FALSE) %>%
   filter(gmin > 0,
+         species != "Molinia caerulea",
          method %in% c("A","C","D")) %>%
-  mutate(method = dplyr::recode(method, A="gcut_isol", C="gcut_seal", D="gmin"))
+  mutate(method = dplyr::recode(method, A="gcut_isol", C="gcut_seal", D="gmin"),
+         gmin = 2 * 10^5 * 10^-3 * gmin / 41,
+         source = "Kerstiens 1996")
 # Notes:
 # A - astomatous cuticle, removed from leaf (permeance)
[…]

### 2017-03-29 R/make_figures.R

@@ -5,8 +5,8 @@ to.pdf(figure_g0g1_cor(lin2015, "Nicolas Martin-StPaul_Quercus ilex_StPaul_Puech
 
 to.pdf(figure_R2g0(lin2015coef, miner), "output/figure3.pdf", height=4, width=8)
 
-to.pdf(figure_gmin_review_2(gdfr, minags), "output/figure4.pdf", height=4, width=4.5)
-to.pdf(figure_gmin_review_3(kerst2), "output/figure5.pdf", height=4, width=3.5)
+to.pdf(figure_gmin_review_2(gdfr), "output/figure4.pdf", height=4, width=4.5)
+to.pdf(figure_gmin_review_3(condreview_ave), "output/figure5.pdf", height=4, width=3.5)
 #to.pdf(figure_amings_lin2015), "output/figure5.pdf", height=5, width=5)
 
 #to.pdf(figure_wtc4_gmin(wtc4gmin, wtc4gdark), "output/figure5.pdf", height=4, width=8)

### 2017-03-29 R/make_phillips_2010.R

@@ -1,0 +1,85 @@
+
+library(stringr)
+
+phil <- as.data.frame(read_excel("data/Cuticular Conductance 6-1-10 (RA).xls", skip=7))
+
+names(phil)[c(1:2, ncol(phil))] <- c("species", "leafage", "leafarea")
+phil <- dplyr::select(phil, species, leafage, leafarea, everything())
+
+# time information into columns; useful for reshape
+names(phil)[4:ncol(phil)] <- paste0("time.", phil[1,4:ncol(phil)])
+
+
+phil <- slice(phil, -c(1:2)) %>%
+  filter(!is.na(species))
+  
+
+phill <- reshape(as.data.frame(phil), varying=4:ncol(phil), direction="long", timevar="Time") %>%
+  rename(weight=time) %>% 
+  mutate(species = str_trim(species),
+         leafage = str_trim(leafage)) %>%
+  arrange(id, Time)
+
+
+#for(i in 1:30)with(subset(phill, id == i), plot(Time, weight, main=i))
+
+badcurves <- c(3,7,23,25,28,29)
+
+phillsub <- filter(phill, 
+                   !id %in% badcurves,
+                   Time < 1400)
+
+
+
+library(segmented)
+x <- subset(phillsub, id ==1)
+fit <- lm(weight ~ Time, data=x)
+seg <- segmented(fit, seg.Z = ~Time)
+
+
+for(i in unique(phillsub$id)){
+  x <- subset(phillsub, id == i)
+  fit <- lm(weight ~ Time, data=x)
+  seg <- segmented(fit, seg.Z = ~Time)
+  
+  with(x, plot(Time, weight, main=i))
+  plot(seg, add=TRUE)
+}
+
+
+phils <- split(phillsub, phillsub$id)
+fits <- lapply(phils, function(x){
+  fit <- lm(weight ~ Time, data=x)
+  return(segmented(fit, seg.Z = ~Time))
+})
+
+p <- sapply(fits, coef)
+
+# coefficients are intercept, slope section 1, slope section 2 as difference from section 1
+# don't know what coefficient 4 is but always zero
+
+# emin in g min-1
+emin <- apply(p, 2, function(x)x[2] + x[3])
+
+emindf <- data.frame(id = as.numeric(names(emin)), emin=-as.vector(emin))
+
+m <- dplyr::select(phillsub, id, species, leafage, leafarea) %>% distinct()
[…]

### 2017-03-29 manuscript.Rmd

--- 2017-03-27 manuscript.Rmd
@@
 
+[@hari1986; @hari1999] does not mention it (their Eq 6 assumes g0 = 0)
+
 g~0~ modifies not just conductance in zero light, but also low light and high VPD/ high T.
@@
  **Cite Figure 1**
+
+We need a non-zero minimum conductance when simulating water fluxes in very dry soil [@blackman2016; @anderegg2012]
+Also @gleason2014
 

### 2017-03-29 manuscript.Rmd
Manuscript organized properly

@@
 ---
-title: "On g~0~"
+title: "On the minimum leaf conductance g~0~"
 author: ''
@@
     theme: flatly
+    number_sections: true
   word_document:
@@
 ```{r echo=FALSE}
-includefigs <- F
+includefigs <- T
 library(png)
@@
 
-# new outline
-
-1. Introduction (short): questions about g0.
-2. Need for inclusion of g0 in models
-3. Comparison of values for various possible definitions
-4. Estimation of g~0~ by fitting 
-5. Nighttime conductance (v. short)
-6. Cuticular conductance (short)
-7. Minimum conductance (longer)
-8. Recommendations and future work
-
-# Abstract
-
-# Introduction
+# Introduction: questions about g0.
 
@@
 
-# Methods
+@manzoni2011 assumes minimum conductance is cuticular only, because CO2 does not diffuse across it but water does.
+Seems to work in his modelling context but appears to be in disagreement with the observation that gmin >> gcut.
 
-## Data
-
-We compare estimates of g~cut~, g~min~, and g~dark~ from the literature.
-Measurements of g~cut~ and g~min~ were taken from the compilation by @kerstiens1996, with additional g~min~ measurements from @schuster2016, @burghardt2003, @cavenderbares2007, and @phillips2010 (for the latter study, raw data were obtained and g~min~ calculated).
-It is customary for g~cut~ and g~min~ to be expressed per unit total leaf surface area, however, since we aim a comparison with g~s~ and g~dark~, all values were converted to per unit projected surface area.
-Units were converted to molar fluxes assuming 41 mol m^-3^ (cf.
-@kerstiens1996).
-For g~dark~, we used a recent compilation of measurements from @lombardozzi2017.
-
-For the combined dataset, we looked up plant growth form and phylogenetic group (angiosperm or gymnosperm) from the TRY categorical traits database [@kattge2011].
-
-We present two case studies demonstrating the acclimation of g~min~ to environmental drivers.
-Both studies use identical methods.
-The first dataset was collected on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature.
-Leaves were sampled from XXX, n=XXX.
-The second case study was on *Hakea* plants.
-Details here (measurement T?).
-Methods for gmin from drying curves here.
-
-## Model
-
-\
-(2)
-$$ A_n = \frac{g_s}{1.6}(C_a - C_i) $$
-
-\
-where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~).
-
-\
-(3)
-$$ A_n = \frac{a \cdot (C_i - \Gamma^{*})}{C_i + b} - R_d$$
-
-where a = V~cmax~ or ~J/4~ (where J goes to Jmax at high light) depending on the limitation to photosynthesis, and b also varies.
-Dependence on temperature enters through V~cmax~, J~max~, R~d~, $\Gamma^{*}$ [@medlyn2002].
-
-Implemented in the *plantecophys* R package [@duursma2015]
-
-# Do we need g~0~ in simulations of water and carbon fluxes?
+# Need for inclusion of g0 in models
 
@@
 
- **Cite Figure 1**
-
-We need a non-zero minimum conductance when simulating water fluxes in very dry soil [@blackman2016; @anderegg2012]
-Also @gleason2014
-
-# g~0~ is poorly estimated from regression
-
-"We have assumed in the following equations that when stomata are closed the leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." [@jarvis1976]
-
-The difficulty of this can be seen from the fact that a) many studies set g0 to something rather than fitting it (e.g.
-@leuning1995 uses 0.01 for all); presumably because unreliable estimates were obtained, b) negative fitted values are common [@leuning1995; @heroult2013; @miner2017].
-
-Barnard and Bauerle said so too but gave no details.
-Even in very good datasets, g0 and g1 are strongly correlated (Figure 2).
-Across datasets, g0 increases when the model fits poorly (Figure 3), and the width of the confidence interval increases when less variation in a ball-berry coefficient.
-Often g0 is even negative [@heroult2013; @leuning1990], so it is often set to some arbitrary value
-
-When fitting a Ball-Berry type model, negative values are quite common.
-A typical approach is to avoid this by setting some value based on guesswork [e.g.
-@leuning1990], which we assume is quite common when inspecting the many rounded values in the review by
-
-g~0~ fitted is higher than minimum g~s~ from datasets.
-If we restric the fittings to those were g~0~ had a relatively narrow width of the confidence interval (< 0.03 mol m-2 s-1), then the average was `r with(subset(lin2015coef, g0_uci - g0_lci < 0.03), mean(g0))` compared to `r mean(lin2015coef$g0)` for all.
-
-Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015]
-
-**Cite Figure 2 and 3**
-
-# Definitions of minimum conductance
-
-- gcut (cuticle without stomata)
-- gmin (minimum conductance in situ, stomata closed)
-- g0 (minimum conductance in the Ball-Berry-type model)
-- gdark (nighttime conductance)
-
-gdark > gmin > gcut
-g0?
-
-**Cite Figure 4**
-
-@schuster2016 thermal tolerance and gmin relationship across species.
-
 @egea2011 Assumes g0 is constant, does not vary with water stress, and that it represents a ‘cuticular conductance’.
 
-@waldencoleman2013 Very interesting!
-gdark is different from gmin.
-Whole-plant WUE correlated with gdark (but not gmin).
-Lots of references on gdark and gmin and their variation.
-Closed system to measure gdark+gmin; gmin on detached leaves.
+\
+(2)
+$$ A_n = \frac{g_s}{1.6}(C_a - C_i) $$
 
-@muchow1989 correlation stomatal density and gmin suggesting leaky stomata
+\
+where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~).
 
-@brodribb2014 59-94% of transpiration during leaf dry down was from stomatal side of the leaf in seven conifer species, suggesting large component of gmin is leaky stomata.
+\
+(3)
+$$ A_n = \frac{a \cdot (C_i - \Gamma^{*})}{C_i + b} - R_d$$
 
-@brodribb2014 gmin across conifer species was correlated with the rainfall of the driest quarter, again suggesting gmin is a plastic trait
+where a = V~cmax~ or ~J/4~ (where J goes to Jmax at high light) depending on the limitation to photosynthesis, and b also varies.
+Dependence on temperature enters through V~cmax~, J~max~, R~d~, $\Gamma^{*}$ [@medlyn2002].
 
-## Cuticular permeance
-
-Not really a conductance, since water does not diffuse as a gas through the cuticle.
-Instead it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle.
-The main barrier to diffusion is actually a thin layer at the leaf surface, and hence gcut does not depend on the thickness of the cuticle  [@riederer2001; @anfodillo2002].
-@onoda2012 found that thicker cuticles are more resistant to tearing (force to tear and tensile strength both approx proportional to cuticle thickness)
-
-gcut is probably too low in the way it is measured (as a permeance), due to low water content of the cuticle in those conditions [@kerstiens1996]
-
-gseal on average not different from gcut - but for individual studies they frequently differ a lot indicating variable role of stomatal contribution (santrucek)
-
-## Night-time conductance
-
-Nighttime conductance not necessarily same as minimum conductance
-We are not reviewing nighttime conductance here, that has been done extensively by others [@caird2007]
-
-gnight not constant 
-@barbour2007 considerable stomatal control during the night; mirrors daytime
-@howard2007 nighttime conductance lower in drought. Nighttime 2-5x higher than cuticular.
-
-@misson2004 g0 decreases with pre-dawn water potential
-
-Stomata simply do not close all the way, might stay open regardless of photosynthesis, or are broken.
-
-@rescodedios2016 early night gs much (much) lower than pre-dawn gs, in E. camaldulensis
-@rescodedios2013 nighttime fluctuation gs independent of environment
-
-@caird2007 For many species, gnight is not stable throughout the night period.
-Endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [lots of refs]
-
-@korner1994 table of minimum conductance. Also points out (without reference) that g0 in droughted leaves is certainly lower than g0 in dark or very high CO2 leaves.
-
-@howard2007 gdark much higher than gmin (Helianthus annuus). In general this is the case [@caird2007]
-
-## Minimum conductance from leaf-dry down curves
-
-Hygen 1951 early work showing MLD method
-
-true cuticular
-MLD method (Figure 6) gives higher values but this is probably more useful for models (since these are whole leaves)
-
-santrucek Direct measurements showing much higher 'cuticular' conductance of the leaf side with stomata, though stomata were closed.
-In agreement with @kerstiens1996 who showed 10x higher gmin for stomatal side vs.
-non-stomatal cuticle.
-
-@riederer2001 Review on cuticular conductance.
-gcut is not correlated with cuticle thickness.
-No reference to stomata.
-Cuticles are even less permeable than polymers of equivalent thickness used in packaging.
-
-@fanourakis2013 gmin (desiccation gmin) much higher in high RH grown plants, at any value of leaf water content (Table 1).
-This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss (Hypothesis 2).
-
-@burghardt2003 gmin was the same as cuticular permeance (Table 3) (Gmin is lowest gs in desiccating leaves, which includes cuticle and closed stomata).
-This demonstrates that stomata were able to completely close, and not lose significant amounts of water (with the exception of Hedera helix).
-
-conductance: cuticle + closed stomata + cracks, broken stomata
-
-- Kerstiens
-@boyer1997 shows that water but not CO~2~ diffuses across non-stomatal cuticles.
-
-@manzoni2011 assumes minimum conductance is cuticular only, because CO2 does not diffuse across it but water does.
-Seems to work in his modelling context but appears to be in disagreement with the observation that gmin >> gcut.
-
-- gmin as a plastic trait. Temperature response generally unresolved [@kerstiens2006]
-- @riederer2001 correlation between temperature and gcut (non-stomatal cuticles)
-- @schuster2016 correlation temperature, gcut AND gmin for Nerium oleander (page 59).
-T range much wider than Blackman though...
-But Schuster gmin not different from gcut at most temperatures, which is a bit strange too.
-page 90 few gmin values from drying curves that we can add.
-
-- @burghardt2003 gmin review
-- @schreiber1996 cuticle review
-
-[@anfodillo2002] gmin high altitude.
-Four-fold increase gmin from February (winter) to August in Picea abies, probably due to incomplete stomatal closure not changes in cuticle properties.
-Cite common finding that gmin increases with altitude (top left page 485).
-
-[@herrick1991] gmin increased with altitude in Picea rubens.
-
-[@jordan2007] large increase gmin in the herb Agastachys odorata with leaf age
-
-Sack et al. 2010 protocol for measurement
-
-[@cavenderbares2007] gdark similar to gmin in oaks (unusual!). Lower gmin in droughted oaks.
-
-[@duarte2016] fitted g0 with BB and BBOpti.
-Lower g0 in heat wave treatment.
-Also lower gdark in heat wave treatment.
-But g1 or m not affected by heat waves (again example where a gmin is a responsive parameter!)
-
-** cite figures 5,6 and 7**
-
-# Future directions
-
-## Figures
+Implemented in the *plantecophys* R package [@duursma2015]
 
@@
 Parameters were, etc.
+
+We need a non-zero minimum conductance when simulating water fluxes in very dry soil [@blackman2016; @anderegg2012]
+Also @gleason2014, @sinclair2000
+
+# Comparison of values for various definitions
+
+**Table 1.** Definitions of minimum conductance
+
+| Variable   |  Definition |
+| ---------- | ----------- |
+| g~cut~     | Conductance of an isolated non-stomatal cuticle         |
+| g~min~     | Minimum conductance measured from a leaf dry-down curve |
+| g~dark~    | Night-time conductance                                  |
+| g~0~       | Intercept in the Ball-Berry-type stomatal conductance model, fitted. |
+
+We compare estimates of g~cut~, g~min~, and g~dark~ from the literature.
+Measurements of g~cut~ and g~min~ were taken from the compilation by @kerstiens1996, with additional g~min~ measurements from @schuster2016, @burghardt2003, @cavenderbares2007, and @phillips2010 (for the latter study, raw data were obtained and g~min~ calculated).
+It is customary for g~cut~ and g~min~ to be expressed per unit total leaf surface area, however, since we aim a comparison with g~s~ and g~dark~, all values were converted to per unit projected surface area.
+Units were converted to molar fluxes assuming 41 mol m^-3^ (cf.
+@kerstiens1996).
+For g~dark~, we used a recent compilation of measurements from @lombardozzi2017.
+
+For the combined dataset, we looked up plant growth form and phylogenetic group (angiosperm or gymnosperm) from the TRY categorical traits database [@kattge2011].
+
+@korner1994 table of minimum conductance. Also points out (without reference) that g0 in droughted leaves is certainly lower than g0 in dark or very high CO2 leaves.
+
+@waldencoleman2013 gdark is much higher than gmin.
+Whole-plant WUE correlated with gdark (but not gmin).
+Lots of references on gdark and gmin and their variation.
+Closed system to measure gdark+gmin; gmin on detached leaves.
+
+```{r fig.width=4,echo=FALSE, eval=includefigs}
+plotfig("output/figure4.png")
+```
+
+**Figure 4**.
+Comparison of various estimates of minimum conductance.
+Error bars are standard errors, grey vertical lines 90% quantiles indicating wide range of variation for any variable.
+g~cut,isol~ : cuticual permeance measured on isolated cuticles that lack stomata (data from Kerstiens 1996), g~min,seal~ : minimum conductance based on leaf dry-down curves where the side of the leaf that bears stomata is sealed (data from Kerstiens 1996), g~min~ : minimum conductance based on leaf-dry-down curves (data from Kerstiens 1996, with a few extra data points - to be added), g~dark : stomatal conductance in the dark at night, measured in field conditions (data from Lombardozzi et al.
+2017), g~0~ : fitted g~0~ in the Medlyn et al.
+(2011) model with the Lin et al.
+(2015) database, g~s,low A~ : for the Lin et al.
+(2015) database, minimum g~s~ when net assimilation rate was less than 2 $mu$ mol m^-2^ s^-1^, calculated once for each dataset.
+Minimum conductance and cuticular conductance are commonly expressed per all surface area but here projected was used because that's more common for g0, gdark.
+
+# Estimation of g~0~ by fitting
+
+"We have assumed in the following equations that when stomata are closed the leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." [@jarvis1976]
+
+The difficulty of this can be seen from the fact that a) many studies set g0 to something rather than fitting it (e.g.
+@leuning1995 uses 0.01 for all); presumably because unreliable estimates were obtained, b) negative fitted values are common [@leuning1995; @heroult2013; @miner2017].
+
+[@barnard2013] mention difficulty of fitting but give no details.
+
+Even in very good datasets, g0 and g1 are strongly correlated (Figure 2).
+Across datasets, g0 increases when the model fits poorly (Figure 3), and the width of the confidence interval increases when less variation in a ball-berry coefficient.
+
+Often g0 is even negative [@heroult2013; @leuning1990], so it is often set to some arbitrary value
+
+When fitting a Ball-Berry type model, negative values are quite common.
+A typical approach is to avoid this by setting some value based on guesswork [e.g.
+@leuning1990], which we assume is quite common when inspecting the many rounded values in the review by
+
+g~0~ fitted is higher than minimum g~s~ from datasets.
+If we restrict the fittings to those were g~0~ had a relatively narrow width of the confidence interval (< 0.03 mol m-2 s-1), then the average was `r with(subset(lin2015coef, g0_uci - g0_lci < 0.03), mean(g0))` compared to `r mean(lin2015coef$g0)` for all.
+
+Relatively few reports on the plasticity of g~0~, perhaps because it is so poorly estimated. Exceptions:
+
+[@duarte2016] fitted g0 with BB and BBOpti.
+Lower g0 in heat wave treatment.
+Also lower gdark in heat wave treatment.
+But g1 or m not affected by heat waves (again example where a gmin is a responsive parameter!)
 
@@
 
-```{r fig.width=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure4.png")
-```
+# Nighttime conductance (v. short)
 
-**Figure 4**.
-Comparison of various estimates of minimum conductance.
-Error bars are standard errors, grey vertical lines 90% quantiles indicating wide range of variation for any variable.
-g~cut,isol~ : cuticual permeance measured on isolated cuticles that lack stomata (data from Kerstiens 1996), g~min,seal~ : minimum conductance based on leaf dry-down curves where the side of the leaf that bears stomata is sealed (data from Kerstiens 1996), g~min~ : minimum conductance based on leaf-dry-down curves (data from Kerstiens 1996, with a few extra data points - to be added), g~dark : stomatal conductance in the dark at night, measured in field conditions (data from Lombardozzi et al.
-2017), g~0~ : fitted g~0~ in the Medlyn et al.
-(2011) model with the Lin et al.
-(2015) database, g~s,low A~ : for the Lin et al.
-(2015) database, minimum g~s~ when net assimilation rate was less than 2 $mu$ mol m^-2^ s^-1^, calculated once for each dataset.
-Minimum conductance and cuticular conductance are commonly expressed per all surface area but here projected was used because that's more common for g0, gdark.
+- gdark is variable during the night
+- gdark is an active process not just a fixed parameter
+- gs very early morning is quite high because in many cases gs increases before the sun comes up
+- reviewed extensively elsewhere
+
+Nighttime conductance not necessarily same as minimum conductance
+We are not reviewing nighttime conductance here, that has been done extensively by others [@caird2007]
+
+gnight not constant 
+@barbour2007 considerable stomatal control during the night; mirrors daytime
+@howard2007 nighttime conductance lower in drought. Nighttime 2-5x higher than cuticular.
+
+@misson2004 g0 decreases with pre-dawn water potential
+
+@rescodedios2016 early night gs much (much) lower than pre-dawn gs, in E. camaldulensis
+@rescodedios2013 nighttime fluctuation gs independent of environment
+
+@caird2007 For many species, gnight is not stable throughout the night period.
+Endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [lots of refs]
+
+@howard2007 gdark much higher than gmin (Helianthus annuus). In general this is the case [@caird2007]
+
+#  Cuticular conductance (short)
+
+- true cuticular conductance is that of isolated non-stomatal peels, and is very low compared to other definitions
+- what is it, and is it variable with anything? yes, temperature, humidity.
+
+Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015]
+
+Details of cuticular composition, modes of transport, and composition has been reviewed 
+[@kerstiens1996; @kerstiens2006; @schreiber1996; @schreiber2001]
+
+Not really a conductance, since water does not diffuse as a gas through the cuticle.
+Instead it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle.
+The main barrier to diffusion is actually a thin layer at the leaf surface, and hence gcut does not depend on the thickness of the cuticle  [@riederer2001; @anfodillo2002].
+@onoda2012 found that thicker cuticles are more resistant to tearing (force to tear and tensile strength both approx proportional to cuticle thickness)
+
+gcut is probably too low in the way it is measured (as a permeance), due to low water content of the cuticle in those conditions [@kerstiens1996]
+
+@riederer2001 Review on cuticular conductance.
+gcut is not correlated with cuticle thickness.
+No reference to stomata.
+Cuticles are even less permeable than polymers of equivalent thickness used in packaging.
+
+- @riederer2001 correlation between temperature and gcut (non-stomatal cuticles)
+
+# Minimum conductance (longer)
+
+[@hygen1951] early work showing MLD method
+
+Sack et al. 2010 protocol for measurement
+
+- similar to minimum gs observed for the Lin dataset
+- plastic trait
+
+One reason gmin is higher than gcut is the role of stomata, even when presumably shut or as shut as they will be.
+
+gseal on average not different from gcut - but for individual studies they frequently differ a lot indicating variable role of stomatal contribution [@santrucek2004; @brodribb2014].
+@muchow1989 correlation stomatal density and gmin suggesting leaky stomata.
+[@santrucek2004] Direct measurements showing much higher conductance of the leaf side with stomata, though stomata were closed.
+
+@brodribb2014 59-94% of transpiration during leaf dry down was from stomatal side of the leaf in seven conifer species, suggesting large component of gmin is leaky stomata.
+
+but not always, in @burghardt2003 gmin was the same as cuticular permeance (Gmin is lowest gs in desiccating leaves, which includes cuticle and closed stomata).
+This demonstrates that stomata were able to completely close, and not lose significant amounts of water (with the exception of Hedera helix).
+
+## Minimum conductance acclimates to growing conditions
+
+@anfodillo2002 gmin high altitude.
+Four-fold increase gmin from February (winter) to August in Picea abies, probably due to incomplete stomatal closure not changes in cuticle properties.
+Cite common finding that gmin increases with altitude (top left page 485).
+
+@herrick1991 gmin increased with altitude in Picea rubens.
+
+Hypothesis that conifers have low gmin may have to be revisited in the light of @brodribb2014 (check what they actually say!)
+
+@jordan2007 large increase gmin in the herb Agastachys odorata with leaf age
+
+@fanourakis2013 gmin (desiccation gmin) much higher in high RH grown plants, at any value of leaf water content (Table 1).
+This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss (Hypothesis 2).
+
+@schuster2016 thermal tolerance and gmin relationship across species.
+
+@brodribb2014 gmin across conifer species was correlated with the rainfall of the driest quarter, again suggesting gmin is a plastic trait
+
+[@cavenderbares2007] gdark similar to gmin in oaks (unusual!). Lower gmin in droughted oaks.
+
+- @schuster2016 correlation temperature, gcut AND gmin for Nerium oleander (page 59).
+T range much wider than Blackman though...
+But Schuster gmin not different from gcut at most temperatures, which is a bit strange too.
+page 90 few gmin values from drying curves that we can add.
+
+We present two case studies demonstrating the acclimation of g~min~ to environmental drivers.
+Both studies use identical methods.
+The first dataset was collected on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature.
+Leaves were sampled from XXX, n=XXX.
+The second case study was on *Hakea* plants.
+Details here (measurement T?).
+Methods for gmin from drying curves here.
 
@@
 
+# Recommendations and future work
+
+g0 at the moment covers everything, but we need separate ones (night, high VPD/T, drought, low PAR during day)
+g0 is poorly estimated from regression - we need to figure out whether it can be measured independently
+one option is to use gmin in models, perhaps as averages by growth form / group, but that approach will have to be tested
+g0 is certainly not the same as gdark despite the fact that on average they have similar magnitudes. We know for sure gs can be much lower than fitted g0 and gdark in field conditions.
+
 # References {.unnumbered}

### 2017-03-29 R/figures.R

@@ -372,7 +372,7 @@ figure_hakea_gmin <- function(lopw){
       mgp=c(2.5, 0.5,0), tcl=0.2, cex.axis=0.9, cex.lab=1.1)
   m <- t(as.matrix(lopw2[,c("gmin.w","gmin.d")]))
   b <- barplot(m, beside=T,
-          ylim=c(0,10),
+          ylim=c(0,20),
           col=bar_cols,
           names.arg=abbrev_hak(lopw2$species),
           ylab=expression(g[min]~~(mmol~m^-2~s^-1)))

### 2017-03-29 R/make_figures.R

@@ -17,14 +17,3 @@ to.pdf(figure_hakea_gmin(lopw), "output/figure7.pdf", height=5, width=6)
 
 convert_all_pdf("output")
 
-
-
-
-
-# pdf("allfits.pdf")
-# for(g in lin2015coef$fitgroup){
-#   figure_g0g1_cor(lin2015, g)
-# }
-# dev.off()
-
-

### 2017-03-29 manuscript.Rmd
kala

@@
 output:
-  html_document:
-    theme: flatly
-    number_sections: true
   word_document:
     reference_docx: manuscriptstyle.docx
+  html_document:
+    number_sections: yes
+    theme: flatly
 csl: global-change-biology.csl
@@
 
-We argue that
-
-- we need g0 to get responses at low light correct, and probably at high T
-- g0 is poorly estimated as the intercept, with the usual regression technique
-- can we estimate g0 independently?
-Some assume that g0 = gdark (also @miner2017 !) but that's rubbish.
-It is also much larger than cuticular conductance, but seems roughly consistent with gmin (minimum conductance from leaf dry-down curves).
-- there is inter-specific variation in gmin that is related to growth form 
-- gmin is a plastic trait as it responds to growth at elevated temperature, and growth in dry soil
-- models should consider night-time conductance and daily minimum conductance as distinctly separate processes.
-
 g~0~ is often defined as g~s~ when A approaches zero when PPFD approaches zero [@leuning1995] but it is clear from Eq.
@@
 Seems to work in his modelling context but appears to be in disagreement with the observation that gmin >> gcut.
+
+@kala2016 g0 effect heat waves etc.
 

### 2017-03-30 manuscript.Rmd
adjusted some figures

@@
 
-The g~0~ parameter is the intercept, i.e. g~s~ when A is zero. Depending on the form of f(D), it can also represent g~s~ when D is very large.
+The origin of Eq.
+1 is @ball1987 , but no g~0~ was mentioned.
+It is introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model.
 
-The objective of this paper is to review sources of variation and common misconceptions on the g~0~ parameter.
+@leuning1990 defined g~0~ as g~s~ when A approaches zero as PPFD approaches the light compensation point, and also points out that it plays a role in predicting A/g~s~ (and C~i~) at low light [see also @leuning1995].
+
+[@collatz1992] shows that including g0 affects response of gs to RH in the coupled leaf gas exchange model.
+And "It seems possible that other conditions such as severe drought stress may affect the minimum value of  g  as well, though we currently have no evidence to support this."
 
@@
 
-One estimate of g~0~ (low PPFD) comes from night-time measurements of conductance, which have been known for a long time to be substantially larger than zero.
-It is however not clear whether daytime g~0~ at low PPFD, for darkened leaves, is comparable in magnitude to g~0~ in the dark at night.
-Evidence from experiments where environmental conditions are held constant suggests that g~s~ varies tremendously throughout the day independent of these drivers, and has therefore been attributed to an internal clock (Resco).
+Though the role and potential for estimation using Eq.
+1 is clear, it is not clear whether and how g~0~ compares to other estimates of g~s~ when photosynthesis is zero.
+It is sometimes assumed that g~0~ must equal g~s~ at night, simply since no photosynthesis occurs at night [@lombardozzi2016; @barnard2013].
+Other studies assume that g~0~ equals a 'cuticular conductance` [@manzoni2011; @egea2011], but it has yet to be determined how various minimum conductance estimates compare.
 
-@manzoni2011 assumes minimum conductance is cuticular only, because CO2 does not diffuse across it but water does.
-Seems to work in his modelling context but appears to be in disagreement with the observation that gmin >> gcut.
-
-@kala2016 g0 effect heat waves etc.
+The objective of this paper is to take a closer look at g~0~, its role in modelling fluxes of carbon and water, estimation of g~0~ from regression of Eq.
+1 against data.
+Finally we review the literature for sources of variation in the actual minimum conductance measured in various ways.
+On the way we try to set straight some common misconceptions about g~0~, and make recommendations for future work.
 
@@
 
-The first Ball-Berry model did not have a g0
-
-@leuning1995 says "g0 > 0, otherwise ci/ca would be constant for all values of Q.
+@leuning1990 and @leuning1995 says "g0 > 0, otherwise ci/ca would be constant for all values of Q.
 The rate of convergence of ci to ca at low A (low Q) depends on the value of g0; when g0 is large, ci deviates from the asymptotic value at higher values of A than when g0 is small".
-
-[@collatz1992] shows that including g0 affects response of gs to RH in the coupled leaf gas exchange model.
-And "It seems possible that other conditions such as severe drought stress may affect the minimum value of  g  as well, though we currently have no evidence to support this."
 
@@
 
-similarly at high T. See also Kala. Figure 1.
+similarly at high T. @kala2016 @urban 
+drake in prep
+duursma2014
 
@@
 @barnard2013 makes the same assumption, without any discussion of the possibility that daytime g0 might be different from nighttime conductance
-
-@dekauwe2015 makes this point as well, their Fig. A2 (can we add that also?)
 
@@
 Measurements of g~cut~ and g~min~ were taken from the compilation by @kerstiens1996, with additional g~min~ measurements from @schuster2016, @burghardt2003, @cavenderbares2007, and @phillips2010 (for the latter study, raw data were obtained and g~min~ calculated).
+We excluded measurements of g~min~ where stomatal leaf sides were sealed, because few data were available and the comparison is not consistent with individual studies where the role of stomata in g~min~ is studied (see SectionXXX).
+
 It is customary for g~cut~ and g~min~ to be expressed per unit total leaf surface area, however, since we aim a comparison with g~s~ and g~dark~, all values were converted to per unit projected surface area.
@@
 
-@waldencoleman2013 gdark is much higher than gmin.
-Whole-plant WUE correlated with gdark (but not gmin).
-Lots of references on gdark and gmin and their variation.
-Closed system to measure gdark+gmin; gmin on detached leaves.
+g~s,minA~ calculated from @lin2015 data. Also g0 fitted using the model of Medlyn et al. 2011, and those from @miner2016 (combined all).
+
+g~s,low A~ much lower than fitted g~0~ or g~dark~ (see also @dekauwe2015, their Fig. A2).
+
+g~dark~ much higher than g~min~, also confirmed by experiments. In soybean, @waldencoleman2013 showed much higher g~dark~  compared to g~min~.
 
@@
 Error bars are standard errors, grey vertical lines 90% quantiles indicating wide range of variation for any variable.
-g~cut,isol~ : cuticual permeance measured on isolated cuticles that lack stomata (data from Kerstiens 1996), g~min,seal~ : minimum conductance based on leaf dry-down curves where the side of the leaf that bears stomata is sealed (data from Kerstiens 1996), g~min~ : minimum conductance based on leaf-dry-down curves (data from Kerstiens 1996, with a few extra data points - to be added), g~dark : stomatal conductance in the dark at night, measured in field conditions (data from Lombardozzi et al.
+g~cut,isol~ : cuticual permeance measured on isolated cuticles that lack stomata (data from Kerstiens 1996), g~min~ : minimum conductance based on leaf-dry-down curves (data from Kerstiens 1996, with a few extra data points - to be added), g~dark : stomatal conductance in the dark at night, measured in field conditions (data from Lombardozzi et al.
 2017), g~0~ : fitted g~0~ in the Medlyn et al.
@@
 # Nighttime conductance (v. short)
+
+One estimate of g~0~ (low PPFD) comes from night-time measurements of conductance, which have been known for a long time to be substantially larger than zero.
+It is however not clear whether daytime g~0~ at low PPFD, for darkened leaves, is comparable in magnitude to g~0~ in the dark at night.
+Evidence from experiments where environmental conditions are held constant suggests that g~s~ varies tremendously throughout the day independent of these drivers, and has therefore been attributed to an internal clock (Resco).
 

### 2017-03-30 R/figures.R

@@ -27,26 +27,31 @@ figure_gmin_review <- function(gdfr){
 
 figure_gmin_review_2 <- function(gdfr){
   par(mar=c(3,5,1,0.5), cex.lab=1.2)
+  
+  # get rid of sealed leaves - small sample size and confusing result
+  gdfr <- droplevels(subset(gdfr, method != "gcut_seal"))
+  
   fit <- lm(log10(gmin) ~ method-1, data=gdfr)
   cis <- confint(fit)
   quans <- sapply(split(gdfr, gdfr$method), function(x)quantile(log10(x$gmin), probs=c(0.05, 0.95)))
   gmins <- with(gdfr, tapply(log10(gmin), method, mean, na.rm=TRUE))
+  ng <- nlevels(gdfr$method)
   
-  plot(1:6, 10^gmins, pch=19, cex=1.2, xlab="",
+  plot(1:ng, 10^gmins, pch=19, cex=1.2, xlab="",
        axes=FALSE,
        #panel.first=segments(x0=1:5, x1=1:5, y0=10^quans[1,], y1=10^quans[2,], col="grey"),
        ylim=c(0,50),
-       xlim=c(0.5,6.5),
+       xlim=c(0.5,ng+0.5),
        ylab=expression(Conductance~~(mmol~m^-2~s^-1)))
-  axis(1, at=1:6, labels=c(expression(g["cut,isol"]),
-                           expression(g["min,seal"]),
+  axis(1, at=1:ng, labels=c(expression(g["cut,isol"]),
+                           #expression(g["min,seal"]),
                            expression(g[min]),
                            expression(g[dark]),
                            expression(g[0]),
                            expression(g["s,low A"])))
   axis(2)
   box()
-  arrows(x0=1:6, x1=1:6, y0=10^cis[,1], y1=10^cis[,2], angle=90, length=0.1, code=3)
+  arrows(x0=1:ng, x1=1:ng, y0=10^cis[,1], y1=10^cis[,2], angle=90, length=0.1, code=3)
 }
 
 
@@ -121,6 +126,7 @@ figure_g0g1_cor <- function(lin2015, group, legend=FALSE){
                ylab=expression(g[s]~(mol~m^-2~s^-1)),
                pch=16, cex=0.8, col="dimgrey"))
   box()
[…]

### 2017-03-30 R/functions.R

@@ -81,3 +81,14 @@ convert_pdf_png <- function(filename, fnout=NULL, overwrite=TRUE, res=600, resiz
   }
 }
 
+# Simple function for placing labels on a figure.
+plotlabel <- function(txt, where, inset=0.08, inset.x=inset, inset.y=inset,...){
+  u <- par()$usr
+  if(grepl("left",where))x <- u[1] + inset.x*(u[2]-u[1])
+  if(grepl("right",where))x <- u[2] - inset.x*(u[2]-u[1])
+  if(grepl("bottom",where))y <- u[3] + inset.y*(u[4]-u[3])
+  if(grepl("top",where))y <- u[4] - inset.y*(u[4]-u[3])
+  
+  text(x,y,txt,font=2,...)
+}
+

### 2017-03-30 manuscript.Rmd
more references

@@
 
+g~0~ usually ignored, for example in recent review of stomatal conductance models [@damour2010]
+
 where f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), g~1~ the slope parameter, and g~0~ an intercept.
@@
 1 is clear, it is not clear whether and how g~0~ compares to other estimates of g~s~ when photosynthesis is zero.
-It is sometimes assumed that g~0~ must equal g~s~ at night, simply since no photosynthesis occurs at night [@lombardozzi2016; @barnard2013].
-Other studies assume that g~0~ equals a 'cuticular conductance` [@manzoni2011; @egea2011], but it has yet to be determined how various minimum conductance estimates compare.
+It is sometimes assumed that g~0~ must equal g~s~ at night, simply since no photosynthesis occurs at night [@lombardozzi2016; @barnard2013; @uddling2005].
+Other studies assume that g~0~ equals a 'cuticular conductance` [@baldocchi1997; @manzoni2011; @egea2011], but it has yet to be determined how various minimum conductance estimates compare.
 
@@
 
-similarly at high T. @kala2016 @urban 
-drake in prep
-duursma2014
+similarly at high T, which is important [@kala2016]
 
-@lombardozzi2017 puts g0 into CLM, notes in introduction that 'conductance during the night (and at other times when A = 0 ) is thus represented using g0'.
-
-@barnard2013 makes the same assumption, without any discussion of the possibility that daytime g0 might be different from nighttime conductance
+The assumption of Eq. 1 is that g~s~ goes to g~0~ both at low light and very high temperature, when A <-- 0 [@duursma2014]. Some studies however show that stomata don't shut at high T [@urban2017].
+Another thing is that the effect is actually larger than g~0~ because this additional conductance allows slightly more stomatal opening (because photosynthesis occurs more because of higher g~s~).
 
@@
 @egea2011 Assumes g0 is constant, does not vary with water stress, and that it represents a ‘cuticular conductance’.
+
+@lombardozzi2017 puts g0 into CLM, notes in introduction that 'conductance during the night (and at other times when A = 0 ) is thus represented using g0'.
+
+@barnard2013 makes the same assumption, without any discussion of the possibility that daytime g0 might be different from nighttime conductance
 

### 2017-05-22 manuscript.Rmd
much writing wow

--- 2017-03-30 manuscript.Rmd
@@
 
-# Introduction: questions about g0.
+# Introduction
 
-Prevalence of the Ball-Berry class of stomatal conductance models characterized by dependence on assimilation rate, CO~2~ and VPD or RH.
+Stomata exert a strong control on the transpiration rate of leaves.
+It is well known that stomata respond quickly to changes in light intensity, humidity and CO~2~ concentration at the leaf surface.
+In doing so, stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus can be expected to close stomata when photosynthetic rates are low.
+To avoid dessiccation and ultimate death, stomata can also be expected to close during periods of severe water stress.
+Plants cannot, however, completely halt water loss even when stomata are closed.
+In this review, we discuss various aspects of the minimum leaf conductance reached when stomata are essentially closed.
+We pay special attention to the role that minimum conductance plays in land-surface models, in particular as it forms part of the much-used 'Ball-Berry'-type class of stomatal conductance models.
+
+Most current-generation process-baed global vegetation models (GVMs) use a version of the Ball-Berry class of stomatal conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity.
 
@@
 
-where f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), g~1~ the slope parameter, and g~0~ an intercept.
-The Medlyn model is actually non-linear but can be expressed in an approximate linear form of Eq.
-1.
+Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), and g~0~ is the value of g~s~ when A~n~ is zero.
+The effects of photosynthetic photon flux density (PPFD) and leaf temperature are both assumed to enter via the dependence of A~n~ on these drivers.
+The g~0~ parameter is another definition of the minimum conductance, but we keep the notation g~0~ to indicate that it is a parameter in the Ball-Berry model.
+Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the inocuous g~0~ parameter has been largely ignored.
+For example, a recent review of stomatal conductance models did not mention the minimum conductance [@damour2010].
+The original description of Eq.
+1 as published by @ball1987 did not include an intercept term.
+It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model.
 
-The origin of Eq.
-1 is @ball1987 , but no g~0~ was mentioned.
-It is introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model.
+The g~0~ parameter not only sets a minimum value of g~s~ in the model, it also modifies the behaviour of Eq.
+1 after it has been coupled to the Farquhar-von Caemmerer-Berry model of photosynthesis (the so-called 'coupled leaf gas exchange model').
+@leuning1990 defined g~0~ as g~s~ when A approaches zero as PPFD approaches the light compensation point, and pointed out that g~0~ is necessary to simulate the increase in A/g~s~ (and C~i~) at low light [see also @leuning1995].
+Similarly, @collatz1992 showed that including g~0~ affects the response of g~s~ to relative humidity in the coupled leaf gas exchange model.
+Despite these early reports, the exact role of g~0~ in models is often overlooked.
 
-@leuning1990 defined g~0~ as g~s~ when A approaches zero as PPFD approaches the light compensation point, and also points out that it plays a role in predicting A/g~s~ (and C~i~) at low light [see also @leuning1995].
+Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero.
+Based on this, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2016; @barnard2013; @uddling2005].
+It is also frequently assumed that g~0~ is the 'cuticular conductance` [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata [@kerstiens1996].
+Another measurement of the minimum conductance is the leaf drying curve: leaves are picked, and gradually dried out and weighed to determine the water loss rate [@hygen1951].
+This estimate of minimum conductance (which we refer to as g~min~) probably most closely reflects the dessiccation dynamics of severely droughted plants [@blackman2016], but has rarely been mentioned in the context of the Ball-Berry model.
 
-[@collatz1992] shows that including g0 affects response of gs to RH in the coupled leaf gas exchange model.
-And "It seems possible that other conditions such as severe drought stress may affect the minimum value of  g  as well, though we currently have no evidence to support this."
+In this review, we aim to compare values of the various estimates of minimum conductance, discuss components and mechanisms of minimum conductance.
+We place additional emphasis on measurements of g~min~, sources of variation in g~min~, and the role that g~min~ can play in models of water use.
+Before we discuss the various definitions and measurements of minimum conductance, we first review the modelling literature on its use of g~0~ in Eq.
+1 (and related approaches), and demonstrate why g~0~ needs to be included in models.
 
-g~0~ is often defined as g~s~ when A approaches zero when PPFD approaches zero [@leuning1995] but it is clear from Eq.
-1 that g~0~ is also attained when A approaches zero due to other environmental drivers, in particular temperature, and drought stress (which may either reduce photosynthetic capacity or g~1~, or both, see Zhou).
-It is not clear nor obvious whether these various minima should equal each other.
+# The use of g~0~ in models
 
-Though the role and potential for estimation using Eq.
-1 is clear, it is not clear whether and how g~0~ compares to other estimates of g~s~ when photosynthesis is zero.
-It is sometimes assumed that g~0~ must equal g~s~ at night, simply since no photosynthesis occurs at night [@lombardozzi2016; @barnard2013].
-Other studies assume that g~0~ equals a 'cuticular conductance` [@manzoni2011; @egea2011], but it has yet to be determined how various minimum conductance estimates compare.
+A popular approach in thinking about how stomata 'should' respond to environmental drivers is the idea that stomatal conductance is varied to optimize the rate of photosynthesis per unit water used [@cowan1977].
+The consequence is that stomata tend to open during periods that are favorable for photosynthesis (high light, optimal temperature), and close when photosynthesis drops to zero (darkness, very high temperature).
+Clearly, in this framework there is no place for g~0~ since it is always sub-optimal to open stomata (i.e.
+spend water) when there is no photosynthetic gain.
+Early work on optimal stomatal conductance models completely ignored the possibility of g~s~ > 0 when A~n~ = 0 [@cowan1977; @cowan1978; @hari1986; @hari1999], whereas more recent work derives the optimal g~s~ and then adds on a g~0~ [@medlyn2011].
 
-The objective of this paper is to take a closer look at g~0~, its role in modelling fluxes of carbon and water, estimation of g~0~ from regression of Eq.
-1 against data.
-Finally we review the literature for sources of variation in the actual minimum conductance measured in various ways.
-On the way we try to set straight some common misconceptions about g~0~, and make recommendations for future work.
+In the following we discuss three reasons to include g~0~ in models: the response of C~i~ at low PPFD, the response of transpiration to high leaf temperature (T~leaf~), and the dessiccation of leaves in severe drought.
 
-# Need for inclusion of g0 in models
+As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) is constant with the photon flux density (PPFD).
+When PPFD approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~.
+To see this point, we can rearrange Eq.
+1 to give,
 
-@leuning1990 and @leuning1995 says "g0 > 0, otherwise ci/ca would be constant for all values of Q.
-The rate of convergence of ci to ca at low A (low Q) depends on the value of g0; when g0 is large, ci deviates from the asymptotic value at higher values of A than when g0 is small".
+\
+(2)
+$$A~n~/g~s~ = {C_a}f(D) / g~1~$$
 
-[@medlyn2011] does not develop theory for g~0~, instead adds it as a minimum conductance because that's how it is!
+and using the diffusion constraint $A~n~ = g~s~ / 1.6 \cdot (C~a~ - C~i~)$ we get an expression for C~i~ / C~a~,
 
-[@hari1986; @hari1999] does not mention it (their Eq 6 assumes g0 = 0)
+\
+(3)
+$$C_i / C_a = 1 - \frac{f(D)}{1.6 \cdot g_1}$$
 
-g~0~ modifies not just conductance in zero light, but also low light and high VPD/ high T.
-It is commonly observed and makes sense that Ci converges to Ca at zero light, but the coupled leaf gas exchange model does not predict that unless a g0 is included [@leuning1995].
+Eq. 3 - derived using g~0~ - thus does not give any dependence on PPFD.
 
-similarly at high T. @kala2016 @urban 
-drake in prep
-duursma2014
+The g~0~ parameter also represents g~s~ when T~leaf~ is very high, because A~n~ goes to zero at high T~leaf~.
+As a result, simulations of vegetation water use during heat waves depends highly on the value of g~0~ used [@kala2016] .
+It should be pointed out that the response of g~s~ to T~leaf~ is likely poorly modelled with a Ball-Berry model, since a number of studies have shown that g~s~ and A~n~ decouple at high T~leaf~, resulting in high rates of g~s~ when A~n~ is near zero [@urban2017] (Drake et al., unpublished).
 
-@lombardozzi2017 puts g0 into CLM, notes in introduction that 'conductance during the night (and at other times when A = 0 ) is thus represented using g0'.
-
-@barnard2013 makes the same assumption, without any discussion of the possibility that daytime g0 might be different from nighttime conductance
-
-@messinger2006 Conclude that Ci response of gs is the same in the dark or when DCMU was applied (which disrupts electron transport).
-However, Fig5A-B indicate that gs was much higher in the dark than when DCMU was applied (although the ‘response’ was similar).
-This indicates gs response to low PAR (darkness) is more than just responding to what photosynthesis does.
-
-@egea2011 Assumes g0 is constant, does not vary with water stress, and that it represents a ‘cuticular conductance’.
+To illustrate the effect of g~0~ on the modelled response of g~s~ to PPFD and T~leaf~, it is necessary to embed the stomatal conductance model in a coupled leaf gas exchange model, which is formed by three equations; Eq.
+1 giving the dependence of g~s~ on A~n~ and some environmental drivers, the diffusion constraint,
 
@@
 \
-where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~).
+where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~). And finally the Farquhar-von Caemmerer-Berry photosynthesis model that gives A~n~ as a function of C~i~,
 
@@
 where a = V~cmax~ or ~J/4~ (where J goes to Jmax at high light) depending on the limitation to photosynthesis, and b also varies.
-Dependence on temperature enters through V~cmax~, J~max~, R~d~, $\Gamma^{*}$ [@medlyn2002].
+The dependence on temperature enters through V~cmax~, J~max~, R~d~, $\Gamma^{*}$, and other parameters [@medlyn2002].
 
-Implemented in the *plantecophys* R package [@duursma2015]
+Using the coupled leaf gas exchange as implemented in the *plantecophys* R package [@duursma2015], we show how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^).
+We also demonstrate the effect of g~0~ on modelling the T~leaf~ response.
+In this simulation, T~leaf~ and D are assumed to co-vary with an empirical relationship as used by @duursma2014.
+Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increase leaf transpiration E~L~, but at a slightly higher rate than just due to g~0~.
+This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq.
+1.
+
+@egea2011 Assumes g0 is constant, does not vary with water stress, and that it represents a ‘cuticular conductance’.
+
+@lombardozzi2017 puts g0 into CLM, notes in introduction that 'conductance during the night (and at other times when A = 0 ) is thus represented using g0'.
+
+@barnard2013 makes the same assumption, without any discussion of the possibility that daytime g0 might be different from nighttime conductance
 
@@
 - reviewed extensively elsewhere
+
+@messinger2006 Conclude that Ci response of gs is the same in the dark or when DCMU was applied (which disrupts electron transport).
+However, Fig5A-B indicate that gs was much higher in the dark than when DCMU was applied (although the ‘response’ was similar).
+This indicates gs response to low PAR (darkness) is more than just responding to what photosynthesis does.
 

### 2017-05-22 manuscript.Rmd
writing (lit review)

@@
 ---
-title: "On the minimum leaf conductance g~0~"
+title: "On the minimum leaf conductance"
 author: ''
@@
 
-```{r echo=FALSE}
-includefigs <- T
+```{r include=FALSE}
+includefigs <- F
 library(png)
@@
 
+\
 Stomata exert a strong control on the transpiration rate of leaves.
@@
 
+\
 Most current-generation process-baed global vegetation models (GVMs) use a version of the Ball-Berry class of stomatal conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity.
@@
 
+\
 Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), and g~0~ is the value of g~s~ when A~n~ is zero.
@@
 The original description of Eq.
-1 as published by @ball1987 did not include an intercept term.
+1 as published by @ballberry1987 did not include an intercept term.
 It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model.
 
+\
 The g~0~ parameter not only sets a minimum value of g~s~ in the model, it also modifies the behaviour of Eq.
@@
 
+\
 Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero.
-Based on this, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2016; @barnard2013; @uddling2005].
+Based on this, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005].
 It is also frequently assumed that g~0~ is the 'cuticular conductance` [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata [@kerstiens1996].
-Another measurement of the minimum conductance is the leaf drying curve: leaves are picked, and gradually dried out and weighed to determine the water loss rate [@hygen1951].
+Another measurement of the minimum conductance arises from leaf drying curve: leaves are picked, and gradually dried out and weighed to determine the water loss rate [@hygen1951].
 This estimate of minimum conductance (which we refer to as g~min~) probably most closely reflects the dessiccation dynamics of severely droughted plants [@blackman2016], but has rarely been mentioned in the context of the Ball-Berry model.
 
+\
 In this review, we aim to compare values of the various estimates of minimum conductance, discuss components and mechanisms of minimum conductance.
-We place additional emphasis on measurements of g~min~, sources of variation in g~min~, and the role that g~min~ can play in models of water use.
+We place additional emphasis on sources of variation in g~min~, and the role that g~min~ can play in models of water use.
 Before we discuss the various definitions and measurements of minimum conductance, we first review the modelling literature on its use of g~0~ in Eq.
@@
 
+\
 A popular approach in thinking about how stomata 'should' respond to environmental drivers is the idea that stomatal conductance is varied to optimize the rate of photosynthesis per unit water used [@cowan1977].
@@
 
+\
 In the following we discuss three reasons to include g~0~ in models: the response of C~i~ at low PPFD, the response of transpiration to high leaf temperature (T~leaf~), and the dessiccation of leaves in severe drought.
 
+\
 As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) is constant with the photon flux density (PPFD).
@@
 (2)
-$$A~n~/g~s~ = {C_a}f(D) / g~1~$$
+$$A_n/g_s = {C_a}f(D) / g_1$$
 
@@
 
+\
 The g~0~ parameter also represents g~s~ when T~leaf~ is very high, because A~n~ goes to zero at high T~leaf~.
@@
 
+\
 To illustrate the effect of g~0~ on the modelled response of g~s~ to PPFD and T~leaf~, it is necessary to embed the stomatal conductance model in a coupled leaf gas exchange model, which is formed by three equations; Eq.
@@
 
-where a = V~cmax~ or ~J/4~ (where J goes to Jmax at high light) depending on the limitation to photosynthesis, and b also varies.
+\
+where a = V~cmax~ or J/4 (where the electron transport rate J approaches Jmax at high light) depending on the limitation to photosynthesis, and b also varies.
 The dependence on temperature enters through V~cmax~, J~max~, R~d~, $\Gamma^{*}$, and other parameters [@medlyn2002].
 
+\
 Using the coupled leaf gas exchange as implemented in the *plantecophys* R package [@duursma2015], we show how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^).
@@
 1.
-
-@egea2011 Assumes g0 is constant, does not vary with water stress, and that it represents a ‘cuticular conductance’.
-
-@lombardozzi2017 puts g0 into CLM, notes in introduction that 'conductance during the night (and at other times when A = 0 ) is thus represented using g0'.
-
-@barnard2013 makes the same assumption, without any discussion of the possibility that daytime g0 might be different from nighttime conductance
 
@@
 
-We need a non-zero minimum conductance when simulating water fluxes in very dry soil [@blackman2016; @anderegg2012]
-Also @gleason2014, @sinclair2000
+\
+The third reason for inclusion of a g~0~ term in models of stomatal conductance arises during severe drought.
+When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata.
+Thus models that aim to predict when plants dessiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000].
+We return to measurements of g~min~ from leaf drying curves in Section XXX.
 
@@
 
-**Table 1.** Definitions of minimum conductance
+**Table 1.** Definitions of minimum conductance.
 
@@
 
-We compare estimates of g~cut~, g~min~, and g~dark~ from the literature.
-Measurements of g~cut~ and g~min~ were taken from the compilation by @kerstiens1996, with additional g~min~ measurements from @schuster2016, @burghardt2003, @cavenderbares2007, and @phillips2010 (for the latter study, raw data were obtained and g~min~ calculated).
+\
+We identify four distinct definitions of minimum conductance, each of which is measured with its own set of methods.
+Before discussing each of the minimum conductances in detail in the following sections, we here compile literature values to compare the relative magnitude of each of the conductances.
+
+\
+Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) and of minimum conductance from leaf dry-down curves (g~min~) were taken from the compilation by @kerstiens1996, with additional g~min~ measurements from @schuster2016, @burghardt2003, @cavenderbares2007, and @phillips2010 (for the latter study, raw data were obtained and g~min~ re-calculated).
 We excluded measurements of g~min~ where stomatal leaf sides were sealed, because few data were available and the comparison is not consistent with individual studies where the role of stomata in g~min~ is studied (see SectionXXX).
+A compilation of night-time (and 'dark') conductance (g~dark~) values was taken from @lombardozzi2017.
+We also estimated g~0~ from Eq.
+1 using the stomatal conductance model of @medlyn2011, using non-linear regression.
+These estimates were combined with the compilation of g~0~ values from @miner2017, who reported on xxx values from a literature compilation (each of those values was estimated with linear regression of the Ball-Berry model).
+Finally, we also estimated minimum values of g~s~ attained from a large database of leaf gas exchange data [@lin2015], by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset.
+This estimate was added as it represents actual minimum values measured, rather than a fitted/extrapolated g~0~ term.
 
-It is customary for g~cut~ and g~min~ to be expressed per unit total leaf surface area, however, since we aim a comparison with g~s~ and g~dark~, all values were converted to per unit projected surface area.
+\
+It is customary for g~cut~ and g~min~ to be expressed per unit total leaf surface area, however, since we aim a comparison with g~0~ and g~dark~, all values were converted to per unit projected surface area.
 Units were converted to molar fluxes assuming 41 mol m^-3^ (cf.
 @kerstiens1996).
-For g~dark~, we used a recent compilation of measurements from @lombardozzi2017.
-
 For the combined dataset, we looked up plant growth form and phylogenetic group (angiosperm or gymnosperm) from the TRY categorical traits database [@kattge2011].
 
-@korner1994 table of minimum conductance. Also points out (without reference) that g0 in droughted leaves is certainly lower than g0 in dark or very high CO2 leaves.
+```{r fig.width=4,echo=FALSE, eval=includefigs}
+plotfig("output/figure4.png")
+```
 
-g~s,minA~ calculated from @lin2015 data. Also g0 fitted using the model of Medlyn et al. 2011, and those from @miner2016 (combined all).
+**Figure 4**. Comparison of various estimates of the minimum conductance. Error bars are standard errors. See Table 1 for definitions of minimum conductance, and text for the data sources.
 
@@
 
-```{r fig.width=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure4.png")
-```
-
-**Figure 4**.
-Comparison of various estimates of minimum conductance.
-Error bars are standard errors, grey vertical lines 90% quantiles indicating wide range of variation for any variable.
-g~cut,isol~ : cuticual permeance measured on isolated cuticles that lack stomata (data from Kerstiens 1996), g~min~ : minimum conductance based on leaf-dry-down curves (data from Kerstiens 1996, with a few extra data points - to be added), g~dark : stomatal conductance in the dark at night, measured in field conditions (data from Lombardozzi et al.
-2017), g~0~ : fitted g~0~ in the Medlyn et al.
-(2011) model with the Lin et al.
-(2015) database, g~s,low A~ : for the Lin et al.
-(2015) database, minimum g~s~ when net assimilation rate was less than 2 $mu$ mol m^-2^ s^-1^, calculated once for each dataset.
-Minimum conductance and cuticular conductance are commonly expressed per all surface area but here projected was used because that's more common for g0, gdark.
+@korner1994 table of minimum conductance. Also points out (without reference) that g0 in droughted leaves is certainly lower than g0 in dark or very high CO2 leaves.
 
@@
 
-"We have assumed in the following equations that when stomata are closed the leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." [@jarvis1976]
+The standard approach in estimating g~0~ for use in Eq.
+1 is to use regression with g~s~ as the response variable, and the right-hand side of Eq.
+1 as the predictor.
+The g~0~ parameter is thus estimated as the intercept.
+Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated in this way.
+The difficulty of the estimation can already be concluded from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g.
+@leuning1995 uses 0.01 for all); presumably because unreliable estimates were obtained, b) negative g~0~ estimates (i.e.
+negative intercepts) are common [@leuning1995; @heroult2013; @miner2017].
+In a pioneering publication which presented the Jarvis model of stomatal conductance, @jarvis1976 states that "we have assumed in the following equations that when stomata are closed the  leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." [@jarvis1976].
+More recently, @barnard2013 mention the difficulty of fitting Eq.
+1 to estimate g~0~ but give no details.
 
-The difficulty of this can be seen from the fact that a) many studies set g0 to something rather than fitting it (e.g.
-@leuning1995 uses 0.01 for all); presumably because unreliable estimates were obtained, b) negative fitted values are common [@leuning1995; @heroult2013; @miner2017].
+The first problem is that generally in linear regression, the intercept and slope are highly correlated.
+In Fig.
+2, we demonstrate this problem.
+The consequence is that it is more difficult to interpret g~1~ and g~0~ independently, because small errors in g~1~ lead to errors in g~0~.
 
-[@barnard2013] mention difficulty of fitting but give no details.
-
-Even in very good datasets, g0 and g1 are strongly correlated (Figure 2).
 Across datasets, g0 increases when the model fits poorly (Figure 3), and the width of the confidence interval increases when less variation in a ball-berry coefficient.
-
-Often g0 is even negative [@heroult2013; @leuning1990], so it is often set to some arbitrary value
-
-When fitting a Ball-Berry type model, negative values are quite common.
-A typical approach is to avoid this by setting some value based on guesswork [e.g.
-@leuning1990], which we assume is quite common when inspecting the many rounded values in the review by
 

### 2017-05-23 manuscript.Rmd
fig update; more writing

@@
 
-The first problem is that generally in linear regression, the intercept and slope are highly correlated.
+It is also telling that there are few reports on the plasticity of g~0~, perhaps because it is so poorly estimated.
+Exceptions are @duarte2016, who found a lower g0 in a heat wave treatment (although g~1~ was unaffected), and @misson2004 who report on a close negative correlation between pre-dawn leaf water potential and g~0~.
+
+\
+Next we demonstrate that the fitting process gives three problems: 1) the estimates of g~0~ and g~1~ are highly correlated, 2) the precision for g~0~ is generally much lower than for g~1~, and 3) for data that have a worse fit overall, the g~0~ estimates are elevated.
+
+\
 In Fig.
-2, we demonstrate this problem.
-The consequence is that it is more difficult to interpret g~1~ and g~0~ independently, because small errors in g~1~ lead to errors in g~0~.
+2, we demonstrate how g~0~ and g~1~ are statistically correlated via a bootstrap sampling approach.
+The correlation between slopes and intercepts is a general aspect of linear models [@stapleton1995].
+The consequence is that it is more difficult to interpret g~1~ and g~0~ independently, and that confidence intervals on each of the parameters should also not be interpreted independently.
+See @becker2007 for more discussion.
 
-Across datasets, g0 increases when the model fits poorly (Figure 3), and the width of the confidence interval increases when less variation in a ball-berry coefficient.
+Across datasets, g0 increases when the model fits poorly (Fig.
+3 a,b), and the standard error (SE) of g~0~ increases when less variation in a ball-berry coefficient (Fig.
+3c).
+Also, SE is very large compared to the average magnitude of g~0~ across studies (compare Fig.
+2 and Fig.
+3).
 
@@
 If we restrict the fittings to those were g~0~ had a relatively narrow width of the confidence interval (< 0.03 mol m-2 s-1), then the average was `r with(subset(lin2015coef, g0_uci - g0_lci < 0.03), mean(g0))` compared to `r mean(lin2015coef$g0)` for all.
-
-Relatively few reports on the plasticity of g~0~, perhaps because it is so poorly estimated. Exceptions:
-
-[@duarte2016] fitted g0 with BB and BBOpti.
-Lower g0 in heat wave treatment.
-Also lower gdark in heat wave treatment.
-But g1 or m not affected by heat waves (again example where a gmin is a responsive parameter!)
 
@@
 
-# Nighttime conductance (v. short)
+# Nighttime conductance
 
-One estimate of g~0~ (low PPFD) comes from night-time measurements of conductance, which have been known for a long time to be substantially larger than zero.
-It is however not clear whether daytime g~0~ at low PPFD, for darkened leaves, is comparable in magnitude to g~0~ in the dark at night.
-Evidence from experiments where environmental conditions are held constant suggests that g~s~ varies tremendously throughout the day independent of these drivers, and has therefore been attributed to an internal clock (Resco).
+A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~).
+However as our literature comparison showed, g~dark~ is substantially higher compared to g~s~ reached at low photosynthetic rates, and g~min~.
+Previous work on g~dark~ have confirmed that night-time conductance is generally much higher than g~min~, when measured on the same species [@howard2007; @caird2007].
+The idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.
 
-- gdark is variable during the night
-- gdark is an active process not just a fixed parameter
-- gs very early morning is quite high because in many cases gs increases before the sun comes up
-- reviewed extensively elsewhere
+\
+First, ample observations suggect that g~dark~ is not a fixed rate but varies tremendously during the night.
+@caird2007 (and references therein) described how for many species, g~dark~ is not stable throughout the night period.
+Instead, endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments.
+@rescodedios2016 showed that, in Eucalyptus camaldulensis, g~dark~ in the period just after sunset was much lower than pre-dawn g~dark~.
+Strong evidence for endogenous regulation of g~dark~ was reported by @rescodedios2013, who showed that g~dark~ fluctuated throughout the night despite environmental conditions being held constant in whole-tree chambers.
 
-@messinger2006 Conclude that Ci response of gs is the same in the dark or when DCMU was applied (which disrupts electron transport).
-However, Fig5A-B indicate that gs was much higher in the dark than when DCMU was applied (although the ‘response’ was similar).
-This indicates gs response to low PAR (darkness) is more than just responding to what photosynthesis does.
+#  Cuticular conductance
 
-Nighttime conductance not necessarily same as minimum conductance
-We are not reviewing nighttime conductance here, that has been done extensively by others [@caird2007]
+The conductance of stomata-free leaf cuticles represents the absolute minimum conductance that leaves can attain.
+It is clear from our literature review, g~cut~ is very much lower than other definitions of minimum conductance - it is on average twice as low as g~min~.
+We admit some caution in this conclusion, because g~cut~ is probably too low in the way it is measured (as a permeance), due to low water content of the cuticle in those conditions [@kerstiens1996].
+We briefly review a few important aspects of cuticular conductance, and to what extent we expect it to vary with species and environments.
+We refer to previous in-depth reviews for a more detailed treatment of other aspects of plant cuticles, such as chemical composition and morphology, and modes of water transport across the cuticle [@kerstiens1996; @kerstiens2006; @schreiber1996; @schreiber2001].
+We return to the question on what component of g~min~ is due to g~cut~ in the next section.
 
-gnight not constant 
-@barbour2007 considerable stomatal control during the night; mirrors daytime
-@howard2007 nighttime conductance lower in drought. Nighttime 2-5x higher than cuticular.
-
-@misson2004 g0 decreases with pre-dawn water potential
-
-@rescodedios2016 early night gs much (much) lower than pre-dawn gs, in E. camaldulensis
-@rescodedios2013 nighttime fluctuation gs independent of environment
-
-@caird2007 For many species, gnight is not stable throughout the night period.
-Endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [lots of refs]
-
-@howard2007 gdark much higher than gmin (Helianthus annuus). In general this is the case [@caird2007]
-
-#  Cuticular conductance (short)
-
-- true cuticular conductance is that of isolated non-stomatal peels, and is very low compared to other definitions
-- what is it, and is it variable with anything? yes, temperature, humidity.
+Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle.
+Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle.
+The main barrier to diffusion is actually a very thin layer at the leaf surface.
+Because most of the resistance is located in such a thin layer, g~cut~ does not depend on the thickness of the cuticle  [@riederer2001; @anfodillo2002].
+One may ask why there is so much variation in cuticle thickness between plant species, if it does not affect g~cut~.
+The answer may lie in mechanical strength rather than water relations, as @onoda2012 reported that thicker cuticles are more resistant to tearing (he force to tear and tensile strength both approx proportional to cuticle thickness).
 
 Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015]
+This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata.
 
-Details of cuticular composition, modes of transport, and composition has been reviewed 
-[@kerstiens1996; @kerstiens2006; @schreiber1996; @schreiber2001]
-
-Not really a conductance, since water does not diffuse as a gas through the cuticle.
-Instead it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle.
-The main barrier to diffusion is actually a thin layer at the leaf surface, and hence gcut does not depend on the thickness of the cuticle  [@riederer2001; @anfodillo2002].
-@onoda2012 found that thicker cuticles are more resistant to tearing (force to tear and tensile strength both approx proportional to cuticle thickness)
-
-gcut is probably too low in the way it is measured (as a permeance), due to low water content of the cuticle in those conditions [@kerstiens1996]
-
-@riederer2001 Review on cuticular conductance.
-gcut is not correlated with cuticle thickness.
-No reference to stomata.
-Cuticles are even less permeable than polymers of equivalent thickness used in packaging.
-
-- @riederer2001 correlation between temperature and gcut (non-stomatal cuticles)
+- @riederer2001 @schuster2016 correlation between temperature and gcut (non-stomatal cuticles)
 

### 2017-05-23 R/figures.R

@@ -102,6 +102,7 @@ figure_gmin_review_3 <- function(dat){
 figure_g0g1_cor <- function(lin2015, group, legend=FALSE){
   
   x <- subset(lin2015, fitgroup == group)
+  x$Cond <- 1000*x$Cond
   
   # for bootCase to work
   assign("x",x,envir=.GlobalEnv)
@@ -118,25 +119,25 @@ figure_g0g1_cor <- function(lin2015, group, legend=FALSE){
       cex.lab=1.1, cex.axis=0.9)
   with(x, plot(BBopti, Cond, 
                xlim=c(0,0.04),
-               ylim=c(0,0.15),
+               ylim=c(0,150),
                # ylim=c(0, max(Cond)),
                # xlim=c(0,max(BBopti)),
                panel.first=add_regres_line(fit),
                xlab=expression(A/(C[a]*sqrt(D))),
-               ylab=expression(g[s]~(mol~m^-2~s^-1)),
+               ylab=expression(g[s]~(mmol~m^-2~s^-1)),
                pch=16, cex=0.8, col="dimgrey"))
   box()
   plotlabel("(a)","topleft")
   
-  plot(el, type='l', lty=3, 
-       xlab=expression(g[0]~~(mol~m^-2~s^-1)), 
+  plot(el[,1], el[,2]/1000, type='l', lty=3, 
+       xlab=expression(g[0]~~(mmol~m^-2~s^-1)), 
        ylab=expression(g[1]~~(kPa^-0.5)),
-       xlim=c(0,0.025),
+       xlim=c(0,25),
        ylim=c(3,5))
        # xlim=c(min(b[,1]), max(b[,1])),
        # ylim=c(min(b[,2]), max(b[,2])))
-  points(b, pch=16, cex=0.3, col="dimgrey")
-  points(coef(fit)[1],coef(fit)[2], pch=19)
+  points(b[,1], b[,2]/1000, pch=16, cex=0.3, col="dimgrey")
+  points(coef(fit)[1],coef(fit)[2]/1000, pch=19)
   plotlabel("(b)","topleft")
   if(legend)legend("topright", group, bty='n', cex=0.6)
 }
@@ -146,28 +147,42 @@ figure_g0g1_cor <- function(lin2015, group, legend=FALSE){
 
 figure_R2g0 <- function(lin2015coef, miner){
   
[…]

### 2017-05-23 R/functions.R

@@ -16,11 +16,15 @@ fits_lin2015 <- function(lin2015a){
   p <- as.data.frame(coef(fits)) %>% 
     rownames_to_column(var = "fitgroup") %>%  
     rename(g0 = `(Intercept)`, g1 = BBopti) %>% 
-    mutate(R2 = sapply(fits, function(x)summary(x)$adj.r.squared),
+    mutate(g0 = 1000 * g0,
+           R2 = sapply(fits, function(x)summary(x)$adj.r.squared),
            rmse = sapply(fits, function(x)summary(x)$sigma),
            cor = sapply(fits, function(x)summary(x, correlation=T)$correlation[2,1]),
-           g0_lci = sapply(fits, function(x)confint(x)[1,1]),
-           g0_uci = sapply(fits, function(x)confint(x)[1,2])) %>% 
+           g0_lci = 1000 * sapply(fits, function(x)confint(x)[1,1]),
+           g0_uci = 1000 * sapply(fits, function(x)confint(x)[1,2]),
+           g0_se = 1000 * sapply(fits, function(x)summary(x)$coefficients[1,2]),
+           g1_se = sapply(fits, function(x)summary(x)$coefficients[2,2])
+           ) %>% 
     left_join(summarize(lin2015a, 
                         sd_bbopti = sd(BBopti, na.rm=TRUE),
                         cv_bbopti = sd_bbopti/mean(BBopti, na.rm=TRUE),

### 2017-05-23 R/load.R

@@ -180,7 +180,8 @@ condreview_ave$LeafPhenology <- as.factor(condreview_ave$LeafPhenology) %>%
 condreview_ave <- ungroup(condreview_ave)
 
 # Miner et al. 2016
-miner <- read.csv("data/Miner_table1.csv")
+miner <- read.csv("data/Miner_table1.csv") %>%
+  mutate(g0 = 1000*g0)
 
 
 # Lombardozzi et al 2017

### 2017-05-23 R/make_figures.R

@@ -3,7 +3,7 @@ to.pdf(figure_sim(), "output/figure1.pdf", height=6, width=6)
 to.pdf(figure_g0g1_cor(lin2015, "Nicolas Martin-StPaul_Quercus ilex_StPaul_Puechabon"), 
        "output/figure2.pdf", height=4, width=8)
 
-to.pdf(figure_R2g0(lin2015coef, miner), "output/figure3.pdf", height=4, width=8)
+to.pdf(figure_R2g0(lin2015coef, miner), "output/figure3.pdf", height=4, width=9)
 
 to.pdf(figure_gmin_review_2(gdfr), "output/figure4.pdf", height=4, width=4.5)
 to.pdf(figure_gmin_review_3(condreview_ave), "output/figure5.pdf", height=4, width=3.5)
@@ -15,5 +15,5 @@ to.pdf(figure_wtc4_gmin_2(wtc4gmin), "output/figure6.pdf", height=4, width=3.5)
 to.pdf(figure_hakea_gmin(lopw), "output/figure7.pdf", height=5, width=6)
 
 
-convert_all_pdf("output")
+#convert_all_pdf("output")
 

### 2017-06-14 manuscript.Rmd
minor fix

@@
 Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), and g~0~ is the value of g~s~ when A~n~ is zero.
+In this paper, we refer to Eq.
+1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987) (NB they used relative humidity).
+
+\
 The effects of photosynthetic photon flux density (PPFD) and leaf temperature are both assumed to enter via the dependence of A~n~ on these drivers.

### 2017-06-27 manuscript.Rmd
notes added

@@
 # Introduction
+
+\
+Notes
+gmin can increase with ozone and or aerosol particles (no ref yet) - potential for interaction with air pollution. Kerstiens?
+
+Hoad et al; strong positive relationship between g0 and RWC - is our segmented linear method hiding changes in gmin over time?
+
+Gardingen & Grace, gcut depends on VPD (and some useful discussion on difference isolated cuticle vs. intact leaf on non-stomatal side)
+
+rawson and clarke 1988; gs increased hours pre-dawn in wheat. Nocturnal gs > cuticular (which was 13-50% of total)
+
+hygen midgaard 1954- no effect of VPD on gmin (eventual decline in RWC is very linear and does not differ with various VPD treatments).
+"Experiments on the effect of darkening indicated that the stomata may remain partly open even during prolonged dark periods"
 
@@
 
+\
 It is also telling that there are few reports on the plasticity of g~0~, perhaps because it is so poorly estimated.
@@
 
+Lower g~min~ result is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time.
+This is not due to a direct effect of temperature on in cuticular conductance, which increases with temperature (Schreiber 2001).
+Our results showed a large  acclimation effect but no evidence for a direct T effect (over the narrow range in T compared to Schreiber).
+
 ```{r fig.height=4,echo=FALSE, eval=includefigs}

### 2017-08-01 figures.Rmd
Use new database; Rmd with all figures; some figures redone with new grouping

---
title: "gmin figures"
author: "Remko Duursma"
date: "1 August 2017"
output: 
  html_document:
    theme: flatly
---

```{r}
source("R/load.R")
pacman::p_load(magicaxis)
```

```{r}
v <- gmindat$gmin
hist(v, breaks=30, xlim=c(0,40), main="", freq=FALSE,
     ylim=c(0,0.15),
     col="lightgrey",
     xlab=expression(g[min]~~(mmol~m^-2~s^-1)))
curve(dlnorm(x, mean(log(v)), sd=sd(log(v))), add=TRUE,
      n=501)
```

```{r}
v <- log10(gmindat$gmin)
par(yaxs="i", mgp=c(3,1,0), cex.lab=1.2)
hist(v, breaks=30, main="", freq=FALSE,
     #ylim=c(0,0.15),
     xlim=c(-1,2),
     col="lightgrey",
     axes=FALSE,
     xlab=expression(g[min]~~(mmol~m^-2~s^-1)))
magaxis(side=1, unlog=1, usepar=TRUE)
axis(2)
curve(dnorm(x, mean=mean(v), sd=sd(v)), add=TRUE,
      n=101)

```

```{r, fig.cap="Figure 1.", fig.width=6, fig.height=6}
figure_sim()
```

```{r, fig.width=9, fig.height=4}
figure_R2g0(lin2015coef, miner)
```

```{r, fig.width=5, fig.height=5}
figure_gmin_review_2(gdfr)
```

```{r, fig.width=3, fig.height=4}
figure_gmin_bygroup(gmindat)
```

```{r, fig.width=3.5, fig.height=4 }
figure_wtc4_gmin_2(wtc4gmin)
```

```{r, fig.width=6, fig.height=5}
figure_hakea_gmin(lopw)
```

```{r, fig.width=6, fig.height=5}
gmin_by_family(gmindat)
```

```{r, fig.width=6, fig.height=5}
gmin_by_order(gmindat)
```

### 2017-08-01 manuscript.Rmd
Use new database; Rmd with all figures; some figures redone with new grouping

@@
 "Experiments on the effect of darkening indicated that the stomata may remain partly open even during prolonged dark periods"
+
+Mackova (Environmental and Experimental Botany Volume 86, February 2013, Pages 70-75): adding ABA to Lepidium sativum increased the chain length of cuticular waxes (but not total amounts).
+Other work has shown that the length of the chain decreases cuticular transpiration (they cite Hauke and Schreiber, 1998 Ontogenetic and seasonal development of wax composition and cuticular transpiration of ivy (Hedera helix L.) sun and shade leaves)
+
+@rosasanderson2014, (2013, thesis), lower gmin in drought-stressed peanut
 

### 2017-08-01 R/figures.R

@@ -26,77 +26,34 @@ figure_gmin_review <- function(gdfr){
 }
 
 figure_gmin_review_2 <- function(gdfr){
-  par(mar=c(3,5,1,0.5), cex.lab=1.2)
   
-  # get rid of sealed leaves - small sample size and confusing result
-  gdfr <- droplevels(subset(gdfr, method != "gcut_seal"))
+  par(mar=c(4,4,1.5,1), mgp=c(2.5,0.5,0), tcl=-0.2, 
+      las=3, yaxs="i",
+      cex.lab=1.2)
+  plotCI2(gmin, method, gdfr, transform_log10=TRUE,
+          label_las=1,
+          ylab=expression(Conductance~(mmol~m^-2~s^-1)),
+          labels=c(expression(g["cut,isol"]),
+                   expression(g["cut,seal"]),
+                   expression(g["min"]),
+                   expression(g["dark"]),
+                   expression(g["0"]),
+                   expression(g["s,low A"])))
   
-  fit <- lm(log10(gmin) ~ method-1, data=gdfr)
-  cis <- confint(fit)
-  quans <- sapply(split(gdfr, gdfr$method), function(x)quantile(log10(x$gmin), probs=c(0.05, 0.95)))
-  gmins <- with(gdfr, tapply(log10(gmin), method, mean, na.rm=TRUE))
-  ng <- nlevels(gdfr$method)
   
-  plot(1:ng, 10^gmins, pch=19, cex=1.2, xlab="",
-       axes=FALSE,
-       #panel.first=segments(x0=1:5, x1=1:5, y0=10^quans[1,], y1=10^quans[2,], col="grey"),
-       ylim=c(0,50),
-       xlim=c(0.5,ng+0.5),
-       ylab=expression(Conductance~~(mmol~m^-2~s^-1)))
-  axis(1, at=1:ng, labels=c(expression(g["cut,isol"]),
-                           #expression(g["min,seal"]),
-                           expression(g[min]),
-                           expression(g[dark]),
-                           expression(g[0]),
-                           expression(g["s,low A"])))
-  axis(2)
-  box()
-  arrows(x0=1:ng, x1=1:ng, y0=10^cis[,1], y1=10^cis[,2], angle=90, length=0.1, code=3)
 }
 
 
[…]

### 2017-08-01 R/functions.R

@@ -96,3 +96,51 @@ plotlabel <- function(txt, where, inset=0.08, inset.x=inset, inset.y=inset,...){
   text(x,y,txt,font=2,...)
 }
 
+
+
+plotCI2 <- function(yvar, group, data, transform_log10=FALSE, 
+                    ylab=NULL, labels=NULL, label_las=3, ...){
+  
+  if(is.null(ylab))ylab <- substitute(yvar)
+  
+  data <- as.data.frame(data)
+  
+  data$Y <- eval(substitute(yvar), data)
+  data$G <- as.factor(eval(substitute(group), data))
+  
+  if(is.null(labels)){
+    labels <- levels(data$G)
+  }
+  
+  if(transform_log10){
+    fit <- lm(log10(Y) ~ G - 1, data=data)
+    ci <- 10^confint(fit)
+    mn <- 10^coef(fit)
+  } else {
+    fit <- lm(Y ~ G - 1, data=data)
+    ci <- confint(fit)
+    mn <- coef(fit)
+  }
+  
+  g <- glht(fit, linfct=mcp(G = "Tukey"))
+  lets <- cld(g)$mcletters$Letters
+  
+  n <- length(mn)
+  
+  plot(1:n, mn, pch=19, cex=1.1, ylim=c(0, max(ci)+0.1*max(ci)), axes=FALSE,
+       xlab="",
+       xlim=c(0.7, n+0.3), ylab=ylab, ...)
+  axis(1, at=1:n, labels=labels, las=label_las)
+  arrows(x0=1:n, x1=1:n, y0=ci[,1], y1=ci[,2], angle=90, length=0.05, code=3)
+  text(x=1:n, y=ci[,2], lets, pos=3, cex=0.8)
+  par(las=1)
+  axis(2)
+  box()
+  
+  tb <- unname(table(data$G))
+  mtext(side=3, at=1:n, line=0.4, text=tb, las=1, cex=0.7)
+  mtext(side=3, at=0.5, line=0.4, text="n = ", las=1, cex=0.7)
+  
+}
+
+

### 2017-08-01 R/load.R

@@ -2,7 +2,7 @@
 pacman::p_load(Hmisc, car, dplyr, tidyr, nlme, nlshelper, 
                forcats, tibble, magicaxis, 
                plantecophys, readxl, multcomp,
-               reporttools)
+               reporttools, Taxonstand, taxize)
 
 
 
@@ -10,26 +10,6 @@ if(!dir.exists("download"))dir.create("download")
 if(!dir.exists("output"))dir.create("output")
 source("R/functions.R")
 source("R/figures.R")
-
-
-# You must find TRY categorical traits yourself and place it in /data
-# I cannot share it here (registration required)
-if(!file.exists("data/trydb.rds")){
-  tryfile <- "data/TRY_Categorical_Traits_Lookup_Table_2012_03_17_TestRelease.xlsx"
-  if(!file.exists(tryfile))stop("Place the TRY look up table in data/.")
-  trydb <- read_excel(tryfile) %>% 
-    rename(species = AccSpeciesName)
-  trydb <- trydb[,c("species","PhylogeneticGroup","PlantGrowthForm","LeafType","LeafPhenology")]
-  saveRDS(trydb, "data/trydb.rds")
-} else {
-  trydb <- readRDS("data/trydb.rds")
-}
-
-# Additional growth form / phylogenetic group from googling missing species
-gf_add <- read.csv("data/growthform_additional.csv", stringsAsFactors = FALSE)
-trydb <- bind_rows(trydb, gf_add)
-
-
 
 # Lin et al. 2015
 linfn <- "download/lin2015.csv"
@@ -54,22 +34,13 @@ lin2015a <- group_by(lin2015, fitgroup) %>%
 lin2015coef <- fits_lin2015(lin2015a)
 
 
-
-
-# Riederer and Schreiber 2001
-rieder <- read.csv("data/riederer2001_table1.csv", stringsAsFactors = FALSE) %>%
-  filter(gmin < 10000) %>%
-  mutate(gmin = 2 * 10^5 * 10^-3 * gmin / 41,
-         source = "Riederer and Schreiber 2001",
-         method = "gcut_isol")
-
 # Kerstiens 1996
 kerst <- read.csv("data/kerstiens1996_table1.csv", stringsAsFactors = FALSE) %>%
   filter(gmin > 0,
          species != "Molinia caerulea",
-         method %in% c("A","C","D")) %>%
[…]

### 2017-08-01 R/make_figures.R

@@ -5,11 +5,11 @@ to.pdf(figure_g0g1_cor(lin2015, "Nicolas Martin-StPaul_Quercus ilex_StPaul_Puech
 
 to.pdf(figure_R2g0(lin2015coef, miner), "output/figure3.pdf", height=4, width=9)
 
-to.pdf(figure_gmin_review_2(gdfr), "output/figure4.pdf", height=4, width=4.5)
-to.pdf(figure_gmin_review_3(condreview_ave), "output/figure5.pdf", height=4, width=3.5)
-#to.pdf(figure_amings_lin2015), "output/figure5.pdf", height=5, width=5)
+to.pdf(figure_gmin_review_2(gdfr), "output/figure4.pdf", height=5, width=5)
 
-#to.pdf(figure_wtc4_gmin(wtc4gmin, wtc4gdark), "output/figure5.pdf", height=4, width=8)
+to.pdf(figure_gmin_bygroup(gmindat), "output/figure5.pdf", height=4, width=3)
+
+
 to.pdf(figure_wtc4_gmin_2(wtc4gmin), "output/figure6.pdf", height=4, width=3.5)
 
 to.pdf(figure_hakea_gmin(lopw), "output/figure7.pdf", height=5, width=6)

### 2017-08-01 kerst_bygrowthform.R
remove old file

@@ -1,23 +1,0 @@
-
-kerst2a <- group_by(kerst2, species, method) %>%
-  summarize(gmin=mean(gmin),
-            PhylogeneticGroup = last(PhylogeneticGroup),
-            PlantGrowthForm = last(PlantGrowthForm),
-            LeafType = last(LeafType))
-
-summ <- filter(kerst2a, 
-                    method == "gmin",
-                    !is.na(PlantGrowthForm)) %>%
-  group_by(PlantGrowthForm) %>%
-  summarize(gmin_mean = mean(gmin),
-            n = n(),
-            se = sd(gmin)/sqrt(n))
-
-
-library(multcomp)
-
-fit <- lm(gmin ~ PlantGrowthForm, data=kerst2a, subset=method=="gmin")
-g <- glht(fit, linfct=mcp(PlantGrowthForm = "Tukey"))
-cld(g)$mcletters$Letters
-
-

### 2017-08-01 manuscript.Rmd
some writing; references added

--- 2017-06-27 manuscript.Rmd
@@
 \
-Notes
-gmin can increase with ozone and or aerosol particles (no ref yet) - potential for interaction with air pollution. Kerstiens?
-
-Hoad et al; strong positive relationship between g0 and RWC - is our segmented linear method hiding changes in gmin over time?
-
-Gardingen & Grace, gcut depends on VPD (and some useful discussion on difference isolated cuticle vs. intact leaf on non-stomatal side)
 
@@
 
-hygen midgaard 1954- no effect of VPD on gmin (eventual decline in RWC is very linear and does not differ with various VPD treatments).
+@hygen1954 - no effect of VPD on gmin (eventual decline in RWC is very linear and does not differ with various VPD treatments).
 "Experiments on the effect of darkening indicated that the stomata may remain partly open even during prolonged dark periods"
+
+@mackova2013 adding ABA to Lepidium sativum increased the chain length of cuticular waxes (but not total amounts).
+
+@rosasanderson2014, (2013, thesis), lower gmin in drought-stressed peanut
 
@@
 In doing so, stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus can be expected to close stomata when photosynthetic rates are low.
-To avoid dessiccation and ultimate death, stomata can also be expected to close during periods of severe water stress.
+To avoid desiccation and ultimate death, stomata can also be expected to close during periods of severe water stress.
 Plants cannot, however, completely halt water loss even when stomata are closed.
@@
 Another measurement of the minimum conductance arises from leaf drying curve: leaves are picked, and gradually dried out and weighed to determine the water loss rate [@hygen1951].
-This estimate of minimum conductance (which we refer to as g~min~) probably most closely reflects the dessiccation dynamics of severely droughted plants [@blackman2016], but has rarely been mentioned in the context of the Ball-Berry model.
+This estimate of minimum conductance (which we refer to as g~min~) probably most closely reflects the desiccation dynamics of severely droughted plants [@blackman2016], but has rarely been mentioned in the context of the Ball-Berry model.
 
 \
-In this review, we aim to compare values of the various estimates of minimum conductance, discuss components and mechanisms of minimum conductance.
-We place additional emphasis on sources of variation in g~min~, and the role that g~min~ can play in models of water use.
+Here we review the various options to estimate g~0~ for use in models of plant water use and carbon uptake.
+We place an emphasis on sources of variation in g~min~, and the role that g~min~ can play in models of water use.
 Before we discuss the various definitions and measurements of minimum conductance, we first review the modelling literature on its use of g~0~ in Eq.
@@
 \
-A popular approach in thinking about how stomata 'should' respond to environmental drivers is the idea that stomatal conductance is varied to optimize the rate of photosynthesis per unit water used [@cowan1977].
+A popular approach in thinking about how stomata 'should' respond to environmental drivers is the idea that stomatal conductance is varied to maximize total photosynthesis for a given amount of water use [@cowan1977].
 The consequence is that stomata tend to open during periods that are favorable for photosynthesis (high light, optimal temperature), and close when photosynthesis drops to zero (darkness, very high temperature).
@@
 spend water) when there is no photosynthetic gain.
-Early work on optimal stomatal conductance models completely ignored the possibility of g~s~ > 0 when A~n~ = 0 [@cowan1977; @cowan1978; @hari1986; @hari1999], whereas more recent work derives the optimal g~s~ and then adds on a g~0~ [@medlyn2011].
+Indeed, early work on optimal stomatal conductance models ignored the possibility of g~s~ > 0 when A~n~ = 0 [@cowan1977; @cowan1978; @hari1986; @hari1999].
+More recent work derives the optimal g~s~, and simply adds a g~0~ to the solution [@medlyn2011].
 
 \
-In the following we discuss three reasons to include g~0~ in models: the response of C~i~ at low PPFD, the response of transpiration to high leaf temperature (T~leaf~), and the dessiccation of leaves in severe drought.
-
-\
+In the following we discuss three reasons to include g~0~ in models: the response of C~i~ at low PPFD, the response of transpiration to high leaf temperature (T~leaf~), and the desiccation of leaves in severe drought.
 As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) is constant with the photon flux density (PPFD).
@@
 When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata.
-Thus models that aim to predict when plants dessiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000].
-We return to measurements of g~min~ from leaf drying curves in Section XXX.
+Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000].
 
-# Comparison of values for various definitions
+# Comparison of various definitions of minimum conductance
 
@@
 | g~cut~     | Conductance of an isolated non-stomatal cuticle         |
-| g~min~     | Minimum conductance measured from a leaf dry-down curve |
+| g~min~     | Minimum conductance measured from weight loss of detached leaves |
 | g~dark~    | Night-time conductance                                  |
@@
 \
-We identify four distinct definitions of minimum conductance, each of which is measured with its own set of methods.
-Before discussing each of the minimum conductances in detail in the following sections, we here compile literature values to compare the relative magnitude of each of the conductances.
-
-\
-Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) and of minimum conductance from leaf dry-down curves (g~min~) were taken from the compilation by @kerstiens1996, with additional g~min~ measurements from @schuster2016, @burghardt2003, @cavenderbares2007, and @phillips2010 (for the latter study, raw data were obtained and g~min~ re-calculated).
-We excluded measurements of g~min~ where stomatal leaf sides were sealed, because few data were available and the comparison is not consistent with individual studies where the role of stomata in g~min~ is studied (see SectionXXX).
+Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) were taken from the compilation by @kerstiens1996.
+Data on minimum conductance (g~min~) from weight loss of detached leaves was compiled from XXX papers, see Appendix.
 A compilation of night-time (and 'dark') conductance (g~dark~) values was taken from @lombardozzi2017.
-We also estimated g~0~ from Eq.
-1 using the stomatal conductance model of @medlyn2011, using non-linear regression.
+We also estimated g~0~ for each study included in a large database of leaf gas exchange data [@lin2015], by fitting Eq.
+1 using the stomatal conductance model of @medlyn2011.
 These estimates were combined with the compilation of g~0~ values from @miner2017, who reported on xxx values from a literature compilation (each of those values was estimated with linear regression of the Ball-Berry model).
-Finally, we also estimated minimum values of g~s~ attained from a large database of leaf gas exchange data [@lin2015], by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset.
-This estimate was added as it represents actual minimum values measured, rather than a fitted/extrapolated g~0~ term.
+Finally, we also estimated minimum values of g~s~ attained from the Lin et al.
+database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset.
+This estimate was added as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.
 
@@
 @kerstiens1996).
-For the combined dataset, we looked up plant growth form and phylogenetic group (angiosperm or gymnosperm) from the TRY categorical traits database [@kattge2011].
+For the combined dataset, we looked up plant growth form and phylogenetic group (angiosperm or gymnosperm) from the TRY categorical traits database [@kattge2011], and taxonomic family and order with the taxize R package [@chamberlain2013].
 
@@
 We briefly review a few important aspects of cuticular conductance, and to what extent we expect it to vary with species and environments.
-We refer to previous in-depth reviews for a more detailed treatment of other aspects of plant cuticles, such as chemical composition and morphology, and modes of water transport across the cuticle [@kerstiens1996; @kerstiens2006; @schreiber1996; @schreiber2001].
-We return to the question on what component of g~min~ is due to g~cut~ in the next section.
+We refer to previous in-depth reviews for a more detailed treatment of other aspects of plant cuticles, such as chemical composition and morphology, and modes of water transport across the cuticle [@kerstiens1996; @kerstiens2006; @schreiber1996; @schreiber2001; @schuster2016b].
 
@@
 One may ask why there is so much variation in cuticle thickness between plant species, if it does not affect g~cut~.
-The answer may lie in mechanical strength rather than water relations, as @onoda2012 reported that thicker cuticles are more resistant to tearing (he force to tear and tensile strength both approx proportional to cuticle thickness).
+The answer may lie in mechanical strength rather than water relations, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness).
+
+Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged.
+Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but this variation could not explain variation in cuticular conductance (gcutisol).
+Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.
+
+@riederer2001 @schuster2016b correlation between temperature and gcut (non-stomatal cuticles)
 
@@
 
-- @riederer2001 @schuster2016 correlation between temperature and gcut (non-stomatal cuticles)
-
 # Minimum conductance (longer)
@@
 [@hygen1951] early work showing MLD method
-
-Sack et al. 2010 protocol for measurement
-
-- similar to minimum gs observed for the Lin dataset
-- plastic trait
 
@@
 But Schuster gmin not different from gcut at most temperatures, which is a bit strange too.
-page 90 few gmin values from drying curves that we can add.
 
@@
 Methods for gmin from drying curves here.
+
+Lower g~min~ result is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time.
+This is not due to a direct effect of temperature on in cuticular conductance, which increases with temperature (Schreiber 2001).
+Our results showed a large  acclimation effect but no evidence for a direct T effect (over the narrow range in T compared to Schreiber).
 
@@
 There is no effect of measurement T on g~min~, but leaves of trees grown in elevated temperature showed consistently lower g~min~.
-
-Lower g~min~ result is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time.
-This is not due to a direct effect of temperature on in cuticular conductance, which increases with temperature (Schreiber 2001).
-Our results showed a large  acclimation effect but no evidence for a direct T effect (over the narrow range in T compared to Schreiber).
 

### 2017-08-04 R/load.R
Some references added; bug fix in load

@@ -2,7 +2,7 @@
 pacman::p_load(Hmisc, car, dplyr, tidyr, nlme, nlshelper, 
                forcats, tibble, magicaxis, 
                plantecophys, readxl, multcomp,
-               reporttools, Taxonstand, taxize)
+               reporttools, Taxonstand, taxize, speciesmap)
 
 
 
@@ -132,8 +132,7 @@ gmindat <- read.csv("c:/repos/gmindatabase/combined/gmindatabase.csv",
 #
 climfile <- "data/species_climate_wcpet.rds"
 if(!file.exists(climfile)){
-  library(speciesmap)
-  set_zomerpet_path("c:/data/zomerpet")
+  options(zomerpetpath="c:/data/zomerpet", worldclimpath="c:/data/worldclim")
   ALA4R::ala_config(cache_directory="c:/data/ALAcache")
   
   sp <- unique(gmindat$species)
@@ -141,15 +140,17 @@ if(!file.exists(climfile)){
   sp <- sp[nf == 2]
   sp <- sort(sp)
   
-  wc <- worldclim_presence(sp, PET=TRUE, database="both", topath="c:/tmp")
+  wc <- climate_presence(sp, vars=c("tavg","prec","bio","pet"), database="both")
   
   saveRDS(wc, climfile)
 } else {
   wc <- readRDS(climfile)
 }
 
-wc$species <- as.character(wc$species)
-gmindat2 <- left_join(gmindat, wc, by="species")
+wc2 <- annualize_clim(wc) %>% aggregate_clim %>%
+  mutate(species = as.character(species))
+
+gmindat2 <- left_join(gmindat, wc2, by="species")
 
 # New grouping variable.
 gmindat$group2 <- NA

### 2017-08-13 manuscript.Rmd
Some references added; bug fix in load
more writing; one ref added

@@
 
+@bi2017 lower residual transpiration rate in drought-acclimated wheat, in 3 out of 4 genotypes.
+Drought caused a change in cuticular wax production and chemical composition, but in a highly genotype-specific way.
+
 @rosasanderson2014, (2013, thesis), lower gmin in drought-stressed peanut
@@
 \
-Here we review the various options to estimate g~0~ for use in models of plant water use and carbon uptake.
-We place an emphasis on sources of variation in g~min~, and the role that g~min~ can play in models of water use.
-Before we discuss the various definitions and measurements of minimum conductance, we first review the modelling literature on its use of g~0~ in Eq.
-1 (and related approaches), and demonstrate why g~0~ needs to be included in models.
+Here we review a) the role and current use of g~0~ in models, b) various definitions of minimum conductance and comparison of literature values, c) sources of variation in minimum conductance.
 
@@
 
-# Comparison of various definitions of minimum conductance
-
-**Table 1.** Definitions of minimum conductance.
-
-| Variable   |  Definition |
-| ---------- | ----------- |
-| g~cut~     | Conductance of an isolated non-stomatal cuticle         |
-| g~min~     | Minimum conductance measured from weight loss of detached leaves |
-| g~dark~    | Night-time conductance                                  |
-| g~0~       | Intercept in the Ball-Berry-type stomatal conductance model, fitted. |
-
-\
-Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) were taken from the compilation by @kerstiens1996.
-Data on minimum conductance (g~min~) from weight loss of detached leaves was compiled from XXX papers, see Appendix.
-A compilation of night-time (and 'dark') conductance (g~dark~) values was taken from @lombardozzi2017.
-We also estimated g~0~ for each study included in a large database of leaf gas exchange data [@lin2015], by fitting Eq.
-1 using the stomatal conductance model of @medlyn2011.
-These estimates were combined with the compilation of g~0~ values from @miner2017, who reported on xxx values from a literature compilation (each of those values was estimated with linear regression of the Ball-Berry model).
-Finally, we also estimated minimum values of g~s~ attained from the Lin et al.
-database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset.
-This estimate was added as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.
-
-\
-It is customary for g~cut~ and g~min~ to be expressed per unit total leaf surface area, however, since we aim a comparison with g~0~ and g~dark~, all values were converted to per unit projected surface area.
-Units were converted to molar fluxes assuming 41 mol m^-3^ (cf.
-@kerstiens1996).
-For the combined dataset, we looked up plant growth form and phylogenetic group (angiosperm or gymnosperm) from the TRY categorical traits database [@kattge2011], and taxonomic family and order with the taxize R package [@chamberlain2013].
-
-```{r fig.width=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure4.png")
-```
-
-**Figure 4**. Comparison of various estimates of the minimum conductance. Error bars are standard errors. See Table 1 for definitions of minimum conductance, and text for the data sources.
-
-g~s,low A~ much lower than fitted g~0~ or g~dark~ (see also @dekauwe2015, their Fig. A2).
-
-g~dark~ much higher than g~min~, also confirmed by experiments. In soybean, @waldencoleman2013 showed much higher g~dark~  compared to g~min~.
-
-@korner1994 table of minimum conductance. Also points out (without reference) that g0 in droughted leaves is certainly lower than g0 in dark or very high CO2 leaves.
-
 # Estimation of g~0~ by fitting
@@
 
+# Comparison of various definitions of minimum conductance
+
+**Table 1.** Definitions of minimum conductance.
+
+| Variable   |  Definition |
+| ---------- | ----------- |
+| g~cut~     | Conductance of an isolated non-stomatal cuticle         |
+| g~min~     | Minimum conductance measured from weight loss of detached leaves |
+| g~dark~    | Night-time conductance                                  |
+| g~0~       | Intercept in the Ball-Berry-type stomatal conductance model, fitted. |
+
+Because estimation of g~0~ is problematic, this begs the question whether suitable estimates of minimum conductance can be arrived at from independent measurements.
+Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories.
+The absolute minimum conductance attained by leaves is that through the cuticle only.
+Measurements of cuticular conductance (g~cut~) are typically made on isolated cuticles [@riederer2001], or on the side of the leaf that lacks stomata ('adaxial' side).
+Previous work has thoroughly summarized mechanisms and sources of variation in g~cut~, which will not be repeated in this review.
+We briefly summarize some important points in SectionXXX.
+
+\
+A more common measurement is to detach leaves, and measure leaf weight over time as the leaf dries out.
+After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration rate'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit.
+Because this method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress, it is perhaps the most promising as an independent measurement of minimum conductance.
+Because night-time conductance (g~dark~) is frequently used as an independent measurement of g~0~, we compare literature values of g~dark~ with g~min~ and g~cut~, as well as the compilation of fitted values of g~0~ (already shown in Figs X and Y).
+A final minimum conductance was estimated from the @lin2015 database by extracting actual minimum daytime alues measured, when photosynthetic rates were low.
+
+\
+Measurements of conductance of isolated, non-stomatal cuticles (g~cut,isol~), and of conductance of detached leaves with the stomatal side sealed (g~cut,seal~) were taken from the compilation by @kerstiens1996.
+Technically speaking this is not a conductance, but a 'permeance' because water does not *diffuse* across cuticle [@kerstiens1996].
+Data on minimum conductance (g~min~) from weight loss of detached leaves was compiled from XXX papers, see Appendix for a list of papers.
+For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources.
+A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca.
+150 species during night-time.
+Actual minimum daytime values were estimated from the Lin et al.
+database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset.
+This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.
+
+\
+All conductance values were converted to per unit projected surface area.
+Units were converted to molar fluxes assuming 41 mol m^-3^ (cf.
+@kerstiens1996).
+For the combined dataset, we looked up plant growth form and phylogenetic group (angiosperm or gymnosperm) from the TRY categorical traits database [@kattge2011] (manually filling in numerous gaps), and taxonomic family and order with the taxize R package [@chamberlain2013].
+
+```{r fig.width=4,echo=FALSE, eval=includefigs}
+plotfig("output/figure4.png")
+```
+
+**Figure 4**. Comparison of various estimates of the minimum conductance. Error bars are standard errors. See Table 1 for definitions of minimum conductance, and text for the data sources.
+
+The broad comparison between the various estimates of conductance (FigXX) highlights a number of key features, discussed below.
+
+### Low conductance of isolated cuticles
+
+Literature values of g~cut~ were much lower than any other estimate of conductance, even compared to measurements based on MLD with stomatal side sealed (g~mld,seal~).
+The conductance of stomata-free leaf cuticles represents the absolute minimum conductance that leaves can attain.
+This result should be viewed with caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions [@kerstiens1996].
+A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, and refer to previous work [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b].
+However, a few key points should be reiterated.
+
+Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle.
+Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle.
+The main barrier to diffusion is actually a very thin layer at the leaf surface.
+Because most of the resistance is located in such a thin layer, g~cut~ does not depend on the thickness of the cuticle  [@riederer2001; @anfodillo2002; @schuster2016b].
+One may ask why there is so much variation in cuticle thickness between plant species, if it does not affect g~cut~.
+Perhaps the answer lies in mechanical strength rather than water relations, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness).
+
+Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged.
+Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but this variation could not explain variation in g~cut~.
+Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.
+
+@riederer2001 @schuster2016b correlation between temperature and gcut (non-stomatal cuticles)
+
+Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015].
+This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata.
+
+2.
+g~s,low A~ much lower than fitted g~0~ or g~dark~.
+In other words, actual daytime conductance is much lower than the presumed minimum estimated from fitted g~0~.
+A similar result was already presented by @dekauwe2015 (their  Fig.
+A2).
+3.
+g~dark~ on average eight-fold higher than g~min~.
+Not many studies measure both g~dark~ and g~min.
+In soybean, @waldencoleman2013 showed much higher g~dark~  compared to g~min~.
+@koerner1994 also states that g~s~ in droughted leaves is certainly lower than g~s~ in the dark.
+
 # Nighttime conductance
@@
 Strong evidence for endogenous regulation of g~dark~ was reported by @rescodedios2013, who showed that g~dark~ fluctuated throughout the night despite environmental conditions being held constant in whole-tree chambers.
-
-#  Cuticular conductance
-
-The conductance of stomata-free leaf cuticles represents the absolute minimum conductance that leaves can attain.
-It is clear from our literature review, g~cut~ is very much lower than other definitions of minimum conductance - it is on average twice as low as g~min~.
-We admit some caution in this conclusion, because g~cut~ is probably too low in the way it is measured (as a permeance), due to low water content of the cuticle in those conditions [@kerstiens1996].
-We briefly review a few important aspects of cuticular conductance, and to what extent we expect it to vary with species and environments.
-We refer to previous in-depth reviews for a more detailed treatment of other aspects of plant cuticles, such as chemical composition and morphology, and modes of water transport across the cuticle [@kerstiens1996; @kerstiens2006; @schreiber1996; @schreiber2001; @schuster2016b].
-
-Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle.
-Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle.
-The main barrier to diffusion is actually a very thin layer at the leaf surface.
-Because most of the resistance is located in such a thin layer, g~cut~ does not depend on the thickness of the cuticle  [@riederer2001; @anfodillo2002].
-One may ask why there is so much variation in cuticle thickness between plant species, if it does not affect g~cut~.
-The answer may lie in mechanical strength rather than water relations, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness).
-
-Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged.
-Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but this variation could not explain variation in cuticular conductance (gcutisol).
-Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.
-
-@riederer2001 @schuster2016b correlation between temperature and gcut (non-stomatal cuticles)
-
-Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015]
-This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata.
 

### 2017-08-16 figures.Rmd
Some references added; bug fix in load
Added crop genotypes figure

@@
 
-```{r, fig.width=6, fig.height=5}
+```{r, fig.width=6, fig.height=5, warning=FALSE}
 gmin_by_family(gmindat)
@@
 
-```{r, fig.width=6, fig.height=5}
+```{r, fig.width=6, fig.height=5, warning=FALSE}
 gmin_by_order(gmindat)
 ```
+
+```{r, fig.width=6, fig.height=5, warning=FALSE}
+figure_crop_genotypes(cropgmin)
+```

### 2017-08-16 R/figures.R

@@ -397,4 +397,29 @@ gmin_by_family <- function(gmindat){
 }
 
 
+figure_crop_genotypes <- function(cropgmin){
+  
+  # average by genotypes
+  cropgmin2 <- summaryBy(. ~ genotype +crop, data=cropgmin, FUN=mean, na.rm=TRUE, keep.names=TRUE)
+  
+  d2 <- summaryBy(gmin ~ crop, FUN=c(mean, min, max, length), data=cropgmin2)
+  d2 <- d2[order(d2$gmin.mean),]
+  n <- nrow(d2)
+  
+  #windows(4,5)
+  par(mar=c(6,5,1,1), las=2, cex.lab=1.1, mgp=c(2.5, 0.5,0), tcl=-0.1)
+  plot(1, type='n', xlim=c(0.4, n + 0.6), ylim=c(0,35), axes=FALSE,
+       xlab="",
+       ylab=expression(g[min]~~(mmol~m^-2~s^-1)))
+  axis(1, at=1:n, labels=d2$crop)
+  axis(2)
+  box()
+  for(i in 1:n){
+    segments(x0=i,x1=i, y0=d2$gmin.min[i], y1=d2$gmin.max[i])
+  }
+  points(1:n, d2$gmin.mean, pch=19)
+  text(x=1:n, y=34, labels=d2$gmin.length, cex=0.8)
+  text(0.5, 34, "n =", cex=0.8)
+  
+}
 

### 2017-08-16 R/functions.R

@@ -144,3 +144,24 @@ plotCI2 <- function(yvar, group, data, transform_log10=FALSE,
 }
 
 
+as_dataframe_cls <- function(cls){
+  
+  fun <- function(x){
+    if(all(is.na(x))){
+      return(NULL)
+    } else {
+      vec <- c(species = x$name[x$rank == 'species'],
+               Family = x$name[x$rank == 'family'],
+               Order = x$name[x$rank == 'order'])
+      if(!"species" %in% names(vec)){
+        return(NULL)
+      } else {
+        return(vec)
+      }
+    }
+  }
+  
+  m <- do.call(rbind, lapply(cls, fun))
+return(as.data.frame(m, stringsAsFactors = FALSE))
+}
+

### 2017-08-16 R/load.R

@@ -2,7 +2,8 @@
 pacman::p_load(Hmisc, car, dplyr, tidyr, nlme, nlshelper, 
                forcats, tibble, magicaxis, 
                plantecophys, readxl, multcomp,
-               reporttools, Taxonstand, taxize, speciesmap)
+               reporttools, Taxonstand, taxize, speciesmap,
+               doBy, stringi)
 
 
 
@@ -125,32 +126,36 @@ gmindat <- read.csv("c:/repos/gmindatabase/combined/gmindatabase.csv",
   filter(gmin > 0) %>%
   group_by(species) %>%
   summarize(gmin = mean(gmin),
-            source = first(source)) %>%
+            datasource = first(datasource)) %>%
   ungroup %>%
   left_join(species_classes, by="species")
 
+cropgmin <- read.csv("c:/repos/gmindatabase/combined/cropgmindatabase.csv",
+                    stringsAsFactors = FALSE)
 #
-climfile <- "data/species_climate_wcpet.rds"
-if(!file.exists(climfile)){
-  options(zomerpetpath="c:/data/zomerpet", worldclimpath="c:/data/worldclim")
-  ALA4R::ala_config(cache_directory="c:/data/ALAcache")
+if(FALSE){
+  climfile <- "data/species_climate_wcpet.rds"
+  if(!file.exists(climfile)){
+    options(zomerpetpath="c:/data/zomerpet", worldclimpath="c:/data/worldclim")
+    ALA4R::ala_config(cache_directory="c:/data/ALAcache")
+    
+    sp <- unique(gmindat$species)
+    nf <- sapply(strsplit(sp, " "), length)
+    sp <- sp[nf == 2]
+    sp <- sort(sp)
+    
+    wc <- climate_presence(sp, vars=c("tavg","prec","bio","pet"), database="both")
+    
+    saveRDS(wc, climfile)
+  } else {
+    wc <- readRDS(climfile)
+  }
   
-  sp <- unique(gmindat$species)
-  nf <- sapply(strsplit(sp, " "), length)
-  sp <- sp[nf == 2]
-  sp <- sort(sp)
+  wc2 <- annualize_clim(wc) %>% aggregate_clim %>%
+    mutate(species = as.character(species))
   
-  wc <- climate_presence(sp, vars=c("tavg","prec","bio","pet"), database="both")
-  
-  saveRDS(wc, climfile)
-} else {
[…]

### 2017-08-17 manuscript.Rmd
Added all references from gmindatabase, cleaning up refs; clark
Lots of writing; figures updated

@@
 \
-Remko A. Duursma^1^, Chris Blackman^1^, Rosana Lopéz^1^, Belinda Medlyn^1^
+Remko A. Duursma^1^, Chris Blackman^1^, Rosana Lopéz^1^, Nicolas Martin-StPaul^2^, Belinda Medlyn^1^
 
@@
 ^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia
+^2^ INRA
 
@@
 # Introduction
-
-\
-
-rawson and clarke 1988; gs increased hours pre-dawn in wheat. Nocturnal gs > cuticular (which was 13-50% of total)
-
-@hygen1954 - no effect of VPD on gmin (eventual decline in RWC is very linear and does not differ with various VPD treatments).
-"Experiments on the effect of darkening indicated that the stomata may remain partly open even during prolonged dark periods"
-
-@mackova2013 adding ABA to Lepidium sativum increased the chain length of cuticular waxes (but not total amounts).
-
-@bi2017 lower residual transpiration rate in drought-acclimated wheat, in 3 out of 4 genotypes.
-Drought caused a change in cuticular wax production and chemical composition, but in a highly genotype-specific way.
-
-@rosasanderson2014, (2013, thesis), lower gmin in drought-stressed peanut
 
@@
 Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero.
-Based on this, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005].
-It is also frequently assumed that g~0~ is the 'cuticular conductance` [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata [@kerstiens1996].
-Another measurement of the minimum conductance arises from leaf drying curve: leaves are picked, and gradually dried out and weighed to determine the water loss rate [@hygen1951].
-This estimate of minimum conductance (which we refer to as g~min~) probably most closely reflects the desiccation dynamics of severely droughted plants [@blackman2016], but has rarely been mentioned in the context of the Ball-Berry model.
+But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, begging the question whether the same g~0~ is reached in each of these cases.
+In the application of the Ball-Berry models, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005].
+It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata [@kerstiens1996].
+Finally, measuring the low water loss rates of detached leaves mimicks drought conditions, yielding an estimate of whole-leaf minimum conductance (g~min~) [@pisekberger1938; @hygen1951].
+Photosynthesis also ceases at very high temperature (ref), but stomata do not appear to close in concert [@drake, @urban], thus we cannot use high temperature to estimate a reasonable g~0~.
 
 \
-Here we review a) the role and current use of g~0~ in models, b) various definitions of minimum conductance and comparison of literature values, c) sources of variation in minimum conductance.
+In this paper, we a) review the role and current use of g~0~ in models, b) perform a quantitative comparison of various measurements of the presumed minimum conductance, including fitting g~0~ with Eq.
+1, night-time conductance (based on a previous compilation), and a new compilation of g~min~ estimates from weight-loss of detached leaves, and c) with use of two case studies and review of the literature, we demonstrate that g~min~ shows considerable acclimation to growing conditions.
+
+# Methods
+
+## Coupled leaf gas exchange model
+
+To illustrate the effect of g~0~ on the modelled response of g~s~ to PPFD and T~leaf~, it is necessary to embed the stomatal conductance model in a coupled leaf gas exchange model, which is formed by three equations; Eq.
+1 giving the dependence of g~s~ on A~n~ and some environmental drivers, the diffusion constraint,
+
+\
+(2)
+$$ A_n = \frac{g_s}{1.6}(C_a - C_i) $$
+
+\
+where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~). And finally the Farquhar-von Caemmerer-Berry photosynthesis model that gives A~n~ as a function of C~i~,
+
+\
+(3)
+$$ A_n = \frac{a \cdot (C_i - \Gamma^{*})}{C_i + b} - R_d$$
+
+\
+where a = V~cmax~ or J/4 (where the electron transport rate J approaches Jmax at high light) depending on the limitation to photosynthesis, and b also varies.
+The dependence on temperature enters through V~cmax~, J~max~, R~d~, $\Gamma^{*}$, and other parameters [@medlyn2002].
+For all simulations with this model, we used the implementation in the *plantecophys* R package [@duursma2015].
+
+## Literature review
+
+For the @lin2015 database, we fit the model of @medlyn2011 using non-linear regression, while @miner2016 presented results from linear regression of the @ballberry1987 model.
+
+\
+Measurements of conductance of isolated, non-stomatal cuticles (g~cut,isol~), and of conductance of detached leaves with the stomatal side sealed (g~cut,seal~) were taken from the compilation by @kerstiens1996.
+
+\
+Data on minimum conductance (g~min~) from weight loss of detached leaves was compiled from 44 original papers, see Appendix A for a list of papers.
+For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources.
+All conductance values were converted to per unit projected surface area.
+Units were converted to molar fluxes assuming 41 mol m^-3^ (cf.
+@kerstiens1996).
+
+\
+A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca.
+150 species during night-time.
+Actual minimum daytime values were estimated from the Lin et al.
+database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset.
+This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.
+
+## Case studies on minimum conductance of detached leaves
+
+Both studies use identical methods.
+The first dataset was collected on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature.
+Leaves were sampled from XXX, n=XXX.
+The second case study was on *Hakea* plants.
+Details here (measurement T?).
+Methods for gmin from drying curves here.
 
@@
 \
-In the following we discuss three reasons to include g~0~ in models: the response of C~i~ at low PPFD, the response of transpiration to high leaf temperature (T~leaf~), and the desiccation of leaves in severe drought.
+In the following we show how inclusion of g~0~ in models affects the response of C~i~ at low PPFD and the response of transpiration to high leaf temperature (T~leaf~).
+Another reason to include a non-zero g~0~ term in models of stomatal conductance arises during severe drought.
+When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata.
+Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000].
+
+\
 As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) is constant with the photon flux density (PPFD).
@@
 \
-To illustrate the effect of g~0~ on the modelled response of g~s~ to PPFD and T~leaf~, it is necessary to embed the stomatal conductance model in a coupled leaf gas exchange model, which is formed by three equations; Eq.
-1 giving the dependence of g~s~ on A~n~ and some environmental drivers, the diffusion constraint,
-
-\
-(2)
-$$ A_n = \frac{g_s}{1.6}(C_a - C_i) $$
-
-\
-where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~). And finally the Farquhar-von Caemmerer-Berry photosynthesis model that gives A~n~ as a function of C~i~,
-
-\
-(3)
-$$ A_n = \frac{a \cdot (C_i - \Gamma^{*})}{C_i + b} - R_d$$
-
-\
-where a = V~cmax~ or J/4 (where the electron transport rate J approaches Jmax at high light) depending on the limitation to photosynthesis, and b also varies.
-The dependence on temperature enters through V~cmax~, J~max~, R~d~, $\Gamma^{*}$, and other parameters [@medlyn2002].
-
-\
-Using the coupled leaf gas exchange as implemented in the *plantecophys* R package [@duursma2015], we show how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^).
+Using the coupled leaf gas exchange model, we show how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^).
 We also demonstrate the effect of g~0~ on modelling the T~leaf~ response.
@@
 1.
+
+# Estimation of g~0~ by fitting
+
+The standard approach in estimating g~0~ for use in Eq.
+1 is to use regression with g~s~ as the response variable, and the right-hand side of Eq.
+1 as the predictor.
+The g~0~ parameter is thus estimated as the intercept.
+Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated in this way.
+The difficulty of the estimation can already be concluded from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g.
+@leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species); presumably because unreliable estimates were obtained, b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though nonsensical.
+
+\
+It is also telling that there are few reports on the plasticity (intra- or interspecific) of g~0~, perhaps because it is so poorly estimated.
+Exceptions are @duarte2016, who found a lower g~0~ in a heat wave treatment in Douglas-fir (although g~1~ was unaffected), but in their case g~0~ was very accurately estimated by careful multi-point light response curves.
+Another exception is @misson2004, who report on a close negative correlation between pre-dawn leaf water potential and g~0~ in Ponderosa pine (again, the slope was unaffected).
+
+\
+In the following we demonstrate that the fitting process is problematic for three reasons: 1) the estimates of g~0~ and g~1~ are highly correlated, 2) the precision for g~0~ is generally much lower than for g~1~, and 3) for data that have a worse fit overall, the g~0~ estimates are elevated.
+We thus conclude that g~0~ should really not be estimated from regression, although this is the most common method applied.
+Recently, @barnard2013 also mention the difficulty of fitting Eq.
+1 to estimate g~0~, but did not present specific details.
+
+\
+In Fig.
+2, we demonstrate how g~0~ and g~1~ are statistically correlated, when fitted for a single study.
+The correlation between slopes and intercepts is not unique to Eq.
+1, but a general property of linear models [@stapleton1995].
+The consequence is that it is more difficult to interpret g~1~ and g~0~ independently, and that confidence intervals on each of the parameters should also not be interpreted independently (see @becker2007 for more discussion on the statistical aspects of comparing slopes and intercepts).
+
+\
+Next we study the values obtained when fitting Eq.
+1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation of fitted g~0~ and m (applying the model of ) from @miner2016.
+In both cases, estimates of g~0~ are inflated when the model fits poorly (Fig.
+3 a,b).
+This can be understood by considering that a poor fit often results in the flattening out of the regression line, thus giving a large value for the intercept.
+A poor fit is often obtained when there is little variation in the right-hand side of Eq.
+1, usually because environmental conditions varied little.
+We confirm this by showing that the standard error (SE) of g~0~ increases when the coefficient of variation the ball-berry coefficient (i.e.
+right-hand side of Eq.
+1) is lower (Fig.
+3c).
+
+# Comparison of various definitions of minimum conductance
+
+**Table 1.** Definitions of minimum conductance.
+
+| Variable   |  Definition |
+| ---------- | ----------- |
+| g~cut~     | Conductance of an isolated non-stomatal cuticle         |
+| g~min~     | Minimum conductance measured from weight loss of detached leaves |
+| g~dark~    | Night-time conductance                                  |
+| g~0~       | Intercept in the Ball-Berry-type stomatal conductance model, fitted. |
+| g~s,low A~ | Cuvette measurements of g~s~ when photosynthesis rates are low. |
+
+Measurements of (presumed) minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories.
+The absolute minimum conductance attained by leaves is that through the cuticle only.
+Measurements of cuticular conductance (g~cut~) are typically made on isolated cuticles [@riederer2001] (g~cut,isol~), or on the side of the leaf that lacks stomata ('adaxial' side), or by sealing the side with stomata (g~cut,seal~).
+Another measurement is to detach leaves, and measure leaf weight over time as the leaf dries out.
+After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration rate'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit.
+This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress.
+Because night-time conductance (g~dark~) is frequently used as an independent measurement of g~0~, we compare literature values of g~dark~ with g~min~ and g~cut~, as well as the compilation of fitted values of g~0~ (already shown in Fig.
+3).
+
+Results shown in Fig. 4.
+
+\
+Conductance in the dark was on average eight-fold higher than g~min~.
+It is no new observation that stomata do not close in the dark, but do close on detached leaves, as already demonstrated by @hygen1954 in Scots pine, and described by @koerner1994.
+More recently, @waldencoleman2013 showed much higher g~dark~ compared to g~min~ in Cotton.
+
+\
+Actual measured stomatal conductance during conditions when photosynthesis was low (g~s,low A~) was also much lower than fitted g~0~ or g~dark~.
+In other words, actual daytime conductance is much lower than the presumed minimum estimated from fitted g~0~.
+A similar result was already presented by @dekauwe2015 (their  Fig.
+A2).
+If we would use the fitted g~0~ in models, simulated g~s~ would never drop below g~0~.
+
+### Low conductance of isolated cuticles
+
+Literature values of g~cut~, especially when measured on isolated cuticles (g~cut,isol~) were much lower than any other estimate of conductance, even compared to measurements based on MLD with stomatal side sealed (g~mld,seal~).
+The conductance of stomata-free leaf cuticles represents the absolute minimum conductance that leaves can attain.
+This result should be viewed with caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions.
+A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these processes have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b].
+However, a few key points should be reiterated as they are relevant to the current discussion.
+
+Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle.
+Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle [@kerstiens1996; @riederer2001; @schreiber1996].
+The main barrier to diffusion is actually a very thin layer at the leaf surface.
+Because most of the resistance is located in such a thin layer, g~cut~ does not correlate with the thickness of the cuticle  [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b].
+Nonetheless there is considerable variation in cuticle thickness between plant species [@schuster2016a], as well as within species in response to drought (cite), along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woordruff2010].
+If the cuticle thickness does not directly affect its conductance, what is the value of increasing the cuticle?
+Perhaps the answer lies in mechanical strength rather than water loss per se, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness).
+
+Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged.
+Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but this variation could not explain variation in g~cut~.
+Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.
+
+@mackova2013 adding ABA to Lepidium sativum increased the chain length of cuticular waxes (but not total amounts).
+
+@riederer2001 @schuster2016b correlation between temperature and gcut (non-stomatal cuticles)
+
+Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015].
+This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata.
+
+# Nighttime conductance
+
+A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~).
+However as our literature comparison showed, g~dark~ is substantially higher compared to g~s~ reached at low photosynthetic rates, and g~min~.
+Previous work on g~dark~ have confirmed that night-time conductance is generally much higher than g~min~, when measured on the same species [@howard2007; @caird2007].
+The idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.
+
+\
+First, ample observations suggect that g~dark~ is not a fixed rate but varies tremendously during the night.
+@caird2007 (and references therein) described how for many species, g~dark~ is not stable throughout the night period.
+Instead, endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [@rawson1988].
+@rescodedios2016 showed that, in *Eucalyptus camaldulensis*, g~dark~ in the period just after sunset was much lower than pre-dawn g~dark~.
+Strong evidence for endogenous regulation of g~dark~ was reported by @rescodedios2013, who showed that g~dark~ fluctuated throughout the night despite environmental conditions being held constant in whole-tree chambers.
+Clearly g~dark~ is an actively controlled process that cannot be adequately summarized by a single constant g~0~.
+
+# Minimum conductance of detached leaves
+
+Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after a while a constant low rate is achieved.
+From this minimum transpiration rate, a minimum conductance (g~min~) is usually estimated, taking into account the vapour pressure deficit [@sinclair1986].
+This method has been termed the 'mass loss of detached leaves' (MLD) [@kerstiens1996].
+Many previous reports use the term 'cuticular conductance' for g~min~, but this is inaccurate as some fraction of the water loss will occur through broken, leaky, or incompletely closed stomata.
+Our literature review was somewhat inconclusive in this regard, as conductance g~min~ measured on leaves where the stomatal side was sealed was similar (across all included studies) to g~min~ of whole leaves (Fig.
+3).
+More conclusive is evidence from individual studies where the stomatal component to g~min~ could be quantified.
+A detailed study of *Hedera helix*  concluded that 35\% of water loss occurred across the stomatal pores (though closed), and 65\% across the non-stomatal part of the cuticle, despite the fact that the stomata presumably covered only a small fraction of the leaf [@santrucek2004].
+For ten Sorghum genotypes, @muchow1989 reported a correlation between stomatal density and g~min~, suggesting leaky stomata contribute substantially to g~min~.
+For seven conifer species, @brodribb2014 reported that 50 - 94\% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky.
+However @santrucek found that the cuticle (solid mass only) of the stomatal side in *Hedera helix* was many times more more permeable than the non-stomatal side, thus providing another potential explanation for this difference.
+Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles.
+
+In the following sections, we review the quite substantial literature on g~min~ measured with the MLD technique, stretching back to the 1930s [@pisekberger], focussing on environmental and ecological determinants.
+We also further tease apart the literature compilation of g~min~ (shown just as a single average value in Fig.
+4).
+
+## Minimum conductance shows little meaningful interspecific variation
+
+Using the literature compilation g~min~ (see SectionX), we asked whether g~min~ shows clear variation due to, species growth form, taxonomic order, and average climate across the species' distribution.
+
+We found not so much! Grasses have a higher gmin. So do crops, and within crops there is a ton of genotypic variation.
+
+Very few studies have found meaningful correlations. Exception is @brodribb2014, showing a correlation between rainfall at species origin (of the driest quarter), and g~min~.
+
+Comparing all available data, we did not find significant relationships between rainfall, aridity, or temperature and g~min~.
+There is, however, a substantial acclimatory response of g~min~ to water availability.
+
+## Minimum conductance acclimates to growing conditions
+
+@bi2017 lower residual transpiration rate in drought-acclimated wheat, in 3 out of 4 genotypes.
+Drought caused a change in cuticular wax production and chemical composition, but in a highly genotype-specific way.
+
+@rosasanderson2014, lower gmin in drought-stressed peanut
+
+@clark1956 lower gmin in drought-stressed soybean
+
+@anfodillo2002 gmin high altitude.
+Four-fold increase gmin from February (winter) to August in Picea abies, probably due to incomplete stomatal closure not changes in cuticle properties.
+Cite common finding that gmin increases with altitude [e.g.
+@delucia1984; @herrick1991].
+
+@jordan2007 large increase gmin in the herb Agastachys odorata with leaf age
+
+@fanourakis2013 gmin (desiccation gmin) much higher in high RH grown plants, at any value of leaf water content (Table 1).
+This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss (Hypothesis 2).
+
+@schuster2016 thermal tolerance and gmin relationship across species.
+
+@cavenderbares2007 gdark similar to gmin in oaks (unusual!). Lower gmin in droughted oaks.
+
+- @schuster2016 correlation temperature, gcut AND gmin for Nerium oleander (page 59).
+T range much wider than Blackman though...
+But Schuster gmin not different from gcut at most temperatures, which is a bit strange too.
+
+We present two case studies demonstrating the acclimation of g~min~ to environmental drivers.
+
+Lower g~min~ result is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time.
+This is not due to a direct effect of temperature on in cuticular conductance, which increases with temperature (Schreiber 2001).
+Our results showed a large  acclimation effect but no evidence for a direct T effect (over the narrow range in T compared to Schreiber).
+
+# Conclusions
+
+In a pioneering publication which presented the Jarvis model of stomatal conductance, @jarvis1976 states that "we have assumed in the following equations that when stomata are closed the leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." Similarly, given the poor statistical properties of g~0~ estimated from regression, the very high values compared to minimum conductance reached for detached leaves, we conclude that g~0~ should not be estimated from regression on leaf gas exchange data for use in models.
+
+Then, how should g~0~ be estimated?
+It is clear from our comparisons that there is no single minimum conductance.
+Leaves maintain much higher g~dark~ (itself an actively controlled process) than the minimum obtained for detached leaves.
+Thus when modelling night-time or low-light conductance, a different g~0~ should be used than when modelling the drought response of plants.
+Likewise, as briefly mentioned in the introduction, yet a different g~0~ is reached when temperature is so high that photosynthesis does not take place.
+It is not our intention to make models more complex, but certainly we cannot avoid the conclusion that g~0~ should be replaced with a more complicated expression.
+
+Our review of g~min~ of detached leaves shows remarkable plasticity to growing conditions.
+The lack of clear species trends, when comparing growing conditions, climate of origin, or growth habit, can perhaps be explained by the very large acclimatory potential of g~min~ to growing conditions.
+Future studies should thus compare g~min~ on many species grown in the same conditions to better understand the adaptive value of the minimum water loss rate of leaves.
+
+# Figure captions
 
@@
 For all simulations it is assumed that T~leaf~ is equal to T~air~.
-Parameters were, etc.
-
-\
-The third reason for inclusion of a g~0~ term in models of stomatal conductance arises during severe drought.
-When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata.
-Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000].
-
-# Estimation of g~0~ by fitting
-
-The standard approach in estimating g~0~ for use in Eq.
-1 is to use regression with g~s~ as the response variable, and the right-hand side of Eq.
-1 as the predictor.
-The g~0~ parameter is thus estimated as the intercept.
-Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated in this way.
-The difficulty of the estimation can already be concluded from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g.
-@leuning1995 uses 0.01 for all); presumably because unreliable estimates were obtained, b) negative g~0~ estimates (i.e.
-negative intercepts) are common [@leuning1995; @heroult2013; @miner2017].
-In a pioneering publication which presented the Jarvis model of stomatal conductance, @jarvis1976 states that "we have assumed in the following equations that when stomata are closed the  leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." [@jarvis1976].
-More recently, @barnard2013 mention the difficulty of fitting Eq.
-1 to estimate g~0~ but give no details.
-
-\
-It is also telling that there are few reports on the plasticity of g~0~, perhaps because it is so poorly estimated.
-Exceptions are @duarte2016, who found a lower g0 in a heat wave treatment (although g~1~ was unaffected), and @misson2004 who report on a close negative correlation between pre-dawn leaf water potential and g~0~.
-
-\
-Next we demonstrate that the fitting process gives three problems: 1) the estimates of g~0~ and g~1~ are highly correlated, 2) the precision for g~0~ is generally much lower than for g~1~, and 3) for data that have a worse fit overall, the g~0~ estimates are elevated.
-
-\
-In Fig.
-2, we demonstrate how g~0~ and g~1~ are statistically correlated via a bootstrap sampling approach.
-The correlation between slopes and intercepts is a general aspect of linear models [@stapleton1995].
-The consequence is that it is more difficult to interpret g~1~ and g~0~ independently, and that confidence intervals on each of the parameters should also not be interpreted independently.
-See @becker2007 for more discussion.
-
-Across datasets, g0 increases when the model fits poorly (Fig.
-3 a,b), and the standard error (SE) of g~0~ increases when less variation in a ball-berry coefficient (Fig.
-3c).
-Also, SE is very large compared to the average magnitude of g~0~ across studies (compare Fig.
-2 and Fig.
-3).
-
-g~0~ fitted is higher than minimum g~s~ from datasets.
-If we restrict the fittings to those were g~0~ had a relatively narrow width of the confidence interval (< 0.03 mol m-2 s-1), then the average was `r with(subset(lin2015coef, g0_uci - g0_lci < 0.03), mean(g0))` compared to `r mean(lin2015coef$g0)` for all.
 
@@
 In the right panel, a very similar association between g~0~ and R^2^ is presented by the review of @miner2017, using the Ball-Berry model.
-
-# Comparison of various definitions of minimum conductance
-
-**Table 1.** Definitions of minimum conductance.
-
-| Variable   |  Definition |
-| ---------- | ----------- |
-| g~cut~     | Conductance of an isolated non-stomatal cuticle         |
-| g~min~     | Minimum conductance measured from weight loss of detached leaves |
-| g~dark~    | Night-time conductance                                  |
-| g~0~       | Intercept in the Ball-Berry-type stomatal conductance model, fitted. |
-
-Because estimation of g~0~ is problematic, this begs the question whether suitable estimates of minimum conductance can be arrived at from independent measurements.
-Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories.
-The absolute minimum conductance attained by leaves is that through the cuticle only.
-Measurements of cuticular conductance (g~cut~) are typically made on isolated cuticles [@riederer2001], or on the side of the leaf that lacks stomata ('adaxial' side).
-Previous work has thoroughly summarized mechanisms and sources of variation in g~cut~, which will not be repeated in this review.
-We briefly summarize some important points in SectionXXX.
-
-\
-A more common measurement is to detach leaves, and measure leaf weight over time as the leaf dries out.
-After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration rate'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit.
-Because this method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress, it is perhaps the most promising as an independent measurement of minimum conductance.
-Because night-time conductance (g~dark~) is frequently used as an independent measurement of g~0~, we compare literature values of g~dark~ with g~min~ and g~cut~, as well as the compilation of fitted values of g~0~ (already shown in Figs X and Y).
-A final minimum conductance was estimated from the @lin2015 database by extracting actual minimum daytime alues measured, when photosynthetic rates were low.
-
-\
-Measurements of conductance of isolated, non-stomatal cuticles (g~cut,isol~), and of conductance of detached leaves with the stomatal side sealed (g~cut,seal~) were taken from the compilation by @kerstiens1996.
-Technically speaking this is not a conductance, but a 'permeance' because water does not *diffuse* across cuticle [@kerstiens1996].
-Data on minimum conductance (g~min~) from weight loss of detached leaves was compiled from XXX papers, see Appendix for a list of papers.
-For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources.
-A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca.
-150 species during night-time.
-Actual minimum daytime values were estimated from the Lin et al.
-database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset.
-This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.
-
-\
-All conductance values were converted to per unit projected surface area.
-Units were converted to molar fluxes assuming 41 mol m^-3^ (cf.
-@kerstiens1996).
-For the combined dataset, we looked up plant growth form and phylogenetic group (angiosperm or gymnosperm) from the TRY categorical traits database [@kattge2011] (manually filling in numerous gaps), and taxonomic family and order with the taxize R package [@chamberlain2013].
+In panel (c) coefficient of variation and R2.
 
@@
 **Figure 4**. Comparison of various estimates of the minimum conductance. Error bars are standard errors. See Table 1 for definitions of minimum conductance, and text for the data sources.
-
-The broad comparison between the various estimates of conductance (FigXX) highlights a number of key features, discussed below.
-
-### Low conductance of isolated cuticles
-
-Literature values of g~cut~ were much lower than any other estimate of conductance, even compared to measurements based on MLD with stomatal side sealed (g~mld,seal~).
-The conductance of stomata-free leaf cuticles represents the absolute minimum conductance that leaves can attain.
-This result should be viewed with caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions [@kerstiens1996].
-A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, and refer to previous work [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b].
-However, a few key points should be reiterated.
-
-Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle.
-Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle.
-The main barrier to diffusion is actually a very thin layer at the leaf surface.
-Because most of the resistance is located in such a thin layer, g~cut~ does not depend on the thickness of the cuticle  [@riederer2001; @anfodillo2002; @schuster2016b].
-One may ask why there is so much variation in cuticle thickness between plant species, if it does not affect g~cut~.
-Perhaps the answer lies in mechanical strength rather than water relations, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness).
-
-Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged.
-Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but this variation could not explain variation in g~cut~.
-Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.
-
-@riederer2001 @schuster2016b correlation between temperature and gcut (non-stomatal cuticles)
-
-Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015].
-This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata.
-
-2.
-g~s,low A~ much lower than fitted g~0~ or g~dark~.
-In other words, actual daytime conductance is much lower than the presumed minimum estimated from fitted g~0~.
-A similar result was already presented by @dekauwe2015 (their  Fig.
-A2).
-3.
-g~dark~ on average eight-fold higher than g~min~.
-Not many studies measure both g~dark~ and g~min.
-In soybean, @waldencoleman2013 showed much higher g~dark~  compared to g~min~.
-@koerner1994 also states that g~s~ in droughted leaves is certainly lower than g~s~ in the dark.
-
-# Nighttime conductance
-
-A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~).
-However as our literature comparison showed, g~dark~ is substantially higher compared to g~s~ reached at low photosynthetic rates, and g~min~.
-Previous work on g~dark~ have confirmed that night-time conductance is generally much higher than g~min~, when measured on the same species [@howard2007; @caird2007].
-The idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.
-
-\
-First, ample observations suggect that g~dark~ is not a fixed rate but varies tremendously during the night.
-@caird2007 (and references therein) described how for many species, g~dark~ is not stable throughout the night period.
-Instead, endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments.
-@rescodedios2016 showed that, in Eucalyptus camaldulensis, g~dark~ in the period just after sunset was much lower than pre-dawn g~dark~.
-Strong evidence for endogenous regulation of g~dark~ was reported by @rescodedios2013, who showed that g~dark~ fluctuated throughout the night despite environmental conditions being held constant in whole-tree chambers.
-
-# Minimum conductance (longer)
-
-[@hygen1951] early work showing MLD method
-
-One reason gmin is higher than gcut is the role of stomata, even when presumably shut or as shut as they will be.
-
-gseal on average not different from gcut - but for individual studies they frequently differ a lot indicating variable role of stomatal contribution [@santrucek2004; @brodribb2014].
-@muchow1989 correlation stomatal density and gmin suggesting leaky stomata.
-[@santrucek2004] Direct measurements showing much higher conductance of the leaf side with stomata, though stomata were closed.
-
-@brodribb2014 59-94% of transpiration during leaf dry down was from stomatal side of the leaf in seven conifer species, suggesting large component of gmin is leaky stomata.
-
-but not always, in @burghardt2003 gmin was the same as cuticular permeance (Gmin is lowest gs in desiccating leaves, which includes cuticle and closed stomata).
-This demonstrates that stomata were able to completely close, and not lose significant amounts of water (with the exception of Hedera helix).
-
-## Minimum conductance acclimates to growing conditions
-
-@anfodillo2002 gmin high altitude.
-Four-fold increase gmin from February (winter) to August in Picea abies, probably due to incomplete stomatal closure not changes in cuticle properties.
-Cite common finding that gmin increases with altitude (top left page 485).
-
-@herrick1991 gmin increased with altitude in Picea rubens.
-
-Hypothesis that conifers have low gmin may have to be revisited in the light of @brodribb2014 (check what they actually say!)
-
-@jordan2007 large increase gmin in the herb Agastachys odorata with leaf age
-
-@fanourakis2013 gmin (desiccation gmin) much higher in high RH grown plants, at any value of leaf water content (Table 1).
-This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss (Hypothesis 2).
-
-@schuster2016 thermal tolerance and gmin relationship across species.
-
-@brodribb2014 gmin across conifer species was correlated with the rainfall of the driest quarter, again suggesting gmin is a plastic trait
-
-[@cavenderbares2007] gdark similar to gmin in oaks (unusual!). Lower gmin in droughted oaks.
-
-- @schuster2016 correlation temperature, gcut AND gmin for Nerium oleander (page 59).
-T range much wider than Blackman though...
-But Schuster gmin not different from gcut at most temperatures, which is a bit strange too.
-
-We present two case studies demonstrating the acclimation of g~min~ to environmental drivers.
-Both studies use identical methods.
-The first dataset was collected on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature.
-Leaves were sampled from XXX, n=XXX.
-The second case study was on *Hakea* plants.
-Details here (measurement T?).
-Methods for gmin from drying curves here.
-
-Lower g~min~ result is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time.
-This is not due to a direct effect of temperature on in cuticular conductance, which increases with temperature (Schreiber 2001).
-Our results showed a large  acclimation effect but no evidence for a direct T effect (over the narrow range in T compared to Schreiber).
 
@@
 
-# Recommendations and future work
-
-g0 at the moment covers everything, but we need separate ones (night, high VPD/T, drought, low PAR during day)
-g0 is poorly estimated from regression - we need to figure out whether it can be measured independently
-one option is to use gmin in models, perhaps as averages by growth form / group, but that approach will have to be tested
-g0 is certainly not the same as gdark despite the fact that on average they have similar magnitudes. We know for sure gs can be much lower than fitted g0 and gdark in field conditions.
-
 # References {.unnumbered}

### 2017-08-17 R/figures.R

@@ -32,6 +32,7 @@ figure_gmin_review_2 <- function(gdfr){
       cex.lab=1.2)
   plotCI2(gmin, method, gdfr, transform_log10=TRUE,
           label_las=1,
+          ylim=c(0,80),
           ylab=expression(Conductance~(mmol~m^-2~s^-1)),
           labels=c(expression(g["cut,isol"]),
                    expression(g["cut,seal"]),
@@ -369,6 +370,22 @@ figure_hakea_gmin <- function(lopw){
 }
 
 
+gmin_loghist <- function(gmindat){
+  v <- log10(gmindat$gmin)
+  par(yaxs="i", mgp=c(3,1,0), cex.lab=1.2)
+  hist(v, breaks=30, main="", freq=FALSE,
+       #ylim=c(0,0.15),
+       xlim=c(-1,2),
+       col="lightgrey",
+       axes=FALSE,
+       xlab=expression(g[min]~~(mmol~m^-2~s^-1)))
+  magaxis(side=1, unlog=1, usepar=TRUE)
+  axis(2)
+  curve(dnorm(x, mean=mean(v), sd=sd(v)), add=TRUE,
+        n=101)
+  
+}
+
 
 gmin_by_order <- function(gmindat){
   
@@ -377,11 +394,25 @@ gmin_by_order <- function(gmindat){
                      Order = reorder(Order, gmin, median)) %>%
     filter(Order != "Other") %>% droplevels
   
-  par(mar=c(6,4,2,2))
+  par(mar=c(6,5,2,2))
   plotCI2(gmin, Order, gmindat3, transform_log10=TRUE,
-          ylab=expression(Conductance~(mmol~m^-2~s^-1)))
+          jit=0.5, datacex=0.5, datacol="grey",
+          ylim=c(0,25),
+          ylab=expression(g[min]~(mmol~m^-2~s^-1)))
 
 }
+
+gmin_3panel <- function(gmindat, cropgmin){
+  par(mfrow=c(1,3), mar=c(6,5,1,1), cex.lab=1.1, mgp=c(2.5, 0.5,0), tcl=-0.1)
+  gmin_loghist(gmindat)
+  par(las=2)
+  gmin_by_order(gmindat)
+  figure_crop_genotypes(cropgmin)
+}
+
+
+
+
 
 gmin_by_family <- function(gmindat){
   
[…]

### 2017-08-17 R/functions.R

@@ -99,7 +99,9 @@ plotlabel <- function(txt, where, inset=0.08, inset.x=inset, inset.y=inset,...){
 
 
 plotCI2 <- function(yvar, group, data, transform_log10=FALSE, 
-                    ylab=NULL, labels=NULL, label_las=3, ...){
+                    ylab=NULL, ylim=NULL, labels=NULL, label_las=3, add_data=TRUE, 
+                    jit=0.6, datacex=0.4, datacol="lightgrey", 
+                    ...){
   
   if(is.null(ylab))ylab <- substitute(yvar)
   
@@ -107,6 +109,8 @@ plotCI2 <- function(yvar, group, data, transform_log10=FALSE,
   
   data$Y <- eval(substitute(yvar), data)
   data$G <- as.factor(eval(substitute(group), data))
+  
+  datl <- split(data, data$G)
   
   if(is.null(labels)){
     labels <- levels(data$G)
@@ -122,15 +126,20 @@ plotCI2 <- function(yvar, group, data, transform_log10=FALSE,
     mn <- coef(fit)
   }
   
+  if(is.null(ylim))ylim <- c(0, max(ci)+0.1*max(ci))
   g <- glht(fit, linfct=mcp(G = "Tukey"))
   lets <- cld(g)$mcletters$Letters
   
   n <- length(mn)
   
-  plot(1:n, mn, pch=19, cex=1.1, ylim=c(0, max(ci)+0.1*max(ci)), axes=FALSE,
+  plot(1:n, mn, pch=19, cex=1.1, ylim=ylim, axes=FALSE,
        xlab="",
+       panel.first={
+         if(add_data)with(data, points(jitter(as.numeric(G), factor=jit), gmin, pch=16, col=datacol, cex=datacex))
+       },
        xlim=c(0.7, n+0.3), ylab=ylab, ...)
   axis(1, at=1:n, labels=labels, las=label_las)
+  
   arrows(x0=1:n, x1=1:n, y0=ci[,1], y1=ci[,2], angle=90, length=0.05, code=3)
   text(x=1:n, y=ci[,2], lets, pos=3, cex=0.8)
   par(las=1)
@@ -139,7 +148,7 @@ plotCI2 <- function(yvar, group, data, transform_log10=FALSE,
   
   tb <- unname(table(data$G))
   mtext(side=3, at=1:n, line=0.4, text=tb, las=1, cex=0.7)
[…]

### 2017-08-17 R/make_figures.R

@@ -7,12 +7,13 @@ to.pdf(figure_R2g0(lin2015coef, miner), "output/figure3.pdf", height=4, width=9)
 
 to.pdf(figure_gmin_review_2(gdfr), "output/figure4.pdf", height=5, width=5)
 
-to.pdf(figure_gmin_bygroup(gmindat), "output/figure5.pdf", height=4, width=3)
-
+to.pdf(gmin_3panel(gmindat, cropgmin), "output/figure5.pdf", height=3, width=8)
 
 to.pdf(figure_wtc4_gmin_2(wtc4gmin), "output/figure6.pdf", height=4, width=3.5)
 
 to.pdf(figure_hakea_gmin(lopw), "output/figure7.pdf", height=5, width=6)
+
+
 
 
 #convert_all_pdf("output")

### 2017-08-23 figures.Rmd
Update figures Rmd; captions written; more references

--- 2017-08-01 figures.Rmd
@@
 ---
-title: "gmin figures"
-author: "Remko Duursma"
+title: "On the minimum leaf conductance"
+author: ""
 date: "1 August 2017"
@@
     theme: flatly
+csl: global-change-biology.csl
+bibliography: references.bib
 ---
 
-```{r}
+```{r, echo=FALSE}
 source("R/load.R")
 pacman::p_load(magicaxis)
+knitr::opts_chunk$set(warning = FALSE, echo=FALSE)
 ```
 
-```{r}
-v <- gmindat$gmin
-hist(v, breaks=30, xlim=c(0,40), main="", freq=FALSE,
-     ylim=c(0,0.15),
-     col="lightgrey",
-     xlab=expression(g[min]~~(mmol~m^-2~s^-1)))
-curve(dlnorm(x, mean(log(v)), sd=sd(log(v))), add=TRUE,
-      n=501)
+```{r, fig.width=6, fig.height=6}
+figure_sim()
 ```
 
-```{r}
-v <- log10(gmindat$gmin)
-par(yaxs="i", mgp=c(3,1,0), cex.lab=1.2)
-hist(v, breaks=30, main="", freq=FALSE,
-     #ylim=c(0,0.15),
-     xlim=c(-1,2),
-     col="lightgrey",
-     axes=FALSE,
-     xlab=expression(g[min]~~(mmol~m^-2~s^-1)))
-magaxis(side=1, unlog=1, usepar=TRUE)
-axis(2)
-curve(dnorm(x, mean=mean(v), sd=sd(v)), add=TRUE,
-      n=101)
+**Figure 1**.
+Simulations  of intrinsic water-use efficiency (A~n~/g~s~), intercellular CO~2~ concentration (C~i~) and leaf transpiration rate (E~L~) with three values of g~0~.
+For the top two panels, all environmental drivers were held constant except the photosynthetic photon flux density (PPFD), demonstrating that the inclusion of g~0~ is necessary to simulate the decrease in A~n~/g~s~ at low PPFD.
+For the bottom two panels, VPD and T~air~ were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions.
+For all simulations it is assumed that T~leaf~ is equal to T~air~.
 
+```{r, fig.width=8, fig.height=4}
+figure_g0g1_cor(lin2015, "Nicolas Martin-StPaul_Quercus ilex_StPaul_Puechabon")
 ```
 
-```{r, fig.cap="Figure 1.", fig.width=6, fig.height=6}
-figure_sim()
-```
+**Figure 2**.
+Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al.
+(2011).
+This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al.
+(2015).
+The solid line is regression line, and the shaded area a 95% confidence interval for the mean.
+The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression.
+The dotted ellipse is a bivariate 95% confidence interval for slope and intercept.
+The symbols represent 1000 bootstrap samples of the coefficients.
 
@@
 
+**Figure 3**.
+Estimates of g~0~ depend on the overall goodness of fit.
+In the left panel, we fitted the linearized form of the Medlyn et al.
+(2011) model to each of the datasets in the Lin et al.
+(2015) database.
+Vertical lines are 95% confidence intervals.
+In the middle panel, a very similar association between g~0~ and R^2^ is presented by the review of @miner2017, using the Ball-Berry model.
+In the right panel, shown is the correlation between the standard error of the estimate of g~0~ and the coefficient of variation of the Ball-Berry coefficient (A/Ca*sqrt(D)), showing that good estimates of g~0~ can only be obtained when there is much variation in the drivers.
+
 ```{r, fig.width=5, fig.height=5}
@@
 
-```{r, fig.width=3, fig.height=4}
-figure_gmin_bygroup(gmindat)
+**Figure 4**.
+Comparison of various estimates of the minimum conductance.
+Error bars are 95\% confidence intervals.
+Different letters denote significant differences (at $\alpha$ = 0.05).
+See Table 1 for definitions of minimum conductance, and text for the data sources.
+g~cut,isol~, cuticular permeance on isolated cuticles, g~cut,seal~, mass loss of detached leaves when stomatal side of the leaf is sealed (*might remove this*), g~min~ minimum conductance with mass loss of detached leaves, g~dark~ nighttime conductance (from Lombardozzi), g~0~ fitted parameter, using the Lin database and Miner 2017, g~s,lowA~ stomatal conductance from the Lin database, calculated for each study where A~n~ < 2.
+
+```{r, fig.height=3, fig.with=8}
+gmin_3panel(gmindat, cropgmin)
 ```
+
+**Figure 5**.
+Literature compilation of minimum conductance (g~min~), measured with mass loss of detached leaves.
+Left panel, distribution across all 220 species values extracted from the literature, showing a log-normal distribution.
+Mean=, sd=.
+Middle panel, g~min~ shown by taxonomic order.
+Note higher g~min~ for grasses (Poales), but otherwise few clear differences.
+Sample size (number of species) shown on top of the figure.
+Right panel: genotypic variation in g~min~ for 9 major crops.
+Sample sizes are the number of distinct crop genotypes on which g~min~ was reported for each crop, bars denote the range in g~min~ across all the data for a crop, illustrating the very large genotypic variation compared to inter-specific variation.
 
@@
 
+**Figure 6**.
+Minimum conductance measured from leaf dry-down curves on *Eucalyptus parramattensis* grown in whole-tree chambers.
+Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C).
+Measurements of g~min~ were carried out at five different temperatures on replicate leaves.
+There is no effect of measurement T on g~min~, but leaves of trees grown in elevated temperature showed consistently lower g~min~.
+
 ```{r, fig.width=6, fig.height=5}
@@
 
-```{r, fig.width=6, fig.height=5}
-gmin_by_family(gmindat)
-```
+**Figure 7**.
+Minimum conductance measured from leaf dry-down curves on a variety *Hakea* species, a genus native to Australia.
+Plants were grown in containers in a grow house, and supplied with ample water or underwent long-term mild drought stress.
+Bars are labelled by leaf form : broadleaf (B) or needle-like (N).
+For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`).
 
-```{r, fig.width=6, fig.height=5}
-gmin_by_order(gmindat)
-```
+# References

### 2017-08-23 manuscript.Rmd
Update figures Rmd; captions written; more references

@@
 
-plotfig <- function(...)grid.raster(readPNG(...))
+source("R/load.R")
+```
 
-source("R/load.R")
-
-pval <- function(...)formatPval(..., includeEquality=TRUE)
-```
+# Abstract
 
@@
 \
-Stomata exert a strong control on the transpiration rate of leaves.
+Stomata exert a strong control on the transpiration rate of leaves [@bange1953].
 It is well known that stomata respond quickly to changes in light intensity, humidity and CO~2~ concentration at the leaf surface.
@@
 \
-Most current-generation process-baed global vegetation models (GVMs) use a version of the Ball-Berry class of stomatal conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity.
+Most current-generation process-based global vegetation models (GVMs) use a version of the Ball-Berry class of stomatal conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity.
 
@@
 If the cuticle thickness does not directly affect its conductance, what is the value of increasing the cuticle?
+
 Perhaps the answer lies in mechanical strength rather than water loss per se, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness).
+Providing support for the mechanical strength rather than water loss idea, thick cuticles may be useful because  exposure to wind has been shown to dislodge cuticular compounds by absrasion [@rogge1993], thereby dramatically increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991].
+Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986].
+
+It is also important to bear in mind other functions of the cuticle, particularly as a barrier for fungal pathogens, and insect herbivory [@kerstiens1996c].
 
@@
 
+@bengtson1978 lower g~min~ in six oat varieties, and increase in cuticular waxes, but no overall relationship between the two and very different responses between different oat genotypes.
+
 ## Minimum conductance acclimates to growing conditions
@@
 Drought caused a change in cuticular wax production and chemical composition, but in a highly genotype-specific way.
+
+@paje1988 lower gmin in droughted soybean genotypes
+
+@saneoka1987 lower g~min~ (but different technique!) in three crop species exposed to low soil water conditions for a month.
 

### 2017-08-23 R/functions.R

@@ -8,6 +8,8 @@ to.pdf <- function(expr, filename, ..., verbose=TRUE) {
   eval.parent(substitute(expr))
 }
 
+plotfig <- function(...)grid.raster(readPNG(...))
+pval <- function(...)formatPval(..., includeEquality=TRUE)
 
 
 fits_lin2015 <- function(lin2015a){

### 2017-08-23 R/make_figures.R

@@ -1,5 +1,8 @@
 
+source("R/load.R")
+
 to.pdf(figure_sim(), "output/figure1.pdf", height=6, width=6)
+
 to.pdf(figure_g0g1_cor(lin2015, "Nicolas Martin-StPaul_Quercus ilex_StPaul_Puechabon"), 
        "output/figure2.pdf", height=4, width=8)
 

### 2017-08-30 manuscript.Rmd
drought effect table added; bunch of refs fixed

@@
 ---
-title: "On the minimum leaf conductance"
+title: "On the minimum leaf conductance: ecological and environmental controls, and its role in models of plant water use"
 author: ''
@@
 
+Stomata exert a strong control on the transpiration rate of leaves.
+During severe drought, or conditions where photosynthesis does not take place, it seems advantageous for plants to completely halt water loss.
+However, water loss continues through the cuticle and incompletely closed stomata, together determined by the leaf minimum conductance (gmin).
+In this review, we illustrate the role of g~min~ in models of leaf gas exchange, critically evaluate sources of variation in g~min~, and quantitatively compare various methods for its estimation.
+A literature compilation of g~min~ as measured by weight loss of detached leaves is presented, which shows much variation in this trait that is not clearly related to species groups, climate of origin, or leaf type.
+Much evidence points to the idea that g~min~ is highly responsive to the growing conditions of the plant, including soil water availability, temperature, and air humidity - as we further demonstrate with two case studies.
+The synthesis presented here provides guidelines for use of g~min~ in ecosystem models, and points to clear research gaps for this drought tolerance trait.
+
 # Introduction
@@
 In this paper, we refer to Eq.
-1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987) (NB they used relative humidity).
+1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987) (they used relative humidity).
 
@@
 Finally, measuring the low water loss rates of detached leaves mimicks drought conditions, yielding an estimate of whole-leaf minimum conductance (g~min~) [@pisekberger1938; @hygen1951].
-Photosynthesis also ceases at very high temperature (ref), but stomata do not appear to close in concert [@drake, @urban], thus we cannot use high temperature to estimate a reasonable g~0~.
+Photosynthesis also ceases at very high temperature (ref), but stomata do not appear to close in concert [@drake, @urban2017], thus we cannot use high temperature to estimate a reasonable g~0~.
 
@@
 
-For the @lin2015 database, we fit the model of @medlyn2011 using non-linear regression, while @miner2016 presented results from linear regression of the @ballberry1987 model.
+For the @lin2015 database, we fit the model of @medlyn2011 using non-linear regression, while @miner2017 presented results from linear regression of the @ballberry1987 model.
 
@@
 It is no new observation that stomata do not close in the dark, but do close on detached leaves, as already demonstrated by @hygen1954 in Scots pine, and described by @koerner1994.
-More recently, @waldencoleman2013 showed much higher g~dark~ compared to g~min~ in Cotton.
+More recently, @walden-coleman2013 showed much higher g~dark~ compared to g~min~ in Cotton.
 
@@
 If we would use the fitted g~0~ in models, simulated g~s~ would never drop below g~0~.
+
+@cavender-bares2007 gdark similar to gmin in oaks (unusual!).
 
@@
 Because most of the resistance is located in such a thin layer, g~cut~ does not correlate with the thickness of the cuticle  [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b].
-Nonetheless there is considerable variation in cuticle thickness between plant species [@schuster2016a], as well as within species in response to drought (cite), along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woordruff2010].
+Nonetheless there is considerable variation in cuticle thickness between plant species [@schuster2016a], as well as within species in response to drought (cite), along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010].
 If the cuticle thickness does not directly affect its conductance, what is the value of increasing the cuticle?
 
-Perhaps the answer lies in mechanical strength rather than water loss per se, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness).
-Providing support for the mechanical strength rather than water loss idea, thick cuticles may be useful because  exposure to wind has been shown to dislodge cuticular compounds by absrasion [@rogge1993], thereby dramatically increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991].
+It is important to bear in mind other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, and various interactions with insects [@mueller2008] [@kerstiens1996c;@riederer2008].
+The thickness of the cuticle per se may also have to do with mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness).
+Thick cuticles may also be useful because  exposure to wind has been shown to dislodge cuticular compounds by absrasion [@rogge1993], thereby dramatically increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991].
+
+@grace1974: in a *Festuca* species, g~min~ more than doubled after exposure to high wind
+
 Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986].
-
-It is also important to bear in mind other functions of the cuticle, particularly as a barrier for fungal pathogens, and insect herbivory [@kerstiens1996c].
+Such exposure to the elements may be responsible for the typical gradual increase in g~min~ with leaf age in field-grown plants [@jordan2007].
 
@@
 For seven conifer species, @brodribb2014 reported that 50 - 94\% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky.
-However @santrucek found that the cuticle (solid mass only) of the stomatal side in *Hedera helix* was many times more more permeable than the non-stomatal side, thus providing another potential explanation for this difference.
+However @santrucek2004 found that the cuticle (solid mass only) of the stomatal side in *Hedera helix* was many times more more permeable than the non-stomatal side, thus providing another potential explanation for this difference.
 Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles.
 
-In the following sections, we review the quite substantial literature on g~min~ measured with the MLD technique, stretching back to the 1930s [@pisekberger], focussing on environmental and ecological determinants.
+In the following sections, we review the quite substantial literature on g~min~ measured with the MLD technique, stretching back to the 1930s [@pisekberger1938], focussing on environmental and ecological determinants.
 We also further tease apart the literature compilation of g~min~ (shown just as a single average value in Fig.
@@
 
-@bengtson1978 lower g~min~ in six oat varieties, and increase in cuticular waxes, but no overall relationship between the two and very different responses between different oat genotypes.
-
 ## Minimum conductance acclimates to growing conditions
 
-@bi2017 lower residual transpiration rate in drought-acclimated wheat, in 3 out of 4 genotypes.
-Drought caused a change in cuticular wax production and chemical composition, but in a highly genotype-specific way.
+### Drought
 
-@paje1988 lower gmin in droughted soybean genotypes
+@bengtson1978 lower g~min~ in six oat varieties in response to drought, and increase in cuticular waxes, but no overall relationship between the two and very different responses between different oat genotypes.
 
-@saneoka1987 lower g~min~ (but different technique!) in three crop species exposed to low soil water conditions for a month.
+@bi2017 Drought caused a change in cuticular wax production and chemical composition, but in a highly genotype-specific way.
 
-@rosasanderson2014, lower gmin in drought-stressed peanut
+@kottapalli2009 higher epicuticular wax load in a drought-tolerant vs. susceptible peanut genotype
 
-@clark1956 lower gmin in drought-stressed soybean
+Ubiquitous response of g~min~ after drought stress, always a decrease. Only study we found that did not find a difference was @vilagrosa2003. All other studies summarized in TableXX.
+
+```{r, echo=FALSE}
+dat <- read.csv("data/drought_gmin_table.csv", stringsAsFactors = FALSE)
+
+knitr::kable(dat, 
+             col.names= c("Study","Species","Experiment","Reduction in g~min~"))
+```
+
+### Altitude
 
@@
 
+### Leaf age
+
+Also discussed above!
+Bunch of papers in gmindatabase have age.
+
 @jordan2007 large increase gmin in the herb Agastachys odorata with leaf age
 
-@fanourakis2013 gmin (desiccation gmin) much higher in high RH grown plants, at any value of leaf water content (Table 1).
+### Humidity and temperature
+
+@fanourakis2013 gmin (MLD) much higher in high RH grown plants, at any value of leaf water content (Table 1).
 This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss (Hypothesis 2).
 
-@schuster2016 thermal tolerance and gmin relationship across species.
-
-@cavenderbares2007 gdark similar to gmin in oaks (unusual!). Lower gmin in droughted oaks.
-
-- @schuster2016 correlation temperature, gcut AND gmin for Nerium oleander (page 59).
+@schuster2016a thermal tolerance and gmin relationship across species.
+@schuster2016a correlation temperature, gcut AND gmin for Nerium oleander (page 59).
 T range much wider than Blackman though...
 But Schuster gmin not different from gcut at most temperatures, which is a bit strange too.
+
+Also see @paje1988 for more complex response to temperature
 

### 2017-09-19 manuscript.Rmd
Switched to tansley review format, lots of reorganizing

@@
 ---
-title: "On the minimum leaf conductance: ecological and environmental controls, and its role in models of plant water use"
+title: "On the minimum leaf conductance: its role in models of plant water use, and ecological and environmental controls"
 author: ''
@@
     theme: flatly
-csl: global-change-biology.csl
+csl: new-phytologist.csl
 bibliography: references.bib
@@
 
-# Abstract
+# Summary
 
@@
 
-# Introduction
+# I. Introduction
 
@@
 Plants cannot, however, completely halt water loss even when stomata are closed.
-In this review, we discuss various aspects of the minimum leaf conductance reached when stomata are essentially closed.
-We pay special attention to the role that minimum conductance plays in land-surface models, in particular as it forms part of the much-used 'Ball-Berry'-type class of stomatal conductance models.
 
@@
 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987) (they used relative humidity).
-
-\
-The effects of photosynthetic photon flux density (PPFD) and leaf temperature are both assumed to enter via the dependence of A~n~ on these drivers.
-The g~0~ parameter is another definition of the minimum conductance, but we keep the notation g~0~ to indicate that it is a parameter in the Ball-Berry model.
-Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the inocuous g~0~ parameter has been largely ignored.
+It should be clear from Eq.
+1 that the Ball-Berry model is actually a model of water-use efficiency, since both g~s~ and A~n~ appear in the equation.
+Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been largely ignored.
 For example, a recent review of stomatal conductance models did not mention the minimum conductance [@damour2010].
@@
 It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model.
+Discuss @barnard2013 briefly (in a positive way :)
 
 \
-The g~0~ parameter not only sets a minimum value of g~s~ in the model, it also modifies the behaviour of Eq.
-1 after it has been coupled to the Farquhar-von Caemmerer-Berry model of photosynthesis (the so-called 'coupled leaf gas exchange model').
-@leuning1990 defined g~0~ as g~s~ when A approaches zero as PPFD approaches the light compensation point, and pointed out that g~0~ is necessary to simulate the increase in A/g~s~ (and C~i~) at low light [see also @leuning1995].
-Similarly, @collatz1992 showed that including g~0~ affects the response of g~s~ to relative humidity in the coupled leaf gas exchange model.
-Despite these early reports, the exact role of g~0~ in models is often overlooked.
+The absolute minimum attainable water loss rate through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf.
+Other measurements allow for the estimation of water loss through incompletely closed and broken stomata.
+Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time.
+This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants that are severely droughted, and includes all surfaces of the leaf.
+Since the seminal review by @kerstiens1996, no attempt has been made to synthesize the literature for values of g~min~, which is timely as models start to incorporate more detailed drought dynamics all the way to the point of mortality by desiccation (Xu, Blackman, Martin-StPaul, Sperry?).
+Instead, many models rely on inaccurately estimated 'minimum' conductance from cuvette data, by applying regression techniques.
+We critically examine this method in Section III.
 
 \
-Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero.
-But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, begging the question whether the same g~0~ is reached in each of these cases.
-In the application of the Ball-Berry models, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005].
-It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata [@kerstiens1996].
-Finally, measuring the low water loss rates of detached leaves mimicks drought conditions, yielding an estimate of whole-leaf minimum conductance (g~min~) [@pisekberger1938; @hygen1951].
-Photosynthesis also ceases at very high temperature (ref), but stomata do not appear to close in concert [@drake, @urban2017], thus we cannot use high temperature to estimate a reasonable g~0~.
+Previews reviews have focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles (riederer, kerstiens).
+Thus there is a disconnect between dizzying level of detailed understanding of the plant cuticle [@riederer2008], and the very simple  assumptions made when applying ecosystem models.
+We aim to improve the connection somewhat, and point at obvious gaps in understanding.
+
+# II. Comparison of various definitions of minimum conductance
+
+**Table 1.** Definitions of minimum conductance.
+
+| Variable   |  Definition |
+| ---------- | ----------- |
+| g~cut~     | Conductance of an isolated non-stomatal cuticle.         |
+| g~min~     | Minimum conductance measured from weight loss of detached leaves. |
+| g~dark~    | Night-time conductance, or conductance after significant dark adaptation.              |
+| g~0~       | Intercept in the Ball-Berry-type stomatal conductance model, i.e. g~s~ when A~n~ approaches zero. |
+| g~s,low A~ | Cuvette measurements of g~s~ when photosynthesis rates are very low. |
+
+Measurements of (presumed) minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories.
+The absolute minimum conductance attained by leaves is that through the cuticle only.
+Measurements of cuticular conductance (g~cut~) are typically made on isolated cuticles [@riederer2001] (g~cut,isol~), or on the side of the leaf that lacks stomata ('adaxial' side), or by sealing the side with stomata (g~cut,seal~).
+Another measurement is to detach leaves, and measure leaf weight over time as the leaf dries out.
+After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration rate'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit [@sinclair1986].
+This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress.
 
 \
-In this paper, we a) review the role and current use of g~0~ in models, b) perform a quantitative comparison of various measurements of the presumed minimum conductance, including fitting g~0~ with Eq.
-1, night-time conductance (based on a previous compilation), and a new compilation of g~min~ estimates from weight-loss of detached leaves, and c) with use of two case studies and review of the literature, we demonstrate that g~min~ shows considerable acclimation to growing conditions.
-
-# Methods
-
-## Coupled leaf gas exchange model
-
-To illustrate the effect of g~0~ on the modelled response of g~s~ to PPFD and T~leaf~, it is necessary to embed the stomatal conductance model in a coupled leaf gas exchange model, which is formed by three equations; Eq.
-1 giving the dependence of g~s~ on A~n~ and some environmental drivers, the diffusion constraint,
-
-\
-(2)
-$$ A_n = \frac{g_s}{1.6}(C_a - C_i) $$
-
-\
-where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~). And finally the Farquhar-von Caemmerer-Berry photosynthesis model that gives A~n~ as a function of C~i~,
-
-\
-(3)
-$$ A_n = \frac{a \cdot (C_i - \Gamma^{*})}{C_i + b} - R_d$$
-
-\
-where a = V~cmax~ or J/4 (where the electron transport rate J approaches Jmax at high light) depending on the limitation to photosynthesis, and b also varies.
-The dependence on temperature enters through V~cmax~, J~max~, R~d~, $\Gamma^{*}$, and other parameters [@medlyn2002].
-For all simulations with this model, we used the implementation in the *plantecophys* R package [@duursma2015].
-
-## Literature review
-
 For the @lin2015 database, we fit the model of @medlyn2011 using non-linear regression, while @miner2017 presented results from linear regression of the @ballberry1987 model.
-
-\
 Measurements of conductance of isolated, non-stomatal cuticles (g~cut,isol~), and of conductance of detached leaves with the stomatal side sealed (g~cut,seal~) were taken from the compilation by @kerstiens1996.
-
-\
-Data on minimum conductance (g~min~) from weight loss of detached leaves was compiled from 44 original papers, see Appendix A for a list of papers.
+Data on minimum conductance (g~min~) from weight loss of detached leaves was compiled from 44+ original papers, see SuppInfo A for a list of papers and additional figures describing this new compilation.
 For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources.
@@
 @kerstiens1996).
-
-\
+Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance, it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful.
 A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca.
@@
 
-## Case studies on minimum conductance of detached leaves
+\
+Conductance in the dark was on average eight-fold higher than g~min~.
+It is no new observation that stomata do not close in the dark, but do close on detached leaves, as already demonstrated by @hygen1954 in Scots pine, and described by @koerner1994.
+More recently, @walden-coleman2013 showed much higher g~dark~ compared to g~min~ in Cotton.
+However, @cavender-bares2007 gdark similar to gmin in oaks.
+Actual measured stomatal conductance during conditions when photosynthesis was low (g~s,low A~) was also much lower than fitted g~0~ or g~dark~.
+In other words, actual daytime conductance is much lower than the presumed minimum estimated from fitted g~0~.
+A similar result was already presented by @dekauwe2015 (their  Fig.
+A2).
 
-Both studies use identical methods.
-The first dataset was collected on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature.
-Leaves were sampled from XXX, n=XXX.
-The second case study was on *Hakea* plants.
-Details here (measurement T?).
-Methods for gmin from drying curves here.
-
-# The use of g~0~ in models
-
-\
-A popular approach in thinking about how stomata 'should' respond to environmental drivers is the idea that stomatal conductance is varied to maximize total photosynthesis for a given amount of water use [@cowan1977].
-The consequence is that stomata tend to open during periods that are favorable for photosynthesis (high light, optimal temperature), and close when photosynthesis drops to zero (darkness, very high temperature).
-Clearly, in this framework there is no place for g~0~ since it is always sub-optimal to open stomata (i.e.
-spend water) when there is no photosynthetic gain.
-Indeed, early work on optimal stomatal conductance models ignored the possibility of g~s~ > 0 when A~n~ = 0 [@cowan1977; @cowan1978; @hari1986; @hari1999].
-More recent work derives the optimal g~s~, and simply adds a g~0~ to the solution [@medlyn2011].
-
-\
-In the following we show how inclusion of g~0~ in models affects the response of C~i~ at low PPFD and the response of transpiration to high leaf temperature (T~leaf~).
-Another reason to include a non-zero g~0~ term in models of stomatal conductance arises during severe drought.
-When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata.
-Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000].
-
-\
-As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) is constant with the photon flux density (PPFD).
-When PPFD approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~.
-To see this point, we can rearrange Eq.
-1 to give,
-
-\
-(2)
-$$A_n/g_s = {C_a}f(D) / g_1$$
-
-and using the diffusion constraint $A~n~ = g~s~ / 1.6 \cdot (C~a~ - C~i~)$ we get an expression for C~i~ / C~a~,
-
-\
-(3)
-$$C_i / C_a = 1 - \frac{f(D)}{1.6 \cdot g_1}$$
-
-Eq. 3 - derived using g~0~ - thus does not give any dependence on PPFD.
-
-\
-The g~0~ parameter also represents g~s~ when T~leaf~ is very high, because A~n~ goes to zero at high T~leaf~.
-As a result, simulations of vegetation water use during heat waves depends highly on the value of g~0~ used [@kala2016] .
-It should be pointed out that the response of g~s~ to T~leaf~ is likely poorly modelled with a Ball-Berry model, since a number of studies have shown that g~s~ and A~n~ decouple at high T~leaf~, resulting in high rates of g~s~ when A~n~ is near zero [@urban2017] (Drake et al., unpublished).
-
-\
-Using the coupled leaf gas exchange model, we show how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^).
-We also demonstrate the effect of g~0~ on modelling the T~leaf~ response.
-In this simulation, T~leaf~ and D are assumed to co-vary with an empirical relationship as used by @duursma2014.
-Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increase leaf transpiration E~L~, but at a slightly higher rate than just due to g~0~.
-This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq.
-1.
-
-# Estimation of g~0~ by fitting
+# III. Estimating minimum conductance from cuvette data
 
@@
 
-# Comparison of various definitions of minimum conductance
+# IV. Use of minimum conductance in models
 
-**Table 1.** Definitions of minimum conductance.
-
-| Variable   |  Definition |
-| ---------- | ----------- |
-| g~cut~     | Conductance of an isolated non-stomatal cuticle         |
-| g~min~     | Minimum conductance measured from weight loss of detached leaves |
-| g~dark~    | Night-time conductance                                  |
-| g~0~       | Intercept in the Ball-Berry-type stomatal conductance model, fitted. |
-| g~s,low A~ | Cuvette measurements of g~s~ when photosynthesis rates are low. |
-
-Measurements of (presumed) minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories.
-The absolute minimum conductance attained by leaves is that through the cuticle only.
-Measurements of cuticular conductance (g~cut~) are typically made on isolated cuticles [@riederer2001] (g~cut,isol~), or on the side of the leaf that lacks stomata ('adaxial' side), or by sealing the side with stomata (g~cut,seal~).
-Another measurement is to detach leaves, and measure leaf weight over time as the leaf dries out.
-After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration rate'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit.
-This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress.
-Because night-time conductance (g~dark~) is frequently used as an independent measurement of g~0~, we compare literature values of g~dark~ with g~min~ and g~cut~, as well as the compilation of fitted values of g~0~ (already shown in Fig.
-3).
-
-Results shown in Fig. 4.
+## 1. Models of water-use efficiency
 
 \
-Conductance in the dark was on average eight-fold higher than g~min~.
-It is no new observation that stomata do not close in the dark, but do close on detached leaves, as already demonstrated by @hygen1954 in Scots pine, and described by @koerner1994.
-More recently, @walden-coleman2013 showed much higher g~dark~ compared to g~min~ in Cotton.
+The effects of photosynthetic photon flux density (PPFD) and leaf temperature on g~s~ are both assumed to enter via the dependence of A~n~ on these drivers.
+The g~0~ parameter not only sets a minimum value of g~s~ in the model, it also modifies the behaviour of Eq.
+1 after it has been coupled to the Farquhar-von Caemmerer-Berry model of photosynthesis (the so-called 'coupled leaf gas exchange model').
+@leuning1990 defined g~0~ as g~s~ when A approaches zero as PPFD approaches the light compensation point, and pointed out that g~0~ is necessary to simulate the increase in A/g~s~ (and C~i~) at low light [see also @leuning1995].
+Similarly, @collatz1992 showed that including g~0~ affects the response of g~s~ to relative humidity in the coupled leaf gas exchange model.
+Despite these early reports, the exact role of g~0~ in models is often overlooked.
 
 \
-Actual measured stomatal conductance during conditions when photosynthesis was low (g~s,low A~) was also much lower than fitted g~0~ or g~dark~.
-In other words, actual daytime conductance is much lower than the presumed minimum estimated from fitted g~0~.
-A similar result was already presented by @dekauwe2015 (their  Fig.
-A2).
-If we would use the fitted g~0~ in models, simulated g~s~ would never drop below g~0~.
+Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero.
+But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, begging the question whether the same g~0~ is reached in each of these cases.
+In the application of the Ball-Berry models, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005].
+It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata [@kerstiens1996].
+Finally, measuring the low water loss rates of detached leaves mimicks drought conditions, yielding an estimate of whole-leaf minimum conductance (g~min~) [@pisekberger1938; @hygen1951].
+Photosynthesis also ceases at very high temperature (ref), but stomata do not appear to close in concert [@drake, @urban2017], thus we cannot use high temperature to estimate a reasonable g~0~.
 
-@cavender-bares2007 gdark similar to gmin in oaks (unusual!).
+\
+A popular approach in thinking about how stomata 'should' respond to environmental drivers is the idea that stomatal conductance is varied to maximize total photosynthesis for a given amount of water use [@cowan1977].
+The consequence is that stomata tend to open during periods that are favorable for photosynthesis (high light, optimal temperature), and close when photosynthesis drops to zero (darkness, very high temperature).
+Clearly, in this framework there is no place for g~0~ since it is always sub-optimal to open stomata (i.e.
+spend water) when there is no photosynthetic gain.
+Indeed, early work on optimal stomatal conductance models ignored the possibility of g~s~ > 0 when A~n~ = 0 [@cowan1977; @cowan1978; @hari1986; @hari1999].
+More recent work derives the optimal g~s~, and simply adds a g~0~ to the solution [@medlyn2011].
 
-### Low conductance of isolated cuticles
+\
+As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) is constant with the photon flux density (PPFD).
+When PPFD approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~.
+To see this point, we can rearrange Eq.
+1 to give,
+
+\
+(2)
+$$A_n/g_s = {C_a}f(D) / g_1$$
+
+and using the diffusion constraint $A~n~ = g~s~ / 1.6 \cdot (C~a~ - C~i~)$ we get an expression for C~i~ / C~a~,
+
+\
+(3)
+$$C_i / C_a = 1 - \frac{f(D)}{1.6 \cdot g_1}$$
+
+Eq. 3 - derived using g~0~ - thus does not give any dependence on PPFD.
+
+\
+The g~0~ parameter also represents g~s~ when T~leaf~ is very high, because A~n~ goes to zero at high T~leaf~.
+As a result, simulations of vegetation water use during heat waves depends highly on the value of g~0~ used [@kala2016] .
+It should be pointed out that the response of g~s~ to T~leaf~ is likely poorly modelled with a Ball-Berry model, since a number of studies have shown that g~s~ and A~n~ decouple at high T~leaf~, resulting in high rates of g~s~ when A~n~ is near zero [@urban2017] (Drake et al., unpublished).
+
+\
+Using the coupled leaf gas exchange model, we show how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^).
+We also demonstrate the effect of g~0~ on modelling the T~leaf~ response.
+In this simulation, T~leaf~ and D are assumed to co-vary with an empirical relationship as used by @duursma2014.
+Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increase leaf transpiration E~L~, but at a slightly higher rate than just due to g~0~.
+This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq.
+1.
+
+## 2. Models of drought response of stomatal conductance
+
+Another reason to include a non-zero g~0~ term in models of stomatal conductance arises during severe drought.
+When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata.
+Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000].
+
+@pisekwinkler1953, early calculation of the length of time needed to desiccate leaves to some water content, given the minimum transpiration rate, and the water content of the leaves.
+
+@sinclair2000 gmin as a drought tolerance trait, can calculate time needed to death during drought phase when plants no longer take up water.
+
+Refined by @blackman2016, incorporating more well-known hydraulics traits into the framework.
+
+Discuss Martin-StPaul (2017); key role of minimum conductance in a model of desiccation dynamics
+
+Various new models that aim to simulate desiccation have a g~min~ term in it, but it is not yet well constrained by data.
+
+# VI. Cuticular conductance
 
@@
 
+\
 Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle.
@@
 
+\
 It is important to bear in mind other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, and various interactions with insects [@mueller2008] [@kerstiens1996c;@riederer2008].
@@
 Thick cuticles may also be useful because  exposure to wind has been shown to dislodge cuticular compounds by absrasion [@rogge1993], thereby dramatically increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991].
-
-@grace1974: in a *Festuca* species, g~min~ more than doubled after exposure to high wind
-
 Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986].
@@
 
+\
 Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged.
@@
 
-@mackova2013 adding ABA to Lepidium sativum increased the chain length of cuticular waxes (but not total amounts).
-
-@riederer2001 @schuster2016b correlation between temperature and gcut (non-stomatal cuticles)
-
+\
 Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015].
 This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata.
+Nonetheless it is an interesting observation that could have major consequences for modelling and interpreting water-use efficiency data during drought conditions.
 
-# Nighttime conductance
+# VII. Nighttime conductance
 
@@
 
-# Minimum conductance of detached leaves
+# VIII. Environmental and ecological variation in minimum conductance
+
+## 1. Contribution of stomata to minimum conductance
 
@@
 
+## 2. Minimum conductance shows little meaningful interspecific variation
+
 In the following sections, we review the quite substantial literature on g~min~ measured with the MLD technique, stretching back to the 1930s [@pisekberger1938], focussing on environmental and ecological determinants.
@@
 4).
-
-## Minimum conductance shows little meaningful interspecific variation
 
@@
 
-Comparing all available data, we did not find significant relationships between rainfall, aridity, or temperature and g~min~.
-There is, however, a substantial acclimatory response of g~min~ to water availability.
+Comparing all available data, we did not find significant relationships between rainfall, aridity, or temperature and g~min~ (see SuppInfo 2).
 
-## Minimum conductance acclimates to growing conditions
-
-### Drought
+## 3. Acclimation to drought
 
 @bengtson1978 lower g~min~ in six oat varieties in response to drought, and increase in cuticular waxes, but no overall relationship between the two and very different responses between different oat genotypes.
+
+@mackova2013 adding ABA to Lepidium sativum increased the chain length of cuticular waxes (but not total amounts).
 
@@
 Ubiquitous response of g~min~ after drought stress, always a decrease. Only study we found that did not find a difference was @vilagrosa2003. All other studies summarized in TableXX.
+
+The second case study was on *Hakea* plants. Details here (measurement T?).
 
@@
 
-### Altitude
-
-@anfodillo2002 gmin high altitude.
-Four-fold increase gmin from February (winter) to August in Picea abies, probably due to incomplete stomatal closure not changes in cuticle properties.
-Cite common finding that gmin increases with altitude [e.g.
-@delucia1984; @herrick1991].
-
-### Leaf age
-
-Also discussed above!
-Bunch of papers in gmindatabase have age.
-
-@jordan2007 large increase gmin in the herb Agastachys odorata with leaf age
-
-### Humidity and temperature
-
-@fanourakis2013 gmin (MLD) much higher in high RH grown plants, at any value of leaf water content (Table 1).
-This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss (Hypothesis 2).
+## 4. Acclimation and instantaneous response to temperature
 
@@
 
+@riederer2001 @schuster2016b correlation between temperature and gcut (non-stomatal cuticles)
+
 We present two case studies demonstrating the acclimation of g~min~ to environmental drivers.
@@
 This is not due to a direct effect of temperature on in cuticular conductance, which increases with temperature (Schreiber 2001).
+
+A case study on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature.
 Our results showed a large  acclimation effect but no evidence for a direct T effect (over the narrow range in T compared to Schreiber).
+
+## 5. Other factors
+
+Also discussed above in Section~(cuticular conductance), the cuticle is constantly exposed to physical (wind, rain) and biotic (insects, fungi) agents which damage the cuticle over time.
+
+@jordan2007 large increase gmin in the herb Agastachys odorata with leaf age
+
+--> more in database
+
+@anfodillo2002 gmin high altitude.
+Four-fold increase gmin from February (winter) to August in Picea abies, probably due to incomplete stomatal closure not changes in cuticle properties.
+Cite common finding that gmin increases with altitude [e.g.
+@delucia1984; @herrick1991].
+
+@fanourakis2013 gmin (MLD) much higher in high RH grown plants, at any value of leaf water content (Table 1).
+This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss (Hypothesis 2).
 
@@
 
+# Appendix or SuppInfo
+
+\
+To illustrate the effect of g~0~ on the modelled response of g~s~ to PPFD and T~leaf~, it is necessary to embed the stomatal conductance model in a coupled leaf gas exchange model, which is formed by three equations; Eq.
+1 giving the dependence of g~s~ on A~n~ and some environmental drivers, the diffusion constraint,
+
+\
+(2)
+$$ A_n = \frac{g_s}{1.6}(C_a - C_i) $$
+
+\
+where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~). And finally the Farquhar-von Caemmerer-Berry photosynthesis model that gives A~n~ as a function of C~i~,
+
+\
+(3)
+$$ A_n = \frac{a \cdot (C_i - \Gamma^{*})}{C_i + b} - R_d$$
+
+\
+where a = V~cmax~ or J/4 (where the electron transport rate J approaches Jmax at high light) depending on the limitation to photosynthesis, and b also varies.
+The dependence on temperature enters through V~cmax~, J~max~, R~d~, $\Gamma^{*}$, and other parameters [@medlyn2002].
+For all simulations with this model, we used the implementation in the *plantecophys* R package [@duursma2015].
+
 # References {.unnumbered}

### 2017-09-19 R/figures.R

@@ -45,6 +45,8 @@ figure_gmin_review_2 <- function(gdfr){
 }
 
 
+
+
 figure_gmin_bygroup <- function(gmindat){
   
   par(mar=c(6.5,4,1.5,1), mgp=c(2.5,0.5,0), tcl=-0.2, 

### 2017-09-19 R/load.R
minor fixes

@@ -169,7 +169,7 @@ gmindat$group2[is.na(gmindat$group2)] <- "Other non-woody"
 clsfile <- "data/cls_output.rds"
 sp_dfr <- data.frame(species=unique(gmindat$species), stringsAsFactors = FALSE)
 if(!file.exists(clsfile)){
-  cls <- classification(sp, db="ncbi")
+  cls <- classification(sp_dfr$species, db="ncbi")
   saveRDS(cls, clsfile)
 } else {
   cls <- readRDS(clsfile)

### 2017-09-21 manuscript.Rmd
some more writing; table of contents Tansley style

@@
 library(grid)
+library(knitr)
 
 source("R/load.R")
+```
+
+**Contents:**
+
+```{r, echo=FALSE}
+tab <- read.csv(text="I. Introduction
+II. Comparison of various definitions of minimum conductance
+III. Problems with estimating minimum conductance from cuvette data
+IV. Use of minimum conductance in models
+\t1. Models of water-use efficiency
+\t2. Models of drought response of stomatal conductance
+V. Cuticular conductance
+VI. Nighttime conductance
+VII. Environmental and ecological variation in minimum conductance 
+\t1. Contribution of stomata to minimum conductance
+\t2. Minimum conductance shows little meaningful interspecific variation
+\t3. Acclimation to drought
+\t4. Acclimation and instantaneous response to temperature
+\t5. Other factors
+VIII. Conclusions", header=F)
+
+knitr::kable(tab[[1]], col.names="")
+
 ```
@@
 Instead, many models rely on inaccurately estimated 'minimum' conductance from cuvette data, by applying regression techniques.
-We critically examine this method in Section III.
 
 \
-Previews reviews have focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles (riederer, kerstiens).
+(*expand*) Previews reviews have focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles (riederer, kerstiens).
 Thus there is a disconnect between dizzying level of detailed understanding of the plant cuticle [@riederer2008], and the very simple  assumptions made when applying ecosystem models.
 We aim to improve the connection somewhat, and point at obvious gaps in understanding.
+Throughout our review, we discuss both the practical implementation of a 'minimum' conductance g~0~ as implemented in e.g.
+Ball-Berry models of stomatal conductance (Eq.
+1) and similar, and a 'true' minimum conductance as estimated on detached leaves (g~min~).
 
@@
 The absolute minimum conductance attained by leaves is that through the cuticle only.
-Measurements of cuticular conductance (g~cut~) are typically made on isolated cuticles [@riederer2001] (g~cut,isol~), or on the side of the leaf that lacks stomata ('adaxial' side), or by sealing the side with stomata (g~cut,seal~).
-Another measurement is to detach leaves, and measure leaf weight over time as the leaf dries out.
-After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration rate'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit [@sinclair1986].
+Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001] (g~cut,isol~), or via gas exchange techniques, or mass loss of detached leaves vby sealing the side with stomata (g~cut,seal~).
+The method we here refer to as minimum conductance is that of mass loss of detached leaves (MLD).
+In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out.
+After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit [@sinclair1986].
 This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress.
+The g~0~ parameter, which is a *modelled* minimum conductance reached when photosynthesis rate is zero, is typically estimated by applying regression to Eq.
+1 (of which various forms exist).
 
 \ 
-For the @lin2015 database, we fit the model of @medlyn2011 using non-linear regression, while @miner2017 presented results from linear regression of the @ballberry1987 model.
+We quantitatively compared each of these different 'minimum' codnductance values by synthesizing existing data.
+To estimate g~0~ from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model.
 Measurements of conductance of isolated, non-stomatal cuticles (g~cut,isol~), and of conductance of detached leaves with the stomatal side sealed (g~cut,seal~) were taken from the compilation by @kerstiens1996.
-Data on minimum conductance (g~min~) from weight loss of detached leaves was compiled from 44+ original papers, see SuppInfo A for a list of papers and additional figures describing this new compilation.
+Data on minimum conductance (g~min~) from MLD was compiled from 44+ original papers (see SuppInfo A for a list of papers and additional figures describing this new compilation).
 For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources.
@@
 \
+(*expand*) The results are presented in Fig.
+1.
 Conductance in the dark was on average eight-fold higher than g~min~.
@@
 
-# III. Estimating minimum conductance from cuvette data
+# III. Problems with estimating minimum conductance from cuvette data
 
@@
 The g~0~ parameter is thus estimated as the intercept.
-Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated in this way.
+Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated by this method.
 The difficulty of the estimation can already be concluded from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g.
-@leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species); presumably because unreliable estimates were obtained, b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though nonsensical.
+@leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species); presumably because unreliable estimates were obtained, b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though it is clearly non-sensical to suggest negative conductance values.
 
@@
 
+Various new models that aim to simulate desiccation have a g~min~ term in it, but it is not yet well constrained by data.
+
 @pisekwinkler1953, early calculation of the length of time needed to desiccate leaves to some water content, given the minimum transpiration rate, and the water content of the leaves.
@@
 
-Discuss Martin-StPaul (2017); key role of minimum conductance in a model of desiccation dynamics
+Discuss Martin-StPaul (2017); key role of minimum conductance in a model of desiccation dynamics.
 
-Various new models that aim to simulate desiccation have a g~min~ term in it, but it is not yet well constrained by data.
+**Could we have a figure showing time to mortality as a function of gmin? I know Nico wants to use this in a separate paper, but it might draw some attention here.**
 
-# VI. Cuticular conductance
+# V. Cuticular conductance
 
@@
 Thick cuticles may also be useful because  exposure to wind has been shown to dislodge cuticular compounds by absrasion [@rogge1993], thereby dramatically increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991].
-Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986].
+Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986], most dramatically demonstrated in an ice storm, leading to much higher g~min~ [@boyce2003].
 Such exposure to the elements may be responsible for the typical gradual increase in g~min~ with leaf age in field-grown plants [@jordan2007].
@@
 
-# VII. Nighttime conductance
+# VI. Nighttime conductance
 
@@
 Strong evidence for endogenous regulation of g~dark~ was reported by @rescodedios2013, who showed that g~dark~ fluctuated throughout the night despite environmental conditions being held constant in whole-tree chambers.
+In addition, g~dark~ can show a clear response to vapour pressure deficit [@barbour2007] during night-time conditions.
 Clearly g~dark~ is an actively controlled process that cannot be adequately summarized by a single constant g~0~.
 
-# VIII. Environmental and ecological variation in minimum conductance
+# VII. Environmental and ecological variation in minimum conductance
 
@@
 
-# Conclusions
+# VIII. Conclusions
 

### 2017-09-22 manuscript.Rmd
Text as of draft zero - sent to coauthors

@@
 ```{r include=FALSE}
-includefigs <- F
+includefigs <- T
 library(png)
@@
 library(knitr)
+library(pander)
 
@@
 
-knitr::kable(tab[[1]], col.names="")
+knitr::kable(tab, col.names="")
 
@@
 During severe drought, or conditions where photosynthesis does not take place, it seems advantageous for plants to completely halt water loss.
-However, water loss continues through the cuticle and incompletely closed stomata, together determined by the leaf minimum conductance (gmin).
+However, water loss continues through the cuticle and incompletely closed stomata, together determined by the leaf minimum conductance (g~min~).
 In this review, we illustrate the role of g~min~ in models of leaf gas exchange, critically evaluate sources of variation in g~min~, and quantitatively compare various methods for its estimation.
@@
 \
+The absolute minimum attainable water loss rate through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf.
+Other measurements allow for the estimation of water loss through incompletely closed and broken stomata.
+Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time.
+This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants that are severely droughted, and includes all surfaces of the leaf.
+Since the seminal review by @kerstiens1996, no attempt has been made to synthesize the literature for values of g~min~, which is timely as models start to incorporate more detailed drought dynamics all the way to the point of mortality by desiccation (Xu, Blackman, Martin-StPaul, Sperry?).
+Instead, many models rely on inaccurately estimated 'minimum' conductance from cuvette data, by applying regression techniques.
+
+\
 Most current-generation process-based global vegetation models (GVMs) use a version of the Ball-Berry class of stomatal conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity.
@@
 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987) (they used relative humidity).
-It should be clear from Eq.
-1 that the Ball-Berry model is actually a model of water-use efficiency, since both g~s~ and A~n~ appear in the equation.
 Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been largely ignored.
-For example, a recent review of stomatal conductance models did not mention the minimum conductance [@damour2010].
+For example, a recent comprehensive review of stomatal conductance models did not mention g~0~ or any similar minimum conductance [@damour2010].
 The original description of Eq.
@@
 It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model.
-Discuss @barnard2013 briefly (in a positive way :)
+In practice, g~0~ is estimated from a regression of Eq.
+1 with leaf gas exchange data, but this may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.
 
 \
-The absolute minimum attainable water loss rate through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf.
-Other measurements allow for the estimation of water loss through incompletely closed and broken stomata.
-Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time.
-This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants that are severely droughted, and includes all surfaces of the leaf.
-Since the seminal review by @kerstiens1996, no attempt has been made to synthesize the literature for values of g~min~, which is timely as models start to incorporate more detailed drought dynamics all the way to the point of mortality by desiccation (Xu, Blackman, Martin-StPaul, Sperry?).
-Instead, many models rely on inaccurately estimated 'minimum' conductance from cuvette data, by applying regression techniques.
+(*expand*) We thus have several different definitions of what constitutes a 'minimum' conductance of leaves.
+In this review, we demonstrate that inclusion of a minimum conductance is necessary both for simulation of the response of water-use efficiency to various environmental drivers, as well as a component of the response of plants to severe drought.
+However, there is no guidance at the moment on how to estimate minimum conductance for use in models, how various definitions compare quantitatively, and whether it is feasible to assign species-specific values to the minimum conductance.
+We do not expect that g~0~, cuticular conductance (g~cut~) and minimum conductance from mass loss of detached leaves (we refer to minimum conductance measured with this method only as g~min~) to give similar values, in fact the contribution of stomata to each of these conductances will be very different.
 
@@
 (*expand*) Previews reviews have focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles (riederer, kerstiens).
+Other work has focused on the role of g~0~ in models of plant water use [e.g.
+@barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance.
 Thus there is a disconnect between dizzying level of detailed understanding of the plant cuticle [@riederer2008], and the very simple  assumptions made when applying ecosystem models.
 We aim to improve the connection somewhat, and point at obvious gaps in understanding.
-Throughout our review, we discuss both the practical implementation of a 'minimum' conductance g~0~ as implemented in e.g.
-Ball-Berry models of stomatal conductance (Eq.
-1) and similar, and a 'true' minimum conductance as estimated on detached leaves (g~min~).
 
@@
 
+\
 Measurements of (presumed) minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories.
@@
 \ 
-We quantitatively compared each of these different 'minimum' codnductance values by synthesizing existing data.
+We quantitatively compared each of these different 'minimum' conductance values by synthesizing existing data.
 To estimate g~0~ from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model.
@@
 
+```{r fig.width=4,echo=FALSE, eval=includefigs}
+plotfig("output/figure1.png")
+```
+
+**Figure 1**.
+Comparison of various estimates of the minimum conductance.
+Error bars are 95/% confidence intervals.
+See Table 1 for the various definitions.
+Different letters denote significant differences (alpha = 0.05).
+Note the logarithmic scale.
+
+**In case you are interested, this is what the same graph looks like with no log scale**
+
+```{r fig.width=4,echo=FALSE, eval=includefigs}
+plotfig("output/figure1_unlog.png")
+```
+
+**Figure 1 Version 2**. As Figure 1, but natural scale.
+
 # III. Problems with estimating minimum conductance from cuvette data
 
+\
 The standard approach in estimating g~0~ for use in Eq.
@@
 Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated by this method.
-The difficulty of the estimation can already be concluded from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g.
+The difficulty of the estimation can already be anticipated from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g.
 @leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species); presumably because unreliable estimates were obtained, b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though it is clearly non-sensical to suggest negative conductance values.
+@barnard2013 also mentioned the difficulty of fitting Eq.
+1 to estimate g~0~, but did not present specific details.
 
@@
 We thus conclude that g~0~ should really not be estimated from regression, although this is the most common method applied.
-Recently, @barnard2013 also mention the difficulty of fitting Eq.
-1 to estimate g~0~, but did not present specific details.
 
@@
 The consequence is that it is more difficult to interpret g~1~ and g~0~ independently, and that confidence intervals on each of the parameters should also not be interpreted independently (see @becker2007 for more discussion on the statistical aspects of comparing slopes and intercepts).
+(*expand*)
+
+```{r fig.height=4,echo=FALSE, eval=includefigs}
+plotfig("output/figure2.png")
+```
+
+**Figure 2**.
+Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al.
+(2011) applied to a leaf gas exchange dataset of Martin-StPaul (cite).
+This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al.
+(2015).
+The solid line is regression line, and the shaded area a 95% confidence interval for the mean.
+The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression.
+The dotted ellipse is a bivariate 95% confidence interval for slope and intercept.
+The symbols represent 1000 bootstrap samples of the coefficients.
 
@@
 Next we study the values obtained when fitting Eq.
-1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation of fitted g~0~ and m (applying the model of ) from @miner2016.
+1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation of fitted g~0~ and m (applying the model of ) from @miner2017.
 In both cases, estimates of g~0~ are inflated when the model fits poorly (Fig.
@@
 
+```{r fig.height=4,echo=FALSE, eval=includefigs}
+plotfig("output/figure3.png")
+```
+
+**Figure 3**.
+(*Might go in SuppInfo?*) Correlation between estimated g~0~ and the R^2^ of the regression.
+In the left panel, we fitted the linearized form of the Medlyn et al.
+(2011) model to each of the datasets in the Lin et al.
+(2015) database.
+Vertical lines are 95% confidence intervals.
+In the right panel, a very similar association between g~0~ and R^2^ is presented by the review of @miner2017, using the Ball-Berry model.
+In panel (c) coefficient of variation and R2.
+
 # IV. Use of minimum conductance in models
@@
 \
-The effects of photosynthetic photon flux density (PPFD) and leaf temperature on g~s~ are both assumed to enter via the dependence of A~n~ on these drivers.
-The g~0~ parameter not only sets a minimum value of g~s~ in the model, it also modifies the behaviour of Eq.
+In the Ball-Berry type framework of stomatal conductance models, the effects of photosynthetic photon flux density (PPFD) and leaf temperature on g~s~ are both assumed to enter via the dependence of A~n~ on these drivers.
+Here we show that the g~0~ parameter not only sets a minimum value of g~s~ in the model (when A~n~ is zero), it also modifies the behaviour of Eq.
 1 after it has been coupled to the Farquhar-von Caemmerer-Berry model of photosynthesis (the so-called 'coupled leaf gas exchange model').
@@
 In the application of the Ball-Berry models, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005].
-It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata [@kerstiens1996].
-Finally, measuring the low water loss rates of detached leaves mimicks drought conditions, yielding an estimate of whole-leaf minimum conductance (g~min~) [@pisekberger1938; @hygen1951].
-Photosynthesis also ceases at very high temperature (ref), but stomata do not appear to close in concert [@drake, @urban2017], thus we cannot use high temperature to estimate a reasonable g~0~.
+It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata (see Section XXX).
+
+\
+Photosynthesis also ceases at very high temperature (ref), but stomata do not appear to close in proportion to the decrease in photosynthesis, if at all [@drake, @urban2017], thus we cannot use high temperature to estimate a reasonable g~0~.
+Simulations of water use during heat waves depends highly on the value of g~0~ used [@kala2016] .
+It should be pointed out that the response of g~s~ to T~leaf~ is likely poorly modelled with a Ball-Berry model, since a number of studies have shown that g~s~ and A~n~ decouple at high T~leaf~, resulting in high rates of g~s~ when A~n~ is near zero [@urban2017] (Drake et al., unpublished).
 
@@
 \
-The g~0~ parameter also represents g~s~ when T~leaf~ is very high, because A~n~ goes to zero at high T~leaf~.
-As a result, simulations of vegetation water use during heat waves depends highly on the value of g~0~ used [@kala2016] .
-It should be pointed out that the response of g~s~ to T~leaf~ is likely poorly modelled with a Ball-Berry model, since a number of studies have shown that g~s~ and A~n~ decouple at high T~leaf~, resulting in high rates of g~s~ when A~n~ is near zero [@urban2017] (Drake et al., unpublished).
-
-\
-Using the coupled leaf gas exchange model, we show how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^).
+Using the coupled leaf gas exchange model, we show in Fig.
+4 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^).
 We also demonstrate the effect of g~0~ on modelling the T~leaf~ response.
@@
 1.
+(*expand*, cite some papers showing these are the responses we expect).
+
+```{r fig.height=6,echo=FALSE, eval=includefigs}
+plotfig("output/figure4.png")
+```
+
+**Figure 4**.
+Simulations  of intrinsic water-use efficiency (A~n~/g~s~), intercellular CO~2~ concentration (C~i~) and leaf transpiration rate (E~L~) with three values of g~0~.
+For the top two panels, all environmental drivers were held constant except the photosynthetic photon flux density (PPFD), demonstrating that the inclusion of g~0~ is necessary to simulate the decrease in A~n~/g~s~ at low PPFD.
+For the bottom two panels, VPD and T~air~ were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions.
+For all simulations it is assumed that T~leaf~ is equal to T~air~.
 
@@
 
+\
 Another reason to include a non-zero g~0~ term in models of stomatal conductance arises during severe drought.
@@
 Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000].
-
-Various new models that aim to simulate desiccation have a g~min~ term in it, but it is not yet well constrained by data.
+Various new models that aim to simulate desiccation have a g~min~ term in it (even when not explicitly defined in the main equations) [@xu2016] (sperry, others), but it is not yet well constrained by data.
 
@@
 
-Literature values of g~cut~, especially when measured on isolated cuticles (g~cut,isol~) were much lower than any other estimate of conductance, even compared to measurements based on MLD with stomatal side sealed (g~mld,seal~).
+\
+Literature values of g~cut~, especially when measured on isolated cuticles (g~cut,isol~) were much lower than any other estimate of conductance (Fig.
+1), in line with previous comparisons [@kerstiens1996].
 The conductance of stomata-free leaf cuticles represents the absolute minimum conductance that leaves can attain.
 This result should be viewed with caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions.
+Indeed, some more recent studies have shown a very good correspondence between g~cut~ and g~min~ (riederer, schuster).
 A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these processes have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b].
-However, a few key points should be reiterated as they are relevant to the current discussion.
+However, a few key points should be reiterated as they are relevant to the current discussion, in particular when we aim to interpret variation in literature values of g~min~.
 
@@
 \
-Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015].
-This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata.
+(*expand*) Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015].
+This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata (see Section XXX).
 Nonetheless it is an interesting observation that could have major consequences for modelling and interpreting water-use efficiency data during drought conditions.
@@
 
-A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~).
-However as our literature comparison showed, g~dark~ is substantially higher compared to g~s~ reached at low photosynthetic rates, and g~min~.
-Previous work on g~dark~ have confirmed that night-time conductance is generally much higher than g~min~, when measured on the same species [@howard2007; @caird2007].
-The idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.
+\
+A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~) [@barnard2013; @lombardozzi2017], simply because it is a condition where no photosynthesis takes place.
+However as our literature comparison showed (Fig.
+1), g~dark~ is substantially higher compared to any other measure of minimum conductance.
+Previous work on g~dark~ have confirmed that night-time conductance is generally much higher than g~min~, when measured on the same plants [@howard2007; @caird2007; @walden-coleman2013; @phillips2010].
+The simple idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.
 
 \
-First, ample observations suggect that g~dark~ is not a fixed rate but varies tremendously during the night.
+First, ample observations suggest that g~dark~ is not a fixed rate but varies tremendously during the night.
 @caird2007 (and references therein) described how for many species, g~dark~ is not stable throughout the night period.
@@
 Many previous reports use the term 'cuticular conductance' for g~min~, but this is inaccurate as some fraction of the water loss will occur through broken, leaky, or incompletely closed stomata.
-Our literature review was somewhat inconclusive in this regard, as conductance g~min~ measured on leaves where the stomatal side was sealed was similar (across all included studies) to g~min~ of whole leaves (Fig.
-3).
-More conclusive is evidence from individual studies where the stomatal component to g~min~ could be quantified.
+A few studies have directly quantified the stomatal component of g~min~.
 A detailed study of *Hedera helix*  concluded that 35\% of water loss occurred across the stomatal pores (though closed), and 65\% across the non-stomatal part of the cuticle, despite the fact that the stomata presumably covered only a small fraction of the leaf [@santrucek2004].
-For ten Sorghum genotypes, @muchow1989 reported a correlation between stomatal density and g~min~, suggesting leaky stomata contribute substantially to g~min~.
+For ten Sorghum genotypes, @muchow1989 reported a strong positive correlation between stomatal density and g~min~, suggesting leaky stomata contribute substantially to g~min~.
 For seven conifer species, @brodribb2014 reported that 50 - 94\% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky.
-However @santrucek2004 found that the cuticle (solid mass only) of the stomatal side in *Hedera helix* was many times more more permeable than the non-stomatal side, thus providing another potential explanation for this difference.
-Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles.
+However @santrucek2004 found that the cuticle of the stomatal side (but excluding stomata itself) in *Hedera helix* was many times more more permeable than the non-stomatal side, thus providing another potential explanation for this difference.
+Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles, and the contribution of leaky stomata, and potentially more permeable cuticle on the abaxial side, likely varies between species.
 
@@
 In the following sections, we review the quite substantial literature on g~min~ measured with the MLD technique, stretching back to the 1930s [@pisekberger1938], focussing on environmental and ecological determinants.
-We also further tease apart the literature compilation of g~min~ (shown just as a single average value in Fig.
+We also further tease apart the literature compilation of g~min~ (which was shown just as a single average value in Fig.
 4).
 
-Using the literature compilation g~min~ (see SectionX), we asked whether g~min~ shows clear variation due to, species growth form, taxonomic order, and average climate across the species' distribution.
-
-We found not so much! Grasses have a higher gmin. So do crops, and within crops there is a ton of genotypic variation.
-
-Very few studies have found meaningful correlations. Exception is @brodribb2014, showing a correlation between rainfall at species origin (of the driest quarter), and g~min~.
+We found it difficult to explain the variation in g~min~ between species.
+Grasses have a higher g~min~, and so do crops.
+Otherwise few differences were significant, with Pinales tending towards lower values of g~min~, and Myrtales (including many *Eucalyptus* species), tending towards higher values.
 
 Comparing all available data, we did not find significant relationships between rainfall, aridity, or temperature and g~min~ (see SuppInfo 2).
+There was a weak correlation between g~min~ and P~50~, but this was driven entirely by the difference in P~50~ between gymnosperms and angiosperms (not shown).
+
+Very few studies have found meaningful correlations between g~min~ and environmental factors or ecological categories.
+One notable exception is @brodribb2014, who showed a correlation between rainfall at species origin (of the driest quarter), and g~min~.
+
+```{r fig.height=5, fig.width=5, echo=FALSE, eval=includefigs}
+plotfig("output/figure5.png")
+```
+
+**Figure 5**.
+Analysis of a new literature compilation of g~min~ estimates, as measured with mass loss of detached leaves.
+Left panel, probability density function of all values (after averaging by species, n=200+), with a normal distribution curve (mean=, sd=, on the log scale).
+Middle panel, g~min~ averaged by phylogenetic order (including only the top ten orders in the database).
+Bars are 95/% confidence intervals.
+Numbers on top refer to the number of species, different letters denote significant differences (at alpha = 0.05, adjusted for multiple comparisons), and grey symbols are individual species.
+Right panel, data for crops only where each point is a genotype.
+Bars denote the range, illustrating the large variation in g~min~ between genotypes for a particular crop species.
+References for panels A and B can be found in the SuppInfo 1, references for panel C are XXX.
+(*have to add some labels*).
 
@@
 
-@bengtson1978 lower g~min~ in six oat varieties in response to drought, and increase in cuticular waxes, but no overall relationship between the two and very different responses between different oat genotypes.
+(*expand*) From the literature we found a general acclimatory response of g~min~ to drought stress, nearly always leading to a decrease in g~min~ (see Table 2 for a summary of the literature); the only exception was @vilagrosa2003 who found no response.
+A number of studies have attempted to attribute this decrease in g~min~ to a change in cuticular composition.
+@bengtson1978 reported lower g~min~ in six oat varieties in response to drought, and increase in cuticular waxes, but no overall relationship between the two and very different responses between different oat genotypes.
+@mackova2013 found that adding ABA (simulating drought stress) to *Lepidium sativum* increased the chain length of cuticular waxes (but not total amounts).
+@bi2017 Drought caused a change in cuticular wax production and chemical composition, but in a highly genotype-specific way.
+@kottapalli2009 higher epicuticular wax load in a drought-tolerant vs.
+susceptible peanut genotype.
 
-@mackova2013 adding ABA to Lepidium sativum increased the chain length of cuticular waxes (but not total amounts).
-
-@bi2017 Drought caused a change in cuticular wax production and chemical composition, but in a highly genotype-specific way.
-
-@kottapalli2009 higher epicuticular wax load in a drought-tolerant vs. susceptible peanut genotype
-
-Ubiquitous response of g~min~ after drought stress, always a decrease. Only study we found that did not find a difference was @vilagrosa2003. All other studies summarized in TableXX.
-
-The second case study was on *Hakea* plants. Details here (measurement T?).
+\
+We also demonstrate this effect via a case study on *Hakea* plants.
+**Measurement details here, or in caption**.
+The results showed a significant decrease in g~min~ with long-term exposure to low soil water availability (average effect X, p-value X).
+(*expand*).
+We also found evidence for a link between g~min~ and leaf shape, as the studied *Hakea* species had either needle-like or broad leaves (see Fig.
+6).
 
@@
 knitr::kable(dat, 
+             caption="**Table 2.** Summary of studies demonstrating a decrease in g~min~ after an acclimation period to low soil water availability",
              col.names= c("Study","Species","Experiment","Reduction in g~min~"))
@@
 
+```{r fig.height=4,echo=FALSE, eval=includefigs}
+plotfig("output/figure6.png")
+```
+
+**Figure 6**.
+Minimum conductance measured from leaf dry-down curves on a variety *Hakea* species, a genus native to Australia.
+Plants were grown in containers in a grow house, and supplied with ample water or underwent long-term mild drought stress.
+Bars are labelled by leaf form : broadleaf (B) or needle-like (N).
+For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`).
+
 ## 4. Acclimation and instantaneous response to temperature
 
-@schuster2016a thermal tolerance and gmin relationship across species.
-@schuster2016a correlation temperature, gcut AND gmin for Nerium oleander (page 59).
-T range much wider than Blackman though...
-But Schuster gmin not different from gcut at most temperatures, which is a bit strange too.
+Perhaps even more significant is the response of g~min~ (and g~cut~) to temperature: it shows both an instantaneous response, as well as an acclimatory response.
+@riederer2001 and @schuster2016b demonstrated a steep non-linear response of g~cut~ to temperature, with a flat response below XC, and a steep response thereafter (*expand*).
 
-Also see @paje1988 for more complex response to temperature
+@schuster2016a reported on a correlation between thermal tolerance and g~min~ relationship across nine species.
 
-@riederer2001 @schuster2016b correlation between temperature and gcut (non-stomatal cuticles)
+Also see @paje1988 for more complex response to temperature (figure out what he really showed).
 
-We present two case studies demonstrating the acclimation of g~min~ to environmental drivers.
+To further study the acclimatory and instantaneous response to temperature, we carried out a case study on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature.
+Our results showed a large acclimation effect but no evidence for a direct T effect (over the narrow range in T compared to Schreiber) (Fig.
+X).
+(*expand stats and description of results*).
 
-Lower g~min~ result is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time.
-This is not due to a direct effect of temperature on in cuticular conductance, which increases with temperature (Schreiber 2001).
+Lower g~min~ in the trees acclimated to high temperature is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time.
+Note that this response is not due to a direct effect of temperature on in cuticular conductance, which increases with temperature (Schreiber 2001).
 
-A case study on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature.
-Our results showed a large  acclimation effect but no evidence for a direct T effect (over the narrow range in T compared to Schreiber).
+```{r fig.height=4,echo=FALSE, eval=includefigs}
+plotfig("output/figure7.png")
+```
+
+**Figure 7**.
+Minimum conductance measured from leaf dry-down curves on *Eucalyptus parramattensis* grown in whole-tree chambers.
+Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C).
+Measurements of g~min~ were carried out at five different temperatures on replicate leaves.
+There is no effect of measurement T on g~min~, but leaves of trees grown in elevated temperature showed consistently lower g~min~.
 
@@
 
-@jordan2007 large increase gmin in the herb Agastachys odorata with leaf age
-
---> more in database
+@jordan2007 large increase gmin in the herb Agastachys odorata with leaf age. (*Several studies in database that have leaf age effects; summarize?*).
 
@@
 This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss (Hypothesis 2).
+(*could go to temperature section, as T and RH are correlated in field conditions*).
 
@@
 
+\
+It is probably unavoidable to improve the Ball-Berry model with an expression like,
+
+$$ g_s = g_{min} + (g_0 - g_{min}) + g_1 \frac{A_n}{C_a}f(D)$$
+
+where g~min~ is estimated via MLD, representing a true whole-leaf minimum that is reached under drought conditions, and g~0~ is the minimum reached in darkness for well-watered plants (which itself should be a function of soil water availability, e.g.@misson2004).
+
 Our review of g~min~ of detached leaves shows remarkable plasticity to growing conditions.
 The lack of clear species trends, when comparing growing conditions, climate of origin, or growth habit, can perhaps be explained by the very large acclimatory potential of g~min~ to growing conditions.
-Future studies should thus compare g~min~ on many species grown in the same conditions to better understand the adaptive value of the minimum water loss rate of leaves.
+We conclude as a result that g~min~ is not a "trait" in the sense that it can be meaningfully summarized for a given species, but has to be linked to growing conditions.
+Future studies should compare g~min~ on many species grown in the same conditions to better understand the adaptive value of the minimum water loss rate of leaves.
 
-# Figure captions
-
-```{r fig.height=6,echo=FALSE, eval=includefigs}
-plotfig("output/figure1.png")
-```
-
-**Figure 1**.
-Simulations  of intrinsic water-use efficiency (A~n~/g~s~), intercellular CO~2~ concentration (C~i~) and leaf transpiration rate (E~L~) with three values of g~0~.
-For the top two panels, all environmental drivers were held constant except the photosynthetic photon flux density (PPFD), demonstrating that the inclusion of g~0~ is necessary to simulate the decrease in A~n~/g~s~ at low PPFD.
-For the bottom two panels, VPD and T~air~ were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions.
-For all simulations it is assumed that T~leaf~ is equal to T~air~.
-
-```{r fig.height=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure2.png")
-```
-
-**Figure 2**.
-Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al.
-(2011).
-This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al.
-(2015).
-The solid line is regression line, and the shaded area a 95% confidence interval for the mean.
-The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression.
-The dotted ellipse is a bivariate 95% confidence interval for slope and intercept.
-The symbols represent 1000 bootstrap samples of the coefficients.
-
-```{r fig.height=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure3.png")
-```
-
-**Figure 3**.
-Correlation between estimated g~0~ and the R^2^ of the regression.
-In the left panel, we fitted the linearized form of the Medlyn et al.
-(2011) model to each of the datasets in the Lin et al.
-(2015) database.
-Vertical lines are 95% confidence intervals.
-In the right panel, a very similar association between g~0~ and R^2^ is presented by the review of @miner2017, using the Ball-Berry model.
-In panel (c) coefficient of variation and R2.
-
-```{r fig.width=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure4.png")
-```
-
-**Figure 4**. Comparison of various estimates of the minimum conductance. Error bars are standard errors. See Table 1 for definitions of minimum conductance, and text for the data sources.
-
-```{r fig.height=5, fig.width=5, echo=FALSE, eval=includefigs}
-plotfig("output/figure5.png")
-```
-
-**Figure 5**.
-Analysis of g~min~ from Figure 4 by plant growth form and phylogenetic group.
-Different letters denote significant differences at $alpha$ = 0.05.
-The groups 'Angiosperm' and 'Gymnosperm' include trees only.
-
-```{r fig.height=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure6.png")
-```
-
-**Figure 6**.
-Minimum conductance measured from leaf dry-down curves on *Eucalyptus parramattensis* grown in whole-tree chambers.
-Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C).
-Measurements of g~min~ were carried out at five different temperatures on replicate leaves.
-There is no effect of measurement T on g~min~, but leaves of trees grown in elevated temperature showed consistently lower g~min~.
-
-```{r fig.height=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure7.png")
-```
-
-**Figure 7**.
-Minimum conductance measured from leaf dry-down curves on a variety *Hakea* species, a genus native to Australia.
-Plants were grown in containers in a grow house, and supplied with ample water or underwent long-term mild drought stress.
-Bars are labelled by leaf form : broadleaf (B) or needle-like (N).
-For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`).
-
-# Appendix or SuppInfo
+# Appendix or SuppInfo or Delete
 

### 2017-09-22 R/figures.R

@@ -2,27 +2,18 @@
 
 figure_gmin_review <- function(gdfr){
   
-  par(mar=c(3,5,1,0.5), cex.lab=1.2)
-  fit <- lm(log10(gmin) ~ method-1, data=gdfr)
-  cis <- confint(fit)
-  quans <- sapply(split(gdfr, gdfr$method), 
-                  function(x)quantile(log10(x$gmin), probs=c(0.05, 0.95)))
-  gmins <- with(gdfr, tapply(log10(gmin), method, mean, na.rm=TRUE))
-  
-  plot(1:5, gmins, pch=19, cex=1.2, axes=FALSE, xlab="",
-       panel.first=segments(x0=1:5, x1=1:5, y0=quans[1,], y1=quans[2,], col="grey"),
-       ylim=c(0,2),
-       xlim=c(0.5,5.5),
-       ylab=expression(Conductance~~(mmol~m^-2~s^-1)))
-  axis(1, at=1:5, labels=c(expression(g["cut,isol"]), 
-                           expression(g["min,seal"]), 
-                           expression(g[min]), 
-                           expression(g[dark]),
-                           expression(g[0])))
-  magaxis(side=2, unlog=2)
-  box()
-  arrows(x0=1:5, x1=1:5, y0=cis[,1], y1=cis[,2], angle=90, length=0.1, code=3)
-  
+  par(mar=c(4,4,1.5,1), mgp=c(2.5,0.5,0), tcl=-0.2, 
+      las=3, yaxs="i",
+      cex.lab=1.2)
+  plotCIlog(gmin, method, gdfr, 
+            ylim=c(-1,2.2),
+            label_las=1,
+            ylab=expression(Conductance~(mmol~m^-2~s^-1)),
+            labels=c(expression(g["cut,isol"]),
+                     expression(g["min"]),
+                     expression(g["dark"]),
+                     expression(g["0"]),
+                     expression(g["s,low A"])))
   
 }
 
@@ -36,7 +27,7 @@ figure_gmin_review_2 <- function(gdfr){
           ylim=c(0,80),
           ylab=expression(Conductance~(mmol~m^-2~s^-1)),
           labels=c(expression(g["cut,isol"]),
-                   expression(g["cut,seal"]),
+                   #expression(g["cut,seal"]),
                    expression(g["min"]),
                    expression(g["dark"]),
[…]

### 2017-09-22 R/functions.R

@@ -131,13 +131,12 @@ plotCI2 <- function(yvar, group, data, transform_log10=FALSE,
   if(is.null(ylim))ylim <- c(0, max(ci)+0.1*max(ci))
   g <- glht(fit, linfct=mcp(G = "Tukey"))
   lets <- cld(g)$mcletters$Letters
-  
   n <- length(mn)
   
   plot(1:n, mn, pch=19, cex=1.1, ylim=ylim, axes=FALSE,
        xlab="",
        panel.first={
-         if(add_data)with(data, points(jitter(as.numeric(G), factor=jit), gmin, pch=16, col=datacol, cex=datacex))
+         if(add_data)with(data, points(jitter(as.numeric(G), factor=jit), data$Y, pch=16, col=datacol, cex=datacex))
        },
        xlim=c(0.7, n+0.3), ylab=ylab, ...)
   axis(1, at=1:n, labels=labels, las=label_las)
@@ -146,6 +145,55 @@ plotCI2 <- function(yvar, group, data, transform_log10=FALSE,
   text(x=1:n, y=ci[,2], lets, pos=3, cex=0.8)
   par(las=1)
   axis(2)
+  box()
+  
+  tb <- unname(table(data$G))
+  mtext(side=3, at=1:n, line=0.4, text=tb, las=1, cex=0.7)
+  mtext(side=3, at=0.25, line=0.4, text="n = ", las=1, cex=0.7)
+  
+}
+
+
+plotCIlog <- function(yvar, group, data, 
+                    ylab=NULL, ylim=NULL, labels=NULL, label_las=3, add_data=TRUE, 
+                    jit=0.6, datacex=0.4, datacol="lightgrey", 
+                    ...){
+  
+  if(is.null(ylab))ylab <- substitute(yvar)
+  
+  data <- as.data.frame(data)
+  
+  data$Y <- eval(substitute(yvar), data)
+  data$G <- as.factor(eval(substitute(group), data))
+  
+  datl <- split(data, data$G)
+  
+  if(is.null(labels)){
+    labels <- levels(data$G)
+  }
+  
+  fit <- lm(log10(Y) ~ G - 1, data=data)
+  ci <- confint(fit)
+  mn <- coef(fit)
+
+  if(is.null(ylim))ylim <- c(0, max(ci)+0.1*max(ci))
+  g <- glht(fit, linfct=mcp(G = "Tukey"))
[…]

### 2017-09-22 R/load.R

@@ -189,7 +189,8 @@ kerst_simple <- group_by(kerst, species, method) %>%
   dplyr::select(gmin, method)
 
 gdfr <- bind_rows(gmindat_simple, kerst_simple, lombar, g0s, minags) %>%
-  mutate(method = factor(method, levels=c("gcut_isol","gcut_seal","gmin","gnight", "g0","gslowA")))
+  filter(method != "gcut_seal") %>%  # I keep changing my mind!!
+  mutate(method = factor(method, levels=c("gcut_isol","gmin","gnight", "g0","gslowA")))
 
 
 

### 2017-09-22 R/make_figures.R

@@ -1,23 +1,23 @@
 
 source("R/load.R")
 
-to.pdf(figure_sim(), "output/figure1.pdf", height=6, width=6)
+
+to.pdf(figure_gmin_review(gdfr), "output/figure1.pdf", height=5, width=5)
+to.pdf(figure_gmin_review_2(gdfr), "output/figure1_unlog.pdf", height=5, width=5)
 
 to.pdf(figure_g0g1_cor(lin2015, "Nicolas Martin-StPaul_Quercus ilex_StPaul_Puechabon"), 
        "output/figure2.pdf", height=4, width=8)
 
 to.pdf(figure_R2g0(lin2015coef, miner), "output/figure3.pdf", height=4, width=9)
 
-to.pdf(figure_gmin_review_2(gdfr), "output/figure4.pdf", height=5, width=5)
+to.pdf(figure_sim(), "output/figure4.pdf", height=6, width=6)
 
 to.pdf(gmin_3panel(gmindat, cropgmin), "output/figure5.pdf", height=3, width=8)
 
-to.pdf(figure_wtc4_gmin_2(wtc4gmin), "output/figure6.pdf", height=4, width=3.5)
+to.pdf(figure_hakea_gmin(lopw), "output/figure6.pdf", height=5, width=6)
 
-to.pdf(figure_hakea_gmin(lopw), "output/figure7.pdf", height=5, width=6)
-
-
+to.pdf(figure_wtc4_gmin_2(wtc4gmin), "output/figure7.pdf", height=4, width=3.5)
 
 
-#convert_all_pdf("output")
+convert_all_pdf("output")
 

### 2017-12-06 R/load.R
analysis of gmin vs hydraulic traits

@@ -1,12 +1,16 @@
 
+
+
+
 if(!require("pacman"))install.packages("pacman")
 pacman::p_load(Hmisc, car, dplyr, tidyr, nlme, nlshelper, 
                forcats, tibble, magicaxis, 
                plantecophys, readxl, multcomp,
-               reporttools, Taxonstand, taxize, speciesmap,
+               reporttools, Taxonstand, taxize, 
                doBy, stringi)
 
-
+#devtools::install_github("remkoduursma/speciesmap")
+library(speciesmap)
 
 if(!dir.exists("download"))dir.create("download")
 if(!dir.exists("output"))dir.create("output")
@@ -69,8 +73,9 @@ wtc4gmin <- read.csv("data/wtc4_gmin_detached.csv") %>%
 wtc4gdark <- read.csv("data/wtc4_gnight.csv")
 
 # Rosana Lopez Hakea data
-lopez <- read.csv("data/lopez_gmin_hakea.csv") %>% group_by(species, treatment) %>%
-  summarize(gmin = mean(gmin)) %>%
+lopez <- read.csv("data/lopez_gmin_hakea.csv") %>% 
+  group_by(species, treatment) %>%
+  dplyr::summarize(gmin = mean(gmin)) %>%
   mutate(gmin = 2*gmin)  # to convert - badly - to projected area
 
 # some meta data for this dataset (leaf form, MAP)
@@ -84,7 +89,7 @@ lopw <- dplyr::select(lopmet, species, leaf.form) %>% distinct() %>%
 
 # Version 2; by population merged with lopmet (because MAP varies by population)
 lopez2 <- read.csv("data/lopez_gmin_hakea.csv") %>% group_by(species, population, treatment) %>%
-  summarize(gmin = mean(gmin)) %>% 
+  dplyr::summarize(gmin = mean(gmin)) %>% 
   mutate(gmin = 2 * gmin) %>%
   inner_join(lopmet, by=c("species","population"))
 
@@ -110,7 +115,7 @@ g0s <- data.frame(gmin=c(lin2015coef$g0, miner$g0),
 
 # From Lin2015, data where A < threshold.
 minags <- group_by(lin2015, fitgroup) %>%
-  summarize(
+  dplyr::summarize(
     Amin = min(Photo, na.rm=TRUE),
     gmin = 1000 * min(Cond, na.rm=TRUE),
     Qrange = max(PARin) - min(PARin),
[…]

### 2017-12-06 R_other/gmindat_Vs_martin.R

@@ -1,0 +1,57 @@
+# useless idea; only 33 species overlap   
+
+
+#https://zenodo.org/record/854700#.Whyx7kqWbIV
+
+martin <- read_excel("data/DataBase.xlsx", sheet=4) %>%
+  rename(species = Species.binomial)
+names(martin)[6] <- "Pgs90"
+
+martin_sub <- dplyr::select(martin, species, P50, Pgs90)
+
+gmindat2 <- inner_join(gmindat, martin_sub, by="species")
+
+with(gmindat2, plot(Pgs90, gmin))
+
+
+
+martin <- read.csv("c:/repos/nlshelper/R_other/martin.csv")
+library(nlshelper)
+martin$P50m <- -martin$P50
+martin$Pgs90m <- -martin$Pgs90
+fit1 <- nls(Pgs90m ~ ymax*(1-exp(-k*P50m)), 
+            start=list(ymax=2, k=1), data=martin)
+plot_nls(fit1)
+
+
+
+
+
+
+
+martin <- read_excel("data/DataBase.xlsx", sheet=4) %>%
+  rename(species = Species.binomial)
+names(martin)[6] <- "Pgs90"
+
+martin <- mutate(martin, 
+                 P50m = -P50,
+                 Pgs90m = -Pgs90) %>%
+  filter(!is.na(Pgs90))
+
+
+fit2 <- nls(Pgs90m ~ SSasympOff(P50m, Asym, lrc, c0), data=martin)
+plot_nls(fit2, xlim=c(0,20), ylim=c(0,5))
+
+
+library(quantreg)
+
+fit_nl1 <- lapply(c(0.1,0.5,0.9),
+                  function(x)nlrq(Pgs90m ~ SSasympOff(P50m, Asym, lrc, c0), 
+                       data=martin,
+                       tau=x))
+                
+
+plot_nls(fit_nl1[[2]], xlim=c(0,20), ylim=c(0,5))
+plot_nls(fit_nl1[[1]], add=T, lty=5)
+plot_nls(fit_nl1[[3]], add=T, lty=5)
+

### 2018-01-08 manuscript.Rmd
analysis of gmin vs hydraulic traits
co-author comments and edits incorporated; main script added; various formatting and other changes

@@
 \
-Remko A. Duursma^1^, Chris Blackman^1^, Rosana Lopéz^1^, Nicolas Martin-StPaul^2^, Belinda Medlyn^1^
+Remko A. Duursma^1^, Chris J. Blackman^1^, Rosana Lopéz^1^, Nicolas K. Martin-StPaul^2^, Hervé Cochard^3^, Belinda E. Medlyn^1^
 
@@
 ^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia
-^2^ INRA
+^2^ URFM, INRA, 84000, Avignon, France
+^3^ Université Clermont-Auvergne, INRA, PIAF, 63000, Clermont-Ferrand, France
 
@@
 ```{r include=FALSE}
-includefigs <- T
+includefigs <- FALSE
 library(png)
@@
 ```
+
+# Notes
+
+@fernandez2017 : more discussion of cuticular composition, increase of gmin with altitude ('unripe' cuticles due to short growing season), foliar uptake of water (and adaptation of the cuticle to facilitate that)
+
+Relationship between gmin and yield (to mention in the text???):
+Clarke et al. 1989 . Genotypes with low gmin yielded 40-150% more in 4 out of 5 driest enviroments in 9 years of study.
+Clarke and McCaig (1982 ) low order but significant positive correlation between yield and water retention (low gmin) in genotypes of Triticum
+
+Perhaps include here changes with ontogeny is species with different juvenile and adult leaves?
+Climent et al.
+2006  and Pardos et al.
+2009 - juvenile needles in some pines have higher rate of water loss than adult needles of the same individuals.
+Darrow et al.
+2002  found also that in some New Zealand heteroblastic sp juveniles had a higher gmin than adults but other sp no significant differences.
+
+@hanson2016 large effect of small gmin on estimates of mesophyll conductance, Ci. This is important but not the focus of this review.
 
@@
 In this paper, we refer to Eq.
-1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987) (they used relative humidity).
+1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987).
 Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been largely ignored.
@@
 However, there is no guidance at the moment on how to estimate minimum conductance for use in models, how various definitions compare quantitatively, and whether it is feasible to assign species-specific values to the minimum conductance.
-We do not expect that g~0~, cuticular conductance (g~cut~) and minimum conductance from mass loss of detached leaves (we refer to minimum conductance measured with this method only as g~min~) to give similar values, in fact the contribution of stomata to each of these conductances will be very different.
+We do not expect that g~0~, cuticular conductance (g~cut~) and minimum conductance from mass loss of detached leaves (we refer to minimum conductance measured with this method only as g~min~) to give similar values, in fact the contribution of stomata to each of these conductances will be very different [@brodribb2014].
 
@@
 @barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance.
-Thus there is a disconnect between dizzying level of detailed understanding of the plant cuticle [@riederer2008], and the very simple  assumptions made when applying ecosystem models.
-We aim to improve the connection somewhat, and point at obvious gaps in understanding.
+Thus there is a disconnect between dizzying level of detailed understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models We aim to improve the connection somewhat, and point at obvious gaps in understanding.
 
@@
 | Variable      |  Definition                                                                                       |
-| ---------- | ----------- |
+| ------------- | ------------------------------------------------------------------------------------------------- |
 | g~cut~        | Conductance of an isolated non-stomatal cuticle.                                                  |
@@
 \
-Measurements of (presumed) minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories.
+Measurements of (presumed) minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories (see Table 1).
 The absolute minimum conductance attained by leaves is that through the cuticle only.
-Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001] (g~cut,isol~), or via gas exchange techniques, or mass loss of detached leaves vby sealing the side with stomata (g~cut,seal~).
-The method we here refer to as minimum conductance is that of mass loss of detached leaves (MLD).
+Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001], via gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata (g~cut,seal~) [@kerstiens1996].
+The method we here refer to as minimum conductance is that of mass loss of intact detached leaves (MLD).
 In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out.
@@
 This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress.
-The g~0~ parameter, which is a *modelled* minimum conductance reached when photosynthesis rate is zero, is typically estimated by applying regression to Eq.
+Finally, the g~0~ parameter, which is a *modelled* minimum conductance reached when photosynthesis rate is zero, is typically estimated by applying regression to Eq.
 1 (of which various forms exist).
@@
 It is no new observation that stomata do not close in the dark, but do close on detached leaves, as already demonstrated by @hygen1954 in Scots pine, and described by @koerner1994.
-More recently, @walden-coleman2013 showed much higher g~dark~ compared to g~min~ in Cotton.
+More recently, @walden-coleman2013 showed much higher g~dark~ compared to g~min~ in cotton.
+
 However, @cavender-bares2007 gdark similar to gmin in oaks.
-Actual measured stomatal conductance during conditions when photosynthesis was low (g~s,low A~) was also much lower than fitted g~0~ or g~dark~.
+
+Actual measured stomatal conductance during low photosynthesis (g~s,low A~) was also much lower than fitted g~0~ or g~dark~.
 In other words, actual daytime conductance is much lower than the presumed minimum estimated from fitted g~0~.
@@
 Comparison of various estimates of the minimum conductance.
-Error bars are 95/% confidence intervals.
-See Table 1 for the various definitions.
-Different letters denote significant differences (alpha = 0.05).
-Note the logarithmic scale.
-
-**In case you are interested, this is what the same graph looks like with no log scale**
-
-```{r fig.width=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure1_unlog.png")
-```
-
-**Figure 1 Version 2**. As Figure 1, but natural scale.
+Error bars are 95\% confidence intervals.
+Different letters denote significant differences (at $\alpha$ = 0.05).
+See Table 1 for definitions of minimum conductance, and text for the data sources.
 
 # III. Problems with estimating minimum conductance from cuvette data
+
+- Measurement error, small fluxes
 
@@
 \
-It is also telling that there are few reports on the plasticity (intra- or interspecific) of g~0~, perhaps because it is so poorly estimated.
+It is also telling that there are few reports on the intra-specific plasticity, or inter-specific variation of g~0~, perhaps because it is so poorly estimated.
 Exceptions are @duarte2016, who found a lower g~0~ in a heat wave treatment in Douglas-fir (although g~1~ was unaffected), but in their case g~0~ was very accurately estimated by careful multi-point light response curves.
@@
 Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al.
-(2011) applied to a leaf gas exchange dataset of Martin-StPaul (cite).
+(2011) applied to a leaf gas exchange dataset of @martinstpaul2012.
 This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al.
@@
 Next we study the values obtained when fitting Eq.
-1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation of fitted g~0~ and m (applying the model of ) from @miner2017.
+1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation of fitted g~0~ and *m* from @miner2017 (who applied the original model of Ball-Berry).
 In both cases, estimates of g~0~ are inflated when the model fits poorly (Fig.
@@
 1, usually because environmental conditions varied little.
-We confirm this by showing that the standard error (SE) of g~0~ increases when the coefficient of variation the ball-berry coefficient (i.e.
-right-hand side of Eq.
-1) is lower (Fig.
+We confirm this by showing that the standard error (SE) of g~0~ increases when the coefficient of variation of the right-hand side of Eq.
+1 is lower (Fig.
 3c).
@@
 **Figure 3**.
-(*Might go in SuppInfo?*) Correlation between estimated g~0~ and the R^2^ of the regression.
+Correlation between estimated g~0~ and the R^2^ of the regression.
 In the left panel, we fitted the linearized form of the Medlyn et al.
@@
 In the right panel, a very similar association between g~0~ and R^2^ is presented by the review of @miner2017, using the Ball-Berry model.
-In panel (c) coefficient of variation and R2.
+In panel (c) coefficient of variation and R^2^.
 
@@
 
-and using the diffusion constraint $A~n~ = g~s~ / 1.6 \cdot (C~a~ - C~i~)$ we get an expression for C~i~ / C~a~,
+and using the diffusion constraint $A_n = g_s / 1.6 (C_a - C_i)$ we get an expression for C~i~ / C~a~,
 
@@
 1.
-(*expand*, cite some papers showing these are the responses we expect).
 
@@
 Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000].
-Various new models that aim to simulate desiccation have a g~min~ term in it (even when not explicitly defined in the main equations) [@xu2016] (sperry, others), but it is not yet well constrained by data.
+This is particularly important given the limited variation in stomatal response to drought among plants @martinstpaul2017.
 
@@
 
-Refined by @blackman2016, incorporating more well-known hydraulics traits into the framework.
+More recently, @gleason2014 and @blackman2016 proposed that embolism resistance along with whole plant capacitance and minimum transpiration rates all contribute to define the time to desiccation.
 
-Discuss Martin-StPaul (2017); key role of minimum conductance in a model of desiccation dynamics.
+@burghardt2006 reported on a correlation between g~min~ and the survival time of leaves.
 
-**Could we have a figure showing time to mortality as a function of gmin? I know Nico wants to use this in a separate paper, but it might draw some attention here.**
+Finally, Martin-StPaul et al 2017 demonstrated in a whole plant model of hydraulic failure that gmin was probably one of the key parameters to explain water potential drops below cavitation thresholds during drought given the relatively conserved range of stomatal closure to drought.
+Relying on the Sureau model presented in @martinstpaul2017 we illustrate the critical role of gmin in defining desiccation time.
 
@@
 This result should be viewed with caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions.
+Boyer et al.
+1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases (Boyer et al., 1997)
+
+\
 Indeed, some more recent studies have shown a very good correspondence between g~cut~ and g~min~ (riederer, schuster).
@@
 The thickness of the cuticle per se may also have to do with mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness).
-Thick cuticles may also be useful because  exposure to wind has been shown to dislodge cuticular compounds by absrasion [@rogge1993], thereby dramatically increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991].
+Thick cuticles may also be useful because  exposure to wind has been shown to dislodge cuticular compounds by abrasion [@rogge1993], thereby increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991].
 Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986], most dramatically demonstrated in an ice storm, leading to much higher g~min~ [@boyce2003].
-Such exposure to the elements may be responsible for the typical gradual increase in g~min~ with leaf age in field-grown plants [@jordan2007].
+Such exposure to the elements was also held responsible for thicker cuticles in coastal, wind-exposed sites compared to inland sites [@blackman2005].
 
@@
 Nonetheless it is an interesting observation that could have major consequences for modelling and interpreting water-use efficiency data during drought conditions.
+Also @hanson2016 showing the effect of this on calculation of Ci, and calculation of gmeso.
 
@@
 Bars denote the range, illustrating the large variation in g~min~ between genotypes for a particular crop species.
-References for panels A and B can be found in the SuppInfo 1, references for panel C are XXX.
-(*have to add some labels*).
 
@@
 @bengtson1978 reported lower g~min~ in six oat varieties in response to drought, and increase in cuticular waxes, but no overall relationship between the two and very different responses between different oat genotypes.
+However, @premachandra1992 found that epicuticular wax load increased on leaves of non-irrigated Sorghum cultivars and was positively correlated with cuticular conductance and cell membrane stability.
 @mackova2013 found that adding ABA (simulating drought stress) to *Lepidium sativum* increased the chain length of cuticular waxes (but not total amounts).
@@
 Also see @paje1988 for more complex response to temperature (figure out what he really showed).
-
-To further study the acclimatory and instantaneous response to temperature, we carried out a case study on *Eucalyptus parramattensis* growing in twelve whole-tree chambers (see Drake for methods), either tracking ambient conditions or exposed to a constant +3 C air temperature.
-Our results showed a large acclimation effect but no evidence for a direct T effect (over the narrow range in T compared to Schreiber) (Fig.
-X).
-(*expand stats and description of results*).
 
@@
 
-# Appendix or SuppInfo or Delete
+- Also: there is little information on water loss rate of non-leaf surfaces [@oren2001], should be focus of more research
 
-\
-To illustrate the effect of g~0~ on the modelled response of g~s~ to PPFD and T~leaf~, it is necessary to embed the stomatal conductance model in a coupled leaf gas exchange model, which is formed by three equations; Eq.
-1 giving the dependence of g~s~ on A~n~ and some environmental drivers, the diffusion constraint,
+# SuppInfo
 
-\
-(2)
-$$ A_n = \frac{g_s}{1.6}(C_a - C_i) $$
+E. parramattensis:
+We determined g~min~ from detached leaves collected before sunrise from each of 6 *Eucalyptus parramattensis* trees grown under ambient and ambient+3C temperature conditions, respectively.
+Leaves were scanned for leaf area determination before being weighed using a precision balance and attached to a ‘clothesline’ suspended inside controlled chambers set at different temperatures (17.5, 20, 22.5, 25, 27.5°C) and common low light (~200 umol m^-2^ s^-1^) and a relative humidity of 60%.
+Inbuilt fans in each chamber blew air gently across the leaf surfaces.
+Leaves were allowed to slowly dehydrate over a period of 7-8 hours during which time leaf mass was measured roughly every hour.
+For each leaf, transpiration was calculated from the slope of the linear part of the leaf mass vs time relationship, and converted to g~min~ following @sinclair1986.
 
-\
-where g~s~ is the conductance to H~2~O (the factor 1.6 converts to conductance to CO~2~). And finally the Farquhar-von Caemmerer-Berry photosynthesis model that gives A~n~ as a function of C~i~,
-
-\
-(3)
-$$ A_n = \frac{a \cdot (C_i - \Gamma^{*})}{C_i + b} - R_d$$
-
-\
-where a = V~cmax~ or J/4 (where the electron transport rate J approaches Jmax at high light) depending on the limitation to photosynthesis, and b also varies.
-The dependence on temperature enters through V~cmax~, J~max~, R~d~, $\Gamma^{*}$, and other parameters [@medlyn2002].
-For all simulations with this model, we used the implementation in the *plantecophys* R package [@duursma2015].
+Hakeas:
+After 8 months of water stress treatment, the epidermal conductance of detached leaves of three plants per species and treatment was estimated by using the rate of water loss.
+Leaves were collected in the early morning and subsequently recut under distilled water, rehydrated in darkness for 6 h at 4 °C and scanned to determine leaf area.
+Leaves were then moved to a controlled environment chamber set at 25 °C, 50% relative humidity and low light (~200 umol) and weighed using a precision balance.
+They were attached to a ‘clothesline’ and allowed to dehydrated slowly.
+Leaves were weighted every 15 min for 3 h and then every hour for 3h more.
+For each sample, the time of stomatal closure was determined as the inflection point of the relationship between leaf mass vs time, and gmin from the slope of the linear part of the curve after stomatal closure, normalized by 2-sided leaf area and adjusted for temperature and humidity within the chamber.
 

### 2018-01-08 R/figures.R

@@ -97,7 +97,7 @@ figure_g0g1_cor <- function(lin2015, group, legend=FALSE){
 
 figure_R2g0 <- function(lin2015coef, miner){
   
-  par(mfrow=c(1,3), mar=c(5,5,1,1), cex.axis=0.9)
+  par(mfrow=c(1,3), mar=c(5,5,1,1), cex.axis=0.9, cex.lab=1.2)
   l <- loess(g0 ~ R2, data=lin2015coef, span=0.8)
   with(lin2015coef, {
     plot(R2, g0, pch=16,
@@ -116,6 +116,7 @@ figure_R2g0 <- function(lin2015coef, miner){
   
   
   legend("topright", "Lin et al. 2015", bty='n', cex=0.8)
+  plotlabel("(a)", "topleft")
   
   l <- loess(g0 ~ R2, data=miner, span=0.8)
   with(miner,plot(R2, g0, pch=16, ylim=c(-100, 300), xlim=c(0,1),
@@ -125,6 +126,7 @@ figure_R2g0 <- function(lin2015coef, miner){
   abline(h=0)
   
   legend("topright", "Miner et al. 2017", bty='n', cex=0.8)
+  plotlabel("(b)", "topleft")
   
   l <- loess(g0_se ~ cv_bbopti, data=lin2015coef, span=0.9)
   with(lin2015coef, plot(cv_bbopti, g0_se, pch=16, ylim=c(0,60),
@@ -132,7 +134,7 @@ figure_R2g0 <- function(lin2015coef, miner){
                          ylab=expression("SE of"~g[0]~~(mmol~m^-2~s^-1)),
                          panel.first=plot_loess(l, add=TRUE, band=TRUE, lwd=2, col="darkgrey")
                          ))
-
+  plotlabel("(c)", "topleft")
   
 }
 
@@ -375,9 +377,11 @@ gmin_loghist <- function(gmindat){
 
 gmin_by_order <- function(gmindat){
   
+  logmean <- function(x)mean(log(x), na.rm=TRUE)
+  
   gmindat3 <- mutate(gmindat, 
                      Order = fct_lump(Order, n=10),
-                     Order = reorder(Order, gmin, median)) %>%
+                     Order = reorder(Order, gmin, logmean)) %>%
     filter(Order != "Other") %>% droplevels
   
[…]

### 2018-01-08 R/howmanyrefs.R

@@ -1,0 +1,7 @@
+
+
+rl <- readLines("manuscript.Rmd")
+
+library(stringr)
+ref <- str_extract_all(rl, "@[a-zA-Z]+[0-9]{4}")
+length(unique(do.call(c, ref)))

### 2018-01-08 R/load.R

@@ -4,7 +4,7 @@ pacman::p_load(Hmisc, car, dplyr, tidyr, nlme, nlshelper,
                forcats, tibble, magicaxis, 
                plantecophys, readxl, multcomp,
                reporttools, Taxonstand, taxize, speciesmap,
-               doBy, stringi)
+               doBy, stringi, rmarkdown)
 
 
 

### 2018-01-08 R/make_figures.R

@@ -3,7 +3,6 @@ source("R/load.R")
 
 
 to.pdf(figure_gmin_review(gdfr), "output/figure1.pdf", height=5, width=5)
-to.pdf(figure_gmin_review_2(gdfr), "output/figure1_unlog.pdf", height=5, width=5)
 
 to.pdf(figure_g0g1_cor(lin2015, "Nicolas Martin-StPaul_Quercus ilex_StPaul_Puechabon"), 
        "output/figure2.pdf", height=4, width=8)

### 2018-01-08 run.R

@@ -1,0 +1,10 @@
+
+# Read & download data, load & install packages.
+source("R/load.R")
+
+# Make figures as PDF (see output/)
+# Conversion to png relies on an installation of ImageMagick (and GhostScript); may fail on your system.
+source("R/make_figures.R")
+
+# Convert manuscript (output is manuscript.docx).
+rmarkdown::render("manuscript.Rmd", encoding = "utf8")

### 2018-01-09 manuscript.Rmd
Nico figure added; some writing

@@
 
-@hanson2016 large effect of small gmin on estimates of mesophyll conductance, Ci. This is important but not the focus of this review.
-
 # Summary
@@
 Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time.
-This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants that are severely droughted, and includes all surfaces of the leaf.
-Since the seminal review by @kerstiens1996, no attempt has been made to synthesize the literature for values of g~min~, which is timely as models start to incorporate more detailed drought dynamics all the way to the point of mortality by desiccation (Xu, Blackman, Martin-StPaul, Sperry?).
-Instead, many models rely on inaccurately estimated 'minimum' conductance from cuvette data, by applying regression techniques.
+This method (mass loss of detached leaves, MLD) is of particular interest because it more closely simulates actual field conditions for plants during severe drought, and includes both surfaces of the leaf - not just the stomata-free side which is typically used for cuticular conductance measurements.
 
@@
 \
-(*expand*) We thus have several different definitions of what constitutes a 'minimum' conductance of leaves.
+We thus have several different definitions of what constitutes a 'minimum' conductance of leaves.
 In this review, we demonstrate that inclusion of a minimum conductance is necessary both for simulation of the response of water-use efficiency to various environmental drivers, as well as a component of the response of plants to severe drought.
-However, there is no guidance at the moment on how to estimate minimum conductance for use in models, how various definitions compare quantitatively, and whether it is feasible to assign species-specific values to the minimum conductance.
-We do not expect that g~0~, cuticular conductance (g~cut~) and minimum conductance from mass loss of detached leaves (we refer to minimum conductance measured with this method only as g~min~) to give similar values, in fact the contribution of stomata to each of these conductances will be very different [@brodribb2014].
+However, there is no guidance at the moment on how to estimate minimum conductance for use in models, how various definitions compare quantitatively, and whether it is feasible to assign species-specific values of the minimum conductance.
+We do not expect that g~0~, cuticular conductance (g~cut~) and minimum conductance from mass loss of detached leaves (we refer to minimum conductance measured with this method further as g~min~) to give similar values, in fact the contribution of stomata to each of these conductances will be very different [@brodribb2014].
 
 \
-(*expand*) Previews reviews have focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles (riederer, kerstiens).
+Previews reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017].
 Other work has focused on the role of g~0~ in models of plant water use [e.g.
 @barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance.
-Thus there is a disconnect between dizzying level of detailed understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models We aim to improve the connection somewhat, and point at obvious gaps in understanding.
+Thus there is a disconnect between the dizzying level of understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models.
+We aim to improve this connection by synthesizing the state of knowledge, thereby pointing out gaps in understanding, and suggest areas for further research.
 
@@
 \
-Measurements of (presumed) minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories (see Table 1).
+Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories (see Table 1).
 The absolute minimum conductance attained by leaves is that through the cuticle only.
@@
 This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress.
-Finally, the g~0~ parameter, which is a *modelled* minimum conductance reached when photosynthesis rate is zero, is typically estimated by applying regression to Eq.
+Finally, the g~0~ parameter, which is a model-based minimum conductance reached when photosynthesis rate is zero, is typically estimated by applying regression to Eq.
 1 (of which various forms exist).
@@
 
-- Measurement error, small fluxes
-
 \
@@
 We thus conclude that g~0~ should really not be estimated from regression, although this is the most common method applied.
+In addition, it is difficult to accurately measure low fluxes with a portable gas exchange system, and great care has to be taken to arrive at reasonable estimates of g~0~ that way.
 
@@
 Despite these early reports, the exact role of g~0~ in models is often overlooked.
+
+\
+In this review we focus on the cuticle as a barrier to water loss, but point out that previous work has shown that the cuticle is much less permeable to CO~2~ compared to H~2~O [@boyer1997; @boyer2015].
+This finding, if indeed generally true, has a large effect on the calculation of various gas exchange parameters [@hanson2006].
+@manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is leaky and incompletely closed stomata (see Section XXX).
+Because g~0~ includes not just the cuticle, but also incompletely closed stomata, and the finding that g~0~ is much larger than estimates of g~min~ or g~cut~ (Fig.
+1), we ignore this effect in the model simulations that follow.
 
@@
 
+```{r fig.height=6, echo=FALSE, eval=includefigs}
+plotfig("output/figure5.png")
+```
+
+**Figure 5**
+
 # V. Cuticular conductance
@@
 \
-Literature values of g~cut~, especially when measured on isolated cuticles (g~cut,isol~) were much lower than any other estimate of conductance (Fig.
-1), in line with previous comparisons [@kerstiens1996].
-The conductance of stomata-free leaf cuticles represents the absolute minimum conductance that leaves can attain.
-This result should be viewed with caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions.
-Boyer et al.
-1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases (Boyer et al., 1997)
+Literature values of g~cut~, especially when measured on isolated cuticles (g~cut~) were much lower than any other estimate of conductance (Fig.
+1), in line with the methods comparison of @kerstiens1996.
+More recently, @schuster2017 argued, based on a data for 8 species, that g~min~ from MLD is comparable to g~cut~.
+However in the same study, a literature compilation of hundreds of values showed that g~cut~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cut~ were pooled in their analysis.
+Nonetheless, the comparison between g~cut~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions.
+@boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.
 
 \
-Indeed, some more recent studies have shown a very good correspondence between g~cut~ and g~min~ (riederer, schuster).
 A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these processes have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b].
@@
 \
-It is important to bear in mind other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, and various interactions with insects [@mueller2008] [@kerstiens1996c;@riederer2008].
-The thickness of the cuticle per se may also have to do with mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness).
+It is important to bear in mind the many other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, various interactions with insects including signalling and herbivory resistance [@mueller2008; @kerstiens1996c;@riederer2008].
+The thickness of the cuticle may also confer mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness).
 Thick cuticles may also be useful because exposure to wind has been shown to dislodge cuticular compounds by abrasion [@rogge1993], thereby increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991].
@@
 Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged.
-Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but this variation could not explain variation in g~cut~.
+Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but this variation explains only some of the wide range of g~cut~.
 Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.
-
-\
-(*expand*) Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015].
-This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata (see Section XXX).
-Nonetheless it is an interesting observation that could have major consequences for modelling and interpreting water-use efficiency data during drought conditions.
-Also @hanson2016 showing the effect of this on calculation of Ci, and calculation of gmeso.
 
@@
 ```{r fig.height=5, fig.width=5, echo=FALSE, eval=includefigs}
-plotfig("output/figure5.png")
+plotfig("output/figure6.png")
 ```
 
-**Figure 5**.
+**Figure 6**.
 Analysis of a new literature compilation of g~min~ estimates, as measured with mass loss of detached leaves.
@@
 ```{r fig.height=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure6.png")
+plotfig("output/figure7.png")
 ```
 
-**Figure 6**.
+**Figure 7**.
 Minimum conductance measured from leaf dry-down curves on a variety *Hakea* species, a genus native to Australia.
@@
 ```{r fig.height=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure7.png")
+plotfig("output/figure8.png")
 ```
 
-**Figure 7**.
+**Figure 8**.
 Minimum conductance measured from leaf dry-down curves on *Eucalyptus parramattensis* grown in whole-tree chambers.

### 2018-01-09 R/figures.R

@@ -451,3 +451,36 @@ figure_crop_genotypes <- function(cropgmin){
   
 }
 
+
+
+figure_sureau <- function(planta, plantb){
+  
+  #windows(9,4)
+  par(mfrow=c(1,3), mar=c(4,4,1,1), mgp=c(2.5, 0.5, 0), tcl=0.2, cex.lab=1.2)
+  
+  cols <- c(rgb(230, 159, 0, maxColorValue = 255), rgb(0, 114, 178, maxColorValue = 255))
+  
+  # panel a
+  with(planta, plot(days, REW, lwd=2, ylim=c(0,1), xlim=c(0,100), type='l', xlab="Days", col=cols[1]))
+  with(plantb, lines(days, REW, lwd=2, ylim=c(0,1), xlim=c(0,100), col=cols[2]))
+  plotlabel("(a)", "bottomright")
+  legend("topright", c("3.3", "1.6"), title=expression(g[min]), bty='n', lty=1, col=cols, lwd=2, cex=1.2)
+  
+  # panel b
+  with(planta, plot(days, Pmin, ylim=c(-7,0), xlim=c(0,100), 
+                    col=cols[1],
+                    ylab= "Water potential (MPa)",
+                    type='l', lwd=2, lty=1, xlab="Days"))
+  with(planta, lines(days, Psoil, lwd=2, col=cols[1], lty=5))
+  with(plantb, lines(days, Pmin, type='l', lwd=2, lty=1, col=cols[2]))
+  with(plantb, lines(days, Psoil, lwd=2, col=cols[2], lty=5))
+  plotlabel("(b)", "bottomright")
+  legend("topright", c("leaf","soil"), lty=c(1, 5), lwd=2, bty='n', cex=1.2)
+  
+  # panel c
+  with(planta, plot(days, PLC*100, xlim=c(0,100), ylab="PLC (%)", type='l', lwd=2, xlab="Days", col=cols[1]))
+  with(plantb, lines(days, PLC*100, xlim=c(0,100), type='l', lwd=2, col=cols[2]))
+  plotlabel("(c)", "bottomright")
+  
+}
+

### 2018-01-09 R/load.R

@@ -209,6 +209,11 @@ gdfr <- bind_rows(gmindat_simple, kerst_simple, schuster2017, lombar, g0s, minag
   mutate(method = factor(method, levels=c("gcut_isol","gmin","gnight", "g0","gslowA")))
 
 
+# Nico's simulations with Sureau
+planta <- read.csv2("data/PlantA-DiurnalOut.csv", dec=".")
+plantb <- read.csv2("data/PlantB-DiurnalOut.csv", dec=".")
+planta$days <- (1:nrow(planta))/10
+plantb$days <- (1:nrow(plantb))/10
 
 
 

### 2018-01-09 R/make_figures.R

@@ -11,11 +11,13 @@ to.pdf(figure_R2g0(lin2015coef, miner), "output/figure3.pdf", height=4, width=9)
 
 to.pdf(figure_sim(), "output/figure4.pdf", height=6, width=6)
 
-to.pdf(gmin_3panel(gmindat, cropgmin), "output/figure5.pdf", height=3, width=8)
+to.pdf(figure_sureau(planta, plantb), "output/figure5.pdf", height=4, width=9)
 
-to.pdf(figure_hakea_gmin(lopw), "output/figure6.pdf", height=5, width=6)
+to.pdf(gmin_3panel(gmindat, cropgmin), "output/figure6.pdf", height=3, width=8)
 
-to.pdf(figure_wtc4_gmin_2(wtc4gmin), "output/figure7.pdf", height=4, width=3.5)
+to.pdf(figure_hakea_gmin(lopw), "output/figure7.pdf", height=5, width=6)
+
+to.pdf(figure_wtc4_gmin_2(wtc4gmin), "output/figure8.pdf", height=4, width=3.5)
 
 
 convert_all_pdf("output")

### 2018-01-10 manuscript.Rmd
much writing

@@
 \t1. Models of water-use efficiency
-\t2. Models of drought response of stomatal conductance
+\t2. Models of plant desiccation
 V. Cuticular conductance
@@
 Plants cannot, however, completely halt water loss even when stomata are closed.
+There is increasing recognition that the minimum conductance plays an important role in estimating water fluxes in plant canopies [@barnard2013], during heat waves [@kala2016], and in models of plant drought response [@blackman2016; @martinstpaul2017].
 
 \
-The absolute minimum attainable water loss rate through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf.
-Other measurements allow for the estimation of water loss through incompletely closed and broken stomata.
-Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time.
-This method (mass loss of detached leaves, MLD) is of particular interest because it more closely simulates actual field conditions for plants during severe drought, and includes both surfaces of the leaf - not just the stomata-free side which is typically used for cuticular conductance measurements.
-
-\
-Most current-generation process-based global vegetation models (GVMs) use a version of the Ball-Berry class of stomatal conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity.
+Most current-generation process-based global vegetation models (GVMs) use a version of the Ball-Berry class of stomatal conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity (Eq.
+1).
 
@@
 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987).
-Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been largely ignored.
+Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been studied in much less detail.
 For example, a recent comprehensive review of stomatal conductance models did not mention g~0~ or any similar minimum conductance [@damour2010].
@@
 \
-We thus have several different definitions of what constitutes a 'minimum' conductance of leaves.
 In this review, we demonstrate that inclusion of a minimum conductance is necessary both for simulation of the response of water-use efficiency to various environmental drivers, as well as a component of the response of plants to severe drought.
-However, there is no guidance at the moment on how to estimate minimum conductance for use in models, how various definitions compare quantitatively, and whether it is feasible to assign species-specific values of the minimum conductance.
-We do not expect that g~0~, cuticular conductance (g~cut~) and minimum conductance from mass loss of detached leaves (we refer to minimum conductance measured with this method further as g~min~) to give similar values, in fact the contribution of stomata to each of these conductances will be very different [@brodribb2014].
+However, there is no guidance at the moment on how to estimate minimum conductance for use in models, how various definitions and measurements compare quantitatively, whether it is feasible to assign species-specific values of the minimum conductance, and what are the ecological and environmental controls on g~min~ [@schuster2017].
 
 \
-Previews reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017].
+The absolute minimum attainable water loss rate is through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf.
+Other measurements allow for the estimation of water loss through incompletely closed and broken stomata.
+Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time.
+This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants during severe drought, and includes both surfaces of the leaf - not just the stomata-free side which is typically used for cuticular conductance measurements.
+A comprehensive compilation of estimates of g~min~ is currently lacking, and is needed not just to parameterize models, but also to study sources of variation in this overlooked plant trait.
+
+\
+Previous reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017].
 Other work has focused on the role of g~0~ in models of plant water use [e.g.
@@
 We aim to improve this connection by synthesizing the state of knowledge, thereby pointing out gaps in understanding, and suggest areas for further research.
+Our goals are to a) compare various definitions of g~min~ quantitively, which have been previously assumed equal, b) demonstrate the need to include a non-zero g~0~ (Eq.
+1), for models of water-use efficiency as well as the drought response, c) present a new compilation of minimum conductance measurements, and demonstrate that the large acclimatory potential of this trait to environmental drivers.
 
@@
 The absolute minimum conductance attained by leaves is that through the cuticle only.
-Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001], via gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata (g~cut,seal~) [@kerstiens1996].
-The method we here refer to as minimum conductance is that of mass loss of intact detached leaves (MLD).
-In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out.
-After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit [@sinclair1986].
-This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress.
-Finally, the g~0~ parameter, which is a model-based minimum conductance reached when photosynthesis rate is zero, is typically estimated by applying regression to Eq.
-1 (of which various forms exist).
+Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001], via special gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata [@kerstiens1996].
+We further discuss cuticular conductance in Section V.
 
 \
-We quantitatively compared each of these different 'minimum' conductance values by synthesizing existing data.
-To estimate g~0~ from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model.
-Measurements of conductance of isolated, non-stomatal cuticles (g~cut,isol~), and of conductance of detached leaves with the stomatal side sealed (g~cut,seal~) were taken from the compilation by @kerstiens1996.
-Data on minimum conductance (g~min~) from MLD was compiled from 44+ original papers (see SuppInfo A for a list of papers and additional figures describing this new compilation).
+The minimum conductance of intact leaves is typically measured by mass loss of detached leaves (MLD).
+In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out.
+Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after a while a constant low rate is achieved.
+From this minimum transpiration rate (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), the minimum conductance (g~min~) can be estimated using measured vapour pressure deficit [@sinclair1986].
+This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress.
+
+\ 
+We quantitatively compared each of the different 'minimum' conductance (Table 1) values by synthesizing existing data.
+To estimate g~0~ (Eq.
+1) from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model.
+Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) were taken from the compilations by @kerstiens1996 and @schuster2017.
+Data on minimum conductance (g~min~) from MLD was compiled from 44+ original papers (see SuppInfo 2 for a list of papers).
 For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources.
@@
 @kerstiens1996).
-Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance, it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful.
+Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance (discussed further in Section VI), it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful.
 A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca.
 150 species during night-time.
-Actual minimum daytime values were estimated from the Lin et al.
-database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset.
+Actual minimum daytime values were estimated from the @lin2015 database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset.
 This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.
@@
 \
-(*expand*) The results are presented in Fig.
+The results of the quantitative comparison are presented in Fig.
 1.
-Conductance in the dark was on average eight-fold higher than g~min~.
-It is no new observation that stomata do not close in the dark, but do close on detached leaves, as already demonstrated by @hygen1954 in Scots pine, and described by @koerner1994.
-More recently, @walden-coleman2013 showed much higher g~dark~ compared to g~min~ in cotton.
+From this compilation, a few striking differences between estimates are apparent.
+Cuticular conductance averaged 0.45 mmol m^-2^ s^-1^, ten-fold lower compared to g~min~ (4.9 mmol m^-2^ s^-1^).
+This result is in line with the methods comparison of @kerstiens1996.
+More recently, @schuster2017 argued, based on a data for 8 species, that g~min~ from MLD is comparable to g~cut~.
+However in the same study, a literature compilation of hundreds of values showed that g~cut~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cut~ were pooled in their analysis.
+Nonetheless, the comparison between g~cut~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions.
+@boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.
 
-However, @cavender-bares2007 gdark similar to gmin in oaks.
+\
+Conductance in the dark (g~dark~) (mean = 41.4 mmol m^-2^ s^-1^ was on average eight-fold higher than g~min~.
+It is no new observation that stomata do not close in the dark (further discussed in Section VI).
+Previous studies have directly compared g~dark~ and g~min~ on the same species, and generally concluded that g~dark~ is much higher than g~min~ [@hygen1954; @koerner1994; @walden-coleman2013], but an exception is @cavender-bares2007 who found the two rates to be similar in oaks.
 
-Actual measured stomatal conductance during low photosynthesis (g~s,low A~) was also much lower than fitted g~0~ or g~dark~.
-In other words, actual daytime conductance is much lower than the presumed minimum estimated from fitted g~0~.
+\
+The fitted minimum conductance parameter, g~0~ (mean = 32.8 mmol m^-2^ s^-1^), was much higher than g~min~, but similar to g~dark~.
+However, we point out the uncertainties associated with g~0~ in Section III, which tend to inflate estimates of g~0~.
+It is telling that actual measured stomatal conductance during low photosynthesis (g~s,low A~) was  much lower than g~0~ or g~dark~.
+In other words, actual daytime conductance during these conditions that cause low photosynthetic rates is much lower than the presumed minimum estimated from fitted g~0~.
 A similar result was already presented by @dekauwe2015 (their Fig.
@@
 **Figure 1**.
-Comparison of various estimates of the minimum conductance.
+Comparison of various estimates of the (presumed) minimum conductance.
 Error bars are 95\% confidence intervals.
+Grey dots are the original data (but some data occurs outside the figure range).
 Different letters denote significant differences (at $\alpha$ = 0.05).
-See Table 1 for definitions of minimum conductance, and text for the data sources.
+g~cut,isol~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~0~ : fitted parameter (Eq.
+1) with regression; g~s,low A~ leaf conductance during periods of very low photosynthesis.
+See text for data sources and methods.
 
@@
 @barnard2013 also mentioned the difficulty of fitting Eq.
-1 to estimate g~0~, but did not present specific details.
+1 to estimate g~0~, but did not present specific details - the following analysis builds on their work by demonstrating statistical uncertainties.
 
@@
 The consequence is that it is more difficult to interpret g~1~ and g~0~ independently, and that confidence intervals on each of the parameters should also not be interpreted independently (see @becker2007 for more discussion on the statistical aspects of comparing slopes and intercepts).
-(*expand*)
 
@@
 **Figure 3**.
-Correlation between estimated g~0~ and the R^2^ of the regression.
-In the left panel, we fitted the linearized form of the Medlyn et al.
-(2011) model to each of the datasets in the Lin et al.
-(2015) database.
+Statistical uncertainty in the estimation of g~0~ from regression, demonstrated with two parameter databases.
+(a) We fitted the linearized form of the @medlyn2011 model to each of the datasets in the @lin2015 leaf gas exchange database, showing that for poorly fitted relationships (low R^2^), inflated estimates of g~0~ are obtained.
 Vertical lines are 95% confidence intervals.
-In the right panel, a very similar association between g~0~ and R^2^ is presented by the review of @miner2017, using the Ball-Berry model.
-In panel (c) coefficient of variation and R^2^.
+The grey line is a fitted loess smoother with 95% confidence interval.
+Note the wide confidence intervals, and frequent negative values.
+(b) Similar to (a), but using the published compilation by @miner2017.
+The grey line is a fitted loess smoother with 95% confidence interval.
+(c) Using the fits from panel (a), a demonstration that the standard error (SE) of g~0~ is much higher when the coefficient of variation (CV) of the predictor (i.e.
+right-hand side of the equation being fitted) is lower.
 
@@
 This finding, if indeed generally true, has a large effect on the calculation of various gas exchange parameters [@hanson2006].
-@manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is leaky and incompletely closed stomata (see Section XXX).
+@manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is leaky and incompletely closed stomata.
 Because g~0~ includes not just the cuticle, but also incompletely closed stomata, and the finding that g~0~ is much larger than estimates of g~min~ or g~cut~ (Fig.
@@
 In the application of the Ball-Berry models, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005].
-It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata (see Section XXX).
-
-\
-Photosynthesis also ceases at very high temperature (ref), but stomata do not appear to close in proportion to the decrease in photosynthesis, if at all [@drake, @urban2017], thus we cannot use high temperature to estimate a reasonable g~0~.
-Simulations of water use during heat waves depends highly on the value of g~0~ used [@kala2016] .
-It should be pointed out that the response of g~s~ to T~leaf~ is likely poorly modelled with a Ball-Berry model, since a number of studies have shown that g~s~ and A~n~ decouple at high T~leaf~, resulting in high rates of g~s~ when A~n~ is near zero [@urban2017] (Drake et al., unpublished).
+It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata.
+Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake, @urban2017].
 
@@
 We also demonstrate the effect of g~0~ on modelling the T~leaf~ response.
-In this simulation, T~leaf~ and D are assumed to co-vary with an empirical relationship as used by @duursma2014.
+In this simulation, T~leaf~ and VPD are assumed to co-vary with an empirical relationship as used by @duursma2014.
 Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increase leaf transpiration E~L~, but at a slightly higher rate than just due to g~0~.
@@
 **Figure 4**.
-Simulations  of intrinsic water-use efficiency (A~n~/g~s~), intercellular CO~2~ concentration (C~i~) and leaf transpiration rate (E~L~) with three values of g~0~.
-For the top two panels, all environmental drivers were held constant except the photosynthetic photon flux density (PPFD), demonstrating that the inclusion of g~0~ is necessary to simulate the decrease in A~n~/g~s~ at low PPFD.
-For the bottom two panels, VPD and T~air~ were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions.
-For all simulations it is assumed that T~leaf~ is equal to T~air~.
+Simulations with a coupled leaf gas exchange model [@duursma2015], demonstrating the effect of inclusion of the g~0~ parameter (Eq.
+1) on leaf fluxes.
+(a) Intrinsic water-use efficiency (A~n~/g~s~) as a function of he photosynthetic photon flux density (PPFD), holding other environmental drivers constant, for three values of g~0~.
+(b) The same simulations as in (a), but showing the intercellular CO~2~ concentration (C~i~).
+(c) Stomatal conductance (g~s~) simulations, where vapour pressure deficit (VPD) and air temperature (T~air~) were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions.
+(d) The same simulations as in (c), but showing C~i~.
+Note how C~i~ increases at high VPD and T~air~, only when g~0~> 0.
+For all simulations it is assumed that T~leaf~ is equal to T~air~, and ignoring the differential permeability of the cuticle to CO~2~ and H~2~O [@hanson2016].
 
-## 2. Models of drought response of stomatal conductance
+## 2. Models of plant desiccation
 
 \
-Another reason to include a non-zero g~0~ term in models of stomatal conductance arises during severe drought.
-When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata.
-Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000].
-This is particularly important given the limited variation in stomatal response to drought among plants @martinstpaul2017.
-
-@pisekwinkler1953, early calculation of the length of time needed to desiccate leaves to some water content, given the minimum transpiration rate, and the water content of the leaves.
-
-@sinclair2000 gmin as a drought tolerance trait, can calculate time needed to death during drought phase when plants no longer take up water.
-
+When plants are sufficiently water-stressed that stomata are mostly closed, water loss still continues at a rate determined by the minimum conductance.
+Thus models that aim to predict when plants desiccate and die must include a minimum conductance term.
+A classic study by @pisekwinkler1953 already calculated the length of time needed to desiccate leaves to some critical low water content, given the minimum transpiration rate, and the saturated water content of the leaves.
+Based on that work, @burghardt2006 reported on a direct correlation between g~min~ and the survival time of leaves.
+@sinclair2000 presented the minimum conductance as a key drought tolerance trait, and used it to as a basis for predicting crop mortality during severe drought.
 More recently, @gleason2014 and @blackman2016 proposed that embolism resistance along with whole plant capacitance and minimum transpiration rates all contribute to define the time to desiccation.
-
-@burghardt2006 reported on a correlation between g~min~ and the survival time of leaves.
-
-Finally, Martin-StPaul et al 2017 demonstrated in a whole plant model of hydraulic failure that gmin was probably one of the key parameters to explain water potential drops below cavitation thresholds during drought given the relatively conserved range of stomatal closure to drought.
-Relying on the Sureau model presented in @martinstpaul2017 we illustrate the critical role of gmin in defining desiccation time.
+Finally, @martinstpaul2017 demonstrated in a whole plant model of hydraulic failure that g~min~ was one of the key parameters to explain the drop in water potential below the cavitation threshold, because stomata generally close well before this threshold.
+Relying on the *Sureau* model presented in @martinstpaul2017 we illustrate the critical role of g~min~ in defining the desiccation time (Fig.
+5).
 
@@
 
-**Figure 5**
+**Figure 5**.
+Simulations with the *Sureau* model demonstrating the effect of g~min~ on desiccation tolerance of plants.
+The *Sureau* model simulates water transport in the soil-plant-atmosphere continuum, and includes a detailed representation of capacitance in stem and leaf tissues.
+(a) Soil relative extractable water (REW; 1 = field capacity, 0 = permanent wilting point) for the two simulations, using a minimum conductance (g~min~) of 1.6 or 3.3 mmol m^-2^ s^-1^ - all other parameters were equal.
+(b) Water potential in the soil and leaf as the dry-down progresses.
+Note the brief increase in soil water potential around days 40 - 70; this represents water released from stem cavitation.
+(c) Progression of percent loss conductivity (PLC) of the xylem.
+If we assume that plant mortality occurs at a PLC of 88%, the plants with the low g~min~ lasted 15 days longer (or nearly 40%).
 
@@
 \
-Literature values of g~cut~, especially when measured on isolated cuticles (g~cut~) were much lower than any other estimate of conductance (Fig.
-1), in line with the methods comparison of @kerstiens1996.
-More recently, @schuster2017 argued, based on a data for 8 species, that g~min~ from MLD is comparable to g~cut~.
-However in the same study, a literature compilation of hundreds of values showed that g~cut~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cut~ were pooled in their analysis.
-Nonetheless, the comparison between g~cut~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions.
-@boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.
-
-\
 A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these processes have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b].
-However, a few key points should be reiterated as they are relevant to the current discussion, in particular when we aim to interpret variation in literature values of g~min~.
+However, a few key points should be summarized as they are relevant to the current discussion, in particular when we aim to interpret variation in literature values of g~min~ (Section VII).
 
@@
 Because most of the resistance is located in such a thin layer, g~cut~ does not correlate with the thickness of the cuticle [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b].
-Nonetheless there is considerable variation in cuticle thickness between plant species [@schuster2016a], as well as within species in response to drought (cite), along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010].
+Nonetheless there is considerable variation in cuticle thickness between plant species [@schuster2016a], as well as within species in response to drought (**cite**), along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010].
 If the cuticle thickness does not directly affect its conductance, what is the value of increasing the cuticle?
@@
 \
-It is important to bear in mind the many other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, various interactions with insects including signalling and herbivory resistance [@mueller2008; @kerstiens1996c;@riederer2008].
-The thickness of the cuticle may also confer mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness).
-Thick cuticles may also be useful because exposure to wind has been shown to dislodge cuticular compounds by abrasion [@rogge1993], thereby increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991].
+It is important to bear in mind the many other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, various interactions with insects including signalling and herbivory resistance [@mueller2008; @kerstiens1996c;@riederer2008], and foliar water uptake [@fernandez2017] The thickness of the cuticle may also confer mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness).
+As such, thick cuticles may be advantageous because prolonged exposure to wind increases the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991], by dislodging cuticular compounds by abrasion [@rogge1993].
 Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986], most dramatically demonstrated in an ice storm, leading to much higher g~min~ [@boyce2003].
@@
 1), g~dark~ is substantially higher compared to any other measure of minimum conductance.
-Previous work on g~dark~ have confirmed that night-time conductance is generally much higher than g~min~, when measured on the same plants [@howard2007; @caird2007; @walden-coleman2013; @phillips2010].
 The simple idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.
@@
 
-Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after a while a constant low rate is achieved.
-From this minimum transpiration rate, a minimum conductance (g~min~) is usually estimated, taking into account the vapour pressure deficit [@sinclair1986].
-This method has been termed the 'mass loss of detached leaves' (MLD) [@kerstiens1996].
-Many previous reports use the term 'cuticular conductance' for g~min~, but this is inaccurate as some fraction of the water loss will occur through broken, leaky, or incompletely closed stomata.
 A few studies have directly quantified the stomatal component of g~min~.
@@
 
+\
 In the following sections, we review the quite substantial literature on g~min~ measured with the MLD technique, stretching back to the 1930s [@pisekberger1938], focussing on environmental and ecological determinants.
-We also further tease apart the literature compilation of g~min~ (which was shown just as a single average value in Fig.
-4).
+We also further analyze the literature compilation of g~min~ (Fig.
+6).
+Comparing all available data, we did not find significant relationships between rainfall, aridity, or temperature (all extracted across the range of the species) and g~min~, nor were there meaningful relationships with other traits (**not shown**).
+We thus found it difficult to explain the variation in g~min~ between species.
+A simple breakdown by taxonomic order (Fig.
+6b) revealed that grasses (*Poales*) have a higher g~min~ compared to other orders, and conifers (especially *Pinales*) tended towards the lower range of values.
 
-We found it difficult to explain the variation in g~min~ between species.
-Grasses have a higher g~min~, and so do crops.
-Otherwise few differences were significant, with Pinales tending towards lower values of g~min~, and Myrtales (including many *Eucalyptus* species), tending towards higher values.
+\
+Very few studies have found meaningful correlations between g~min~ and environmental factors or ecological categories.
+Based on their own compilation of the literature, @schuster2017 concluded no significant differences in g~min~ / g~cut~ by plant growth form.
+One notable exception is @brodribb2014, who showed a correlation between rainfall at species origin (of the driest quarter), and g~min~.
+Such correlation did not bear out across all species in our database, and interesting to note is that *Eucalyptus* species (n = 11, included in *Myrtales*) - all measured on the driest continent of Australia - have g~min~ values slightly higher than the average (Fig.
+6b).
 
-Comparing all available data, we did not find significant relationships between rainfall, aridity, or temperature and g~min~ (see SuppInfo 2).
-There was a weak correlation between g~min~ and P~50~, but this was driven entirely by the difference in P~50~ between gymnosperms and angiosperms (not shown).
-
-Very few studies have found meaningful correlations between g~min~ and environmental factors or ecological categories.
-One notable exception is @brodribb2014, who showed a correlation between rainfall at species origin (of the driest quarter), and g~min~.
+\
+**crops**
 
@@
 **Figure 6**.
-Analysis of a new literature compilation of g~min~ estimates, as measured with mass loss of detached leaves.
-Left panel, probability density function of all values (after averaging by species, n=200+), with a normal distribution curve (mean=, sd=, on the log scale).
-Middle panel, g~min~ averaged by phylogenetic order (including only the top ten orders in the database).
+Analysis of a literature compilation of g~min~ estimates, as measured with mass loss of detached leaves.
+(a) Histogram (probability density) of all estimates (after averaging by species, n=`r nrow(gmindat)`), with a log-normal distribution curve (mean = `r round(exp(mean(log(gmindat$gmin))),2)`, sd = `r round(exp(sd(log(gmindat$gmin))),2)`).
+(b) g~min~ averaged by phylogenetic order (including only the top ten orders in the database).
 Bars are 95/% confidence intervals.
-Numbers on top refer to the number of species, different letters denote significant differences (at alpha = 0.05, adjusted for multiple comparisons), and grey symbols are individual species.
-Right panel, data for crops only where each point is a genotype.
-Bars denote the range, illustrating the large variation in g~min~ between genotypes for a particular crop species.
+Numbers above the figure refer to the number of species.
+Different letters denote significant differences (at $\alpha$ = 0.05, adjusted for multiple comparisons), and grey symbols are species-level data.
+(c) g~min~ estimates for crops only, averaged by genotype.
+Bars denote the range, illustrating the wide range in g~min~ between genotypes for a particular crop species.
+Numbers above the figure refer to the number of genotypes included.
 
@@
 
-(*expand*) From the literature we found a general acclimatory response of g~min~ to drought stress, nearly always leading to a decrease in g~min~ (see Table 2 for a summary of the literature); the only exception was @vilagrosa2003 who found no response.
-A number of studies have attempted to attribute this decrease in g~min~ to a change in cuticular composition.
+\
+From the literature we found a general acclimatory response of g~min~ to drought stress, nearly always leading to a significant  decrease in g~min~ (see Table 2 for a summary of the literature), with just one exception.
+In each of the studies summarized, plants were grown in well-watered or droughted conditions, and in one case a difference in g~min~ was demonstrated after just four days of drought exposure [@bengtson1978].
+We also demonstrate drought acclimation via a case study on 11 *Hakea* species grown in two watering treatments (Fig.
+7).
+
+\
+A number of studies have attempted to attribute this drought acclimation in g~min~ to a change in the chemical composition of the cuticle.
 @bengtson1978 reported lower g~min~ in six oat varieties in response to drought, and increase in cuticular waxes, but no overall relationship between the two and very different responses between different oat genotypes.
@@
 @mackova2013 found that adding ABA (simulating drought stress) to *Lepidium sativum* increased the chain length of cuticular waxes (but not total amounts).
-@bi2017 Drought caused a change in cuticular wax production and chemical composition, but in a highly genotype-specific way.
-@kottapalli2009 higher epicuticular wax load in a drought-tolerant vs.
-susceptible peanut genotype.
-
-\
-We also demonstrate this effect via a case study on *Hakea* plants.
-**Measurement details here, or in caption**.
-The results showed a significant decrease in g~min~ with long-term exposure to low soil water availability (average effect X, p-value X).
-(*expand*).
-We also found evidence for a link between g~min~ and leaf shape, as the studied *Hakea* species had either needle-like or broad leaves (see Fig.
-6).
+@bi2017 also reported that drought caused a change in cuticular wax production and chemical composition, but again in a highly genotype-specific way.
+Thus it can be concluded that the cuticle indeed changes substantially after a change in plant available water, but in a complex, species-specific manner that is yet to be connected directly to g~min~.
 
@@
 **Figure 7**.
-Minimum conductance measured from leaf dry-down curves on a variety *Hakea* species, a genus native to Australia.
-Plants were grown in containers in a grow house, and supplied with ample water or underwent long-term mild drought stress.
-Bars are labelled by leaf form : broadleaf (B) or needle-like (N).
+Minimum conductance (g~min~) measured by mass loss of detached leaves on a variety *Hakea* species, a genus native to Australia.
+Plants were grown in containers in a grow house, and supplied with ample water or subjected to long-term (8 months) mild drought stress.
+Bars are labelled by leaf form : broadleaf (B) or needle-like (N), and ordered by g~min~ in the  well-watered treatment.
 For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`).
@@
 
-Perhaps even more significant is the response of g~min~ (and g~cut~) to temperature: it shows both an instantaneous response, as well as an acclimatory response.
-@riederer2001 and @schuster2016b demonstrated a steep non-linear response of g~cut~ to temperature, with a flat response below XC, and a steep response thereafter (*expand*).
+The response of minimum conductance to temperature is also striking; it shows both an instantaneous response to increased temperature, as well as an acclimatory response to growth temperatures.
+@riederer2001 and @schuster2016b have both demonstrated a steep non-linear instantaneous response of g~cut~ to temperature, with the response becoming especially steep at higher temperatures.
 
@@
 **Figure 8**.
-Minimum conductance measured from leaf dry-down curves on *Eucalyptus parramattensis* grown in whole-tree chambers.
-Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C).
+Minimum conductance (g~min~) measured with mass loss of detached leaves on *Eucalyptus parramattensis* grown in whole-tree chambers.
+Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C) [see @drake].
 Measurements of g~min~ were carried out at five different temperatures on replicate leaves.
-There is no effect of measurement T on g~min~, but leaves of trees grown in elevated temperature showed consistently lower g~min~.
+There was no consistent effect of measurement T on g~min~, but leaves of trees grown in elevated temperature showed on average a 45% reduction in g~min~ (p < 0.01, mixed-effects model).
 

### 2018-01-10 R/figures.R

@@ -213,7 +213,7 @@ figure_sim <- function(){
   
   with(t0, plot(Tleaf, 1000*GS, type='l', 
                 xlab=expression(T[leaf]~~(degree*C)),
-                ylab=expression(G[S]~~(mmol~m^-2~s^-1)),
+                ylab=expression(g[S]~~(mmol~m^-2~s^-1)),
                 col=col0, lty=lty0,
                 ylim=c(0,400)))
   with(t1, lines(Tleaf, GS, col=col1, lty=lty1))
@@ -478,7 +478,12 @@ figure_sureau <- function(planta, plantb){
   legend("topright", c("leaf","soil"), lty=c(1, 5), lwd=2, bty='n', cex=1.2)
   
   # panel c
-  with(planta, plot(days, PLC*100, xlim=c(0,100), ylab="PLC (%)", type='l', lwd=2, xlab="Days", col=cols[1]))
+  with(planta, 
+       plot(days, PLC*100, xlim=c(0,100), 
+            ylab="PLC (%)", type='l', lwd=2, 
+            xlab="Days", col=cols[1],
+            panel.first= abline(h = 88, lty=2, col="darkgrey"))
+  )
   with(plantb, lines(days, PLC*100, xlim=c(0,100), type='l', lwd=2, col=cols[2]))
   plotlabel("(c)", "bottomright")
   

### 2018-01-10 R_other/stats.R

@@ -1,0 +1,17 @@
+
+library(lme4)
+library(lmerTest)
+
+fit1 <- lmer(log(gmin) ~ growth_T + factor(ch_temp) + (1|chamber), data=wtc4gmin)
+anova(fit1, test = "F")
+car::Anova(fit1)
+
+group_by(wtc4gmin, growth_T) %>%
+  summarize(gmin = mean(gmin))
+
+
+
+group_by(gdfr, method) %>%
+  summarize(gmin = exp(mean(log(gmin))))
+
+table(gdfr$method)

### 2018-01-11 manuscript.Rmd
moved table 2 to suppinfo; suppinfo with references and two plots

@@
 Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) were taken from the compilations by @kerstiens1996 and @schuster2017.
-Data on minimum conductance (g~min~) from MLD was compiled from 44+ original papers (see SuppInfo 2 for a list of papers).
+Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see SuppInfo 2 for references).
 For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources.
@@
 \
-From the literature we found a general acclimatory response of g~min~ to drought stress, nearly always leading to a significant  decrease in g~min~ (see Table 2 for a summary of the literature), with just one exception.
+From the literature we found a general acclimatory response of g~min~ to drought stress, nearly always leading to a significant  decrease in g~min~ (see Table S2 for a quantitative summary of the literature), with just one exception.
+The magnitude of the decrease of g~min~ with acclimation to drought stress varied from -4 to -70%; no attempt at a meta analysis was made since no drought is created equal.
 In each of the studies summarized, plants were grown in well-watered or droughted conditions, and in one case a difference in g~min~ was demonstrated after just four days of drought exposure [@bengtson1978].
@@
 
-```{r, echo=FALSE}
-dat <- read.csv("data/drought_gmin_table.csv", stringsAsFactors = FALSE)
-
-knitr::kable(dat, 
-             caption="**Table 2.** Summary of studies demonstrating a decrease in g~min~ after an acclimation period to low soil water availability",
-             col.names= c("Study","Species","Experiment","Reduction in g~min~"))
-```
+\
+atmospheric drought
+@fanourakis2013 gmin (MLD) much higher in high RH grown plants, at any value of leaf water content.
+This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss.
 
@@
 
-## 5. Other factors
+## 5. Increased minimum conductance with leaf age
 
+@jordan2007 large increase gmin in the herb Agastachys odorata with leaf age.
+See SuppInfo 3 for a summary of six studies on leaf age effects.
+All but one species show an increase in g~min~ with leaf age.
 Also discussed above in Section~(cuticular conductance), the cuticle is constantly exposed to physical (wind, rain) and biotic (insects, fungi) agents which damage the cuticle over time.
 
-@jordan2007 large increase gmin in the herb Agastachys odorata with leaf age. (*Several studies in database that have leaf age effects; summarize?*).
+- do any of those papers say why? summarize
 
@@
 @delucia1984; @herrick1991].
-
-@fanourakis2013 gmin (MLD) much higher in high RH grown plants, at any value of leaf water content (Table 1).
-This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss (Hypothesis 2).
-(*could go to temperature section, as T and RH are correlated in field conditions*).
+@fernandez2017 summarized the literature on this topic and argue that the short growing seasons at high altitude cause 'unripe' cuticles.
 

### 2018-01-11 manuscript_suppinfo.Rmd
moved table 2 to suppinfo; suppinfo with references and two plots

---
title: "Supporting Information"
author: ''
date: ''
output:
  word_document:
    reference_docx: manuscriptstyle.docx
csl: new-phytologist.csl
bibliography: references.bib
---

```{r include=FALSE}

library(knitr)
library(pander)

source("R/load.R")
```

# 1. Literature compilation of g~min~ estimates

\
Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers. For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources. All conductance values were converted to per unit projected surface area. Total surface area for conifer leaves was converted to projected area following @barclay2000. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). In many cases, minimum transpiration rates were reported instead of a conductance (termed 'residual transpiration', 'epidermal transpiration', 'minimum transpiration', etc.). For all papers included in the compilation, we were able to convert these transpiration rates to conductances, but unfortunately dozens of papers could not be included because of missing information on vapour pressure deficit during measurements. All g~min~ estimates from original papers were entered in the database, but for this paper we only used 'control' treatments (filtering for drought, elevated temperature, and various other treatments). Where practical, other reported variables were also entered for each study. The complete database is available from github.com/remkoduursma/gmindatabase.

\
Taxonomic order and family was looked up with the taxize R package [@taxize2013]. Climate variables across the species range were extracted from WorldClim, using species occurrence records from ALA and GBIF. The speciesmap R package was used for this extraction (github.com/remkoduursma/speciesmap).

```{r gminfig1, echo=FALSE, dpi=300, fig.width=6, fig.height=5}
figure_gmin_bygroup(gmindat)
```
**Figure S1**.

```{r gminfig2, echo=FALSE, dpi=300, fig.width=6, fig.height=5}
# climate plot
```
**Figure S2**.

## References

```{r echo=FALSE, eval=FALSE}
 library(RefManageR)
r <- ReadBib("c:/repos/gmindatabase/references/references.bib")
paste(paste0("@", names(r)), collapse="; ")
```

[@anfodillo2002; @araus1991; @arnold2007; @bengtson1978; @boyce2003; @boyce2000; @brodribb2014; @burghardt2003; @cavender-bares2007; @cornwell2007; @drake2013; @vangardingen1991; @grace1974; @hamerlynck1996; @hao2010; @heinsoo1998; @holbrook1992; @iogna2013; @james2008; @jordan2007; @kerstiens1995; @kerstiens1989; @li2012; @muchow1989; @nagarajah1979; @nazrul-islam1983; @pasquet-kok2010; @phillips2010; @pisekberger1938; @rawson1988; @sack2003; @saito2010; @schuster2016; @scoffoni2011; @sinclair1986; @smith2006; @vilagrosa2003; @villar-salvador1999; @walden-coleman2013; @warren2011; @pallardy1993; @newton1996; @fish2009; @rosas-anderson2014]

# 2. Effects of drought stress on g~min~

```{r, echo=FALSE}
dat <- read.csv("data/drought_gmin_table.csv", stringsAsFactors = FALSE)

knitr::kable(dat, 
             caption="**Table S1.** Summary of studies demonstrating a decrease in g~min~ after an acclimation period to low soil water availability",
             col.names= c("Study","Species","Experiment","Reduction in g~min~"))
```

# 2. Effects of leaf age on g~min~

```{r leafagefig, echo=FALSE, dpi=300, fig.width=8, fig.height=6}
figure_leafage(gminall)
```

**Figure S3**. Correlation between g~min~ and leaf age for six studies in the g~min~ database. Leaf age is expressed in years (0 = current year), with the exception of @phillips2010, who measured juvenile and adult foliage on *Eucalyptus*. Citations with species studied are @boyce2000 (*Abies lasiocarpa*, *Picea engelmanii*), @boyce2003 (*Tsuga canadensis*, *Pinus strobus*, *Pinus resinosa*, *Picea rubens*), @heinsoo1998 (*Picea abies*), @jordan2007 (*Agastachys odorata*), @phillips2010 (*Eucalyptus saligna*, *E. sideroxylon*, *E. cladocalyx*).

# References {.unnumbered}

### 2018-01-11 R/figures.R

@@ -489,3 +489,43 @@ figure_sureau <- function(planta, plantb){
   
 }
 
+
+
+figure_leafage <- function(gminall){
+  
+  ii <- which(sapply(gminall, function(x)"leafage" %in% names(x)))
+  datsub <- gminall[ii]
+  
+  sel <- function(x){
+    x <- x[,c("species","leafage","gmin","datasource")]
+    x$leafage <- as.character(x$leafage)
+    x
+  }
+  datsub <- lapply(datsub, sel)
+  
+  dat <- bind_rows(datsub) %>%
+    group_by(species, leafage, datasource) %>%
+    summarize(gmin = mean(gmin, na.rm=TRUE)) %>%
+    ungroup
+  
+  
+  dat$leafage[dat$leafage %in% c("current year","current")] <- "0"
+  dat$leafage[dat$leafage %in% c("one year old","one-year-old")] <- "1"
+  dat$leafage[dat$leafage == "C"] <- "0"
+  dat$leafage[dat$leafage == "C+1"] <- "1"
+  dat$leafage[dat$leafage == "C+2"] <- "2"
+  dat$leafage[dat$leafage == "C+3"] <- "3"
+  
+  dat$leafage <- factor(dat$leafage, levels=c("young","old","0","1","2","3","4","5","10"))
+  
+  ggplot(dat, aes(x=leafage, y=gmin)) +
+    geom_line(aes(group=species)) + 
+    geom_point() + 
+    geom_hline(yintercept=0, alpha=0) + 
+    facet_wrap(~datasource, drop=TRUE, scales= "free") +
+    theme_bw() +
+    scale_colour_manual(values = rainbow(12)) +
+    labs(x="", y=expression(g[min]~~(mmol~m^-2~s^-1)))
+  
+}
+

### 2018-01-11 R/load.R

@@ -7,7 +7,7 @@ pacman::p_load(Hmisc, car, dplyr, tidyr, nlme, nlshelper,
                forcats, tibble, magicaxis, 
                plantecophys, readxl, multcomp,
                reporttools, Taxonstand, taxize, 
-               doBy, stringi, rmarkdown)
+               doBy, stringi, rmarkdown, ggplot2)
 
 
 #devtools::install_github("remkoduursma/speciesmap")

### 2018-01-11 run.R

@@ -6,5 +6,6 @@ source("R/load.R")
 # Conversion to png relies on an installation of ImageMagick (and GhostScript); may fail on your system.
 source("R/make_figures.R")
 
-# Convert manuscript (output is manuscript.docx).
+# Convert manuscript (output is manuscript.docx & manuscript_suppinfo.docx).
 rmarkdown::render("manuscript.Rmd", encoding = "utf8")
+rmarkdown::render("manuscript_suppinfo.Rmd", encoding = "utf8")

### 2018-01-11 manuscript_suppinfo.Rmd
SuppInfo draft complete

@@
 In many cases, minimum transpiration rates were reported instead of a conductance (termed 'residual transpiration', 'epidermal transpiration', 'minimum transpiration', etc.).
-For all papers included in the compilation, we were able to convert these transpiration rates to conductances, but unfortunately dozens of papers could not be included because of missing information on vapour pressure deficit during measurements.
+For all papers included in the compilation, we were able to convert these transpiration rates to conductances, but unfortunately dozens of papers could not be included because of missing information on vapour pressure deficit or relative humidity / air temperature during measurements.
 All g~min~ estimates from original papers were entered in the database, but for this paper we only used 'control' treatments (filtering for drought, elevated temperature, and various other treatments).
@@
 
-```{r gminfig1, echo=FALSE, dpi=300, fig.width=6, fig.height=5}
-figure_gmin_bygroup(gmindat)
-```
-**Figure S1**.
-
-```{r gminfig2, echo=FALSE, dpi=300, fig.width=6, fig.height=5}
-# climate plot
-```
-**Figure S2**.
-
-## References
-
 ```{r echo=FALSE, eval=FALSE}
@@
 
-[@anfodillo2002; @araus1991; @arnold2007; @bengtson1978; @boyce2003; @boyce2000; @brodribb2014; @burghardt2003; @cavender-bares2007; @cornwell2007; @drake2013; @vangardingen1991; @grace1974; @hamerlynck1996; @hao2010; @heinsoo1998; @holbrook1992; @iogna2013; @james2008; @jordan2007; @kerstiens1995; @kerstiens1989; @li2012; @muchow1989; @nagarajah1979; @nazrul-islam1983; @pasquet-kok2010; @phillips2010; @pisekberger1938; @rawson1988; @sack2003; @saito2010; @schuster2016; @scoffoni2011; @sinclair1986; @smith2006; @vilagrosa2003; @villar-salvador1999; @walden-coleman2013; @warren2011; @pallardy1993; @newton1996; @fish2009; @rosas-anderson2014]
+\
+The complete list of references used to construct the database can be found at the end of this document [@anfodillo2002; @araus1991; @arnold2007; @bengtson1978; @boyce2003; @boyce2000; @brodribb2014; @burghardt2003; @cavender-bares2007; @cornwell2007; @drake2013; @vangardingen1991; @grace1974; @hamerlynck1996; @hao2010; @heinsoo1998; @holbrook1992; @iogna2013; @james2008; @jordan2007; @kerstiens1995; @kerstiens1989; @li2012; @muchow1989; @nagarajah1979; @nazrul-islam1983; @pasquet-kok2010; @phillips2010; @pisekberger1938; @rawson1988; @sack2003; @saito2010; @schuster2016a; @scoffoni2011; @sinclair1986; @smith2006; @vilagrosa2003; @villar-salvador1999; @walden-coleman2013; @warren2011; @pallardy1993; @newton1996; @fish2009; @rosas-anderson2014].
+
+```{r gminfig1, echo=FALSE, dpi=300, fig.width=6, fig.height=5}
+figure_gmin_bygroup(gmindat)
+```
+**Figure S1**.
+Average minimum conductance (g~min~) for crops, graminoids, woody and other non-woody plant species.
+Grey symbols are original datapoints.
+Numbers above the figure refer to the number of species in each group.
+Different letters indicate significant differences.
+Error bars are 95% confidence intervals for the mean (calculated for log-transformed g~min~, and back-transformed).
+
+```{r gminfig2, echo=FALSE, dpi=300, fig.width=6, fig.height=5}
+figure_pet(gmindat2)
+```
+**Figure S2**.
+No relationship between g~min~ and potential evapotranspiration across the range of the species.
+Climate variables were calculated across the distribution of each species; we tested many other variables (precipitation, temperature, etc.), but no relationship with climate could be determined.
 
@@
 **Figure S3**.
-Correlation between g~min~ and leaf age for six studies in the g~min~ database.
+Correlation between g~min~ and leaf age for five studies in the g~min~ database.
 Leaf age is expressed in years (0 = current year), with the exception of @phillips2010, who measured juvenile and adult foliage on *Eucalyptus*.

### 2018-01-11 R/figures.R

@@ -529,3 +529,17 @@ figure_leafage <- function(gminall){
   
 }
 
+figure_pet <- function(gmindat2){
+  
+  par(mar=c(4,4,1,1), cex.lab=1.2, mgp=c(2.5, 1, 0), tcl=0.2)
+  with(gmindat2, plot(PET_mean, gmin, 
+                      ylim=c(0,40),
+                      xlim=c(500,2000),
+                      xlab="Potential evapotranspiration (mm)",
+                      ylab=expression(g[min]~~(mmol~m^-2~s^-1)),
+                      pch=c(1,19)[as.factor(Woody)]))
+  legend("topright", c("Non-woody","Woody"), pch=c(1, 19), inset=0.01)
+
+}
+
+

### 2018-01-11 R/load.R

@@ -149,7 +149,7 @@ cropgmin <- read.csv("c:/repos/gmindatabase/combined/cropgmindatabase.csv",
 gminall <- dget("c:/repos/gmindatabase/combined/gminall.rdput")
 
 #
-if(FALSE){
+if(TRUE){
   climfile <- "data/species_climate_wcpet.rds"
   if(!file.exists(climfile)){
     options(zomerpetpath="c:/data/zomerpet", worldclimpath="c:/data/worldclim")

### 2018-01-12 manuscript.Rmd
completed draft without notes or dangles

@@
 ```
-
-# Notes
-
-@fernandez2017 : more discussion of cuticular composition, increase of gmin with altitude ('unripe' cuticles due to short growing season), foliar uptake of water (and adaptation of the cuticle to facilitate that)
-
-Relationship between gmin and yield (to mention in the text???):
-Clarke et al. 1989 . Genotypes with low gmin yielded 40-150% more in 4 out of 5 driest enviroments in 9 years of study.
-Clarke and McCaig (1982 ) low order but significant positive correlation between yield and water retention (low gmin) in genotypes of Triticum
-
-Perhaps include here changes with ontogeny is species with different juvenile and adult leaves?
-Climent et al.
-2006  and Pardos et al.
-2009 - juvenile needles in some pines have higher rate of water loss than adult needles of the same individuals.
-Darrow et al.
-2002  found also that in some New Zealand heteroblastic sp juveniles had a higher gmin than adults but other sp no significant differences.
 
@@
 
-## 3. Acclimation to drought
+## 3. Acclimation to the environment
 
 \
-From the literature we found a general acclimatory response of g~min~ to drought stress, nearly always leading to a significant  decrease in g~min~ (see Table S2 for a quantitative summary of the literature), with just one exception.
-The magnitude of the decrease of g~min~ with acclimation to drought stress varied from -4 to -70%; no attempt at a meta analysis was made since no drought is created equal.
-In each of the studies summarized, plants were grown in well-watered or droughted conditions, and in one case a difference in g~min~ was demonstrated after just four days of drought exposure [@bengtson1978].
-We also demonstrate drought acclimation via a case study on 11 *Hakea* species grown in two watering treatments (Fig.
-7).
+While there is clearly plenty of variation in g~min~ between species, a number of lines of evidence suggest that g~min~ also has great potential to vary within species *wow rewrite*.
+Here we summarize the literature on the acclimation of g~min~ to drought conditions, to changes in temperature and humidity, and the change in g~min~ with leaf age and with altitude.
+This discussion is directly relevant to the use of g~min~ in models, because if intra-specific variation is large, it complicates model parameterization.
 
 \
-A number of studies have attempted to attribute this drought acclimation in g~min~ to a change in the chemical composition of the cuticle.
-@bengtson1978 reported lower g~min~ in six oat varieties in response to drought, and increase in cuticular waxes, but no overall relationship between the two and very different responses between different oat genotypes.
+There is a general tendency for a decreased g~min~ in plants acclimated to drought stress.
+The magnitude of the decrease of g~min~ with acclimation to drought stress varied from -4 to -70% (across 10 studies, see Table S2), with a typical decrease in the order of 30-40%.
+In each of the studies summarized, plants were grown in well-watered or droughted conditions, and in one case a difference in g~min~ was demonstrated after just four days of drought exposure [@bengtson1978].
+In a similar vein, @fanourakis2013 reported much lower g~min~ in plants grown in low vs.
+high relative humidity, again demonstrating that g~min~ is reduced in response to conditions that cause water limitation.
+We also demonstrate drought acclimation via a case study on 11 *Hakea* species grown in two watering treatments (Fig.
+7) - all 11 species showed a decrease in g~min~ in the drought treatment.
+This significant change in g~min~ with drought acclimation is likely an important component of the overall drought hardening of plants.
+
+\
+A number of studies have attempted to attribute the drought-induced acclimation in g~min~ to a change in the chemical composition of the cuticle.
+@bengtson1978 reported lower g~min~ in six oat varieties in response to drought, and an increase in the amount of cuticular waxes, but could not find a relationship between the two and the response was highly genotype-specific.
 However, @premachandra1992 found that epicuticular wax load increased on leaves of non-irrigated Sorghum cultivars and was positively correlated with cuticular conductance and cell membrane stability.
@@
 @bi2017 also reported that drought caused a change in cuticular wax production and chemical composition, but again in a highly genotype-specific way.
-Thus it can be concluded that the cuticle indeed changes substantially after a change in plant available water, but in a complex, species-specific manner that is yet to be connected directly to g~min~.
-
-\
-atmospheric drought
-@fanourakis2013 gmin (MLD) much higher in high RH grown plants, at any value of leaf water content.
-This may suggest that gmin is useful for the plant, and is reduced actively in response to water loss.
+Thus it can be concluded that the cuticle indeed changes substantially after a change in plant available water, but in a complex, species-specific manner that is yet to be connected directly to changes g~min~.
+This opens the possibility that leaky stomata are a significant component of the drought response, not just the cuticular composition.
 
@@
 
-## 4. Acclimation and instantaneous response to temperature
+\
+The response of g~min~ to temperature is more complex; it shows both a response to instantaneous changes in temperature, as well as an acclimatory response to growth temperature.
+@riederer2001 and @schuster2016b have both demonstrated a steep non-linear instantaneous response of g~cut~ to temperature, with the response becoming especially steep at higher temperatures (> 40C).
+This high temperature increase may well be a crucial component of the ability of plants to tolerate heat waves [@drake].
+In support of this idea, @schuster2016a reported on a correlation between thermal tolerance and g~min~ across nine species.
+The mechanism of the temperature response is complex, highly species-specific, and we refer to @schuster2016b for a detailed investigation.
 
-The response of minimum conductance to temperature is also striking; it shows both an instantaneous response to increased temperature, as well as an acclimatory response to growth temperatures.
-@riederer2001 and @schuster2016b have both demonstrated a steep non-linear instantaneous response of g~cut~ to temperature, with the response becoming especially steep at higher temperatures.
+\
+Less well established is the acclimatory potential of g~min~ to changes in growth temperature.
+We present a case study on *Eucalyptus parramattensis* grown in whole-tree chambers.
+The chambers either tracked ambient conditions, or were subjected to a persistent +3C warming treatment.
+After months of growth in the treatments, g~min~ was measured at various instantaneous temperatures, ranging from 17.5 to 27.5C.
+We found a 56% decrease in g~min~ in the elevated temperature trees, but there was no clear pattern with measurement temperature (Fig.
+8).
+This decrease in g~min~ is consistent with the drought response, because again g~min~ is reduced in leaves that are subjected to conditions that increase evaporative demand (vapour pressure deficit was higher in the elevated temperature treatment, see @drake).
+The direction of this response is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time.
+Responses across temperature gradients are more complex, in particular the change in g~min~ with altitude has been well studied.
+Commonly it is reported that g~min~ increases with altitude [e.g.
+@delucia1984; @herrick1991; @anfodillo2002].
+@fernandez2017 discuss altitude responses in detail.
+and argue that the short growing seasons at high altitude are insufficient for complete maturation of leaf cuticles.
 
-@schuster2016a reported on a correlation between thermal tolerance and g~min~ relationship across nine species.
-
-Also see @paje1988 for more complex response to temperature (figure out what he really showed).
-
-Lower g~min~ in the trees acclimated to high temperature is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time.
-Note that this response is not due to a direct effect of temperature on in cuticular conductance, which increases with temperature (Schreiber 2001).
+\
+Finally we mention the striking effect of leaf age on g~min~.
+We have summarized five studies in Fig.
+S3, all of which reported on increased g~min~ for older leaves (but not for all species studied).
+The data reported by @jordan2007 is particularly impressive, as g~min~ in the herb *Agastachys odorata* gradually increased as leaves age.
+The most likely explanation for this change in g~min~ is the continued exposure to wind, rain, and abrasives, which have been shown to damage the cuticle and increase its conductance (see Section III).
+In plants that maintain several cohorts of leaves, finding an appropriate value of g~min~ for use in models will have to take into account the leaf age effect.
 
@@
 
-## 5. Increased minimum conductance with leaf age
-
-@jordan2007 large increase gmin in the herb Agastachys odorata with leaf age.
-See SuppInfo 3 for a summary of six studies on leaf age effects.
-All but one species show an increase in g~min~ with leaf age.
-Also discussed above in Section~(cuticular conductance), the cuticle is constantly exposed to physical (wind, rain) and biotic (insects, fungi) agents which damage the cuticle over time.
-
-- do any of those papers say why? summarize
-
-@anfodillo2002 gmin high altitude.
-Four-fold increase gmin from February (winter) to August in Picea abies, probably due to incomplete stomatal closure not changes in cuticle properties.
-Cite common finding that gmin increases with altitude [e.g.
-@delucia1984; @herrick1991].
-@fernandez2017 summarized the literature on this topic and argue that the short growing seasons at high altitude cause 'unripe' cuticles.
-
 # VIII. Conclusions
 
+\
 In a pioneering publication which presented the Jarvis model of stomatal conductance, @jarvis1976 states that "we have assumed in the following equations that when stomata are closed the leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." Similarly, given the poor statistical properties of g~0~ estimated from regression, the very high values compared to minimum conductance reached for detached leaves, we conclude that g~0~ should not be estimated from regression on leaf gas exchange data for use in models.
-
 Then, how should g~0~ be estimated?
@@
 Thus when modelling night-time or low-light conductance, a different g~0~ should be used than when modelling the drought response of plants.
-Likewise, as briefly mentioned in the introduction, yet a different g~0~ is reached when temperature is so high that photosynthesis does not take place.
-It is not our intention to make models more complex, but certainly we cannot avoid the conclusion that g~0~ should be replaced with a more complicated expression.
 
 \
-It is probably unavoidable to improve the Ball-Berry model with an expression like,
+Our first conclusion is that g~0~, g~min~ and g~dark~ should not be assumed equal in any modelling study.
+Indeed, a productive way forward is to view each of these conductances as distinct, and they should be parameterized separately.
+For example, when considering modelling stomatal conductance with Eq.
+1, we suggest that g~min~ and g~0~ can be combined:
 
@@
 
-where g~min~ is estimated via MLD, representing a true whole-leaf minimum that is reached under drought conditions, and g~0~ is the minimum reached in darkness for well-watered plants (which itself should be a function of soil water availability, e.g.@misson2004).
+where g~min~ is parameterized with data obtained from mass loss of detached leaves (Fig.
+6), representing a true whole-leaf minimum that is reached under drought conditions, and g~0~ is the minimum reached during low photosynthesis (due to low light or other factors) for well-watered plants.
+During drought, g~0~ will decrease [@misson2004], so that the true minimum reached for Eq.
+1 is g~min~.
 
-Our review of g~min~ of detached leaves shows remarkable plasticity to growing conditions.
-The lack of clear species trends, when comparing growing conditions, climate of origin, or growth habit, can perhaps be explained by the very large acclimatory potential of g~min~ to growing conditions.
+\
+It is however an open question on how g~min~ should be parameterized for different species, as we found both large variation between species that could not be explained by traits, and remarkable plasticity to growing conditions.
+Perhaps this plasticity is the reason that g~min~ did not vary predictably between species.
 We conclude as a result that g~min~ is not a "trait" in the sense that it can be meaningfully summarized for a given species, but has to be linked to growing conditions.
@@
 
-- Also: there is little information on water loss rate of non-leaf surfaces [@oren2001], should be focus of more research
-
-# SuppInfo
-
-E. parramattensis:
-We determined g~min~ from detached leaves collected before sunrise from each of 6 *Eucalyptus parramattensis* trees grown under ambient and ambient+3C temperature conditions, respectively.
-Leaves were scanned for leaf area determination before being weighed using a precision balance and attached to a ‘clothesline’ suspended inside controlled chambers set at different temperatures (17.5, 20, 22.5, 25, 27.5°C) and common low light (~200 umol m^-2^ s^-1^) and a relative humidity of 60%.
-Inbuilt fans in each chamber blew air gently across the leaf surfaces.
-Leaves were allowed to slowly dehydrate over a period of 7-8 hours during which time leaf mass was measured roughly every hour.
-For each leaf, transpiration was calculated from the slope of the linear part of the leaf mass vs time relationship, and converted to g~min~ following @sinclair1986.
-
-Hakeas:
-After 8 months of water stress treatment, the epidermal conductance of detached leaves of three plants per species and treatment was estimated by using the rate of water loss.
-Leaves were collected in the early morning and subsequently recut under distilled water, rehydrated in darkness for 6 h at 4 °C and scanned to determine leaf area.
-Leaves were then moved to a controlled environment chamber set at 25 °C, 50% relative humidity and low light (~200 umol) and weighed using a precision balance.
-They were attached to a ‘clothesline’ and allowed to dehydrated slowly.
-Leaves were weighted every 15 min for 3 h and then every hour for 3h more.
-For each sample, the time of stomatal closure was determined as the inflection point of the relationship between leaf mass vs time, and gmin from the slope of the linear part of the curve after stomatal closure, normalized by 2-sided leaf area and adjusted for temperature and humidity within the chamber.
-
 # References {.unnumbered}
-
-Sack, L., Scoffoni, C., PrometheusWiki contributors, 2010.
-"Minimum epidermal conductance (gmin, a.k.a.
-cuticular conductance)," PrometheusWiki, http://www.publish.csiro.au/prometheuswiki/tiki-pagehistory.php?page=Minimum epidermal conductance (gmin, a.k.a.
-cuticular conductance)&preview=7 (accessed March 23, 2017).

### 2018-01-12 manuscript_suppinfo.Rmd
completed draft without notes or dangles

@@
 
-# 2. Effects of drought stress on g~min~
+# 2. Methods description for case studies
+
+**combine and edit**
+
+E. parramattensis:
+We determined g~min~ from detached leaves collected before sunrise from each of 6 *Eucalyptus parramattensis* trees grown under ambient and ambient+3C temperature conditions, respectively.
+Leaves were scanned for leaf area determination before being weighed using a precision balance and attached to a ‘clothesline’ suspended inside controlled chambers set at different temperatures (17.5, 20, 22.5, 25, 27.5°C) and common low light (~200 umol m^-2^ s^-1^) and a relative humidity of 60%.
+Inbuilt fans in each chamber blew air gently across the leaf surfaces.
+Leaves were allowed to slowly dehydrate over a period of 7-8 hours during which time leaf mass was measured roughly every hour.
+For each leaf, transpiration was calculated from the slope of the linear part of the leaf mass vs time relationship, and converted to g~min~ following @sinclair1986.
+
+Hakeas:
+After 8 months of water stress treatment, the epidermal conductance of detached leaves of three plants per species and treatment was estimated by using the rate of water loss.
+Leaves were collected in the early morning and subsequently recut under distilled water, rehydrated in darkness for 6 h at 4 °C and scanned to determine leaf area.
+Leaves were then moved to a controlled environment chamber set at 25 °C, 50% relative humidity and low light (~200 umol) and weighed using a precision balance.
+They were attached to a ‘clothesline’ and allowed to dehydrated slowly.
+Leaves were weighted every 15 min for 3 h and then every hour for 3h more.
+For each sample, the time of stomatal closure was determined as the inflection point of the relationship between leaf mass vs time, and gmin from the slope of the linear part of the curve after stomatal closure, normalized by 2-sided leaf area and adjusted for temperature and humidity within the chamber.
+
+# 3. Effects of drought stress on g~min~
 
@@
 
-# 2. Effects of leaf age on g~min~
+# 4. Effects of leaf age on g~min~
 

### 2018-01-12 run.R

@@ -3,7 +3,8 @@
 source("R/load.R")
 
 # Make figures as PDF (see output/)
-# Conversion to png relies on an installation of ImageMagick (and GhostScript); may fail on your system.
+# Conversion to png relies on an installation of ImageMagick (and GhostScript),
+# may fail on your system (but you can ignore these errors and continue).
 source("R/make_figures.R")
 
 # Convert manuscript (output is manuscript.docx & manuscript_suppinfo.docx).

### 2018-01-15 manuscript.Rmd
more edits; version sent to co-authors 15 Jan

@@
 II. Comparison of various definitions of minimum conductance
-III. Problems with estimating minimum conductance from cuvette data
+III. Problems with estimating minimum conductance from regression
 IV. Use of minimum conductance in models
@@
 \t2. Minimum conductance shows little meaningful interspecific variation
-\t3. Acclimation to drought
-\t4. Acclimation and instantaneous response to temperature
-\t5. Other factors
+\t3. Acclimation to the environment
 VIII. Conclusions", header=F)
@@
 In this review, we illustrate the role of g~min~ in models of leaf gas exchange, critically evaluate sources of variation in g~min~, and quantitatively compare various methods for its estimation.
+We pay special attention to the role of the minimum conductance in the Ball-Berry model of stomatal conductance, and caution against the usual regression-based method for its estimation.
 A literature compilation of g~min~ as measured by weight loss of detached leaves is presented, which shows much variation in this trait that is not clearly related to species groups, climate of origin, or leaf type.
@@
 Thus there is a disconnect between the dizzying level of understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models.
-We aim to improve this connection by synthesizing the state of knowledge, thereby pointing out gaps in understanding, and suggest areas for further research.
+We aim to improve this connection by synthesizing the state of knowledge.
 Our goals are to a) compare various definitions of g~min~ quantitively, which have been previously assumed equal, b) demonstrate the need to include a non-zero g~0~ (Eq.
@@
 Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see SuppInfo 2 for references).
-For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources.
+For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Supporting Information for full details of the compilation).
 All conductance values were converted to per unit projected surface area.
-Units were converted to molar fluxes assuming 41 mol m^-3^ (cf.
-@kerstiens1996).
 Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance (discussed further in Section VI), it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful.
@@
 150 species during night-time.
-Actual minimum daytime values were estimated from the @lin2015 database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset.
-This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.
+Actual minimum daytime values were estimated from the @lin2015 database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the 52 studies included in that dataset.
+This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted or extrapolated g~0~ term.
 
@@
 This result is in line with the methods comparison of @kerstiens1996.
-More recently, @schuster2017 argued, based on a data for 8 species, that g~min~ from MLD is comparable to g~cut~.
+More recently, @schuster2017 argued, based on data for 8 species, that g~min~ from MLD is comparable to g~cut~.
 However in the same study, a literature compilation of hundreds of values showed that g~cut~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cut~ were pooled in their analysis.
@@
 \
-Conductance in the dark (g~dark~) (mean = 41.4 mmol m^-2^ s^-1^ was on average eight-fold higher than g~min~.
+Conductance in the dark (g~dark~) (mean = 41.4 mmol m^-2^ s^-1^) was on average eight-fold higher than g~min~.
 It is no new observation that stomata do not close in the dark (further discussed in Section VI).
@@
 
+\
+(Figure 1 here)
+
+\
 **Figure 1**.
@@
 Error bars are 95\% confidence intervals.
-Grey dots are the original data (but some data occurs outside the figure range).
+Grey dots are the original data (but a few data points occur outside the figure range).
 Different letters denote significant differences (at $\alpha$ = 0.05).
-g~cut,isol~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~0~ : fitted parameter (Eq.
+g~cut~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~0~ : fitted parameter (Eq.
 1) with regression; g~s,low A~ leaf conductance during periods of very low photosynthesis.
@@
 
-# III. Problems with estimating minimum conductance from cuvette data
+# III. Problems with estimating minimum conductance from regression
 
@@
 Exceptions are @duarte2016, who found a lower g~0~ in a heat wave treatment in Douglas-fir (although g~1~ was unaffected), but in their case g~0~ was very accurately estimated by careful multi-point light response curves.
-Another exception is @misson2004, who report on a close negative correlation between pre-dawn leaf water potential and g~0~ in Ponderosa pine (again, the slope was unaffected).
+Another exception is @misson2004, who report on a close negative correlation between pre-dawn leaf water potential and g~0~ in Ponderosa pine (again, the g~1~ parameter was unaffected).
 
@@
 
+\
+(Figure 2 here)
+
+\
 **Figure 2**.
@@
 A poor fit is often obtained when there is little variation in the right-hand side of Eq.
-1, usually because environmental conditions varied little.
+1, usually because there was low variation in environmental conditions (humidity, light, temperature).
 We confirm this by showing that the standard error (SE) of g~0~ increases when the coefficient of variation of the right-hand side of Eq.
@@
 
+\
+(Figure 3 here)
+
+\
 **Figure 3**.
@@
 In this review we focus on the cuticle as a barrier to water loss, but point out that previous work has shown that the cuticle is much less permeable to CO~2~ compared to H~2~O [@boyer1997; @boyer2015].
-This finding, if indeed generally true, has a large effect on the calculation of various gas exchange parameters [@hanson2006].
+This finding, if indeed generally true, has a large effect on the calculation of various gas exchange parameters [@hanson2016].
 @manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is leaky and incompletely closed stomata.
@@
 But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, begging the question whether the same g~0~ is reached in each of these cases.
-In the application of the Ball-Berry models, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005].
+In the application of the Ball-Berry model, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005].
 It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata.
-Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake, @urban2017].
+Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake2018, @urban2017].
 
@@
 A popular approach in thinking about how stomata 'should' respond to environmental drivers is the idea that stomatal conductance is varied to maximize total photosynthesis for a given amount of water use [@cowan1977].
-The consequence is that stomata tend to open during periods that are favorable for photosynthesis (high light, optimal temperature), and close when photosynthesis drops to zero (darkness, very high temperature).
+The consequence is that stomata tend to open during periods that are favourable for photosynthesis (high light, optimal temperature), and close when photosynthesis drops to zero (darkness, very high temperature).
 Clearly, in this framework there is no place for g~0~ since it is always sub-optimal to open stomata (i.e.
@@
 \
-As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) is constant with the photon flux density (PPFD).
+As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) does not vary with the photon flux density (PPFD), as is typical in leaf gas exchange data.
 When PPFD approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~.
@@
 
-Eq. 3 - derived using g~0~ - thus does not give any dependence on PPFD.
+Eq. 3 - derived using zero g~0~ - thus does not give any dependence on PPFD.
 
@@
 In this simulation, T~leaf~ and VPD are assumed to co-vary with an empirical relationship as used by @duursma2014.
-Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increase leaf transpiration E~L~, but at a slightly higher rate than just due to g~0~.
+Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increases leaf transpiration, but at a slightly higher rate than just due to g~0~.
 This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq.
@@
 
+\
+(Figure 4 here)
+
+\
 **Figure 4**.
@@
 \
-When plants are sufficiently water-stressed that stomata are mostly closed, water loss still continues at a rate determined by the minimum conductance.
+When plants are sufficiently water-stressed so that stomata are mostly closed, water loss still continues at a rate determined by the minimum conductance.
 Thus models that aim to predict when plants desiccate and die must include a minimum conductance term.
@@
 
+\
+(Figure 5 here)
+
+\
 **Figure 5**.
@@
 (c) Progression of percent loss conductivity (PLC) of the xylem.
-If we assume that plant mortality occurs at a PLC of 88%, the plants with the low g~min~ lasted 15 days longer (or nearly 40%).
+If we assume that plant mortality occurs at a PLC of 88%, the plants with the low g~min~ lasted 15 days longer.
 
@@
 Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged.
-Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but this variation explains only some of the wide range of g~cut~.
+Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but was able to explain only some of the wide range of g~cut~ measured across different plant species.
 Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.
@@
 \
-A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~) [@barnard2013; @lombardozzi2017], simply because it is a condition where no photosynthesis takes place.
+A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~) [@barnard2013; @lombardozzi2017], simply because it is a condition when net photosynthesis is zero.
 However as our literature comparison showed (Fig.
@@
 In addition, g~dark~ can show a clear response to vapour pressure deficit [@barbour2007] during night-time conditions.
-Clearly g~dark~ is an actively controlled process that cannot be adequately summarized by a single constant g~0~.
+Clearly g~dark~ is an actively controlled process that cannot be adequately summarized by a single constant g~0~, and should be modelled in a separate framework that is yet to be identified.
 
@@
 
+In order to understand variation in g~min~, we have to know more about the role of the cuticle vs.
+leaky or incompletely closed stomata.
 A few studies have directly quantified the stomatal component of g~min~.
@@
 6).
-Comparing all available data, we did not find significant relationships between rainfall, aridity, or temperature (all extracted across the range of the species) and g~min~, nor were there meaningful relationships with other traits (**not shown**).
+Comparing all available data, we did not find significant relationships between g~min~ and climate of origin, nor were there meaningful relationships with other traits (see Supporting Information).
 We thus found it difficult to explain the variation in g~min~ between species.
 A simple breakdown by taxonomic order (Fig.
-6b) revealed that grasses (*Poales*) have a higher g~min~ compared to other orders, and conifers (especially *Pinales*) tended towards the lower range of values.
+6b) revealed that grasses (*Poales*) have a higher g~min~ compared to other orders, and conifers (especially *Pinales*) tended towards the lower range of values (but were only significantly different from *Poales*).
 
@@
 Very few studies have found meaningful correlations between g~min~ and environmental factors or ecological categories.
-Based on their own compilation of the literature, @schuster2017 concluded no significant differences in g~min~ / g~cut~ by plant growth form.
+Based on their own compilation of the literature, @schuster2017 concluded no significant differences in g~min~ or g~cut~ (pooled) by plant growth form.
 One notable exception is @brodribb2014, who showed a correlation between rainfall at species origin (of the driest quarter), and g~min~.
@@
 \
-**crops**
+In crop science, g~min~ has long been identified as a key drought tolerance trait [@sinclair1986].
+A number of studies have targeted g~min~ as a key trait for breeding more drought-tolerant crops, leading to comparisons of g~min~ across genotypes grown in the same conditions.
+For example, @james2008 compared g~min~ in 58 soybean (*Glycine max*) genotypes, and found > 2-fold variation that could not be easily explained by other traits.
+In Fig.
+6, we have compiled a number of studies in crops, demonstrating not just variation in g~min~ between crops, but also the wide range in g~min~ between genotypes.
+Again, it is striking that the wide variation in g~min~ cannot be easily explained by variation in leaf or other traits or chemical and structural components of the cuticle [@bengtson1978; @james2008; @saito2010].
 
@@
 
+\
+(Figure 6 here)
+
+\
 **Figure 6**.
@@
 (b) g~min~ averaged by phylogenetic order (including only the top ten orders in the database).
-Bars are 95/% confidence intervals.
+Bars are 95% confidence intervals.
 Numbers above the figure refer to the number of species.
@@
 \
-While there is clearly plenty of variation in g~min~ between species, a number of lines of evidence suggest that g~min~ also has great potential to vary within species *wow rewrite*.
+While there is clearly plenty of variation in g~min~ between species, a number of lines of evidence suggest that g~min~ also has great potential to vary within species.
 Here we summarize the literature on the acclimation of g~min~ to drought conditions, to changes in temperature and humidity, and the change in g~min~ with leaf age and with altitude.
@@
 \
-There is a general tendency for a decreased g~min~ in plants acclimated to drought stress.
+There is a general tendency for a decreased g~min~ in plants acclimated to drought stress [@james2008].
 The magnitude of the decrease of g~min~ with acclimation to drought stress varied from -4 to -70% (across 10 studies, see Table S2), with a typical decrease in the order of 30-40%.
@@
 We also demonstrate drought acclimation via a case study on 11 *Hakea* species grown in two watering treatments (Fig.
-7) - all 11 species showed a decrease in g~min~ in the drought treatment.
+7) - all 11 species showed a decrease in g~min~ in the drought treatment (see SuppInfo 3 for methods details).
 This significant change in g~min~ with drought acclimation is likely an important component of the overall drought hardening of plants.
@@
 Thus it can be concluded that the cuticle indeed changes substantially after a change in plant available water, but in a complex, species-specific manner that is yet to be connected directly to changes g~min~.
-This opens the possibility that leaky stomata are a significant component of the drought response, not just the cuticular composition.
+This opens up the possibility that leaky stomata are a significant component of the drought response, not just the cuticular composition.
 
@@
 
+\
+(Figure 7 here)
+
+\
 **Figure 7**.
@@
 Bars are labelled by leaf form : broadleaf (B) or needle-like (N), and ordered by g~min~ in the  well-watered treatment.
-For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`).
+For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`.
 
@@
 @riederer2001 and @schuster2016b have both demonstrated a steep non-linear instantaneous response of g~cut~ to temperature, with the response becoming especially steep at higher temperatures (> 40C).
-This high temperature increase may well be a crucial component of the ability of plants to tolerate heat waves [@drake].
-In support of this idea, @schuster2016a reported on a correlation between thermal tolerance and g~min~ across nine species.
 The mechanism of the temperature response is complex, highly species-specific, and we refer to @schuster2016b for a detailed investigation.
+The quick increase in g~min~ at high temperature may well be a crucial component of the ability of plants to tolerate heat waves [@drake2018].
+In support of the link between g~min~ and heat tolerance, @schuster2016a reported on a negative between thermal tolerance and g~min~ across nine species, so that species with improved tolerance to very high temperature had a low g~min~.
 
@@
 Less well established is the acclimatory potential of g~min~ to changes in growth temperature.
-We present a case study on *Eucalyptus parramattensis* grown in whole-tree chambers.
+We present a case study on *Eucalyptus parramattensis* grown in whole-tree chambers (see SuppInfo 3 for methods details).
 The chambers either tracked ambient conditions, or were subjected to a persistent +3C warming treatment.
@@
 8).
-This decrease in g~min~ is consistent with the drought response, because again g~min~ is reduced in leaves that are subjected to conditions that increase evaporative demand (vapour pressure deficit was higher in the elevated temperature treatment, see @drake).
+This decrease in g~min~ is consistent with the drought response, because again g~min~ is reduced in leaves that are subjected to conditions that increase evaporative demand (vapour pressure deficit was higher in the elevated temperature treatment, see @drake2018).
 The direction of this response is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time.
 Responses across temperature gradients are more complex, in particular the change in g~min~ with altitude has been well studied.
-Commonly it is reported that g~min~ increases with altitude [e.g.
+It is commonly  reported that g~min~ increases with altitude [e.g.
 @delucia1984; @herrick1991; @anfodillo2002].
-@fernandez2017 discuss altitude responses in detail.
-and argue that the short growing seasons at high altitude are insufficient for complete maturation of leaf cuticles.
+@fernandez2017 discuss the literature on altitude responses in detail and argue that the short growing seasons at high altitude are insufficient for complete maturation of leaf cuticles.
 
@@
 We have summarized five studies in Fig.
-S3, all of which reported on increased g~min~ for older leaves (but not for all species studied).
-The data reported by @jordan2007 is particularly impressive, as g~min~ in the herb *Agastachys odorata* gradually increased as leaves age.
+S3, all of which reported on increased g~min~ for older leaves for the majority of species studied.
+The data reported by @jordan2007 is particularly impressive, as g~min~ in the herb *Agastachys odorata* gradually increased with leaf age up to ca.
+10 years.
 The most likely explanation for this change in g~min~ is the continued exposure to wind, rain, and abrasives, which have been shown to damage the cuticle and increase its conductance (see Section III).
@@
 
+\
+(Figure 8 here)
+
+\
 **Figure 8**.
 Minimum conductance (g~min~) measured with mass loss of detached leaves on *Eucalyptus parramattensis* grown in whole-tree chambers.
-Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C) [see @drake].
+Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C) [see @drake2018].
 Measurements of g~min~ were carried out at five different temperatures on replicate leaves.
@@
 Then, how should g~0~ be estimated?
-It is clear from our comparisons that there is no single minimum conductance.
-Leaves maintain much higher g~dark~ (itself an actively controlled process) than the minimum obtained for detached leaves.
+It is clear from our review and synthesis of available data that there is no single minimum conductance.
+Leaves maintain much higher g~dark~ (itself an actively controlled process) than the minimum conductance measured on intact detached leaves.
 Thus when modelling night-time or low-light conductance, a different g~0~ should be used than when modelling the drought response of plants.

### 2018-01-15 manuscript_suppinfo.Rmd
more edits; version sent to co-authors 15 Jan

@@
 The speciesmap R package was used for this extraction (github.com/remkoduursma/speciesmap).
+We found no significant relationships between g~min~ and any climate variable.
+See Fig.
+S2 for an example relationship, using here the average potential evapotranspiration across the native range of the species as the predictor variable.
+We also tested leaf traits including leaf mass per area and leaf lifespan (both extracted from TRY), but found no significant relationships (and only 30% of the species overlapped between databases).
 
@@
 
-**combine and edit**
+## Drought acclimation in *Hakea*
 
-E. parramattensis:
-We determined g~min~ from detached leaves collected before sunrise from each of 6 *Eucalyptus parramattensis* trees grown under ambient and ambient+3C temperature conditions, respectively.
-Leaves were scanned for leaf area determination before being weighed using a precision balance and attached to a ‘clothesline’ suspended inside controlled chambers set at different temperatures (17.5, 20, 22.5, 25, 27.5°C) and common low light (~200 umol m^-2^ s^-1^) and a relative humidity of 60%.
-Inbuilt fans in each chamber blew air gently across the leaf surfaces.
-Leaves were allowed to slowly dehydrate over a period of 7-8 hours during which time leaf mass was measured roughly every hour.
-For each leaf, transpiration was calculated from the slope of the linear part of the leaf mass vs time relationship, and converted to g~min~ following @sinclair1986.
-
-Hakeas:
+\
 After 8 months of water stress treatment, the epidermal conductance of detached leaves of three plants per species and treatment was estimated by using the rate of water loss.
@@
 Leaves were weighted every 15 min for 3 h and then every hour for 3h more.
-For each sample, the time of stomatal closure was determined as the inflection point of the relationship between leaf mass vs time, and gmin from the slope of the linear part of the curve after stomatal closure, normalized by 2-sided leaf area and adjusted for temperature and humidity within the chamber.
+For each sample, the time of stomatal closure was determined as the inflection point of the relationship between leaf mass vs time, and g~min~ from the slope of the linear part of the curve after stomatal closure.
+
+## Temperature acclimation in *Eucalyptus parramattensis*
+
+\
+We determined g~min~ from detached leaves collected before sunrise from each of 6 *Eucalyptus parramattensis* trees grown under ambient and ambient+3C temperature conditions, respectively.
+Leaves were scanned for leaf area determination before being weighed using a precision balance and attached to a ‘clothesline’ suspended inside controlled chambers set at different temperatures (17.5, 20, 22.5, 25, 27.5°C) and common low light (~200 umol m^-2^ s^-1^) and a relative humidity of 60%.
+Inbuilt fans in each chamber blew air gently across the leaf surfaces.
+Leaves were allowed to slowly dehydrate over a period of 7-8 hours during which time leaf mass was measured roughly every hour.
+For each leaf, transpiration was calculated from the slope of the linear part of the leaf mass vs time relationship, and converted to g~min~.
 

### 2018-01-25 manuscript.Rmd
added chris's methods bit and a new SI figure

@@
 ```
+
+# Notes
+
+@xu1995 lower cuticular transpiration in droughted tomato plants, and increased wax content. Similar response to high salinity, which also reduces water availability.
+
+@sack2003 lower gmin in sun leaves (for 2 out of 4 species), again in the same direction as evaporative demand for sun leaves is frequently higher.
+
+@mengel1989 increased gmin in P. abies after treatment with acidic fog, the epicuticular wax layer showing a 'melted appearance'.
 
@@
 Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time.
-This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants during severe drought, and includes both surfaces of the leaf - not just the stomata-free side which is typically used for cuticular conductance measurements.
+This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants during severe drought, and includes both surfaces of the leaf - not just the stomata-free side.
 A comprehensive compilation of estimates of g~min~ is currently lacking, and is needed not just to parameterize models, but also to study sources of variation in this overlooked plant trait.
@@
 Our goals are to a) compare various definitions of g~min~ quantitively, which have been previously assumed equal, b) demonstrate the need to include a non-zero g~0~ (Eq.
-1), for models of water-use efficiency as well as the drought response, c) present a new compilation of minimum conductance measurements, and demonstrate that the large acclimatory potential of this trait to environmental drivers.
+1), for models of water-use efficiency as well as the drought response, c) present a new compilation of minimum conductance measurements, and demonstrate the large acclimatory potential of this trait to environmental drivers.
 
@@
 In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out.
-Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after a while a constant low rate is achieved.
+Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after some time a constant low rate is achieved.
 From this minimum transpiration rate (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), the minimum conductance (g~min~) can be estimated using measured vapour pressure deficit [@sinclair1986].
 This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress.
+We do not discuss the methodology of this measurement in detail, but note that care needs to be taken to yield reliable estimates (see SuppInfo for more discussion and illustration).
 
@@
 1) from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model.
-Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) were taken from the compilations by @kerstiens1996 and @schuster2017.
+Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but were not used since they included no metadata).
 Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see SuppInfo 2 for references).
@@
 However as our literature comparison showed (Fig.
-1), g~dark~ is substantially higher compared to any other measure of minimum conductance.
+1), g~dark~ is substantially higher compared to any other measure of minimum conductance (except fitted g~0~).
 The simple idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.
@@
 A few studies have directly quantified the stomatal component of g~min~.
-A detailed study of *Hedera helix*  concluded that 35\% of water loss occurred across the stomatal pores (though closed), and 65\% across the non-stomatal part of the cuticle, despite the fact that the stomata presumably covered only a small fraction of the leaf [@santrucek2004].
+A detailed study of *Hedera helix*  concluded that 35% of water loss occurred across the stomatal pores (though closed), and 65% across the non-stomatal part of the cuticle, despite the fact that the stomata presumably covered only a small fraction of the leaf [@santrucek2004].
 For ten Sorghum genotypes, @muchow1989 reported a strong positive correlation between stomatal density and g~min~, suggesting leaky stomata contribute substantially to g~min~.
-For seven conifer species, @brodribb2014 reported that 50 - 94\% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky.
+For seven conifer species, @brodribb2014 reported that 50 - 94% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky.
 However @santrucek2004 found that the cuticle of the stomatal side (but excluding stomata itself) in *Hedera helix* was many times more more permeable than the non-stomatal side, thus providing another potential explanation for this difference.

### 2018-01-25 manuscript_suppinfo.Rmd
added chris's methods bit and a new SI figure

@@
 
-# 2. Methods description for case studies
+# 2. Methodological issues with measurement of g~min~
+
+For measurements of gmin in detached leaves, the relationship between the change in mass (due to water loss) and time is typically curvilinear and characterised by three phases: 1)the rate of water loss initially increases as stomata temporally open in response to being placed under light; 2) the rate of water loss declines as stomata begin to close in response to increasing leaf water deficit; and 3) once stomata have closed (as much as possible) the rate of water loss is at minimum and is constant over time until leaf relative water content nears zero.
+In this case, g~min~ is calculated from the linear portion of the curve during phase 3.
+To achieve this type of response, it is important to ensure that the environmental conditions under which g~min~ is measured are conducive to stomatal closure.
+From experience, leaves placed under high light and/or high VPD conditions are often unable to effectively close stomata in response to desiccation, possibly as a result of rapid epidermal drying and hydro-passive stomatal opening [@buckley2005].
+These conditions lead to a rapid decline in leaf water status and an overestimation of g~min~, as compared to control leaves from the same species.
+
+\
+Additionally, sometimes the stomatal response of replicate leaves from the same species can be highly variable, even under the same environmental conditions (see Figure S3).
+In light of these potential issues, we recommend measurements of g~min~ on detached leaves to be conducted under mild conditions (low light and low VPD), or at least that the measurement conditions are tested for different species and adjusted if necessary.
+We also note that previous studies have shown that g~min~ decreases substantially with very low relative water content [@gardingen1992]; again it is beneficial to dry the leaf slowly.
+
+```{r goodbadfig, echo=FALSE, dpi=300, fig.width=8, fig.height=4}
+figure_goodandbad(goodbadcurves)
+```
+**Figure S3**.
+Two example curves for determination of minimum conductance (g~min~) via mass loss of detached leaves (MLD).
+Leaf weights are expressed as relative water content (RWC).
+The good quality curve shows a distinct linear portion at the end of the drying curve, from which minimum transpiration can be calculated (using actual leaf weight, and leaf area).
+The poor quality curve shows a quicker decrease towards the end of the drying curve (> 400min), indicating an *increased* transpiration rate despite the low water content.
+We speculate that this increase - which we have observed in a number of curves - is due to the epidermis drying so quickly that the stomata are 'pulled open', increasing water loss.
+
+# 3. Methods description for case studies
 
@@
 
-# 3. Effects of drought stress on g~min~
+# 4. Effects of drought stress on g~min~
 
@@
 
-# 4. Effects of leaf age on g~min~
+# 5. Effects of leaf age on g~min~
 
@@
 
-**Figure S3**.
+**Figure S4**.
 Correlation between g~min~ and leaf age for five studies in the g~min~ database.

### 2018-01-25 R/figures.R

@@ -543,3 +543,26 @@ figure_pet <- function(gmindat2){
 }
 
 
+figure_goodandbad <- function(dat){
+  par(mfrow=c(1,2), mar=c(4,4,1,1), 
+      mgp=c(2.5, 0.5, 0), cex.lab=1.1, cex.axis=0.9)
+  
+  with(dat, plot(time_mins, RWC_good.leaf, 
+                 ylim=c(0,1),
+                 xlab="Time (mins)",
+                 panel.first={
+                   datsub <- subset(dat, time_mins > 200)
+                   abline(lm(RWC_good.leaf ~ time_mins, data=datsub), lty=5)
+                 },
+                 ylab="Relative water content",
+                 type='o', pch=21, bg="white"))
+  
+  title(main="Good quality curve", font.main=3, adj=0, cex.main=0.8)
+  with(dat, plot(time_mins, RWC_bad.leaf, 
+                 xlab="Time (mins)",
+                 ylab="Relative water content",
+                 ylim=c(0,1),
+                 type='o', pch=21, bg="white"))
+  title(main="Poor quality curve", font.main=3, adj=0, cex.main=0.8)
+
+}

### 2018-01-25 R/load.R

@@ -216,6 +216,8 @@ planta$days <- (1:nrow(planta))/10
 plantb$days <- (1:nrow(plantb))/10
 
 
+# Chris's example drying curves.
+goodbadcurves <- read.csv("data/good and bad gmin.csv")
 
 
 

### 2018-01-31 manuscript.Rmd
LP in ackno
updated sureau figure

@@
 The *Sureau* model simulates water transport in the soil-plant-atmosphere continuum, and includes a detailed representation of capacitance in stem and leaf tissues.
-(a) Soil relative extractable water (REW; 1 = field capacity, 0 = permanent wilting point) for the two simulations, using a minimum conductance (g~min~) of 1.6 or 3.3 mmol m^-2^ s^-1^ - all other parameters were equal.
+(a) Soil relative extractable water (REW; 1 = field capacity, 0 = permanent wilting point) for the two simulations, using a minimum conductance (g~min~) of 2 or 4 mmol m^-2^ s^-1^ - all other parameters were equal.
 (b) Water potential in the soil and leaf as the dry-down progresses.
-Note the brief increase in soil water potential around days 40 - 70; this represents water released from stem cavitation.
 (c) Progression of percent loss conductivity (PLC) of the xylem.
-If we assume that plant mortality occurs at a PLC of 88%, the plants with the low g~min~ lasted 15 days longer.
+Dashed line is at a PLC of 88%, indicating possible mortality.
 
@@
 Perhaps this plasticity is the reason that g~min~ did not vary predictably between species.
-We conclude as a result that g~min~ is not a "trait" in the sense that it can be meaningfully summarized for a given species, but has to be linked to growing conditions.
 Future studies should compare g~min~ on many species grown in the same conditions to better understand the adaptive value of the minimum water loss rate of leaves.
 
+# Acknowledgments
+
+This work was made possible by the ARC Linkage Project LP140100232.
+
 # References {.unnumbered}

### 2018-01-31 R/figures.R

@@ -461,30 +461,30 @@ figure_sureau <- function(planta, plantb){
   cols <- c(rgb(230, 159, 0, maxColorValue = 255), rgb(0, 114, 178, maxColorValue = 255))
   
   # panel a
-  with(planta, plot(days, REW, lwd=2, ylim=c(0,1), xlim=c(0,100), type='l', xlab="Days", col=cols[1]))
-  with(plantb, lines(days, REW, lwd=2, ylim=c(0,1), xlim=c(0,100), col=cols[2]))
+  with(planta, plot(Days, REW, lwd=2, ylim=c(0,1), xlim=c(0,80), type='l', xlab="Days", col=cols[1]))
+  with(plantb, lines(Days, REW, lwd=2, ylim=c(0,1), xlim=c(0,80), col=cols[2]))
   plotlabel("(a)", "bottomright")
-  legend("topright", c("3.3", "1.6"), title=expression(g[min]), bty='n', lty=1, col=cols, lwd=2, cex=1.2)
+  legend("topright", c("2", "4"), title=expression(g[min]), bty='n', lty=1, col=cols, lwd=2, cex=1.2)
   
   # panel b
-  with(planta, plot(days, Pmin, ylim=c(-7,0), xlim=c(0,100), 
+  with(planta, plot(Days, P_leaf_symplasmic, ylim=c(-7,0), xlim=c(0,80), 
                     col=cols[1],
                     ylab= "Water potential (MPa)",
                     type='l', lwd=2, lty=1, xlab="Days"))
-  with(planta, lines(days, Psoil, lwd=2, col=cols[1], lty=5))
-  with(plantb, lines(days, Pmin, type='l', lwd=2, lty=1, col=cols[2]))
-  with(plantb, lines(days, Psoil, lwd=2, col=cols[2], lty=5))
+  with(planta, lines(Days, P_soil, lwd=2, col=cols[1], lty=5))
+  with(plantb, lines(Days, P_leaf_s, type='l', lwd=2, lty=1, col=cols[2]))
+  with(plantb, lines(Days, P_soil, lwd=2, col=cols[2], lty=5))
   plotlabel("(b)", "bottomright")
   legend("topright", c("leaf","soil"), lty=c(1, 5), lwd=2, bty='n', cex=1.2)
   
   # panel c
   with(planta, 
-       plot(days, PLC*100, xlim=c(0,100), 
+       plot(Days, PLC_branch, xlim=c(0,80), 
[…]

### 2018-01-31 R/load.R

@@ -209,11 +209,9 @@ gdfr <- bind_rows(gmindat_simple, kerst_simple, schuster2017, lombar, g0s, minag
   mutate(method = factor(method, levels=c("gcut_isol","gmin","gnight", "g0","gslowA")))
 
 
-# Nico's simulations with Sureau
-planta <- read.csv2("data/PlantA-DiurnalOut.csv", dec=".")
-plantb <- read.csv2("data/PlantB-DiurnalOut.csv", dec=".")
-planta$days <- (1:nrow(planta))/10
-plantb$days <- (1:nrow(plantb))/10
+# Herve's simulations with Sureau
+planta <- read.csv("data/plant_a_sureau.csv", skip=1)
+plantb <- read.csv("data/plant_b_sureau.csv", skip=1)
 
 
 # Chris's example drying curves.

### 2018-02-02 manuscript.Rmd
minor edits; adjusted figures; eamus reference

@@
 # Notes
+
+- bel looks for data on low PAR; maybe replace gslowA with gslowPAR, so we can argue that g0 is much higher than gslowPAR. 
+- some data behind figure 4a (and maybe Arnaud 4d?)
+- min(g0, gs) ('Uddling', has he used it?)  vs. g0 + gs.
+
+We could say that plant dilemma is to be both impermeable to water loss (because they are saturated with water and leave are in an extremely dry air environment and hence need a external layer to prevent dessication= cuticle) AND need to be permeable to CO2 to grow.
+The dilemma is that leaves cannot be permeable to CO2 and not to H20.
+Hence the evolution of stomata that disrupt the cuticle layer enabling CO2 entre but cannot avoid some CO2 exit but can also close to limit H20.
+loss.
+
+  It could be interested to discuss somewhere in the text the relevance of the conductance (gmin) versus the flow (Emin).
+Obviously gmin is more an intrinsic property of the leaf but what matters for the leaf water relation of the flow.
+In terms of drought resistance and water balance (and water relations) Ecuti is more relevant than gmin I think.
+At night for instance gdark is high but vpd so the flow might not be so different than at gmin during the day.
+
+finding of Schuster e al (2016) who found a high proportion of triterpenoid in the cuticule of a desertic thermotolerant species. This compound would act as to stabilize the cuticule.
 
@@
 | ------------- | ------------------------------------------------------------------------------------------------- |
-| g~cut~        | Conductance of an isolated non-stomatal cuticle.                                                  |
+| g~cuti~        | Conductance of an isolated non-stomatal cuticle.                                                  |
 | g~min~        | Minimum conductance measured from weight loss of detached leaves.                                 |
@@
 The absolute minimum conductance attained by leaves is that through the cuticle only.
-Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001], via special gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata [@kerstiens1996].
+Measurements of cuticular conductance (g~cuti~) are typically made on isolated non-stomatal cuticles [@riederer2001], via special gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata [@kerstiens1996].
 We further discuss cuticular conductance in Section V.
@@
 1) from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model.
-Measurements of conductance of isolated, non-stomatal cuticles (g~cut~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but were not used since they included no metadata).
+Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but were not used since they included no metadata).
 Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see SuppInfo 2 for references).
@@
 This result is in line with the methods comparison of @kerstiens1996.
-More recently, @schuster2017 argued, based on data for 8 species, that g~min~ from MLD is comparable to g~cut~.
-However in the same study, a literature compilation of hundreds of values showed that g~cut~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cut~ were pooled in their analysis.
-Nonetheless, the comparison between g~cut~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions.
+More recently, @schuster2017 argued, based on data for 8 species, that g~min~ from MLD is comparable to g~cuti~.
+However in the same study, a literature compilation of hundreds of values showed that g~cuti~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cuti~ were pooled in their analysis.
+Nonetheless, the comparison between g~cuti~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cuti~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions.
 @boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.
@@
 Different letters denote significant differences (at $\alpha$ = 0.05).
-g~cut~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~0~ : fitted parameter (Eq.
+g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~0~ : fitted parameter (Eq.
 1) with regression; g~s,low A~ leaf conductance during periods of very low photosynthesis.
@@
 1 after it has been coupled to the Farquhar-von Caemmerer-Berry model of photosynthesis (the so-called 'coupled leaf gas exchange model').
-@leuning1990 defined g~0~ as g~s~ when A approaches zero as PPFD approaches the light compensation point, and pointed out that g~0~ is necessary to simulate the increase in A/g~s~ (and C~i~) at low light [see also @leuning1995].
+@leuning1990 defined g~0~ as g~s~ when A~n~ approaches zero as PPFD approaches the light compensation point, and pointed out that g~0~ is necessary to simulate the increase in A~n~/g~s~ (and C~i~) at low light [see also @leuning1995].
 Similarly, @collatz1992 showed that including g~0~ affects the response of g~s~ to relative humidity in the coupled leaf gas exchange model.
@@
 @manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is leaky and incompletely closed stomata.
-Because g~0~ includes not just the cuticle, but also incompletely closed stomata, and the finding that g~0~ is much larger than estimates of g~min~ or g~cut~ (Fig.
+Because g~0~ includes not just the cuticle, but also incompletely closed stomata, and the finding that g~0~ is much larger than estimates of g~min~ or g~cuti~ (Fig.
 1), we ignore this effect in the model simulations that follow.
@@
 In the application of the Ball-Berry model, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005].
-It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata.
+It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cuti~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata.
 Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake2018, @urban2017].
@@
 (b) The same simulations as in (a), but showing the intercellular CO~2~ concentration (C~i~).
-(c) Stomatal conductance (g~s~) simulations, where vapour pressure deficit (VPD) and air temperature (T~air~) were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions.
+(c) Leaf transpiration (E~L~) simulations, where vapour pressure deficit (VPD) and air temperature (T~air~) were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions.
 (d) The same simulations as in (c), but showing C~i~.
@@
 The main barrier to diffusion is actually a very thin layer at the leaf surface.
-Because most of the resistance is located in such a thin layer, g~cut~ does not correlate with the thickness of the cuticle [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b].
+Because most of the resistance is located in such a thin layer, g~cuti~ does not correlate with the thickness of the cuticle [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b].
 Nonetheless there is considerable variation in cuticle thickness between plant species [@schuster2016a], as well as within species in response to drought (**cite**), along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010].
@@
 Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged.
-Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but was able to explain only some of the wide range of g~cut~ measured across different plant species.
-Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.
+Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but was able to explain only some of the wide range of g~cuti~ measured across different plant species.
+Similarly, @hauke1998 found no relationship between the gradual decrease in g~cuti~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.
 
@@
 However as our literature comparison showed (Fig.
-1), g~dark~ is substantially higher compared to any other measure of minimum conductance (except fitted g~0~).
+1), g~dark~ is substantially higher compared to all other variables except fitted g~0~.
 The simple idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.
@@
 Very few studies have found meaningful correlations between g~min~ and environmental factors or ecological categories.
-Based on their own compilation of the literature, @schuster2017 concluded no significant differences in g~min~ or g~cut~ (pooled) by plant growth form.
+Based on their own compilation of the literature, @schuster2017 concluded no significant differences in g~min~ or g~cuti~ (pooled) by plant growth form.
 One notable exception is @brodribb2014, who showed a correlation between rainfall at species origin (of the driest quarter), and g~min~.
@@
 \
-While there is clearly plenty of variation in g~min~ between species, a number of lines of evidence suggest that g~min~ also has great potential to vary within species.
+While there is clearly plenty of variation in g~min~ between species, a number of lines of evidence suggest that g~min~ also has great potential for plasticity.
 Here we summarize the literature on the acclimation of g~min~ to drought conditions, to changes in temperature and humidity, and the change in g~min~ with leaf age and with altitude.
-This discussion is directly relevant to the use of g~min~ in models, because if intra-specific variation is large, it complicates model parameterization.
+This discussion is directly relevant to the use of g~min~ in models, because if the degree of plasticity is large, it complicates model parameterization.
 
@@
 Thus it can be concluded that the cuticle indeed changes substantially after a change in plant available water, but in a complex, species-specific manner that is yet to be connected directly to changes g~min~.
-This opens up the possibility that leaky stomata are a significant component of the drought response, not just the cuticular composition.
+This opens up the possibility that leaky stomata are a significant component of the drought response, not just the composition of the cuticle.
 
@@
 The response of g~min~ to temperature is more complex; it shows both a response to instantaneous changes in temperature, as well as an acclimatory response to growth temperature.
-@riederer2001 and @schuster2016b have both demonstrated a steep non-linear instantaneous response of g~cut~ to temperature, with the response becoming especially steep at higher temperatures (> 40C).
-The mechanism of the temperature response is complex, highly species-specific, and we refer to @schuster2016b for a detailed investigation.
+@riederer2001 and @schuster2016b have both demonstrated a steep non-linear instantaneous response of g~cuti~ to temperature, with the response becoming especially steep at higher temperatures (> 40C).
+In *Eucalyptus haemastoma*, the response of g~min~ to temperature was so steep, that the proportion of cuticular to total transpiration increased from 2-3% at 20C, to 40% at 38C [@eamus2008].
+The mechanism of the instantaneous temperature response is complex, highly species-specific, and we refer to @schuster2016b for a detailed investigation.
 The quick increase in g~min~ at high temperature may well be a crucial component of the ability of plants to tolerate heat waves [@drake2018].
@@
 The chambers either tracked ambient conditions, or were subjected to a persistent +3C warming treatment.
-After months of growth in the treatments, g~min~ was measured at various instantaneous temperatures, ranging from 17.5 to 27.5C.
+After several months of growth in the treatments, g~min~ was measured at various temperatures, ranging from 17.5 to 27.5C.
 We found a 56% decrease in g~min~ in the elevated temperature trees, but there was no clear pattern with measurement temperature (Fig.
@@
 
-$$ g_s = g_{min} + (g_0 - g_{min}) + g_1 \frac{A_n}{C_a}f(D)$$
+$$ g_s = min(g_{min}, g_0) + g_1 \frac{A_n}{C_a}f(D)$$
 

### 2018-02-02 R/figures.R

@@ -75,7 +75,7 @@ figure_g0g1_cor <- function(lin2015, group, legend=FALSE){
                xlim=c(0,0.04),
                ylim=c(0,150),
                panel.first=add_regres_line(fit),
-               xlab=expression(A/(C[a]*sqrt(D))),
+               xlab=expression(A[n]/(C[a]*sqrt(D))),
                ylab=expression(g[s]~(mmol~m^-2~s^-1)),
                pch=16, cex=0.8, col="dimgrey"))
   box()
@@ -211,13 +211,13 @@ figure_sim <- function(){
                  g1=g1, g0=g0_2, Vcmax=70, Jmax=140,
                  VPD=vpds, Tleaf=tleafs, Ca=400)
   
-  with(t0, plot(Tleaf, 1000*GS, type='l', 
+  with(t0, plot(Tleaf, ELEAF, type='l', 
                 xlab=expression(T[leaf]~~(degree*C)),
-                ylab=expression(g[S]~~(mmol~m^-2~s^-1)),
+                ylab=expression(E[L]~~(mmol~m^-2~s^-1)),
                 col=col0, lty=lty0,
-                ylim=c(0,400)))
-  with(t1, lines(Tleaf, GS, col=col1, lty=lty1))
-  with(t2, lines(Tleaf, GS, col=col2, lty=lty2))
+                ylim=c(0,8)))
+  with(t1, lines(Tleaf, ELEAF, col=col1, lty=lty1))
+  with(t2, lines(Tleaf, ELEAF, col=col2, lty=lty2))
   plotlabel("(c)", "topright")
   
   with(t0, plot(VPD, Ci, type='l', 

### 2018-02-02 R/load.R

@@ -40,6 +40,7 @@ lin2015a <- group_by(lin2015, fitgroup) %>%
 # Lin 2015 fits
 lin2015coef <- fits_lin2015(lin2015a)
 
+lin2015coef <- subset(lin2015coef, R2 > 0.5)
 
 # Kerstiens 1996
 kerst <- read.csv("data/kerstiens1996_table1.csv", stringsAsFactors = FALSE) %>%

### 2018-02-02 R/make_figures.R

@@ -22,3 +22,4 @@ to.pdf(figure_wtc4_gmin_2(wtc4gmin), "output/figure8.pdf", height=4, width=3.5)
 
 convert_all_pdf("output")
 
+

### 2018-02-13 manuscript.Rmd
workflow tidied; many edits especially introduction; suppinfo guidelines followed

@@
 
-# Notes
-
-- bel looks for data on low PAR; maybe replace gslowA with gslowPAR, so we can argue that g0 is much higher than gslowPAR. 
-- some data behind figure 4a (and maybe Arnaud 4d?)
-- min(g0, gs) ('Uddling', has he used it?)  vs. g0 + gs.
-
-We could say that plant dilemma is to be both impermeable to water loss (because they are saturated with water and leave are in an extremely dry air environment and hence need a external layer to prevent dessication= cuticle) AND need to be permeable to CO2 to grow.
-The dilemma is that leaves cannot be permeable to CO2 and not to H20.
-Hence the evolution of stomata that disrupt the cuticle layer enabling CO2 entre but cannot avoid some CO2 exit but can also close to limit H20.
-loss.
-
-  It could be interested to discuss somewhere in the text the relevance of the conductance (gmin) versus the flow (Emin).
-Obviously gmin is more an intrinsic property of the leaf but what matters for the leaf water relation of the flow.
-In terms of drought resistance and water balance (and water relations) Ecuti is more relevant than gmin I think.
-At night for instance gdark is high but vpd so the flow might not be so different than at gmin during the day.
-
-finding of Schuster e al (2016) who found a high proportion of triterpenoid in the cuticule of a desertic thermotolerant species. This compound would act as to stabilize the cuticule.
-
-@xu1995 lower cuticular transpiration in droughted tomato plants, and increased wax content. Similar response to high salinity, which also reduces water availability.
-
-@sack2003 lower gmin in sun leaves (for 2 out of 4 species), again in the same direction as evaporative demand for sun leaves is frequently higher.
-
-@mengel1989 increased gmin in P. abies after treatment with acidic fog, the epicuticular wax layer showing a 'melted appearance'.
-
 **Contents:**
@@
 \
-Stomata exert a strong control on the transpiration rate of leaves [@bange1953].
-It is well known that stomata respond quickly to changes in light intensity, humidity and CO~2~ concentration at the leaf surface.
+Plants face a dilemma in constructing leaves that minimize water loss while allowing the uptake of CO~2~.
+Since a membrane that is permeable to CO~2~ but not H~2~O has never evolved, all land plants have stomata in their leaves, which disrupt the cuticle and allow CO~2~ uptake.
+It is well known that stomatal open and close in response to changes in light intensity, humidity and CO~2~ concentration at the leaf surface.
 In doing so, stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus can be expected to close stomata when photosynthetic rates are low.
 To avoid desiccation and ultimate death, stomata can also be expected to close during periods of severe water stress.
-Plants cannot, however, completely halt water loss even when stomata are closed.
+When stomata are closed, water loss continues at a greatly diminished rate through the cuticle.
+After accounting for evaporative demand, this rate of water loss is expressed as the minimum conductance of a leaf.
 There is increasing recognition that the minimum conductance plays an important role in estimating water fluxes in plant canopies [@barnard2013], during heat waves [@kala2016], and in models of plant drought response [@blackman2016; @martinstpaul2017].
@@
 \
-Most current-generation process-based global vegetation models (GVMs) use a version of the Ball-Berry class of stomatal conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity (Eq.
+The absolute minimum attainable water loss rate is through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf.
+Other measurements allow for the estimation of water loss through incompletely closed and broken stomata.
+Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time.
+This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants during severe drought, and includes both surfaces of the leaf - not just the stomata-free side.
+A comprehensive compilation of estimates of g~min~ is currently lacking, and is needed not just to parameterize models, but also to study sources of variation in this overlooked plant trait.
+
+\
+In this review, we discuss the role of the minimum conductance in models of plant water use, and critically evaluate sources of variation in this overlooked parameter.
+In this context, it is useful to briefly summarize the approach taken by the majority of vegetation model.
+Most current-generation process-based global vegetation models (GVMs) use a version of the Ball-Berry class of 'stomatal' (technically including both stomata and the cuticle) conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity (Eq.
 1).
@@
 1 with leaf gas exchange data, but this may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.
-
-\
-In this review, we demonstrate that inclusion of a minimum conductance is necessary both for simulation of the response of water-use efficiency to various environmental drivers, as well as a component of the response of plants to severe drought.
-However, there is no guidance at the moment on how to estimate minimum conductance for use in models, how various definitions and measurements compare quantitatively, whether it is feasible to assign species-specific values of the minimum conductance, and what are the ecological and environmental controls on g~min~ [@schuster2017].
-
-\
-The absolute minimum attainable water loss rate is through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf.
-Other measurements allow for the estimation of water loss through incompletely closed and broken stomata.
-Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time.
-This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants during severe drought, and includes both surfaces of the leaf - not just the stomata-free side.
-A comprehensive compilation of estimates of g~min~ is currently lacking, and is needed not just to parameterize models, but also to study sources of variation in this overlooked plant trait.
 
@@
 This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress.
-We do not discuss the methodology of this measurement in detail, but note that care needs to be taken to yield reliable estimates (see SuppInfo for more discussion and illustration).
+We do not discuss the methodology of this measurement in detail, but note that care needs to be taken to yield reliable estimates (see Methods S2 for further information).
 
@@
 Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but were not used since they included no metadata).
-Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see SuppInfo 2 for references).
+Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see Methods S1 for description and references).
 For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Supporting Information for full details of the compilation).
@@
 Because most of the resistance is located in such a thin layer, g~cuti~ does not correlate with the thickness of the cuticle [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b].
-Nonetheless there is considerable variation in cuticle thickness between plant species [@schuster2016a], as well as within species in response to drought (**cite**), along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010].
+Nonetheless there is considerable variation in cuticle thickness between plant species [@schuster2016a], along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010].
 If the cuticle thickness does not directly affect its conductance, what is the value of increasing the cuticle?
@@
 \
-It is important to bear in mind the many other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, various interactions with insects including signalling and herbivory resistance [@mueller2008; @kerstiens1996c;@riederer2008], and foliar water uptake [@fernandez2017] The thickness of the cuticle may also confer mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness).
-As such, thick cuticles may be advantageous because prolonged exposure to wind increases the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991], by dislodging cuticular compounds by abrasion [@rogge1993].
+It is important to bear in mind the many other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, various interactions with insects including signalling and herbivory resistance [@mueller2008; @kerstiens1996c;@riederer2008], and foliar water uptake [@fernandez2017].
+The thickness of the cuticle may also confer mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear was proportional to the cuticle thickness).
+As such, thick cuticles may be advantageous in exposed environments to avoid excessive damage [@blackman2005].
+Prolonged exposure to wind has been shown to increase the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991], by dislodging cuticular compounds by abrasion [@rogge1993].
 Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986], most dramatically demonstrated in an ice storm, leading to much higher g~min~ [@boyce2003].
-Such exposure to the elements was also held responsible for thicker cuticles in coastal, wind-exposed sites compared to inland sites [@blackman2005].
 
@@
 6).
-Comparing all available data, we did not find significant relationships between g~min~ and climate of origin, nor were there meaningful relationships with other traits (see Supporting Information).
+Comparing all available data, we did not find significant relationships between g~min~ and climate of origin, nor were there meaningful relationships with other traits (see Methods S1).
 We thus found it difficult to explain the variation in g~min~ between species.
@@
 Very few studies have found meaningful correlations between g~min~ and environmental factors or ecological categories.
-Based on their own compilation of the literature, @schuster2017 concluded no significant differences in g~min~ or g~cuti~ (pooled) by plant growth form.
+Based on their own compilation of the literature, @schuster2017 concluded no significant differences in g~min~ or g~cuti~ (analyzed together) by plant growth form.
 One notable exception is @brodribb2014, who showed a correlation between rainfall at species origin (of the driest quarter), and g~min~.
@@
 In Fig.
-6, we have compiled a number of studies in crops, demonstrating not just variation in g~min~ between crops, but also the wide range in g~min~ between genotypes.
+6c, we have compiled a number of studies in crops, demonstrating not just variation in g~min~ between crops, but also the wide range in g~min~ between genotypes.
 Again, it is striking that the wide variation in g~min~ cannot be easily explained by variation in leaf or other traits or chemical and structural components of the cuticle [@bengtson1978; @james2008; @saito2010].
@@
 While there is clearly plenty of variation in g~min~ between species, a number of lines of evidence suggest that g~min~ also has great potential for plasticity.
-Here we summarize the literature on the acclimation of g~min~ to drought conditions, to changes in temperature and humidity, and the change in g~min~ with leaf age and with altitude.
+Here we summarize the literature on the acclimation of g~min~ to drought conditions, to changes in temperature and humidity, and the change in g~min~ with leaf age, altitude and other factors.
 This discussion is directly relevant to the use of g~min~ in models, because if the degree of plasticity is large, it complicates model parameterization.
@@
 In each of the studies summarized, plants were grown in well-watered or droughted conditions, and in one case a difference in g~min~ was demonstrated after just four days of drought exposure [@bengtson1978].
-In a similar vein, @fanourakis2013 reported much lower g~min~ in plants grown in low vs.
-high relative humidity, again demonstrating that g~min~ is reduced in response to conditions that cause water limitation.
 We also demonstrate drought acclimation via a case study on 11 *Hakea* species grown in two watering treatments (Fig.
-7) - all 11 species showed a decrease in g~min~ in the drought treatment (see SuppInfo 3 for methods details).
+7) - all 11 species showed a decrease in g~min~ in the drought treatment (see Methods S3 for experimental details).
 This significant change in g~min~ with drought acclimation is likely an important component of the overall drought hardening of plants.
+
+\
+The idea that water limitation causes a reduction in g~min~ (via changes in the chemical composition of the cuticle) can be tested by inspecting the response to factors that increase evaporative demand.
+As one of the very few studies testing this idea, @fanourakis2013 reported much lower g~min~ in *Rosa* sp.
+plants grown in 60% vs.
+95% relative humidity.
+The difference could be attributed largely to a change in stomatal anatomy and lack of closure during desiccation, not changes in the cuticle *per se*.
+Also relevant is @sack2003, who reported lower g~min~ in sun leaves (for 2 out of 4 species) compared to shade leaves, again a lower g~min~ for leaves acclimated to high evaporative conditions.
 
@@
 @riederer2001 and @schuster2016b have both demonstrated a steep non-linear instantaneous response of g~cuti~ to temperature, with the response becoming especially steep at higher temperatures (> 40C).
-In *Eucalyptus haemastoma*, the response of g~min~ to temperature was so steep, that the proportion of cuticular to total transpiration increased from 2-3% at 20C, to 40% at 38C [@eamus2008].
+In *Eucalyptus haemastoma*, the response of g~min~ to temperature was so steep, that the proportion of cuticular to total transpiration increased from 2-3% at 20°C, to 40% at 38°C [@eamus2008].
 The mechanism of the instantaneous temperature response is complex, highly species-specific, and we refer to @schuster2016b for a detailed investigation.
@@
 Less well established is the acclimatory potential of g~min~ to changes in growth temperature.
-We present a case study on *Eucalyptus parramattensis* grown in whole-tree chambers (see SuppInfo 3 for methods details).
-The chambers either tracked ambient conditions, or were subjected to a persistent +3C warming treatment.
-After several months of growth in the treatments, g~min~ was measured at various temperatures, ranging from 17.5 to 27.5C.
+We present a case study on *Eucalyptus parramattensis* grown in whole-tree chambers (see Methods S3 for experimental details).
+The chambers either tracked ambient conditions, or were subjected to a persistent +3°C warming treatment.
+After several months of growth in the treatments, g~min~ was measured at various temperatures, ranging from 17.5 to 27.5°C.
 We found a 56% decrease in g~min~ in the elevated temperature trees, but there was no clear pattern with measurement temperature (Fig.
@@
 This decrease in g~min~ is consistent with the drought response, because again g~min~ is reduced in leaves that are subjected to conditions that increase evaporative demand (vapour pressure deficit was higher in the elevated temperature treatment, see @drake2018).
-The direction of this response is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment (Douglas-fir), which persisted for some time.
+The direction of this response is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment in *Pseudotsuga menziesii*, which persisted for some time.
 Responses across temperature gradients are more complex, in particular the change in g~min~ with altitude has been well studied.
@@
 S3, all of which reported on increased g~min~ for older leaves for the majority of species studied.
-The data reported by @jordan2007 is particularly impressive, as g~min~ in the herb *Agastachys odorata* gradually increased with leaf age up to ca.
+The data reported by @jordan2007 is particularly impressive, as g~min~ in the woody shrub *Agastachys odorata* gradually increased with leaf age up to ca.
 10 years.
@@
 Minimum conductance (g~min~) measured with mass loss of detached leaves on *Eucalyptus parramattensis* grown in whole-tree chambers.
-Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C) [see @drake2018].
+Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 °C) [see @drake2018].
 Measurements of g~min~ were carried out at five different temperatures on replicate leaves.
-There was no consistent effect of measurement T on g~min~, but leaves of trees grown in elevated temperature showed on average a 45% reduction in g~min~ (p < 0.01, mixed-effects model).
+There was no consistent effect of measurement temperature on g~min~, but leaves of trees grown in elevated temperature showed on average a 45% reduction in g~min~ (p < 0.01, mixed-effects model).
 
@@
 where g~min~ is parameterized with data obtained from mass loss of detached leaves (Fig.
-6), representing a true whole-leaf minimum that is reached under drought conditions, and g~0~ is the minimum reached during low photosynthesis (due to low light or other factors) for well-watered plants.
+6), representing a true whole-leaf minimum that is reached under conditions of severe drought, and g~0~ is the minimum reached during low photosynthesis (due to low light or other factors) for well-watered plants.
 During drought, g~0~ will decrease [@misson2004], so that the true minimum reached for Eq.
@@
 
+# Supporting Information
+
+**Methods S1** Literature compilation of g~min~ estimates
+**Methods S2** Methodological issues with measurement of g~min~
+**Methods S3** Methods description for case studies
+**Table S1** Effects of drought stress on g~min~
+**Figure S1** Effects of leaf age on g~min~
+
 # References {.unnumbered}

### 2018-02-13 manuscript_suppinfo.Rmd
workflow tidied; many edits especially introduction; suppinfo guidelines followed

@@
 
-# 1. Literature compilation of g~min~ estimates
+# Methods S1. Literature compilation of g~min~ estimates
 
@@
 ```
-**Figure S1**.
+**Figure 1**.
 Average minimum conductance (g~min~) for crops, graminoids, woody and other non-woody plant species.
@@
 ```
-**Figure S2**.
+**Figure 2**.
 No relationship between g~min~ and potential evapotranspiration across the range of the species.
@@
 
-# 2. Methodological issues with measurement of g~min~
+# Methods S2. Methodological issues with measurement of g~min~
 
@@
 ```
-**Figure S3**.
+
+**Figure 3**.
 Two example curves for determination of minimum conductance (g~min~) via mass loss of detached leaves (MLD).
@@
 
-# 3. Methods description for case studies
+# Methods S3. Methods description for case studies
 
-## Drought acclimation in *Hakea*
+## Drought acclimation in *Hakea* spp.
 
@@
 \
-We determined g~min~ from detached leaves collected before sunrise from each of 6 *Eucalyptus parramattensis* trees grown under ambient and ambient+3C temperature conditions, respectively.
+We determined g~min~ from detached leaves collected before sunrise from each of 6 *Eucalyptus parramattensis* E.C.
+Hall trees grown under ambient and ambient+3C temperature conditions, respectively.
 Leaves were scanned for leaf area determination before being weighed using a precision balance and attached to a ‘clothesline’ suspended inside controlled chambers set at different temperatures (17.5, 20, 22.5, 25, 27.5°C) and common low light (~200 umol m^-2^ s^-1^) and a relative humidity of 60%.
@@
 
-# 4. Effects of drought stress on g~min~
+# Table S1. Effects of drought stress on g~min~
 
@@
 
-# 5. Effects of leaf age on g~min~
+# Figure S1. Effects of leaf age on g~min~
 
@@
 
-**Figure S4**.
+**Figure S1**.
 Correlation between g~min~ and leaf age for five studies in the g~min~ database.

### 2018-02-13 R/figures.R

@@ -9,7 +9,7 @@ figure_gmin_review <- function(gdfr){
             ylim=c(-1,2.2),
             label_las=1,
             ylab=expression(Conductance~(mmol~m^-2~s^-1)),
-            labels=c(expression(g["cut,isol"]),
+            labels=c(expression(g["cuti"]),
                      expression(g["min"]),
                      expression(g["dark"]),
                      expression(g["0"]),
@@ -26,7 +26,7 @@ figure_gmin_review_2 <- function(gdfr){
           label_las=1,
           ylim=c(0,80),
           ylab=expression(Conductance~(mmol~m^-2~s^-1)),
-          labels=c(expression(g["cut,isol"]),
+          labels=c(expression(g["cut"]),
                    #expression(g["cut,seal"]),
                    expression(g["min"]),
                    expression(g["dark"]),

### 2018-02-13 R/functions.R

@@ -1,4 +1,4 @@
-to.pdf <- function(expr, filename, ..., verbose=TRUE) {
+to.pdf <- function(expr, filename, ..., verbose=FALSE) {
   if(!file.exists(dirname(filename)))
     dir.create(dirname(filename), recursive=TRUE)
   if ( verbose )
@@ -69,13 +69,13 @@ plotlabel <- function(txt, where, inset=0.08, inset.x=inset, inset.y=inset,...){
 
 
 convert_all_pdf <- function(path, ...){
+  message("Converting PDFs to PNG...")
   o <- getwd()
   on.exit(setwd(o))
   setwd(path)
   pdfs <- dir(pattern="[.]pdf")
   for(i in seq_along(pdfs)){
     convert_pdf_png(pdfs[i])
-    message(i)
   }
 }
 
@@ -130,7 +130,7 @@ plotCI2 <- function(yvar, group, data, transform_log10=FALSE,
   
   if(is.null(ylim))ylim <- c(0, max(ci)+0.1*max(ci))
   g <- glht(fit, linfct=mcp(G = "Tukey"))
-  lets <- cld(g)$mcletters$Letters
+  lets <- suppressWarnings(cld(g)$mcletters$Letters)
   n <- length(mn)
   
   plot(1:n, mn, pch=19, cex=1.1, ylim=ylim, axes=FALSE,
@@ -178,7 +178,7 @@ plotCIlog <- function(yvar, group, data,
 
   if(is.null(ylim))ylim <- c(0, max(ci)+0.1*max(ci))
   g <- glht(fit, linfct=mcp(G = "Tukey"))
-  lets <- cld(g)$mcletters$Letters
+  lets <- suppressWarnings(cld(g)$mcletters$Letters)
   
   n <- length(mn)
   

### 2018-02-13 R/load.R

@@ -2,16 +2,17 @@
 
 
 
-if(!require("pacman"))install.packages("pacman")
-pacman::p_load(Hmisc, car, dplyr, tidyr, nlme, nlshelper, 
+if(!require("pacman", quietly=TRUE))install.packages("pacman")
+suppressPackageStartupMessages(
+  pacman::p_load(Hmisc, car, dplyr, tidyr, nlme, nlshelper, 
                forcats, tibble, magicaxis, 
                plantecophys, readxl, multcomp,
                reporttools, Taxonstand, taxize, 
                doBy, stringi, rmarkdown, ggplot2)
-
+)
 
 #devtools::install_github("remkoduursma/speciesmap")
-library(speciesmap)
+suppressPackageStartupMessages(library(speciesmap))
 
 if(!dir.exists("download"))dir.create("download")
 if(!dir.exists("output"))dir.create("output")
@@ -39,8 +40,6 @@ lin2015a <- group_by(lin2015, fitgroup) %>%
 
 # Lin 2015 fits
 lin2015coef <- fits_lin2015(lin2015a)
-
-lin2015coef <- subset(lin2015coef, R2 > 0.5)
 
 # Kerstiens 1996
 kerst <- read.csv("data/kerstiens1996_table1.csv", stringsAsFactors = FALSE) %>%

### 2018-02-13 run.R

@@ -1,12 +1,15 @@
 
 # Read & download data, load & install packages.
+message("Loading packages, reading data...")
 source("R/load.R")
 
 # Make figures as PDF (see output/)
 # Conversion to png relies on an installation of ImageMagick (and GhostScript),
 # may fail on your system (but you can ignore these errors and continue).
+message("Making figures...")
 source("R/make_figures.R")
 
 # Convert manuscript (output is manuscript.docx & manuscript_suppinfo.docx).
-rmarkdown::render("manuscript.Rmd", encoding = "utf8")
-rmarkdown::render("manuscript_suppinfo.Rmd", encoding = "utf8")
+message("Rendering documents...")
+rmarkdown::render("manuscript.Rmd", encoding = "utf8", quiet=TRUE)
+rmarkdown::render("manuscript_suppinfo.Rmd", encoding = "utf8", quiet=TRUE)

### 2018-02-15 manuscript.Rmd
references cleanup

@@
 Since a membrane that is permeable to CO~2~ but not H~2~O has never evolved, all land plants have stomata in their leaves, which disrupt the cuticle and allow CO~2~ uptake.
-It is well known that stomatal open and close in response to changes in light intensity, humidity and CO~2~ concentration at the leaf surface.
+It is well known that stomata open and close in response to changes in light intensity, humidity and CO~2~ concentration at the leaf surface.
 In doing so, stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus can be expected to close stomata when photosynthetic rates are low.
@@
 **Methods S1** Literature compilation of g~min~ estimates
+
 **Methods S2** Methodological issues with measurement of g~min~
+
 **Methods S3** Methods description for case studies
+
 **Table S1** Effects of drought stress on g~min~
+
 **Figure S1** Effects of leaf age on g~min~

### 2018-02-28 manuscript.Rmd
fixed comparison and description thereof. Nearly there!

@@
 In this review, we discuss the role of the minimum conductance in models of plant water use, and critically evaluate sources of variation in this overlooked parameter.
-In this context, it is useful to briefly summarize the approach taken by the majority of vegetation model.
+In this context, it is useful to briefly summarize the approach taken by the majority of vegetation models.
 Most current-generation process-based global vegetation models (GVMs) use a version of the Ball-Berry class of 'stomatal' (technically including both stomata and the cuticle) conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity (Eq.
@@
 We aim to improve this connection by synthesizing the state of knowledge.
-Our goals are to a) compare various definitions of g~min~ quantitively, which have been previously assumed equal, b) demonstrate the need to include a non-zero g~0~ (Eq.
-1), for models of water-use efficiency as well as the drought response, c) present a new compilation of minimum conductance measurements, and demonstrate the large acclimatory potential of this trait to environmental drivers.
+Our goals are to a) compare various definitions of g~min~ quantitively, which have sometimes been previously assumed equal, b) demonstrate the need to include a non-zero g~0~ (Eq.
+1), for models of water-use efficiency as well as the drought response, c) present a new compilation of minimum conductance measurements, test for ecologically meaningful patterns, and demonstrate the large acclimatory potential of this trait to environmental drivers.
 
@@
 | g~0~          | Intercept in the Ball-Berry-type stomatal conductance model, i.e. g~s~ when A~n~ approaches zero. |
-| g~s,low A~    | Cuvette measurements of g~s~ when photosynthesis rates are very low.                              |
 
@@
 \
-We quantitatively compared each of the different 'minimum' conductance (Table 1) values by synthesizing existing data.
-To estimate g~0~ (Eq.
-1) from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model.
+We quantitatively compared each of the different 'minimum' conductance (Table 1) definitions by synthesizing existing data.
 Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but were not used since they included no metadata).
@@
 For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Supporting Information for full details of the compilation).
-All conductance values were converted to per unit projected surface area.
-Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance (discussed further in Section VI), it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful.
-A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca.
-150 species during night-time.
-Actual minimum daytime values were estimated from the @lin2015 database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the 52 studies included in that dataset.
-This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted or extrapolated g~0~ term.
+All conductance values were converted to per unit projected surface area, allowing direct comparison to stomatal conductance data which are typically presented as such.
+We contrasted g~cuti~ and g~min~ with three estimates of stomatal conductance when photosynthesis rates are low or zero during non-drought conditions: night-time conductance (g~dark~) (further discussed in Section VI), conductance at low photosynthetically active radiation (PAR) (0 - 40 $\mu$ mol m^-2^ s^-1^) during the day, and conductance during conditions when photosynthesis rates are very low (< 1 $\mu$ mol m^-2^ s^-1^) (but excluding low PAR and drought).
+Estimates of g~dark~ were taken from @lombardozzi2017, who compiled measurements of ca.
+150 species during the night (we selected only gas exchange based measurements from their database).
+Daytime values (g~low PAR~ and g~low A~) were estimated from an update to the @lin2015 database by taking the appropriate subsets, and averaging the g~s~ values by species within study.
 
@@
 Previous studies have directly compared g~dark~ and g~min~ on the same species, and generally concluded that g~dark~ is much higher than g~min~ [@hygen1954; @koerner1994; @walden-coleman2013], but an exception is @cavender-bares2007 who found the two rates to be similar in oaks.
-
-\
-The fitted minimum conductance parameter, g~0~ (mean = 32.8 mmol m^-2^ s^-1^), was much higher than g~min~, but similar to g~dark~.
-However, we point out the uncertainties associated with g~0~ in Section III, which tend to inflate estimates of g~0~.
-It is telling that actual measured stomatal conductance during low photosynthesis (g~s,low A~) was  much lower than g~0~ or g~dark~.
-In other words, actual daytime conductance during these conditions that cause low photosynthetic rates is much lower than the presumed minimum estimated from fitted g~0~.
-A similar result was already presented by @dekauwe2015 (their Fig.
-A2).
+In turn, g~dark~ could not be differentiated from stomatal conductance measurements during daytime wither in very low light (mostly just after dawn, or before sunset) (g~low PAR~), or during conditions when photosynthetic rates were very low (due to very high temperature, VPD, or other factors) (g~low A~).
+These comparisons demonstrate that g~s~ during non-drought conditions when photosynthetic rates are zero or negligible are much higher than the minimum reached in simulated drought conditions (g~min~).
 
@@
 Different letters denote significant differences (at $\alpha$ = 0.05).
-g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~0~ : fitted parameter (Eq.
-1) with regression; g~s,low A~ leaf conductance during periods of very low photosynthesis.
+g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~low PAR~ leaf conductance during low light (photosynthetically active radiation, PAR); g~low A~ leaf conductance during periods of very low photosynthesis.
 See text for data sources and methods.
@@
 \
-As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) does not vary with the photon flux density (PPFD), as is typical in leaf gas exchange data.
-When PPFD approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~.
+As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) does not vary with PAR, as is typical in leaf gas exchange data.
+When PAR approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~.
 To see this point, we can rearrange Eq.
@@
 
-Eq. 3 - derived using zero g~0~ - thus does not give any dependence on PPFD.
+Eq. 3 - derived using zero g~0~ - thus does not give any dependence on PAR.
 
@@
 A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~) [@barnard2013; @lombardozzi2017], simply because it is a condition when net photosynthesis is zero.
-However as our literature comparison showed (Fig.
-1), g~dark~ is substantially higher compared to all other variables except fitted g~0~.
-The simple idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.
-
-\
+We provide some caution to the assumption that g~dark~ can be used in models of daytime leaf conductance.
 First, ample observations suggest that g~dark~ is not a fixed rate but varies tremendously during the night.
@@
 Measurements of g~min~ were carried out at five different temperatures on replicate leaves.
-There was no consistent effect of measurement temperature on g~min~, but leaves of trees grown in elevated temperature showed on average a 45% reduction in g~min~ (p < 0.01, mixed-effects model).
+There was no consistent effect of measurement temperature on g~min~, but leaves of trees grown in elevated temperature showed on average a 45% reduction in g~min~ (p < 0.01, linear mixed-effects model).
 
@@
 \
-In a pioneering publication which presented the Jarvis model of stomatal conductance, @jarvis1976 states that "we have assumed in the following equations that when stomata are closed the leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." Similarly, given the poor statistical properties of g~0~ estimated from regression, the very high values compared to minimum conductance reached for detached leaves, we conclude that g~0~ should not be estimated from regression on leaf gas exchange data for use in models.
+In a pioneering publication which presented the Jarvis model of stomatal conductance, @jarvis1976 states that "we have assumed in the following equations that when stomata are closed the leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." Similarly, given the poor statistical properties of g~0~ estimated from regression we conclude that g~0~ should not be estimated from regression on leaf gas exchange data for use in models.
 Then, how should g~0~ be estimated?
@@
 \
-Our first conclusion is that g~0~, g~min~ and g~dark~ should not be assumed equal in any modelling study.
-Indeed, a productive way forward is to view each of these conductances as distinct, and they should be parameterized separately.
-For example, when considering modelling stomatal conductance with Eq.
-1, we suggest that g~min~ and g~0~ can be combined:
+We suggest that the minimum conductance in the Ball-Berry model should include both a g~0~ and a g~min~ term, since g~0~ represents the minimum reached during low light and conditions of low photosynthesis, and g~min~ represents the minimum reached during severe drought.
+Moreover, we suggest that the minimum conductance is not simply added to the photosynthesis-dependent term (right-hand side in Eq.
+1), but used as an actual minimum.
+Thus,
 
-$$ g_s = min(g_{min}, g_0) + g_1 \frac{A_n}{C_a}f(D)$$
-
-where g~min~ is parameterized with data obtained from mass loss of detached leaves (Fig.
-6), representing a true whole-leaf minimum that is reached under conditions of severe drought, and g~0~ is the minimum reached during low photosynthesis (due to low light or other factors) for well-watered plants.
-During drought, g~0~ will decrease [@misson2004], so that the true minimum reached for Eq.
-1 is g~min~.
+$$ g_s = max( min(g_{min}, g_0), g_1 \frac{A_n}{C_a}f(D))$$
 
 \
-It is however an open question on how g~min~ should be parameterized for different species, as we found both large variation between species that could not be explained by traits, and remarkable plasticity to growing conditions.
+This model for g~s~ will converge to g~0~ during periods of low photosynthesis, and to g~min~ during drought if we further include a model for the dependence of g~0~ on water availability [see @misson2004].
+Another advantage is that independent estimates of g~0~ can be used, and not with regression which produces the undesirable correlation with estimates of g~1~.
+The above formulation is yet to be tested against data, but we propose this test to be done combining data from drought and non-drought conditions.
+
+\
+Finally, we conclude that g~min~ displays a large amount of variation between species that could not be explained by traits, and remarkable plasticity to growing conditions.
 Perhaps this plasticity is the reason that g~min~ did not vary predictably between species.

### 2018-02-28 R/figures.R

@@ -12,7 +12,6 @@ figure_gmin_review <- function(gdfr){
             labels=c(expression(g["cuti"]),
                      expression(g["min"]),
                      expression(g["dark"]),
-                     expression(g["0"]),
                      expression(g["low PAR"]),
                      expression(g["low A"])
                      ))
@@ -132,7 +131,7 @@ figure_R2g0 <- function(lin2015coef, miner){
   
   l <- loess(g0_se ~ cv_bbopti, data=lin2015coef, span=0.9)
   with(lin2015coef, plot(cv_bbopti, g0_se, pch=16, ylim=c(0,60),
-                         xlab=expression("CV of"~A[n]/(C[a] * sqrt(VPD))),
+                         xlab=expression("CV of"~A[n]/(C[a] * sqrt(D))),
                          ylab=expression("SE of"~g[0]~~(mmol~m^-2~s^-1)),
                          panel.first=plot_loess(l, add=TRUE, band=TRUE, lwd=2, col="darkgrey")
                          ))

### 2018-02-28 R/functions.R

@@ -31,6 +31,7 @@ fits_lin2015 <- function(lin2015a){
                         sd_bbopti = sd(BBopti, na.rm=TRUE),
                         cv_bbopti = sd_bbopti/mean(BBopti, na.rm=TRUE),
                         sd_cond = sd(Cond, na.rm=TRUE),
+                        gsmax = max(Cond, na.rm=TRUE),
                         Qrange = max(PARin) - min(PARin),
                         Drange = max(VPD) - min(VPD),
                         Trange = max(Tleaf) - min(Tleaf)), 

### 2018-02-28 R/load.R

@@ -104,7 +104,7 @@ miner <- read.csv("data/Miner_table1.csv") %>%
 # Compilation of nighttime conductance.
 lombar <- read.csv("data/lombardozzi_gnight.csv", stringsAsFactors = FALSE) %>%
   rename(gmin = gnight) %>%
-  filter(gmin > 0, gmin < 1000,
+  filter(gmin > 0, gmin < 1000, 
          method %in% c("ge","gas exchange","Li6400","IRGA")) %>%
   mutate(method="gnight")
 
@@ -113,19 +113,6 @@ g0s <- data.frame(gmin=c(lin2015coef$g0, miner$g0),
                   method="g0", stringsAsFactors = FALSE) %>%
   filter(!is.na(gmin),
          gmin > 0)
-
-
-# From Lin2015, data where A < threshold.
-minags <- group_by(lin2015, fitgroup) %>%
-  dplyr::summarize(
-    Amin = min(Photo, na.rm=TRUE),
-    gmin = 1000 * min(Cond, na.rm=TRUE),
-    Qrange = max(PARin) - min(PARin),
-    Pathway = unique(Pathway)[1]
-  ) %>%
-  filter(Amin < 2) %>%
-  mutate(method="gslowA") %>%
-  dplyr::select(gmin, method)
 
 
 #------ gmindatabase
@@ -216,9 +203,9 @@ kerst_simple <- group_by(kerst, species, method) %>%
   ungroup %>%
   dplyr::select(gmin, method)
 
-gdfr <- bind_rows(gmindat_simple, kerst_simple, schuster2017, lombar, g0s, gs_low_a, gs_low_par) %>%
-  filter(method != "gcut_seal") %>%  # I keep changing my mind!!
-  mutate(method = factor(method, levels=c("gcut_isol","gmin","gnight", "g0","gslowPAR","gslowA")))
+gdfr <- bind_rows(gmindat_simple, kerst_simple, schuster2017, lombar, gs_low_a, gs_low_par) %>%
+  filter(method != "gcut_seal") %>%  
+  mutate(method = factor(method, levels=c("gcut_isol", "gmin", "gnight", "gslowPAR", "gslowA")))
 
 
 # Herve's simulations with Sureau

### 2018-02-28 R_other/cond_low_a_low_par.R

@@ -1,0 +1,23 @@
+
+# This works only in the repository bitbucket.org/bmedlyn/leafgasexchange_ongoing
+
+source("make_database.R")
+
+library(doBy)
+library(dplyr)
+
+dat <- filter(alldata, opt %in% c("opt","low PAR","lower canopy"),
+              PARin > 0,
+              PARin < 40) %>%
+  mutate(dataset_fitgroup = paste(Datacontrib, Location, fitgroup))
+
+
+cond_low_par <- summaryBy(Cond ~ dataset_fitgroup, FUN=mean, data=dat, keep.names=TRUE)
+
+dat <- filter(alldata, PARin > 40, Photo < 1, opt == "opt") %>%
+  mutate(dataset_fitgroup = paste(Datacontrib, Location, fitgroup))
+cond_low_a <- summaryBy(Cond ~ dataset_fitgroup, FUN=mean, data=dat)
+
+
+write.csv(cond_low_par, "cond_low_par.csv", row.names = FALSE)
+write.csv(cond_low_a, "cond_low_a.csv", row.names = FALSE)

### 2018-05-23 manuscript.Rmd
Minor tweaks. Version as submitted 1 March 2018.
Revision completed, as sent to coauthors on 20180523

@@
 tab <- read.csv(text="I. Introduction
-II. Comparison of various definitions of minimum conductance
-III. Problems with estimating minimum conductance from regression
-IV. Use of minimum conductance in models
+II. Use of minimum conductance in models
 \t1. Models of water-use efficiency
 \t2. Models of plant desiccation
+\t3. Towards a new model formulation
+III. Comparison of various definitions of minimum conductance
+IV. Problems with estimating Minimum conductance from regression
 V. Cuticular conductance
-VI. Nighttime conductance
-VII. Environmental and ecological variation in minimum conductance 
+VI. Contribution of stomata
+VII. Nighttime conductance
+VIII. Environmental and ecological variation in minimum conductance 
 \t1. Contribution of stomata to minimum conductance
@@
 \t3. Acclimation to the environment
-VIII. Conclusions", header=F)
+IX. Conclusions", header=F)
 
@@
 
-Stomata exert a strong control on the transpiration rate of leaves.
-During severe drought, or conditions where photosynthesis does not take place, it seems advantageous for plants to completely halt water loss.
-However, water loss continues through the cuticle and incompletely closed stomata, together determined by the leaf minimum conductance (g~min~).
+During conditions where photosynthesis does not take place, such as severe drought, extreme temperature or low light, it seems advantageous for plants to close stomata and completely halt water loss.
+However, water loss continues through the cuticle and incompletely closed stomata, together constituting the leaf minimum conductance (g~min~).
 In this review, we illustrate the role of g~min~ in models of leaf gas exchange, critically evaluate sources of variation in g~min~, and quantitatively compare various methods for its estimation.
@@
 It is well known that stomata open and close in response to changes in light intensity, humidity and CO~2~ concentration at the leaf surface.
-In doing so, stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus can be expected to close stomata when photosynthetic rates are low.
-To avoid desiccation and ultimate death, stomata can also be expected to close during periods of severe water stress.
+Stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus to close stomata when photosynthetic rates are low.
+To avoid desiccation and ultimate death, stomata typically close during periods of severe water stress.
 When stomata are closed, water loss continues at a greatly diminished rate through the cuticle.
@@
 Other measurements allow for the estimation of water loss through incompletely closed and broken stomata.
-Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time.
-This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants during severe drought, and includes both surfaces of the leaf - not just the stomata-free side.
+Of particular interest because of its simplicity is the method where leaves are detached from the plant, and their weight loss monitored over time.
+This method (mass loss of detached leaves, MLD) aims to simulate field conditions during severe drought, when water supply to the leaf has practically ceased, however proceeds much quicker (typically 0.5-2 days).
+This measurement also includes both surfaces of the leaf - not just the stomata-free side.
 A comprehensive compilation of estimates of g~min~ is currently lacking, and is needed not just to parameterize models, but also to study sources of variation in this overlooked plant trait.
@@
 In this review, we discuss the role of the minimum conductance in models of plant water use, and critically evaluate sources of variation in this overlooked parameter.
-In this context, it is useful to briefly summarize the approach taken by the majority of vegetation models.
-Most current-generation process-based global vegetation models (GVMs) use a version of the Ball-Berry class of 'stomatal' (technically including both stomata and the cuticle) conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity (Eq.
-1).
+Previous reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017].
+Other work has focused on a different definition of the minimum conductance (g~0~): the value that should be used in models of plant water use [e.g.
+@barnard2013; @lombardozzi2017], and includes both g~min~ and a 'stomatal residual' due to the fact that stomata do not completely close during periods of zero photosynthesis [@dekauwe2015].
+This discussion often centers around leaf conductance during the night [@lombardozzi2017], but we must also consider appropriate values for the minimum conductance during the day, for example during very high (or low) temperature, transient low light, and extreme drought.
+Thus there is a disconnect between the dizzying level of understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models.
 
 \
-(1)
-$$ g_s = g_0 + g_1 \frac{A_n}{C_a}f(D)$$
-
-\
-Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), and g~0~ is the value of g~s~ when A~n~ is zero.
-In this paper, we refer to Eq.
-1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987).
-Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been studied in much less detail.
-For example, a recent comprehensive review of stomatal conductance models did not mention g~0~ or any similar minimum conductance [@damour2010].
-The original description of Eq.
-1 as published by @ballberry1987 did not include an intercept term.
-It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model.
-In practice, g~0~ is estimated from a regression of Eq.
-1 with leaf gas exchange data, but this may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.
-
-\
-Previous reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017].
-Other work has focused on the role of g~0~ in models of plant water use [e.g.
-@barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance.
-Thus there is a disconnect between the dizzying level of understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models.
-We aim to improve this connection by synthesizing the state of knowledge.
-Our goals are to a) compare various definitions of g~min~ quantitively, which have sometimes been previously assumed equal, b) demonstrate the need to include a non-zero g~0~ (Eq.
-1), for models of water-use efficiency as well as the drought response, c) present a new compilation of minimum conductance measurements, test for ecologically meaningful patterns, and demonstrate the large acclimatory potential of this trait to environmental drivers.
-
-# II. Comparison of various definitions of minimum conductance
+We aim to improve the connection between the physiology of minimum conductance, empirical approaches, and model implementations by synthesizing the state of knowledge.
+Our goals are to a) demonstrate the need to include a non-zero minimum conductance in models of water-use efficiency and drought responses, and discuss current practices b) compare various definitions of g~min~ quantitively, which have sometimes been previously assumed equal, c) present a new compilation of minimum conductance measurements, test for ecologically meaningful patterns, and demonstrate the large acclimatory potential of this trait to environmental drivers.
 
@@
 
-\
-Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories (see Table 1).
-The absolute minimum conductance attained by leaves is that through the cuticle only.
-Measurements of cuticular conductance (g~cuti~) are typically made on isolated non-stomatal cuticles [@riederer2001], via special gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata [@kerstiens1996].
-We further discuss cuticular conductance in Section V.
+# II. Use of minimum conductance in models
+
+## 1. Models of water-use efficiency
+
+Most current-generation process-based global vegetation models (GVMs) use a version of the Ball-Berry class of 'stomatal' (technically including both stomata and the cuticle) conductance (g~s~) models characterized by the dependence on assimilation rate, CO~2~ and humidity (Eq.
+1).
 
 \
-The minimum conductance of intact leaves is typically measured by mass loss of detached leaves (MLD).
-In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out.
-Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after some time a constant low rate is achieved.
-From this minimum transpiration rate (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), the minimum conductance (g~min~) can be estimated using measured vapour pressure deficit [@sinclair1986].
-This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress.
-We do not discuss the methodology of this measurement in detail, but note that care needs to be taken to yield reliable estimates (see Methods S2 for further information).
+(1)
+$$ g_s = g_0 + g_1 \frac{A_n}{C_a}f(D)$$
 
 \
-We quantitatively compared each of the different 'minimum' conductance (Table 1) definitions by synthesizing existing data.
-Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but were not used since they included no metadata).
-Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see Methods S1 for description and references).
-For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Supporting Information for full details of the compilation).
-All conductance values were converted to per unit projected surface area, allowing direct comparison to stomatal conductance data which are typically presented as such.
-We contrasted g~cuti~ and g~min~ with three estimates of stomatal conductance when photosynthesis rates are low or zero during non-drought conditions: night-time conductance (g~dark~) (further discussed in Section VI), conductance at low photosynthetically active radiation (PAR) (0 - 40 $\mu$ mol m^-2^ s^-1^) during the day, and conductance during conditions when photosynthesis rates are very low (< 1 $\mu$ mol m^-2^ s^-1^) (but excluding low PAR and drought).
-Estimates of g~dark~ were taken from @lombardozzi2017, who compiled measurements of ca.
-150 species during the night (we selected only gas exchange based measurements from their database).
-Daytime values (g~low PAR~ and g~low A~) were estimated from an update to the @lin2015 database by taking the appropriate subsets, and averaging the g~s~ values by species within study.
-
-\
-The results of the quantitative comparison are presented in Fig.
-1.
-From this compilation, a few striking differences between estimates are apparent.
-Cuticular conductance averaged 0.45 mmol m^-2^ s^-1^, ten-fold lower compared to g~min~ (4.9 mmol m^-2^ s^-1^).
-This result is in line with the methods comparison of @kerstiens1996.
-More recently, @schuster2017 argued, based on data for 8 species, that g~min~ from MLD is comparable to g~cuti~.
-However in the same study, a literature compilation of hundreds of values showed that g~cuti~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cuti~ were pooled in their analysis.
-Nonetheless, the comparison between g~cuti~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cuti~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions.
-@boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.
-
-\
-Conductance in the dark (g~dark~) (mean = 41.4 mmol m^-2^ s^-1^) was on average eight-fold higher than g~min~.
-It is no new observation that stomata do not close in the dark (further discussed in Section VI).
-Previous studies have directly compared g~dark~ and g~min~ on the same species, and generally concluded that g~dark~ is much higher than g~min~ [@hygen1954; @koerner1994; @walden-coleman2013], but an exception is @cavender-bares2007 who found the two rates to be similar in oaks.
-In turn, g~dark~ could not be differentiated from stomatal conductance measurements during daytime wither in very low light (mostly just after dawn, or before sunset) (g~low PAR~), or during conditions when photosynthetic rates were very low (due to very high temperature, VPD, or other factors) (g~low A~).
-These comparisons demonstrate that g~s~ during non-drought conditions when photosynthetic rates are zero or negligible are much higher than the minimum reached in simulated drought conditions (g~min~).
-
-```{r fig.width=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure1.png")
-```
-
-\
-(Figure 1 here)
-
-\
-**Figure 1**.
-Comparison of various estimates of the (presumed) minimum conductance.
-Error bars are 95\% confidence intervals.
-Grey dots are the original data (but a few data points occur outside the figure range).
-Different letters denote significant differences (at $\alpha$ = 0.05).
-g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~low PAR~ leaf conductance during low light (photosynthetically active radiation, PAR); g~low A~ leaf conductance during periods of very low photosynthesis.
-See text for data sources and methods.
-
-# III. Problems with estimating minimum conductance from regression
-
-\
-The standard approach in estimating g~0~ for use in Eq.
-1 is to use regression with g~s~ as the response variable, and the right-hand side of Eq.
-1 as the predictor.
-The g~0~ parameter is thus estimated as the intercept.
-Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated by this method.
-The difficulty of the estimation can already be anticipated from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g.
-@leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species); presumably because unreliable estimates were obtained, b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though it is clearly non-sensical to suggest negative conductance values.
-@barnard2013 also mentioned the difficulty of fitting Eq.
-1 to estimate g~0~, but did not present specific details - the following analysis builds on their work by demonstrating statistical uncertainties.
-
-\
-It is also telling that there are few reports on the intra-specific plasticity, or inter-specific variation of g~0~, perhaps because it is so poorly estimated.
-Exceptions are @duarte2016, who found a lower g~0~ in a heat wave treatment in Douglas-fir (although g~1~ was unaffected), but in their case g~0~ was very accurately estimated by careful multi-point light response curves.
-Another exception is @misson2004, who report on a close negative correlation between pre-dawn leaf water potential and g~0~ in Ponderosa pine (again, the g~1~ parameter was unaffected).
-
-\
-In the following we demonstrate that the fitting process is problematic for three reasons: 1) the estimates of g~0~ and g~1~ are highly correlated, 2) the precision for g~0~ is generally much lower than for g~1~, and 3) for data that have a worse fit overall, the g~0~ estimates are elevated.
-We thus conclude that g~0~ should really not be estimated from regression, although this is the most common method applied.
-In addition, it is difficult to accurately measure low fluxes with a portable gas exchange system, and great care has to be taken to arrive at reasonable estimates of g~0~ that way.
-
-\
-In Fig.
-2, we demonstrate how g~0~ and g~1~ are statistically correlated, when fitted for a single study.
-The correlation between slopes and intercepts is not unique to Eq.
-1, but a general property of linear models [@stapleton1995].
-The consequence is that it is more difficult to interpret g~1~ and g~0~ independently, and that confidence intervals on each of the parameters should also not be interpreted independently (see @becker2007 for more discussion on the statistical aspects of comparing slopes and intercepts).
-
-```{r fig.height=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure2.png")
-```
-
-\
-(Figure 2 here)
-
-\
-**Figure 2**.
-Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al.
-(2011) applied to a leaf gas exchange dataset of @martinstpaul2012.
-This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al.
-(2015).
-The solid line is regression line, and the shaded area a 95% confidence interval for the mean.
-The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression.
-The dotted ellipse is a bivariate 95% confidence interval for slope and intercept.
-The symbols represent 1000 bootstrap samples of the coefficients.
-
-\
-Next we study the values obtained when fitting Eq.
-1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation of fitted g~0~ and *m* from @miner2017 (who applied the original model of Ball-Berry).
-In both cases, estimates of g~0~ are inflated when the model fits poorly (Fig.
-3 a,b).
-This can be understood by considering that a poor fit often results in the flattening out of the regression line, thus giving a large value for the intercept.
-A poor fit is often obtained when there is little variation in the right-hand side of Eq.
-1, usually because there was low variation in environmental conditions (humidity, light, temperature).
-We confirm this by showing that the standard error (SE) of g~0~ increases when the coefficient of variation of the right-hand side of Eq.
-1 is lower (Fig.
-3c).
-
-```{r fig.height=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure3.png")
-```
-
-\
-(Figure 3 here)
-
-\
-**Figure 3**.
-Statistical uncertainty in the estimation of g~0~ from regression, demonstrated with two parameter databases.
-(a) We fitted the linearized form of the @medlyn2011 model to each of the datasets in the @lin2015 leaf gas exchange database, showing that for poorly fitted relationships (low R^2^), inflated estimates of g~0~ are obtained.
-Vertical lines are 95% confidence intervals.
-The grey line is a fitted loess smoother with 95% confidence interval.
-Note the wide confidence intervals, and frequent negative values.
-(b) Similar to (a), but using the published compilation by @miner2017.
-The grey line is a fitted loess smoother with 95% confidence interval.
-(c) Using the fits from panel (a), a demonstration that the standard error (SE) of g~0~ is much higher when the coefficient of variation (CV) of the predictor (i.e.
-right-hand side of the equation being fitted) is lower.
-
-# IV. Use of minimum conductance in models
-
-## 1. Models of water-use efficiency
+Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987, see @damour2010 for a list of functions), and g~0~ is the value of g~s~ when A~n~ is zero.
+The g~1~ parameter is directly related to the water-use efficiency: large values of g~1~ indicate low water-use efficiency.
+In this paper, we refer to Eq.
+1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987).
+Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been studied in much less detail.
+For example, a recent comprehensive review of stomatal conductance models did not mention g~0~ or any similar minimum conductance [@damour2010].
+The original description of Eq.
+1 as published by @ballberry1987 did not include an intercept term.
+It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model.
+In practice, g~0~ is estimated from a regression of Eq.
+1 with leaf gas exchange data, but as we will show, this may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.
 
@@
 This finding, if indeed generally true, has a large effect on the calculation of various gas exchange parameters [@hanson2016].
-@manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is leaky and incompletely closed stomata.
-Because g~0~ includes not just the cuticle, but also incompletely closed stomata, and the finding that g~0~ is much larger than estimates of g~min~ or g~cuti~ (Fig.
-1), we ignore this effect in the model simulations that follow.
+@manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is the effect of leaky and incompletely closed stomata.
+Because g~0~ includes not just the cuticle, but also incompletely closed stomata, we ignore this effect in the model simulations that follow.
 
@@
 Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero.
-But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, begging the question whether the same g~0~ is reached in each of these cases.
+But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, and it is unclear whether the same g~0~ is reached in each of these cases.
 In the application of the Ball-Berry model, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005].
 It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cuti~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata.
-Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake2018, @urban2017].
+Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake2018; @urban2017].
 
@@
 The consequence is that stomata tend to open during periods that are favourable for photosynthesis (high light, optimal temperature), and close when photosynthesis drops to zero (darkness, very high temperature).
-Clearly, in this framework there is no place for g~0~ since it is always sub-optimal to open stomata (i.e.
+Clearly, in this optimality framework there is no place for g~0~ since it is always sub-optimal to open stomata (i.e.
 spend water) when there is no photosynthetic gain.
@@
 \
-As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) does not vary with PAR, as is typical in leaf gas exchange data.
+As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) does not vary with PAR, as is typical in leaf gas exchange data [though other mechanisms can also be employed to simulate this pattern, see @dewar2018].
 When PAR approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~.
@@
 
-and using the diffusion constraint $A_n = g_s / 1.6 (C_a - C_i)$ we get an expression for C~i~ / C~a~,
+and using the diffusion constraint (Fick's law) $A_n = g_s / 1.6 (C_a - C_i)$ we get an expression for C~i~ / C~a~,
 
@@
 Using the coupled leaf gas exchange model, we show in Fig.
-4 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^).
+1 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^).
 We also demonstrate the effect of g~0~ on modelling the T~leaf~ response.
@@
 ```{r fig.height=6,echo=FALSE, eval=includefigs}
-plotfig("output/figure4.png")
+plotfig("output/figure1.png")
 ```
@@
 \
-(Figure 4 here)
+(Figure 1 here)
 
 \
-**Figure 4**.
+**Figure 1**.
 Simulations with a coupled leaf gas exchange model [@duursma2015], demonstrating the effect of inclusion of the g~0~ parameter (Eq.
@@
 Relying on the *Sureau* model presented in @martinstpaul2017 we illustrate the critical role of g~min~ in defining the desiccation time (Fig.
-5).
+2).
 
 ```{r fig.height=6, echo=FALSE, eval=includefigs}
+plotfig("output/figure2.png")
+```
+
+\
+(Figure 2 here)
+
+\
+**Figure 2**.
+Simulations with the *Sureau* model demonstrating the effect of g~min~ on desiccation tolerance of plants.
+The *Sureau* model simulates water transport in the soil-plant-atmosphere continuum, and includes a detailed representation of capacitance in stem and leaf tissues.
+(a) Soil relative extractable water (REW; 1 = field capacity, 0 = permanent wilting point) for the two simulations, using a minimum conductance (g~min~) of 2 or 4 mmol m^-2^ s^-1^ - all other parameters were equal.
+(b) Water potential in the soil and leaf as the dry-down progresses.
+(c) Progression of percent loss conductivity (PLC) of the xylem.
+Dashed line is at a PLC of 88%, indicating possible mortality.
+
+## 3. Towards a new model formulation
+
+\
+We suggest that the minimum conductance in the Ball-Berry model should include both a g~0~ and a g~min~ term, since g~0~ represents the minimum reached during low light and conditions of low photosynthesis, and g~min~ represents the minimum reached during severe drought.
+Moreover, we suggest that the minimum conductance is not simply added to the photosynthesis-dependent term (right-hand side in Eq.
+1), but used as an actual minimum.
+Thus,
+
+$$ g_s = max \big[ max(g_{min}, g_0), g_1 \frac{A_n}{C_a}f(D) \big]$$
+
+\
+This model for g~s~ will converge to g~0~ during periods of low photosynthesis, and to g~min~ during drought if we further include a model for the dependence of g~0~ on water availability [see e.g.
+@misson2004].
+Another advantage is that independent estimates of g~0~ can be used, and not via regression, which produces the undesirable correlation with estimates of g~1~.
+The above formulation is yet to be tested against data, but we propose this test to be done with data from drought and non-drought conditions.
+
+# III. Comparison of various definitions of minimum conductance
+
+\
+Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories.
+The absolute minimum conductance attained by leaves is that through the cuticle only.
+Measurements of cuticular conductance (g~cuti~) are typically made on isolated non-stomatal cuticles [@riederer2001], via special gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata [@kerstiens1996].
+We further discuss some important aspects of cuticular conductance in Section V.
+
+\
+The minimum conductance of intact leaves is typically measured by mass loss of detached leaves (MLD).
+In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out.
+Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after some time a constant low rate is achieved.
+From this minimum transpiration rate (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), the minimum conductance (g~min~) can be estimated using measured vapour pressure deficit [@sinclair1986].
+Although this method resembles conditions that plants may experience during a dry-down or periods of extreme stress, some uncertainties regarding the methods for measurement remain.
+In our own work, we have found that in some leaves, water loss rate increases (rather than decreases) some time after leaf detachment (see Methods S2).
+It is likely that the relatively rapid dry-down sometimes causes artefacts, and must be carefully avoided [see also @heinsoo1998].
+
+\
+We contrasted g~cuti~ and g~min~ with three estimates of stomatal conductance when photosynthesis rates are low or zero during non-drought conditions: night-time conductance (g~dark~) (further discussed in Section VII), conductance at low photosynthetically active radiation (PAR) (0 - 40 $\mu$ mol m^-2^ s^-1^) during the day, and conductance during conditions when photosynthesis rates are very low (< 1 $\mu$ mol m^-2^ s^-1^) (but excluding low PAR and drought).
+
+\
+We quantitatively compared each of the different conductances by synthesizing existing data.
+Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but were not used since they included no metadata).
+Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see Methods S1 for description and references).
+For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Methods S1 for full details of the compilation).
+The database includes a total of 221 species (136 Angiosperm, 49 Gymnosperm, 1 Pteridophyte), from 57 taxonomic families, with woody species making up roughly 2/3 of the data.
+
+\
+All conductance values were converted to per unit projected surface area, allowing direct comparison to stomatal conductance data which are typically presented as such.
+Estimates of g~dark~ were taken from @lombardozzi2017, who compiled measurements of ca.
+150 species during the night (we selected only gas exchange based measurements from their database).
+Daytime values (g~low PAR~ and g~low A~) were estimated from an update to the @lin2015 database by taking the appropriate subsets, and averaging the g~s~ values by species within study.
+All data and code to reproduce the database and analyses are available online (see Acknowledgments).
+
+\
+The results of the quantitative comparison are presented in Fig.
+3.
+From this compilation, a few striking differences between estimates are apparent.
+Cuticular conductance averaged 0.45 mmol m^-2^ s^-1^, ten-fold lower compared to g~min~ (4.9 mmol m^-2^ s^-1^).
+This result is in line with the methods comparison of @kerstiens1996.
+More recently, @schuster2017 argued, based on data for 8 species, that g~min~ from MLD is comparable to g~cuti~.
+However in the same study, a literature compilation of hundreds of values showed that g~cuti~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cuti~ were pooled in their analysis.
+Nonetheless, the comparison between g~cuti~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cuti~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions.
+@boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.
+
+\
+Conductance in the dark (g~dark~) (mean = 41.4 mmol m^-2^ s^-1^) was on average eight-fold higher than g~min~.
+It is no new observation that stomata do not close fully in the dark (further discussed in Section VII).
+Previous studies have directly compared g~dark~ and g~min~ on the same species, and generally concluded that g~dark~ is much higher than g~min~ [@hygen1954; @koerner1994; @walden-coleman2013], but an exception is @cavender-bares2007 who found the two rates to be similar in oaks.
+In turn, g~dark~ could not be differentiated from stomatal conductance measurements during daytime in very low light (mostly just after dawn, or before sunset) (g~low PAR~), or during conditions when photosynthetic rates were very low (due to very high temperature, VPD, or other factors) (g~low A~).
+These comparisons demonstrate that g~s~ during non-drought conditions when photosynthetic rates are zero or negligible are much higher than the minimum reached in simulated drought conditions (g~min~).
+
+```{r fig.width=4,echo=FALSE, eval=includefigs}
+plotfig("output/figure3.png")
+```
+
+\
+(Figure 3 here)
+
+\
+**Figure 3**.
+Comparison of various estimates of the (presumed) minimum conductance.
+Error bars are 95\% confidence intervals.
+Grey dots are the original data (but a few data points occur outside the figure range).
+Different letters denote significant differences (at $\alpha$ = 0.05).
+g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~low PAR~ leaf conductance during low light (photosynthetically active radiation, PAR); g~low A~ leaf conductance during periods of very low photosynthesis.
+See text for data sources and methods.
+
+# IV. Problems with estimating minimum conductance from regression
+
+The previous section discussed and compared methods for more or less direct measurements of various definitions of minimum conductance.
+The approach taken by the majority of vegetation models is however very different.
+Usually, g~0~ (for use in Eq.
+1) is estimated from regression with g~s~ as the response variable, and the right-hand side of Eq.
+1 as the predictor (a combination of measured photosynthesis rate, air humidity, and CO~2~ concentration).
+The g~0~ parameter is thus estimated as the intercept.
+Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated by this method.
+The difficulty of the estimation can already be anticipated from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g.
+@leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species); presumably because unreliable estimates were obtained, b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though it is clearly non-sensical to suggest negative conductance values.
+@barnard2013 also mentioned the difficulty of fitting Eq.
+1 to estimate g~0~, but did not present specific details - the following analysis builds on their work by demonstrating statistical uncertainties.
+
+\
+It is also telling that there are few reports on the intra-specific plasticity, or inter-specific variation of g~0~, perhaps because it is so poorly estimated.
+Exceptions are @duarte2016, who found a lower g~0~ in a heat wave treatment in Douglas-fir (although g~1~ was unaffected), but in their case g~0~ was very accurately estimated by careful multi-point light response curves.
+Another exception is @misson2004, who report on a close negative correlation between pre-dawn leaf water potential and g~0~ in Ponderosa pine (again, the g~1~ parameter was unaffected).
+
+\
+In the following we demonstrate that the fitting process is problematic for three reasons: 1) the estimates of g~0~ and g~1~ are highly correlated, 2) the precision for g~0~ is generally much lower than for g~1~, and 3) for data that have a worse fit overall, the g~0~ estimates are elevated.
+We thus conclude that g~0~ should really not be estimated from regression, although this is the most common method applied.
+In addition, it is difficult - and generally not recommended - to accurately measure low fluxes with a portable gas exchange system, and great care has to be taken to arrive at reasonable estimates of g~0~ that way.
+
+\
+Typically, gas exchange data are collected across a range of conditions, and used to plot g~s~ versus a combined term including photosynthesis rate, CO~2~ concentration, and air humidity.
+An example dataset is shown in Fig.
+4a, together with a fitted linear regression line.
+We show in Fig.
+4b that, estimates of g~0~ and g~1~ are statistically correlated, that is, their confidence intervals are not independent.
+Thus large estimated values for g~1~ lead to low estimates for g~0~ and vice versa.
+The consequence is that we cannot use estimates of g~0~ from this approach in a compilation, because these estimates depend on g~1~.
+The correlation between slopes and intercepts is not unique to Eq.
+1, but a general property of linear models [@stapleton1995, @becker2007].
+
+```{r fig.height=4,echo=FALSE, eval=includefigs}
+plotfig("output/figure4.png")
+```
+
+\
+(Figure 4 here)
+
+\
+**Figure 4**.
+Example linear regression of stomatal conductance (g~s~) against a combination of terms, in this case the linearized version of the model of Medlyn et al.
+(2011) applied to a leaf gas exchange dataset of @martinstpaul2012.
+This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al.
+(2015).
+The solid line is regression line, and the shaded area a 95% confidence interval for the mean.
+The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression.
+The dotted ellipse is a bivariate 95% confidence interval for slope and intercept.
+The symbols represent 1000 bootstrap samples of the coefficients.
+
+\
+Next we study the values obtained when fitting Eq.
+1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation @miner2017.
+The (updated version of) the Lin et al.
+(2015) database includes more than 15 thousand gas exchange measurements on over 300 species.
+After selecting species/site combinations with more than n=15, we produced 78 estimates of g~1~ and g~0~ with non-linear regression of the @medlyn2011 model of stomatal conductance.
+The compilation by @miner2017 includes 233 estimates of g~0~ and *m* (they compiled parameters for original model of Ball-Berry, equivalent to g~1~) for 172 species (including woody plants and crops).
+
+\
+For both databases, estimates of g~0~ are inflated when the model fits poorly (Fig.
+5 a,b).
+This can be understood by considering that a poor fit often results in the flattening out of the regression line, thus giving a large value for the intercept.
+A poor fit is often obtained when there is little variation in the right-hand side of Eq.
+1, usually because there was low variation in environmental conditions (humidity, light, temperature).
+We confirm this by showing that the standard error (SE) of g~0~ increases when the coefficient of variation of the right-hand side of Eq.
+1 is lower (Fig.
+5c).
+
+```{r fig.height=4,echo=FALSE, eval=includefigs}
 plotfig("output/figure5.png")
@@
 **Figure 5**.
-Simulations with the *Sureau* model demonstrating the effect of g~min~ on desiccation tolerance of plants.
-The *Sureau* model simulates water transport in the soil-plant-atmosphere continuum, and includes a detailed representation of capacitance in stem and leaf tissues.
-(a) Soil relative extractable water (REW; 1 = field capacity, 0 = permanent wilting point) for the two simulations, using a minimum conductance (g~min~) of 2 or 4 mmol m^-2^ s^-1^ - all other parameters were equal.
-(b) Water potential in the soil and leaf as the dry-down progresses.
-(c) Progression of percent loss conductivity (PLC) of the xylem.
-Dashed line is at a PLC of 88%, indicating possible mortality.
+Statistical uncertainty in the estimation of g~0~ from regression, demonstrated with two parameter databases.
+(a) We fitted the linearized form of the @medlyn2011 model to each of the datasets in the @lin2015 leaf gas exchange database, showing that for poorly fitted relationships (low R^2^), inflated estimates of g~0~ are obtained.
+Vertical lines are 95% confidence intervals.
+The grey line is a fitted loess smoother with 95% confidence interval.
+Note the wide confidence intervals, and frequent negative values.
+(b) Similar to (a), but using the published compilation by @miner2017.
+The grey line is a fitted loess smoother with 95% confidence interval.
+(c) Using the fits from panel (a), a demonstration that the standard error (SE) of g~0~ is much higher when the coefficient of variation (CV) of the predictor (i.e.
+right-hand side of the equation being fitted) is lower.
 
@@
 A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these processes have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b].
-However, a few key points should be summarized as they are relevant to the current discussion, in particular when we aim to interpret variation in literature values of g~min~ (Section VII).
+However, a few key points should be summarized as they are relevant to the current discussion, in particular when we aim to interpret variation in literature values of g~min~ (Section VIII).
 
@@
 Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle [@kerstiens1996; @riederer2001; @schreiber1996].
-The main barrier to diffusion is actually a very thin layer at the leaf surface.
+The main barrier to diffusion is actually a very thin layer of wax at the leaf surface.
 Because most of the resistance is located in such a thin layer, g~cuti~ does not correlate with the thickness of the cuticle [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b].
@@
 
-# VI. Nighttime conductance
+# VI. Contribution of stomata
+
+The minimum leaf conductance includes two pathways: across the cuticle, and through the (potentially incompletely closed) stomata.
+Only a few studies have directly quantified the stomatal component of minimum conductance, in contrast to the wealth of information on the cuticular component.
+A detailed study of *Hedera helix*  concluded that 35% of water loss occurred across the stomatal pores (though closed), and 65% across the non-stomatal part of the cuticle, despite the fact that the stomata presumably covered only a small fraction of the leaf [@santrucek2004].
+For ten *Sorghum* genotypes, @muchow1989 reported a strong positive correlation between stomatal density and g~min~, suggesting leaky stomata contribute substantially to g~min~.
+For seven conifer species, @brodribb2014 reported that 50 - 94% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky, and likely incompletely closed.
+However @santrucek2004 found that the cuticle of the stomatal side (but excluding stomata itself) in *Hedera helix* was many times more more permeable than the non-stomatal side, thus providing another potential explanation for this difference.
+Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles, and the contribution of leaky stomata, and potentially more permeable cuticle on the abaxial side, likely varies between species.
+
+\
+One key question is whether incomplete stomatal closure is under the plant's control, or whether it is an unavoidable consequence of imperfect stomata, damage, blocking by particles, etc.
+For example, endophytic fungi - which commonly colonize plant leaves, have been shown to prevent stomatal closure and greatly increase water use  [@arnold2007].
+As we discuss in the next section, there is ample evidence to suggest that stomata not only stay in the dark, but that plants actively control stomatal conductance during the night.
+During severe drought, there is less evidence for such active control, and it seems likely incomplete closure is not under the plant's control as there is no obvious reason to keep stomata open.
+Some conditions (especially rapid drying conditions) may lead to excessive drying of the epidermis, which can physically pull apart the stomata ['mechanical advantage', @buckley2005].
+
+# VII. Nighttime conductance
 
@@
 
-# VII. Environmental and ecological variation in minimum conductance
+# VIII. Environmental and ecological variation in minimum conductance
 
-## 1. Contribution of stomata to minimum conductance
-
-In order to understand variation in g~min~, we have to know more about the role of the cuticle vs.
-leaky or incompletely closed stomata.
-A few studies have directly quantified the stomatal component of g~min~.
-A detailed study of *Hedera helix*  concluded that 35% of water loss occurred across the stomatal pores (though closed), and 65% across the non-stomatal part of the cuticle, despite the fact that the stomata presumably covered only a small fraction of the leaf [@santrucek2004].
-For ten Sorghum genotypes, @muchow1989 reported a strong positive correlation between stomatal density and g~min~, suggesting leaky stomata contribute substantially to g~min~.
-For seven conifer species, @brodribb2014 reported that 50 - 94% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky.
-However @santrucek2004 found that the cuticle of the stomatal side (but excluding stomata itself) in *Hedera helix* was many times more more permeable than the non-stomatal side, thus providing another potential explanation for this difference.
-Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles, and the contribution of leaky stomata, and potentially more permeable cuticle on the abaxial side, likely varies between species.
-
-## 2. Minimum conductance shows little meaningful interspecific variation
+## 1. Minimum conductance shows little meaningful interspecific variation
 
 \
-In the following sections, we review the quite substantial literature on g~min~ measured with the MLD technique, stretching back to the 1930s [@pisekberger1938], focussing on environmental and ecological determinants.
+In the following sections, we review the quite substantial literature on g~min~ measured with the mass loss of detached leaves (MLD) technique, stretching back to the 1930s [@pisekberger1938], focussing on environmental and ecological determinants.
 We also further analyze the literature compilation of g~min~ (Fig.
-6).
+6), and separately analyze crop species and their genotypic variation.
 Comparing all available data, we did not find significant relationships between g~min~ and climate of origin, nor were there meaningful relationships with other traits (see Methods S1).
@@
 Based on their own compilation of the literature, @schuster2017 concluded no significant differences in g~min~ or g~cuti~ (analyzed together) by plant growth form.
-One notable exception is @brodribb2014, who showed a correlation between rainfall at species origin (of the driest quarter), and g~min~.
+One notable exception is @brodribb2014, who showed a correlation between rainfall at species origin (of the driest quarter), but used a gas-exchange approach (over many weeks), not MLD.
 Such correlation did not bear out across all species in our database, and interesting to note is that *Eucalyptus* species (n = 11, included in *Myrtales*) - all measured on the driest continent of Australia - have g~min~ values slightly higher than the average (Fig.
@@
 6c, we have compiled a number of studies in crops, demonstrating not just variation in g~min~ between crops, but also the wide range in g~min~ between genotypes.
-Again, it is striking that the wide variation in g~min~ cannot be easily explained by variation in leaf or other traits or chemical and structural components of the cuticle [@bengtson1978; @james2008; @saito2010].
+Again, it is striking that the wide variation in g~min~ cannot be easily explained by variation in leaf or other traits or chemical and structural components of the cuticle [@bengtson1978; @james2008; @saito2010], suggesting a significant role of incomplete stomatal closure.
 
@@
 The quick increase in g~min~ at high temperature may well be a crucial component of the ability of plants to tolerate heat waves [@drake2018].
-In support of the link between g~min~ and heat tolerance, @schuster2016a reported on a negative between thermal tolerance and g~min~ across nine species, so that species with improved tolerance to very high temperature had a low g~min~.
+In support of the link between g~min~ and heat tolerance, @schuster2016a reported on a negative relationship between thermal tolerance and g~min~ across nine species, so that species with improved tolerance to very high temperature had a low g~min~.
 
@@
 10 years.
-The most likely explanation for this change in g~min~ is the continued exposure to wind, rain, and abrasives, which have been shown to damage the cuticle and increase its conductance (see Section III).
+If this effect is due to properties of the cuticle alone, a possible explanation for the increase in g~min~ is the continued exposure to wind, rain, and abrasives, which have been shown to damage the cuticle and increase its conductance (see discussion in Section V).
+Another possibility is that the contribution of stomata to g~min~ increases with leaf age, as reported by @jordan2007.
 In plants that maintain several cohorts of leaves, finding an appropriate value of g~min~ for use in models will have to take into account the leaf age effect.
@@
 
-# VIII. Conclusions
+# IX. Conclusions
 
@@
 Thus when modelling night-time or low-light conductance, a different g~0~ should be used than when modelling the drought response of plants.
-
-\
-We suggest that the minimum conductance in the Ball-Berry model should include both a g~0~ and a g~min~ term, since g~0~ represents the minimum reached during low light and conditions of low photosynthesis, and g~min~ represents the minimum reached during severe drought.
-Moreover, we suggest that the minimum conductance is not simply added to the photosynthesis-dependent term (right-hand side in Eq.
-1), but used as an actual minimum.
-Thus,
-
-$$ g_s = max( min(g_{min}, g_0), g_1 \frac{A_n}{C_a}f(D))$$
-
-\
-This model for g~s~ will converge to g~0~ during periods of low photosynthesis, and to g~min~ during drought if we further include a model for the dependence of g~0~ on water availability [see @misson2004].
-Another advantage is that independent estimates of g~0~ can be used, and not with regression which produces the undesirable correlation with estimates of g~1~.
-The above formulation is yet to be tested against data, but we propose this test to be done combining data from drought and non-drought conditions.
+We suggest a new model form that includes both g~0~ and g~min~ with some desirable properties in Section II.
 
@@
 Perhaps this plasticity is the reason that g~min~ did not vary predictably between species.
+Another possibility is that the lack of standardized methods for measurement preclude clear comparisons between species.
 Future studies should compare g~min~ on many species grown in the same conditions to better understand the adaptive value of the minimum water loss rate of leaves.
@@
 This work was made possible by the ARC Linkage Project LP140100232.
+All raw data, code and a fully reproducible workflow which generates the figures, this manuscript and all analyses is available at https://www.github.com/remkoduursma/g0paper.
 

### 2018-05-23 R/figures.R

@@ -67,7 +67,7 @@ figure_g0g1_cor <- function(lin2015, group, legend=FALSE){
   el <- confidenceEllipse(fit, draw=FALSE)
   set.seed(123)
   
-  b <- car::Boot(fit)$t
+  b <- suppressMessages(car::Boot(fit)$t)
   
   rm(x, envir=.GlobalEnv)
   

### 2018-05-23 R/load.R

@@ -119,8 +119,6 @@ species_classes <- read.csv("data/Species_classifications.csv", stringsAsFactors
 # Data from gmindatabase
 gmindat <- read.csv("https://raw.githubusercontent.com/RemkoDuursma/gmindatabase/master/combined/gmindatabase.csv",
                      stringsAsFactors = FALSE) %>%
-# gmindat <- read.csv("c:/repos/gmindatabase/combined/gmindatabase.csv",
-#                     stringsAsFactors = FALSE) %>%
   filter(gmin > 0) %>%
   group_by(species) %>%
   dplyr::summarize(gmin = mean(gmin),
@@ -134,7 +132,6 @@ cropgmin <- read.csv(paste0("https://raw.githubusercontent.com/RemkoDuursma/",
 
 gminall <- dget(paste0("https://raw.githubusercontent.com/",
                        "RemkoDuursma/gmindatabase/master/combined/gminall.rdput"))
-#gminall <- dget("c:/repos/gmindatabase/combined/gminall.rdput")
 
 #
 if(TRUE){
@@ -179,6 +176,8 @@ if(!file.exists(clsfile)){
   cls <- readRDS(clsfile)
 }
 cls <- as_dataframe_cls(cls)
+rownames(cls) <- NULL
+cls <- cls[!duplicated(cls$species),]
 
 gmindat <- left_join(gmindat,  cls, by="species")
 

### 2018-05-23 R/make_figures.R

@@ -1,17 +1,16 @@
 
 source("R/load.R")
 
+to.pdf(figure_sim(), "output/figure1.pdf", height=6, width=6)
 
-to.pdf(figure_gmin_review(gdfr), "output/figure1.pdf", height=5, width=5)
+to.pdf(figure_sureau(planta, plantb), "output/figure2.pdf", height=4, width=9)
+
+to.pdf(figure_gmin_review(gdfr), "output/figure3.pdf", height=5, width=5)
 
 to.pdf(figure_g0g1_cor(lin2015, "Nicolas Martin-StPaul_Quercus ilex_StPaul_Puechabon"), 
-       "output/figure2.pdf", height=4, width=8)
+       "output/figure4.pdf", height=4, width=8)
 
-to.pdf(figure_R2g0(lin2015coef, miner), "output/figure3.pdf", height=4, width=9)
-
-to.pdf(figure_sim(), "output/figure4.pdf", height=6, width=6)
-
-to.pdf(figure_sureau(planta, plantb), "output/figure5.pdf", height=4, width=9)
+to.pdf(figure_R2g0(lin2015coef, miner), "output/figure5.pdf", height=4, width=9)
 
 to.pdf(gmin_3panel(gmindat, cropgmin), "output/figure6.pdf", height=3, width=8)
 

### 2018-06-02 manuscript.Rmd
update rosana contact and acknowledgment
co-author comments incorporated, order changed again; png removed

@@
 \
-Remko A. Duursma^1^, Chris J. Blackman^1^, Rosana Lopéz^1^, Nicolas K. Martin-StPaul^2^, Hervé Cochard^3^, Belinda E. Medlyn^1^
+Remko A. Duursma^1^, Chris J. Blackman^1^, Rosana Lopéz^1,3^, Nicolas K. Martin-StPaul^2^, Hervé Cochard^3^, Belinda E. Medlyn^1^
 
@@
 ```{r include=FALSE}
-includefigs <- FALSE
-library(png)
-library(grid)
 library(knitr)
@@
 ```{r, echo=FALSE}
-tab <- read.csv(text="I. Introduction
-II. Use of minimum conductance in models
+tab <- read.csv(text=
+
+    "I. Introduction                                                        
+    II. Comparison of various definitions of minimum conductance           
+    III. Cuticular conductance                                             
+    IV. Contribution of stomata                                            
+    V. Environmental and ecological variation in minimum conductance       
+    \t1. Minimum conductance shows little meaningful interspecific variation
+    \t2. Acclimation to the environment                                     
+    VI. Use of minimum conductance in models                               
     \t1. Models of water-use efficiency                                     
     \t2. Models of plant desiccation                                        
-\t3. Towards a new model formulation
-III. Comparison of various definitions of minimum conductance
-IV. Problems with estimating Minimum conductance from regression
-V. Cuticular conductance
-VI. Contribution of stomata
-VII. Nighttime conductance
-VIII. Environmental and ecological variation in minimum conductance 
-\t1. Contribution of stomata to minimum conductance
-\t2. Minimum conductance shows little meaningful interspecific variation
-\t3. Acclimation to the environment
-IX. Conclusions", header=F)
-
+    \t3. Problems with estimating g~0~ from regression                      
+    \t4. Nighttime conductance                                              
+    \t5. Towards a new model formulation                                    
+    VII. Conclusions",
+    header=F)
 knitr::kable(tab, col.names="")
@@
 However, water loss continues through the cuticle and incompletely closed stomata, together constituting the leaf minimum conductance (g~min~).
-In this review, we illustrate the role of g~min~ in models of leaf gas exchange, critically evaluate sources of variation in g~min~, and quantitatively compare various methods for its estimation.
-We pay special attention to the role of the minimum conductance in the Ball-Berry model of stomatal conductance, and caution against the usual regression-based method for its estimation.
+In this review, we critically evaluate sources of variation in g~min~, quantitatively compare various methods for its estimation, and illustrate the role of g~min~ in models of leaf gas exchange.
 A literature compilation of g~min~ as measured by weight loss of detached leaves is presented, which shows much variation in this trait that is not clearly related to species groups, climate of origin, or leaf type.
 Much evidence points to the idea that g~min~ is highly responsive to the growing conditions of the plant, including soil water availability, temperature, and air humidity - as we further demonstrate with two case studies.
+We pay special attention to the role of the minimum conductance in the Ball-Berry model of stomatal conductance, and caution against the usual regression-based method for its estimation.
 The synthesis presented here provides guidelines for use of g~min~ in ecosystem models, and points to clear research gaps for this drought tolerance trait.
@@
 Other measurements allow for the estimation of water loss through incompletely closed and broken stomata.
-Of particular interest because of its simplicity is the method where leaves are detached from the plant, and their weight loss monitored over time.
-This method (mass loss of detached leaves, MLD) aims to simulate field conditions during severe drought, when water supply to the leaf has practically ceased, however proceeds much quicker (typically 0.5-2 days).
-This measurement also includes both surfaces of the leaf - not just the stomata-free side.
+Of particular interest because of its simplicity is the method where leaves are detached from the plant, their weight loss monitored over time, and expressed as the minimum conductance (g~min~).
+This method (mass loss of detached leaves, MLD) aims to simulate field conditions during severe drought, when water supply to the leaf has practically ceased, although it proceeds much more quickly (typically 0.5-2 days, compared to weeks or even months in field conditions).
+This measurement also includes both surfaces of the leaf - not just the adaxial side.
 A comprehensive compilation of estimates of g~min~ is currently lacking, and is needed not just to parameterize models, but also to study sources of variation in this overlooked plant trait.
@@
 \
-In this review, we discuss the role of the minimum conductance in models of plant water use, and critically evaluate sources of variation in this overlooked parameter.
-Previous reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017].
+In this review, we discuss the role of the minimum conductance in models of plant water use, and critically evaluate sources of variation in this parameter.
+Previous reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008; @schuster2017; @kerstiens1996; @fernandez2017].
 Other work has focused on a different definition of the minimum conductance (g~0~): the value that should be used in models of plant water use [e.g.
-@barnard2013; @lombardozzi2017], and includes both g~min~ and a 'stomatal residual' due to the fact that stomata do not completely close during periods of zero photosynthesis [@dekauwe2015].
+@barnard2013; @lombardozzi2017], which includes both g~min~ and a 'stomatal residual' due to the fact that stomata do not completely close during periods of zero photosynthesis [@dekauwe2015].
 This discussion often centers around leaf conductance during the night [@lombardozzi2017], but we must also consider appropriate values for the minimum conductance during the day, for example during very high (or low) temperature, transient low light, and extreme drought.
-Thus there is a disconnect between the dizzying level of understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models.
+Thus there is a disconnect between the detailed understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models.
 
@@
 We aim to improve the connection between the physiology of minimum conductance, empirical approaches, and model implementations by synthesizing the state of knowledge.
-Our goals are to a) demonstrate the need to include a non-zero minimum conductance in models of water-use efficiency and drought responses, and discuss current practices b) compare various definitions of g~min~ quantitively, which have sometimes been previously assumed equal, c) present a new compilation of minimum conductance measurements, test for ecologically meaningful patterns, and demonstrate the large acclimatory potential of this trait to environmental drivers.
+Our goals are to a) quantitively compare various definitions of g~min~, which have sometimes been assumed equal, b) present a new compilation of minimum conductance measurements, which we use to test for ecologically meaningful patterns, and demonstrate the large acclimatory potential of this trait to environmental drivers, and c) demonstrate the need to include a non-zero minimum conductance in models of water-use efficiency and drought responses.
 
+\
 **Table 1.** Definitions of minimum conductance.
@@
 
-# II. Use of minimum conductance in models
+# II. Comparison of various definitions of minimum conductance
+
+\
+Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories.
+The absolute minimum conductance attained by leaves is that through the cuticle only.
+Measurements of cuticular conductance (g~cuti~) are typically made on isolated non-stomatal cuticles [@riederer2001], via special gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata [@kerstiens1996].
+We further discuss some important aspects of cuticular conductance in Section III.
+
+\
+The minimum conductance of intact leaves is typically measured by mass loss of detached leaves (MLD).
+In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out.
+Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after some time a constant low rate is achieved.
+From this minimum transpiration rate (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), the minimum conductance (g~min~) can be estimated using measured vapour pressure deficit [@sinclair1986].
+Although this method resembles conditions that plants may experience during a dry-down or periods of extreme stress, some uncertainties regarding the methods for measurement remain.
+In our own work, we have found that in some leaves, water loss rate increases (rather than decreases) some time after leaf detachment (see Methods S2).
+It is likely that the relatively rapid dry-down sometimes causes artefacts, and must be carefully avoided [see also @heinsoo1998].
+
+\
+We contrasted g~cuti~ and g~min~ with three estimates of stomatal conductance when photosynthesis rates are low or zero during non-drought conditions: night-time conductance (g~dark~) (further discussed in Section VI.4), conductance at low photosynthetically active radiation (PAR) (0 - 40 $\mu$ mol m^-2^ s^-1^) during the day, and conductance during conditions when photosynthesis rates are very low (< 1 $\mu$ mol m^-2^ s^-1^) (but excluding low PAR and drought).
+
+\
+We quantitatively compared each of the different conductances by synthesizing existing data.
+Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but these were not used since they included no metadata).
+Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see Methods S1 for description and references).
+For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Methods S1 for full details of the compilation).
+The database includes a total of 221 species (136 Angiosperm, 49 Gymnosperm, 1 Pteridophyte), from 57 taxonomic families, with woody species making up roughly 2/3 of the data.
+
+\
+All conductance values were converted to per unit projected surface area, allowing direct comparison to stomatal conductance data which are typically presented in these units.
+Estimates of g~dark~ were taken from @lombardozzi2017, who compiled measurements of ca.
+150 species during the night (we selected only gas exchange based measurements from their database).
+Daytime values (g~low PAR~ and g~low A~) were estimated from an update to the @lin2015 database by taking the appropriate subsets, and averaging the g~s~ values by species within study.
+All data and code to reproduce the database and analyses are available online (see Acknowledgments).
+
+\
+The results of the quantitative comparison are presented in Fig.
+1.
+From this compilation, a few striking differences among estimates are apparent.
+Cuticular conductance averaged 0.45 mmol m^-2^ s^-1^, ten-fold lower compared to g~min~ (4.9 mmol m^-2^ s^-1^).
+This result is in line with the methods comparison of @kerstiens1996.
+More recently, @schuster2017 argued, based on data for 8 species, that g~min~ from MLD is comparable to g~cuti~.
+However in the same study, a literature compilation of hundreds of values showed that g~cuti~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cuti~ were pooled in their analysis.
+Nonetheless, the comparison between g~cuti~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cuti~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions.
+@boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.
+
+\
+Conductance in the dark (g~dark~) (mean = 41.4 mmol m^-2^ s^-1^) was on average eight-fold higher than g~min~.
+It is no new observation that stomata do not close fully in the dark (further discussed in Section VI.4).
+Previous studies have directly compared g~dark~ and g~min~ on the same species, and generally concluded that g~dark~ is much higher than g~min~ [@hygen1954; @koerner1994; @walden-coleman2013], but an exception is @cavender-bares2007 who found the two rates to be similar in oaks.
+In turn, g~dark~ could not be differentiated from stomatal conductance measurements during daytime in very low light (mostly just after dawn, or before sunset) (g~low PAR~), or during conditions when photosynthetic rates were very low (due to very high temperature, VPD, or other factors) (g~low A~).
+These comparisons demonstrate that g~s~ during non-drought conditions when photosynthetic rates are zero or negligible are much higher than the minimum reached in simulated drought conditions (g~min~).
+
+\
+(Figure 1 here)
+
+\
+**Figure 1**.
+Comparison of various estimates of the (presumed) minimum conductance.
+Error bars are 95\% confidence intervals.
+Grey dots are the original data (but a few data points occur outside the figure range).
+Different letters denote significant differences (at $\alpha$ = 0.05).
+g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~low PAR~ leaf conductance during low light (photosynthetically active radiation, PAR); g~low A~ leaf conductance during periods of very low photosynthesis.
+See text for data sources and methods.
+
+# III. Cuticular conductance
+
+\
+A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these topics have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b].
+However, a few key points should be summarized as they are relevant to the current discussion, in particular when we aim to interpret variation in literature values of g~min~ (Section V).
+
+\
+Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle.
+Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle [@kerstiens1996; @riederer2001; @schreiber1996].
+The main barrier to diffusion is actually a very thin layer of wax at the leaf surface.
+Because most of the resistance is located in such a thin layer, g~cuti~ does not correlate with the thickness of the cuticle [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b].
+Nonetheless there is considerable variation in cuticle thickness among plant species [@schuster2016a], along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010].
+If the cuticle thickness does not directly affect its conductance, what is the value of increasing the cuticle?
+
+\
+It is important to bear in mind the many other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, various interactions with insects including signalling and herbivory resistance [@mueller2008; @kerstiens1996c;@riederer2008], and foliar water uptake [@fernandez2017].
+The thickness of the cuticle may also confer mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear was proportional to the cuticle thickness).
+As such, thick cuticles may be advantageous in exposed environments to avoid excessive damage [@blackman2005].
+Prolonged exposure to wind has been shown to increase the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991], by dislodging cuticular compounds by abrasion [@rogge1993].
+Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986], most dramatically demonstrated in an ice storm, leading to much higher g~min~ [@boyce2003].
+
+\
+Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged.
+Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but was able to explain only some of the wide range of g~cuti~ measured across different plant species.
+Similarly, @hauke1998 found no relationship between the gradual decrease in g~cuti~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.
+
+# IV. Contribution of stomata
+
+The minimum leaf conductance includes two pathways: across the cuticle, and through the (potentially incompletely closed) stomata.
+Only a few studies have directly quantified the stomatal component of minimum conductance, in contrast to the wealth of information on the cuticular component.
+A detailed study of *Hedera helix*  concluded that 35% of water loss occurred across the stomatal pores (though closed), and 65% across the non-stomatal part of the cuticle, despite the fact that the stomata presumably covered only a small fraction of the leaf [@santrucek2004].
+For ten *Sorghum* genotypes, @muchow1989 reported a strong positive correlation between stomatal density and g~min~, suggesting leaky stomata contribute substantially to g~min~.
+For seven conifer species, @brodribb2014 reported that 50 - 94% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky, and likely incompletely closed.
+However @santrucek2004 found that the cuticle of the stomatal side (but excluding stomata itself) in *Hedera helix* was many times more permeable than the non-stomatal side, thus providing another potential explanation for this difference.
+Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles, and the contribution of leaky stomata, and potentially more permeable cuticle on the abaxial side, likely varies among species.
+
+\
+One key question is whether incomplete stomatal closure is under the plant's control, or whether it is an unavoidable consequence of imperfect stomata, damage, blocking by particles, etc.
+For example, endophytic fungi - which commonly colonize plant leaves, have been shown to prevent stomatal closure and greatly increase water use  [@arnold2007].
+As we discuss in the next section, there is ample evidence to suggest that stomata not only stay open in the dark, but that plants actively control stomatal conductance during the night.
+During severe drought, there is less evidence for such active control, and it seems likely incomplete closure is not under the plant's control as there is no obvious reason to keep stomata open.
+Some conditions (especially rapid drying conditions) may lead to excessive drying of the epidermis, which can physically pull apart the stomata ['mechanical advantage', @buckley2005].
+
+# V. Environmental and ecological variation in minimum conductance
+
+## 1. Minimum conductance shows little meaningful interspecific variation
+
+\
+In the following sections, we review the quite substantial literature on g~min~ measured with the mass loss of detached leaves (MLD) technique, stretching back to the 1930s [@pisekberger1938], focussing on environmental and ecological determinants.
+We also further analyze the literature compilation of g~min~ (Fig.
+2), and separately analyze crop species and their genotypic variation.
+Comparing all available data, we did not find significant relationships between g~min~ and climate of origin, nor were there meaningful relationships with other traits (see Methods S1).
+We thus found it difficult to explain the variation in g~min~ among species.
+A simple breakdown by taxonomic order (Fig.
+2b) revealed that grasses (*Poales*) have a higher g~min~ compared to other orders, and conifers (especially *Pinales*) tended towards the lower range of values (but were only significantly different from *Poales*).
+
+\
+Very few studies have found meaningful correlations between g~min~ and environmental factors or ecological categories.
+Based on their own compilation of the literature, @schuster2017 concluded no significant differences in g~min~ or g~cuti~ (analyzed together) by plant growth form.
+One notable exception is @brodribb2014, who showed a correlation between rainfall at species origin (of the driest quarter), but these authors used a gas-exchange approach (over many weeks), not MLD.
+This correlation did not hold across the species in our database.
+Of particular note, *Eucalyptus* species (n = 11, included in *Myrtales*) - all measured on the driest continent of Australia - have g~min~ values slightly higher than the average (Fig.
+2b).
+
+\
+In crop science, g~min~ has long been identified as a key drought tolerance trait [@sinclair1986].
+A number of studies have targeted g~min~ as a key trait for breeding more drought-tolerant crops, leading to comparisons of g~min~ across genotypes grown in the same conditions.
+For example, @james2008 compared g~min~ in 58 soybean (*Glycine max*) genotypes, and found > 2-fold variation that could not be easily explained by other traits.
+In Fig.
+2c, we have compiled a number of studies in crops, demonstrating not just variation in g~min~ among crops, but also the wide range in g~min~ among genotypes.
+Again, it is striking that the wide variation in g~min~ cannot be easily explained by variation in leaf or other traits or chemical and structural components of the cuticle [@bengtson1978; @james2008; @saito2010], suggesting a significant role of incomplete stomatal closure.
+
+\
+(Figure 2 here)
+
+\
+**Figure 2**.
+Analysis of a literature compilation of g~min~ estimates, as measured with mass loss of detached leaves.
+(a) Histogram (probability density) of all estimates (after averaging by species, n=`r nrow(gmindat)`), with a log-normal distribution curve (mean = `r round(exp(mean(log(gmindat$gmin))),2)`, sd = `r round(exp(sd(log(gmindat$gmin))),2)`).
+(b) g~min~ averaged by phylogenetic order (including only the top ten orders in the database).
+Bars are 95% confidence intervals.
+Numbers above the figure refer to the number of species.
+Different letters denote significant differences (at $\alpha$ = 0.05, adjusted for multiple comparisons), and grey symbols are species-level data.
+(c) g~min~ estimates for crops only, averaged by genotype.
+Bars denote the range, illustrating the wide range in g~min~ among genotypes for a particular crop species.
+Numbers above the figure refer to the number of genotypes included.
+
+## 2. Acclimation to the environment
+
+\
+While there is clearly considerable variation in g~min~ among species, a number of lines of evidence suggest that g~min~ also has great potential for plasticity.
+Here we summarize the literature on the acclimation of g~min~ to drought conditions, to changes in temperature and humidity, and the change in g~min~ with leaf age, altitude and other factors.
+This discussion is directly relevant to the use of g~min~ in models, because if the degree of plasticity is large, it complicates model parameterization.
+
+\
+There is a general tendency for a decreased g~min~ in plants acclimated to drought stress [@james2008].
+The magnitude of the decrease of g~min~ with acclimation to drought stress varied from -4 to -70% (across 10 studies, see Table S2), with a typical decrease in the order of 30-40%.
+In each of the studies summarized, plants were grown in well-watered or droughted conditions, and in one case a difference in g~min~ was demonstrated after just four days of drought exposure [@bengtson1978].
+We also demonstrate drought acclimation via a case study on 11 *Hakea* species grown in two watering treatments (Fig.
+3).
+All 11 species showed a decrease in g~min~ in the drought treatment (see Methods S3 for experimental details).
+This significant change in g~min~ with drought acclimation is likely an important component of the overall drought hardening of plants.
+
+\
+The idea that water limitation causes a reduction in g~min~ (via changes in the chemical composition of the cuticle) can be tested by inspecting the response to factors that increase evaporative demand.
+As one of the very few studies testing this idea, @fanourakis2013 reported much lower g~min~ in *Rosa* sp.
+plants grown in 60% vs.
+95% relative humidity.
+The difference could be attributed largely to a change in stomatal anatomy and lack of closure during desiccation, not changes in the cuticle *per se*.
+Also relevant is @sack2003, who reported lower g~min~ in sun leaves (for 2 out of 4 species) compared to shade leaves, again a lower g~min~ for leaves acclimated to high evaporative conditions.
+
+\
+A number of studies have attempted to attribute the drought-induced acclimation in g~min~ to a change in the chemical composition of the cuticle.
+@bengtson1978 reported lower g~min~ in six oat varieties in response to drought, and an increase in the amount of cuticular waxes, but could not find a relationship between the two and the response was highly genotype-specific.
+However, @premachandra1992 found that epicuticular wax load increased on leaves of non-irrigated Sorghum cultivars and was positively correlated with cuticular conductance and cell membrane stability.
+@mackova2013 found that adding ABA (simulating drought stress) to *Lepidium sativum* increased the chain length of cuticular waxes (but not total amounts).
+@bi2017 also reported that drought caused a change in cuticular wax production and chemical composition, but again in a highly genotype-specific way.
+Thus it can be concluded that the cuticle indeed changes substantially after a change in plant available water, but in a complex, species-specific manner that is yet to be connected directly to changes in g~min~.
+
+\
+(Figure 3 here)
+
+\
+**Figure 3**.
+Minimum conductance (g~min~) measured by mass loss of detached leaves on a variety *Hakea* species, a genus native to Australia.
+Plants were grown in containers in a grow house, and supplied with ample water or subjected to long-term (8 months) mild drought stress.
+Bars are labelled by leaf form : broadleaf (B) or needle-like (N), and ordered by g~min~ in the  well-watered treatment.
+For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`).
+
+\
+The response of g~min~ to temperature is more complex; it shows both a response to instantaneous changes in temperature, as well as an acclimatory response to growth temperature.
+@riederer2001 and @schuster2016b have both demonstrated a steep non-linear instantaneous response of g~cuti~ to temperature, with the response becoming especially steep at higher temperatures (> 40C).
+In *Eucalyptus haemastoma*, the response of g~min~ to temperature was so steep, that the proportion of cuticular to total transpiration increased from 2-3% at 20°C, to 40% at 38°C [@eamus2008].
+The mechanism of the instantaneous temperature response is complex, highly species-specific, and we refer to @schuster2016b for a detailed investigation.
+The rapid increase in g~min~ at high temperature may well be a crucial component of the ability of plants to tolerate heat waves [@drake2018].
+In support of the link between g~min~ and heat tolerance, @schuster2016a reported on a negative relationship between thermal tolerance and g~min~ across nine species, so that species with improved tolerance to very high temperature had a low g~min~.
+
+\
+Less well established is the acclimatory potential of g~min~ to changes in growth temperature.
+We present a case study on *Eucalyptus parramattensis* grown in whole-tree chambers (see Methods S3 for experimental details).
+The chambers either tracked ambient conditions, or were subjected to a +3°C warming treatment.
+After several months of growth in the treatments, g~min~ was measured at various temperatures, ranging from 17.5 to 27.5°C.
+We found a 56% decrease in g~min~ in the elevated temperature trees, but there was no clear pattern with measurement temperature (Fig.
+4).
+This decrease in g~min~ is consistent with the drought response, because again g~min~ is reduced in leaves that are subjected to conditions that increase evaporative demand [vapour pressure deficit was higher in the elevated temperature treatment, see @drake2018].
+The direction of this response is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment in *Pseudotsuga menziesii*, which persisted for some time.
+Responses across temperature gradients are more complex.
+In particular, the change in g~min~ with altitude has been well studied, and it is commonly  reported that g~min~ increases with altitude [e.g.
+@delucia1984; @herrick1991; @anfodillo2002].
+@fernandez2017 discuss the literature on altitude responses in detail and argue that the short growing seasons at high altitude are insufficient for complete maturation of leaf cuticles.
+
+\
+Finally we mention the striking effect of leaf age on g~min~.
+We have summarized five studies in Fig.
+S3, all of which reported an increased g~min~ for older leaves for the majority of species studied.
+The data reported by @jordan2007 are particularly impressive, as g~min~ in the woody shrub *Agastachys odorata* gradually increased with leaf age up to ca.
+10 years.
+If this effect is due to properties of the cuticle alone, a possible explanation for the increase in g~min~ is the continued exposure to wind, rain, and abrasives, which have been shown to damage the cuticle and increase its conductance (see discussion in Section III).
+Another possibility is that the contribution of stomata to g~min~ increases with leaf age, as reported by @jordan2007.
+In plants that maintain several cohorts of leaves, finding an appropriate value of g~min~ for use in models will have to take into account the leaf age effect.
+
+\
+(Figure 4 here)
+
+\
+**Figure 4**.
+Minimum conductance (g~min~) measured with mass loss of detached leaves on *Eucalyptus parramattensis* grown in whole-tree chambers.
+Trees were grown following ambient conditions, or in an elevated temperature treatment (+3 °C) [see @drake2018].
+Measurements of g~min~ were carried out at five different temperatures on replicate leaves.
+There was no consistent effect of measurement temperature on g~min~, but leaves of trees grown in elevated temperature showed on average a 45% reduction in g~min~ (p < 0.01, linear mixed-effects model).
+
+# VI. Use of minimum conductance in models
 
@@
 \
-Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987, see @damour2010 for a list of functions), and g~0~ is the value of g~s~ when A~n~ is zero.
+Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987, see @damour2010 for a list of functions), and g~0~ is the value of g~s~ when A~n~ is zero.
 The g~1~ parameter is directly related to the water-use efficiency: large values of g~1~ indicate low water-use efficiency.
@@
 It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model.
-In practice, g~0~ is estimated from a regression of Eq.
-1 with leaf gas exchange data, but as we will show, this may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.
+In practice, g~0~ is usually estimated from a regression of Eq.
+1 with leaf gas exchange data, but as we will show, this approach may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.
 
@@
 Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero.
-But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, and it is unclear whether the same g~0~ is reached in each of these cases.
+However, A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, and it is unclear whether the same g~0~ is reached in each of these cases.
 In the application of the Ball-Berry model, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005].
-It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cuti~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata.
+An alternative common assumption is that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cuti~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata.
 Photosynthesis also ceases at very high temperature, but stomata do not appear to always close in proportion to this decrease in photosynthesis, if at all [@drake2018; @urban2017].
@@
 Using the coupled leaf gas exchange model, we show in Fig.
-1 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^).
+5 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^).
 We also demonstrate the effect of g~0~ on modelling the T~leaf~ response.
@@
 
-```{r fig.height=6,echo=FALSE, eval=includefigs}
-plotfig("output/figure1.png")
-```
+\
+(Figure 5 here)
 
 \
-(Figure 1 here)
-
-\
-**Figure 1**.
+**Figure 5**.
 Simulations with a coupled leaf gas exchange model [@duursma2015], demonstrating the effect of inclusion of the g~0~ parameter (Eq.
@@
 Thus models that aim to predict when plants desiccate and die must include a minimum conductance term.
-A classic study by @pisekwinkler1953 already calculated the length of time needed to desiccate leaves to some critical low water content, given the minimum transpiration rate, and the saturated water content of the leaves.
+A classic study by @pisekwinkler1953 calculated the length of time needed to desiccate leaves to some critical low water content, given the minimum transpiration rate, and the saturated water content of the leaves.
 Based on that work, @burghardt2006 reported on a direct correlation between g~min~ and the survival time of leaves.
@@
 More recently, @gleason2014 and @blackman2016 proposed that embolism resistance along with whole plant capacitance and minimum transpiration rates all contribute to define the time to desiccation.
-Finally, @martinstpaul2017 demonstrated in a whole plant model of hydraulic failure that g~min~ was one of the key parameters to explain the drop in water potential below the cavitation threshold, because stomata generally close well before this threshold.
-Relying on the *Sureau* model presented in @martinstpaul2017 we illustrate the critical role of g~min~ in defining the desiccation time (Fig.
-2).
-
-```{r fig.height=6, echo=FALSE, eval=includefigs}
-plotfig("output/figure2.png")
-```
+Building on this work, @martinstpaul2017 demonstrated in a whole plant model of hydraulic failure that g~min~ was one of the key parameters to explain the drop in water potential below the cavitation threshold, because stomata generally close well before this threshold.
+Applying the *Sureau* model presented in @martinstpaul2017 we illustrate the critical role of g~min~ in defining the desiccation time (Fig.
+6).
 
 \
-(Figure 2 here)
+(Figure 6 here)
 
 \
-**Figure 2**.
+**Figure 6**.
 Simulations with the *Sureau* model demonstrating the effect of g~min~ on desiccation tolerance of plants.
@@
 
-## 3. Towards a new model formulation
-
-\
-We suggest that the minimum conductance in the Ball-Berry model should include both a g~0~ and a g~min~ term, since g~0~ represents the minimum reached during low light and conditions of low photosynthesis, and g~min~ represents the minimum reached during severe drought.
-Moreover, we suggest that the minimum conductance is not simply added to the photosynthesis-dependent term (right-hand side in Eq.
-1), but used as an actual minimum.
-Thus,
-
-$$ g_s = max \big[ max(g_{min}, g_0), g_1 \frac{A_n}{C_a}f(D) \big]$$
-
-\
-This model for g~s~ will converge to g~0~ during periods of low photosynthesis, and to g~min~ during drought if we further include a model for the dependence of g~0~ on water availability [see e.g.
-@misson2004].
-Another advantage is that independent estimates of g~0~ can be used, and not via regression, which produces the undesirable correlation with estimates of g~1~.
-The above formulation is yet to be tested against data, but we propose this test to be done with data from drought and non-drought conditions.
-
-# III. Comparison of various definitions of minimum conductance
-
-\
-Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories.
-The absolute minimum conductance attained by leaves is that through the cuticle only.
-Measurements of cuticular conductance (g~cuti~) are typically made on isolated non-stomatal cuticles [@riederer2001], via special gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata [@kerstiens1996].
-We further discuss some important aspects of cuticular conductance in Section V.
-
-\
-The minimum conductance of intact leaves is typically measured by mass loss of detached leaves (MLD).
-In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out.
-Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after some time a constant low rate is achieved.
-From this minimum transpiration rate (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), the minimum conductance (g~min~) can be estimated using measured vapour pressure deficit [@sinclair1986].
-Although this method resembles conditions that plants may experience during a dry-down or periods of extreme stress, some uncertainties regarding the methods for measurement remain.
-In our own work, we have found that in some leaves, water loss rate increases (rather than decreases) some time after leaf detachment (see Methods S2).
-It is likely that the relatively rapid dry-down sometimes causes artefacts, and must be carefully avoided [see also @heinsoo1998].
-
-\
-We contrasted g~cuti~ and g~min~ with three estimates of stomatal conductance when photosynthesis rates are low or zero during non-drought conditions: night-time conductance (g~dark~) (further discussed in Section VII), conductance at low photosynthetically active radiation (PAR) (0 - 40 $\mu$ mol m^-2^ s^-1^) during the day, and conductance during conditions when photosynthesis rates are very low (< 1 $\mu$ mol m^-2^ s^-1^) (but excluding low PAR and drought).
-
-\
-We quantitatively compared each of the different conductances by synthesizing existing data.
-Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but were not used since they included no metadata).
-Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see Methods S1 for description and references).
-For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Methods S1 for full details of the compilation).
-The database includes a total of 221 species (136 Angiosperm, 49 Gymnosperm, 1 Pteridophyte), from 57 taxonomic families, with woody species making up roughly 2/3 of the data.
-
-\
-All conductance values were converted to per unit projected surface area, allowing direct comparison to stomatal conductance data which are typically presented as such.
-Estimates of g~dark~ were taken from @lombardozzi2017, who compiled measurements of ca.
-150 species during the night (we selected only gas exchange based measurements from their database).
-Daytime values (g~low PAR~ and g~low A~) were estimated from an update to the @lin2015 database by taking the appropriate subsets, and averaging the g~s~ values by species within study.
-All data and code to reproduce the database and analyses are available online (see Acknowledgments).
-
-\
-The results of the quantitative comparison are presented in Fig.
-3.
-From this compilation, a few striking differences between estimates are apparent.
-Cuticular conductance averaged 0.45 mmol m^-2^ s^-1^, ten-fold lower compared to g~min~ (4.9 mmol m^-2^ s^-1^).
-This result is in line with the methods comparison of @kerstiens1996.
-More recently, @schuster2017 argued, based on data for 8 species, that g~min~ from MLD is comparable to g~cuti~.
-However in the same study, a literature compilation of hundreds of values showed that g~cuti~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cuti~ were pooled in their analysis.
-Nonetheless, the comparison between g~cuti~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cuti~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions.
-@boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.
-
-\
-Conductance in the dark (g~dark~) (mean = 41.4 mmol m^-2^ s^-1^) was on average eight-fold higher than g~min~.
-It is no new observation that stomata do not close fully in the dark (further discussed in Section VII).
-Previous studies have directly compared g~dark~ and g~min~ on the same species, and generally concluded that g~dark~ is much higher than g~min~ [@hygen1954; @koerner1994; @walden-coleman2013], but an exception is @cavender-bares2007 who found the two rates to be similar in oaks.
-In turn, g~dark~ could not be differentiated from stomatal conductance measurements during daytime in very low light (mostly just after dawn, or before sunset) (g~low PAR~), or during conditions when photosynthetic rates were very low (due to very high temperature, VPD, or other factors) (g~low A~).
-These comparisons demonstrate that g~s~ during non-drought conditions when photosynthetic rates are zero or negligible are much higher than the minimum reached in simulated drought conditions (g~min~).
-
-```{r fig.width=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure3.png")
-```
-
-\
-(Figure 3 here)
-
-\
-**Figure 3**.
-Comparison of various estimates of the (presumed) minimum conductance.
-Error bars are 95\% confidence intervals.
-Grey dots are the original data (but a few data points occur outside the figure range).
-Different letters denote significant differences (at $\alpha$ = 0.05).
-g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~low PAR~ leaf conductance during low light (photosynthetically active radiation, PAR); g~low A~ leaf conductance during periods of very low photosynthesis.
-See text for data sources and methods.
-
-# IV. Problems with estimating minimum conductance from regression
+## 3. Problems with estimating g~0~ from regression
 
@@
 The difficulty of the estimation can already be anticipated from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g.
-@leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species); presumably because unreliable estimates were obtained, b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though it is clearly non-sensical to suggest negative conductance values.
+@leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species, presumably because unreliable estimates were obtained), b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though it is clearly non-sensical to suggest negative conductance values.
 @barnard2013 also mentioned the difficulty of fitting Eq.
-1 to estimate g~0~, but did not present specific details - the following analysis builds on their work by demonstrating statistical uncertainties.
+1 to estimate g~0~, but did not present specific details.
+The following analysis builds on their work by demonstrating statistical uncertainties.
 
@@
 It is also telling that there are few reports on the intra-specific plasticity, or inter-specific variation of g~0~, perhaps because it is so poorly estimated.
-Exceptions are @duarte2016, who found a lower g~0~ in a heat wave treatment in Douglas-fir (although g~1~ was unaffected), but in their case g~0~ was very accurately estimated by careful multi-point light response curves.
+Exceptions are @duarte2016, who found a lower g~0~ in a heat wave treatment in *Pseudotsuga menziesii* (although g~1~ was unaffected), but in their case g~0~ was very accurately estimated by careful multi-point light response curves.
 Another exception is @misson2004, who report on a close negative correlation between pre-dawn leaf water potential and g~0~ in Ponderosa pine (again, the g~1~ parameter was unaffected).
@@
 In the following we demonstrate that the fitting process is problematic for three reasons: 1) the estimates of g~0~ and g~1~ are highly correlated, 2) the precision for g~0~ is generally much lower than for g~1~, and 3) for data that have a worse fit overall, the g~0~ estimates are elevated.
-We thus conclude that g~0~ should really not be estimated from regression, although this is the most common method applied.
+We thus conclude that g~0~ should not be estimated from regression, although this is the most common method applied.
 In addition, it is difficult - and generally not recommended - to accurately measure low fluxes with a portable gas exchange system, and great care has to be taken to arrive at reasonable estimates of g~0~ that way.
@@
 An example dataset is shown in Fig.
-4a, together with a fitted linear regression line.
+7a, together with a fitted linear regression line.
 We show in Fig.
-4b that, estimates of g~0~ and g~1~ are statistically correlated, that is, their confidence intervals are not independent.
+7b that, estimates of g~0~ and g~1~ are statistically correlated, that is, their confidence intervals are not independent.
 Thus large estimated values for g~1~ lead to low estimates for g~0~ and vice versa.
@@
 The correlation between slopes and intercepts is not unique to Eq.
-1, but a general property of linear models [@stapleton1995, @becker2007].
-
-```{r fig.height=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure4.png")
-```
+1, but a general property of linear models [@stapleton1995; @becker2007].
 
 \
-(Figure 4 here)
+(Figure 7 here)
 
 \
-**Figure 4**.
+**Figure 7**.
 Example linear regression of stomatal conductance (g~s~) against a combination of terms, in this case the linearized version of the model of Medlyn et al.
@@
 Next we study the values obtained when fitting Eq.
-1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation @miner2017.
+1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation of @miner2017.
 The (updated version of) the Lin et al.
 (2015) database includes more than 15 thousand gas exchange measurements on over 300 species.
-After selecting species/site combinations with more than n=15, we produced 78 estimates of g~1~ and g~0~ with non-linear regression of the @medlyn2011 model of stomatal conductance.
+After selecting species/site combinations with n > 15, we produced 78 estimates of g~1~ and g~0~ with non-linear regression of the @medlyn2011 model of stomatal conductance.
 The compilation by @miner2017 includes 233 estimates of g~0~ and *m* (they compiled parameters for original model of Ball-Berry, equivalent to g~1~) for 172 species (including woody plants and crops).
@@
 For both databases, estimates of g~0~ are inflated when the model fits poorly (Fig.
-5 a,b).
+8 a,b).
 This can be understood by considering that a poor fit often results in the flattening out of the regression line, thus giving a large value for the intercept.
@@
 1 is lower (Fig.
-5c).
-
-```{r fig.height=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure5.png")
-```
+8c).
 
 \
-(Figure 5 here)
+(Figure 8 here)
 
 \
-**Figure 5**.
+**Figure 8**.
 Statistical uncertainty in the estimation of g~0~ from regression, demonstrated with two parameter databases.
@@
 
-# V. Cuticular conductance
-
-\
-A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these processes have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b].
-However, a few key points should be summarized as they are relevant to the current discussion, in particular when we aim to interpret variation in literature values of g~min~ (Section VIII).
-
-\
-Although we use the term cuticular conductance freely, this transport pathway does not represent a true conductance since water does not diffuse as a gas through the cuticle.
-Instead, it dissolves into the medium of the cuticle, diffuses through the solid matrix, and is desorped at the outer edge of the cuticle [@kerstiens1996; @riederer2001; @schreiber1996].
-The main barrier to diffusion is actually a very thin layer of wax at the leaf surface.
-Because most of the resistance is located in such a thin layer, g~cuti~ does not correlate with the thickness of the cuticle [@priestley1943; @riederer2001; @anfodillo2002; @schuster2016b].
-Nonetheless there is considerable variation in cuticle thickness between plant species [@schuster2016a], along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010].
-If the cuticle thickness does not directly affect its conductance, what is the value of increasing the cuticle?
-
-\
-It is important to bear in mind the many other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, various interactions with insects including signalling and herbivory resistance [@mueller2008; @kerstiens1996c;@riederer2008], and foliar water uptake [@fernandez2017].
-The thickness of the cuticle may also confer mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear was proportional to the cuticle thickness).
-As such, thick cuticles may be advantageous in exposed environments to avoid excessive damage [@blackman2005].
-Prolonged exposure to wind has been shown to increase the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991], by dislodging cuticular compounds by abrasion [@rogge1993].
-Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986], most dramatically demonstrated in an ice storm, leading to much higher g~min~ [@boyce2003].
-
-\
-Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged.
-Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but was able to explain only some of the wide range of g~cuti~ measured across different plant species.
-Similarly, @hauke1998 found no relationship between the gradual decrease in g~cuti~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.
-
-# VI. Contribution of stomata
-
-The minimum leaf conductance includes two pathways: across the cuticle, and through the (potentially incompletely closed) stomata.
-Only a few studies have directly quantified the stomatal component of minimum conductance, in contrast to the wealth of information on the cuticular component.
-A detailed study of *Hedera helix*  concluded that 35% of water loss occurred across the stomatal pores (though closed), and 65% across the non-stomatal part of the cuticle, despite the fact that the stomata presumably covered only a small fraction of the leaf [@santrucek2004].
-For ten *Sorghum* genotypes, @muchow1989 reported a strong positive correlation between stomatal density and g~min~, suggesting leaky stomata contribute substantially to g~min~.
-For seven conifer species, @brodribb2014 reported that 50 - 94% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky, and likely incompletely closed.
-However @santrucek2004 found that the cuticle of the stomatal side (but excluding stomata itself) in *Hedera helix* was many times more more permeable than the non-stomatal side, thus providing another potential explanation for this difference.
-Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles, and the contribution of leaky stomata, and potentially more permeable cuticle on the abaxial side, likely varies between species.
-
-\
-One key question is whether incomplete stomatal closure is under the plant's control, or whether it is an unavoidable consequence of imperfect stomata, damage, blocking by particles, etc.
-For example, endophytic fungi - which commonly colonize plant leaves, have been shown to prevent stomatal closure and greatly increase water use  [@arnold2007].
-As we discuss in the next section, there is ample evidence to suggest that stomata not only stay in the dark, but that plants actively control stomatal conductance during the night.
-During severe drought, there is less evidence for such active control, and it seems likely incomplete closure is not under the plant's control as there is no obvious reason to keep stomata open.
-Some conditions (especially rapid drying conditions) may lead to excessive drying of the epidermis, which can physically pull apart the stomata ['mechanical advantage', @buckley2005].
-
-# VII. Nighttime conductance
+## 4. Night-time conductance
 
@@
 
-# VIII. Environmental and ecological variation in minimum conductance
-
-## 1. Minimum conductance shows little meaningful interspecific variation
+## 5. Towards a new model formulation
 
 \
-In the following sections, we review the quite substantial literature on g~min~ measured with the mass loss of detached leaves (MLD) technique, stretching back to the 1930s [@pisekberger1938], focussing on environmental and ecological determinants.
-We also further analyze the literature compilation of g~min~ (Fig.
-6), and separately analyze crop species and their genotypic variation.
-Comparing all available data, we did not find significant relationships between g~min~ and climate of origin, nor were there meaningful relationships with other traits (see Methods S1).
-We thus found it difficult to explain the variation in g~min~ between species.
-A simple breakdown by taxonomic order (Fig.
-6b) revealed that grasses (*Poales*) have a higher g~min~ compared to other orders, and conifers (especially *Pinales*) tended towards the lower range of values (but were only significantly different from *Poales*).
+We suggest that the minimum conductance in the Ball-Berry model should include both a g~0~ and a g~min~ term, since g~0~ represents the minimum reached during low light and conditions of low photosynthesis, and g~min~ represents the minimum reached during severe drought.
+Moreover, we suggest that the minimum conductance is not simply added to the photosynthesis-dependent term (right-hand side in Eq.
+1), but used as an actual minimum.
+Thus,
 
 \
-Very few studies have found meaningful correlations between g~min~ and environmental factors or ecological categories.
-Based on their own compilation of the literature, @schuster2017 concluded no significant differences in g~min~ or g~cuti~ (analyzed together) by plant growth form.
-One notable exception is @brodribb2014, who showed a correlation between rainfall at species origin (of the driest quarter), but used a gas-exchange approach (over many weeks), not MLD.
-Such correlation did not bear out across all species in our database, and interesting to note is that *Eucalyptus* species (n = 11, included in *Myrtales*) - all measured on the driest continent of Australia - have g~min~ values slightly higher than the average (Fig.
-6b).
+(4)
+$$ g_s = max \big[ max(g_{min}, g_0), g_1 \frac{A_n}{C_a}f(D) \big]$$
 
 \
-In crop science, g~min~ has long been identified as a key drought tolerance trait [@sinclair1986].
-A number of studies have targeted g~min~ as a key trait for breeding more drought-tolerant crops, leading to comparisons of g~min~ across genotypes grown in the same conditions.
-For example, @james2008 compared g~min~ in 58 soybean (*Glycine max*) genotypes, and found > 2-fold variation that could not be easily explained by other traits.
-In Fig.
-6c, we have compiled a number of studies in crops, demonstrating not just variation in g~min~ between crops, but also the wide range in g~min~ between genotypes.
-Again, it is striking that the wide variation in g~min~ cannot be easily explained by variation in leaf or other traits or chemical and structural components of the cuticle [@bengtson1978; @james2008; @saito2010], suggesting a significant role of incomplete stomatal closure.
+This model for g~s~ will converge to g~0~ during periods of low photosynthesis, and to g~min~ during drought if we further include a model for the dependence of g~0~ on water availability [see e.g.
+@misson2004].
+Another advantage is that independent estimates of g~0~ can be used, not those obtained via regression, which produces the undesirable correlation with estimates of g~1~.
+The above formulation is yet to be tested against data, but we propose this test to be done with data from drought and non-drought conditions.
 
-```{r fig.height=5, fig.width=5, echo=FALSE, eval=includefigs}
-plotfig("output/figure6.png")
-```
+# VII. Conclusions
 
 \
-(Figure 6 here)
-
-\
-**Figure 6**.
-Analysis of a literature compilation of g~min~ estimates, as measured with mass loss of detached leaves.
-(a) Histogram (probability density) of all estimates (after averaging by species, n=`r nrow(gmindat)`), with a log-normal distribution curve (mean = `r round(exp(mean(log(gmindat$gmin))),2)`, sd = `r round(exp(sd(log(gmindat$gmin))),2)`).
-(b) g~min~ averaged by phylogenetic order (including only the top ten orders in the database).
-Bars are 95% confidence intervals.
-Numbers above the figure refer to the number of species.
-Different letters denote significant differences (at $\alpha$ = 0.05, adjusted for multiple comparisons), and grey symbols are species-level data.
-(c) g~min~ estimates for crops only, averaged by genotype.
-Bars denote the range, illustrating the wide range in g~min~ between genotypes for a particular crop species.
-Numbers above the figure refer to the number of genotypes included.
-
-## 3. Acclimation to the environment
-
-\
-While there is clearly plenty of variation in g~min~ between species, a number of lines of evidence suggest that g~min~ also has great potential for plasticity.
-Here we summarize the literature on the acclimation of g~min~ to drought conditions, to changes in temperature and humidity, and the change in g~min~ with leaf age, altitude and other factors.
-This discussion is directly relevant to the use of g~min~ in models, because if the degree of plasticity is large, it complicates model parameterization.
-
-\
-There is a general tendency for a decreased g~min~ in plants acclimated to drought stress [@james2008].
-The magnitude of the decrease of g~min~ with acclimation to drought stress varied from -4 to -70% (across 10 studies, see Table S2), with a typical decrease in the order of 30-40%.
-In each of the studies summarized, plants were grown in well-watered or droughted conditions, and in one case a difference in g~min~ was demonstrated after just four days of drought exposure [@bengtson1978].
-We also demonstrate drought acclimation via a case study on 11 *Hakea* species grown in two watering treatments (Fig.
-7) - all 11 species showed a decrease in g~min~ in the drought treatment (see Methods S3 for experimental details).
-This significant change in g~min~ with drought acclimation is likely an important component of the overall drought hardening of plants.
-
-\
-The idea that water limitation causes a reduction in g~min~ (via changes in the chemical composition of the cuticle) can be tested by inspecting the response to factors that increase evaporative demand.
-As one of the very few studies testing this idea, @fanourakis2013 reported much lower g~min~ in *Rosa* sp.
-plants grown in 60% vs.
-95% relative humidity.
-The difference could be attributed largely to a change in stomatal anatomy and lack of closure during desiccation, not changes in the cuticle *per se*.
-Also relevant is @sack2003, who reported lower g~min~ in sun leaves (for 2 out of 4 species) compared to shade leaves, again a lower g~min~ for leaves acclimated to high evaporative conditions.
-
-\
-A number of studies have attempted to attribute the drought-induced acclimation in g~min~ to a change in the chemical composition of the cuticle.
-@bengtson1978 reported lower g~min~ in six oat varieties in response to drought, and an increase in the amount of cuticular waxes, but could not find a relationship between the two and the response was highly genotype-specific.
-However, @premachandra1992 found that epicuticular wax load increased on leaves of non-irrigated Sorghum cultivars and was positively correlated with cuticular conductance and cell membrane stability.
-@mackova2013 found that adding ABA (simulating drought stress) to *Lepidium sativum* increased the chain length of cuticular waxes (but not total amounts).
-@bi2017 also reported that drought caused a change in cuticular wax production and chemical composition, but again in a highly genotype-specific way.
-Thus it can be concluded that the cuticle indeed changes substantially after a change in plant available water, but in a complex, species-specific manner that is yet to be connected directly to changes g~min~.
-This opens up the possibility that leaky stomata are a significant component of the drought response, not just the composition of the cuticle.
-
-```{r fig.height=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure7.png")
-```
-
-\
-(Figure 7 here)
-
-\
-**Figure 7**.
-Minimum conductance (g~min~) measured by mass loss of detached leaves on a variety *Hakea* species, a genus native to Australia.
-Plants were grown in containers in a grow house, and supplied with ample water or subjected to long-term (8 months) mild drought stress.
-Bars are labelled by leaf form : broadleaf (B) or needle-like (N), and ordered by g~min~ in the  well-watered treatment.
-For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`.
-
-\
-The response of g~min~ to temperature is more complex; it shows both a response to instantaneous changes in temperature, as well as an acclimatory response to growth temperature.
-@riederer2001 and @schuster2016b have both demonstrated a steep non-linear instantaneous response of g~cuti~ to temperature, with the response becoming especially steep at higher temperatures (> 40C).
-In *Eucalyptus haemastoma*, the response of g~min~ to temperature was so steep, that the proportion of cuticular to total transpiration increased from 2-3% at 20°C, to 40% at 38°C [@eamus2008].
-The mechanism of the instantaneous temperature response is complex, highly species-specific, and we refer to @schuster2016b for a detailed investigation.
-The quick increase in g~min~ at high temperature may well be a crucial component of the ability of plants to tolerate heat waves [@drake2018].
-In support of the link between g~min~ and heat tolerance, @schuster2016a reported on a negative relationship between thermal tolerance and g~min~ across nine species, so that species with improved tolerance to very high temperature had a low g~min~.
-
-\
-Less well established is the acclimatory potential of g~min~ to changes in growth temperature.
-We present a case study on *Eucalyptus parramattensis* grown in whole-tree chambers (see Methods S3 for experimental details).
-The chambers either tracked ambient conditions, or were subjected to a persistent +3°C warming treatment.
-After several months of growth in the treatments, g~min~ was measured at various temperatures, ranging from 17.5 to 27.5°C.
-We found a 56% decrease in g~min~ in the elevated temperature trees, but there was no clear pattern with measurement temperature (Fig.
-8).
-This decrease in g~min~ is consistent with the drought response, because again g~min~ is reduced in leaves that are subjected to conditions that increase evaporative demand (vapour pressure deficit was higher in the elevated temperature treatment, see @drake2018).
-The direction of this response is consistent with @duarte2016, who reported lower g~0~ and g~dark~ in a heat wave treatment in *Pseudotsuga menziesii*, which persisted for some time.
-Responses across temperature gradients are more complex, in particular the change in g~min~ with altitude has been well studied.
-It is commonly  reported that g~min~ increases with altitude [e.g.
-@delucia1984; @herrick1991; @anfodillo2002].
-@fernandez2017 discuss the literature on altitude responses in detail and argue that the short growing seasons at high altitude are insufficient for complete maturation of leaf cuticles.
-
-\
-Finally we mention the striking effect of leaf age on g~min~.
-We have summarized five studies in Fig.
-S3, all of which reported on increased g~min~ for older leaves for the majority of species studied.
-The data reported by @jordan2007 is particularly impressive, as g~min~ in the woody shrub *Agastachys odorata* gradually increased with leaf age up to ca.
-10 years.
-If this effect is due to properties of the cuticle alone, a possible explanation for the increase in g~min~ is the continued exposure to wind, rain, and abrasives, which have been shown to damage the cuticle and increase its conductance (see discussion in Section V).
-Another possibility is that the contribution of stomata to g~min~ increases with leaf age, as reported by @jordan2007.
-In plants that maintain several cohorts of leaves, finding an appropriate value of g~min~ for use in models will have to take into account the leaf age effect.
-
-```{r fig.height=4,echo=FALSE, eval=includefigs}
-plotfig("output/figure8.png")
-```
-
-\
-(Figure 8 here)
-
-\
-**Figure 8**.
-Minimum conductance (g~min~) measured with mass loss of detached leaves on *Eucalyptus parramattensis* grown in whole-tree chambers.
-Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 °C) [see @drake2018].
-Measurements of g~min~ were carried out at five different temperatures on replicate leaves.
-There was no consistent effect of measurement temperature on g~min~, but leaves of trees grown in elevated temperature showed on average a 45% reduction in g~min~ (p < 0.01, linear mixed-effects model).
-
-# IX. Conclusions
-
-\
-In a pioneering publication which presented the Jarvis model of stomatal conductance, @jarvis1976 states that "we have assumed in the following equations that when stomata are closed the leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." Similarly, given the poor statistical properties of g~0~ estimated from regression we conclude that g~0~ should not be estimated from regression on leaf gas exchange data for use in models.
+In a pioneering publication on stomatal conductance, @jarvis1976 states that "we have assumed in the following equations that when stomata are closed the leaf conductance is zero because field data are generally inadequate to define a cuticular conductance." Similarly, given the poor statistical properties of g~0~ estimated from regression we conclude that g~0~ should not be estimated from regression on leaf gas exchange data for use in models.
 Then, how should g~0~ be estimated?
@@
 Thus when modelling night-time or low-light conductance, a different g~0~ should be used than when modelling the drought response of plants.
-We suggest a new model form that includes both g~0~ and g~min~ with some desirable properties in Section II.
+We suggest a new model form that includes both g~0~ and g~min~ with some desirable properties in Section VI.
 
 \
-Finally, we conclude that g~min~ displays a large amount of variation between species that could not be explained by traits, and remarkable plasticity to growing conditions.
-Perhaps this plasticity is the reason that g~min~ did not vary predictably between species.
-Another possibility is that the lack of standardized methods for measurement preclude clear comparisons between species.
+Finally, we conclude that g~min~ displays a large amount of variation among species that could not be explained by traits, and remarkable plasticity to growing conditions.
+Perhaps this plasticity is the reason that g~min~ did not vary predictably among species.
+Another possibility is that the lack of standardized methods for measurement preclude clear comparisons among species.
 Future studies should compare g~min~ on many species grown in the same conditions to better understand the adaptive value of the minimum water loss rate of leaves.
@@
 This work was made possible by the ARC Linkage Project LP140100232.
-All raw data, code and a fully reproducible workflow which generates the figures, this manuscript and all analyses is available at https://www.github.com/remkoduursma/g0paper.
+RL was supported by a Marie Curie fellowship (FP7-IOF 624473).
+All raw data, code and a fully reproducible workflow which generates the figures, this manuscript and all analyses are available at https://www.github.com/remkoduursma/g0paper.
 

### 2018-06-02 R/make_figures.R

@@ -1,24 +1,22 @@
 
 source("R/load.R")
 
-to.pdf(figure_sim(), "output/figure1.pdf", height=6, width=6)
+to.pdf(figure_gmin_review(gdfr), "output/figure1.pdf", height=5, width=5)
 
-to.pdf(figure_sureau(planta, plantb), "output/figure2.pdf", height=4, width=9)
+to.pdf(gmin_3panel(gmindat, cropgmin), "output/figure2.pdf", height=3, width=8)
 
-to.pdf(figure_gmin_review(gdfr), "output/figure3.pdf", height=5, width=5)
+to.pdf(figure_hakea_gmin(lopw), "output/figure3.pdf", height=5, width=6)
+
+to.pdf(figure_wtc4_gmin_2(wtc4gmin), "output/figure4.pdf", height=4, width=3.5)
+
+to.pdf(figure_sim(), "output/figure5.pdf", height=6, width=6)
+
+to.pdf(figure_sureau(planta, plantb), "output/figure6.pdf", height=4, width=9)
 
 to.pdf(figure_g0g1_cor(lin2015, "Nicolas Martin-StPaul_Quercus ilex_StPaul_Puechabon"), 
-       "output/figure4.pdf", height=4, width=8)
+       "output/figure7.pdf", height=4, width=8)
 
-to.pdf(figure_R2g0(lin2015coef, miner), "output/figure5.pdf", height=4, width=9)
+to.pdf(figure_R2g0(lin2015coef, miner), "output/figure8.pdf", height=4, width=9)
 
-to.pdf(gmin_3panel(gmindat, cropgmin), "output/figure6.pdf", height=3, width=8)
-
-to.pdf(figure_hakea_gmin(lopw), "output/figure7.pdf", height=5, width=6)
-
-to.pdf(figure_wtc4_gmin_2(wtc4gmin), "output/figure8.pdf", height=4, width=3.5)
-
-
-#convert_all_pdf("output")
 
 

### 2018-06-02 manuscript.Rmd
escape space in subscript

@@
     \t3. Problems with estimating g~0~ from regression                      
-    \t4. Nighttime conductance                                              
+    \t4. Night-time conductance                                              
     \t5. Towards a new model formulation                                    
@@
 150 species during the night (we selected only gas exchange based measurements from their database).
-Daytime values (g~low PAR~ and g~low A~) were estimated from an update to the @lin2015 database by taking the appropriate subsets, and averaging the g~s~ values by species within study.
+Daytime values (g~low\ PAR~ and g~low\ A~) were estimated from an update to the @lin2015 database by taking the appropriate subsets, and averaging the g~s~ values by species within study.
 All data and code to reproduce the database and analyses are available online (see Acknowledgments).
@@
 Previous studies have directly compared g~dark~ and g~min~ on the same species, and generally concluded that g~dark~ is much higher than g~min~ [@hygen1954; @koerner1994; @walden-coleman2013], but an exception is @cavender-bares2007 who found the two rates to be similar in oaks.
-In turn, g~dark~ could not be differentiated from stomatal conductance measurements during daytime in very low light (mostly just after dawn, or before sunset) (g~low PAR~), or during conditions when photosynthetic rates were very low (due to very high temperature, VPD, or other factors) (g~low A~).
+In turn, g~dark~ could not be differentiated from stomatal conductance measurements during daytime in very low light (mostly just after dawn, or before sunset) (g~low\ PAR~), or during conditions when photosynthetic rates were very low (due to very high temperature, VPD, or other factors) (g~low\ A~).
 These comparisons demonstrate that g~s~ during non-drought conditions when photosynthetic rates are zero or negligible are much higher than the minimum reached in simulated drought conditions (g~min~).
@@
 Different letters denote significant differences (at $\alpha$ = 0.05).
-g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~low PAR~ leaf conductance during low light (photosynthetically active radiation, PAR); g~low A~ leaf conductance during periods of very low photosynthesis.
+g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~low\ PAR~ leaf conductance during low light (photosynthetically active radiation, PAR); g~low\ A~ leaf conductance during periods of very low photosynthesis.
 See text for data sources and methods.

### 2018-06-02 manuscript.Rmd
zenodo template

@@
 RL was supported by a Marie Curie fellowship (FP7-IOF 624473).
-All raw data, code and a fully reproducible workflow which generates the figures, this manuscript and all analyses are available at https://www.github.com/remkoduursma/g0paper.
+All raw data, code and a fully reproducible workflow which generates the figures, this manuscript and all analyses are available at https://www.github.com/remkoduursma/g0paper, archived at Zenodo (doi:XXX).
 

### 2018-06-02 manuscript.Rmd
Nico revisions added

@@
 \
+A comprehensive review of g~min~ with the goal to improve functional model representation is made difficult by the fact that literature data arise from various methods, and represent distinct processes.
 The absolute minimum attainable water loss rate is through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf.
@@
 For seven conifer species, @brodribb2014 reported that 50 - 94% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky, and likely incompletely closed.
-However @santrucek2004 found that the cuticle of the stomatal side (but excluding stomata itself) in *Hedera helix* was many times more permeable than the non-stomatal side, thus providing another potential explanation for this difference.
+However, @santrucek2004 found that the cuticle of the stomatal side (but excluding stomata itself) in *Hedera helix* was many times more permeable than the non-stomatal side, thus providing another potential explanation for this difference.
 Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles, and the contribution of leaky stomata, and potentially more permeable cuticle on the abaxial side, likely varies among species.
@@
 As pointed out by @leuning1990 and @leuning1995, g~0~ needs to be > 0, otherwise the ratio of intercellular CO~2~ concentration to atmospheric (C~i~ / C~a~) does not vary with PAR, as is typical in leaf gas exchange data [though other mechanisms can also be employed to simulate this pattern, see @dewar2018].
-When PAR approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~a~.
+When PAR approaches the light compensation point, clearly C~i~ needs to approach C~a~, since no photosynthesis is occurring that draws down the C~i~.
 To see this point, we can rearrange Eq.
@@
 
-The previous section discussed and compared methods for more or less direct measurements of various definitions of minimum conductance.
+The previous section discussed and compared methods for more or less direct measurements of minimum conductance according to various definitions.
 The approach taken by the majority of vegetation models is however very different.
@@
 This model for g~s~ will converge to g~0~ during periods of low photosynthesis, and to g~min~ during drought if we further include a model for the dependence of g~0~ on water availability [see e.g.
-@misson2004].
+@misson2004] - as long as the right-hand term (including g~1~) is reduced under drought as well [@zhou2013].
 Another advantage is that independent estimates of g~0~ can be used, not those obtained via regression, which produces the undesirable correlation with estimates of g~1~.

### 2018-07-10 manuscript.Rmd
Final revisions

@@
 
-During conditions where photosynthesis does not take place, such as severe drought, extreme temperature or low light, it seems advantageous for plants to close stomata and completely halt water loss.
+When the rate of photosynthesis is greatly diminished, such as during severe drought, extreme temperature or low light, it seems advantageous for plants to close stomata and completely halt water loss.
 However, water loss continues through the cuticle and incompletely closed stomata, together constituting the leaf minimum conductance (g~min~).
@@
 It is well known that stomata open and close in response to changes in light intensity, humidity and CO~2~ concentration at the leaf surface.
-Stomata can be thought to balance the uptake of CO~2~ with the loss of water, and thus to close stomata when photosynthetic rates are low.
-To avoid desiccation and ultimate death, stomata typically close during periods of severe water stress.
+To avoid desiccation and ultimate death, stomata typically close during periods of water stress.
 When stomata are closed, water loss continues at a greatly diminished rate through the cuticle.
@@
 
-# II. Comparison of various definitions of minimum conductance
+# II. Comparison of various definitions and measurement techniques of minimum conductance
 
 \
-Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories.
+Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories: conductance of the cuticle only, conductance of detached leaves, and gas-exchange measurements during conditions leading to presumed stomatal closure.
 The absolute minimum conductance attained by leaves is that through the cuticle only.
@@
 \
-We contrasted g~cuti~ and g~min~ with three estimates of stomatal conductance when photosynthesis rates are low or zero during non-drought conditions: night-time conductance (g~dark~) (further discussed in Section VI.4), conductance at low photosynthetically active radiation (PAR) (0 - 40 $\mu$ mol m^-2^ s^-1^) during the day, and conductance during conditions when photosynthesis rates are very low (< 1 $\mu$ mol m^-2^ s^-1^) (but excluding low PAR and drought).
+We quantitatively compared g~cuti~ and g~min~ by synthesizing existing data.
+In addition, we compiled data on leaf conductance when photosynthesis rates are low or zero during non-drought conditions: night-time conductance (g~dark~) (further discussed in Section VI.4), conductance at low photosynthetically active radiation (PAR) (0 - 40 $\mu$ mol m^-2^ s^-1^) during the day, and conductance during conditions when photosynthesis rates are very low (< 1 $\mu$ mol m^-2^ s^-1^) (but excluding low PAR and drought).
 
 \
-We quantitatively compared each of the different conductances by synthesizing existing data.
 Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but these were not used since they included no metadata).
@@
 Nonetheless there is considerable variation in cuticle thickness among plant species [@schuster2016a], along altitudinal transects [@delucia1984], and even with increasing height in the canopy of very tall trees [@woodruff2010].
-If the cuticle thickness does not directly affect its conductance, what is the value of increasing the cuticle?
+If the cuticle thickness does not directly affect its conductance, what is the value of a thicker cuticle?
 
@@
 
+\
 The minimum leaf conductance includes two pathways: across the cuticle, and through the (potentially incompletely closed) stomata.
@@
 
-## 1. Minimum conductance shows little meaningful interspecific variation
+## 1. Minimum conductance is highly variable among species
 
@@
 @bengtson1978 reported lower g~min~ in six oat varieties in response to drought, and an increase in the amount of cuticular waxes, but could not find a relationship between the two and the response was highly genotype-specific.
-However, @premachandra1992 found that epicuticular wax load increased on leaves of non-irrigated Sorghum cultivars and was positively correlated with cuticular conductance and cell membrane stability.
+However, @premachandra1992 found that epicuticular wax load increased on leaves of non-irrigated *Sorghum* cultivars and was positively correlated with cuticular conductance and cell membrane stability.
 @mackova2013 found that adding ABA (simulating drought stress) to *Lepidium sativum* increased the chain length of cuticular waxes (but not total amounts).
@@
 In addition, g~dark~ can show a clear response to vapour pressure deficit [@barbour2007] during night-time conditions.
+Finally, g~dark~ is under strong genetic control independent of day-time water use - at least in grapevine [@coupelledru2016].
 Clearly g~dark~ is an actively controlled process that cannot be adequately summarized by a single constant g~0~, and should be modelled in a separate framework that is yet to be identified.

### 2018-07-10 R/load.R

@@ -183,6 +183,7 @@ gmindat <- left_join(gmindat,  cls, by="species")
 
 # gs at low PAR, gs at low A
 # Both from updated Lin database. Extracted for convenience.
+# See R_other/cond_low_a_low_par.R
 gs_low_a <- read.csv("data/cond_low_a.csv") %>%
   dplyr::select(gmin = Cond.mean) %>%
   mutate(method = "gslowA",

### 2018-07-10 manuscript.Rmd
Added Zenodo DOI

@@
 RL was supported by a Marie Curie fellowship (FP7-IOF 624473).
-All raw data, code and a fully reproducible workflow which generates the figures, this manuscript and all analyses are available at https://www.github.com/remkoduursma/g0paper, archived at Zenodo (doi:XXX).
+All raw data, code and a fully reproducible workflow which generates the figures, this manuscript and all analyses are available at https://www.github.com/remkoduursma/g0paper, archived at Zenodo (doi:10.5281/zenodo.1308628).
 

### 2018-07-18 manuscript.Rmd
final minor edits for copyeditor

@@
 \
-Remko A. Duursma^1^, Chris J. Blackman^1^, Rosana Lopéz^1,3^, Nicolas K. Martin-StPaul^2^, Hervé Cochard^3^, Belinda E. Medlyn^1^
+Remko A.
+Duursma^1^ (ORCID: 0000-0002-8499-5580), Christopher J.
+Blackman^1^, Rosana Lopéz^1,3^, Nicolas K.
+Martin-StPaul^2^ (ORCID: 0000-0001-7574-0108), Hervé Cochard^3^ (ORCID: 0000-0002-2727-7072), Belinda E.
+Medlyn^1^ (ORCID: 0000-0001-5728-9827)
 
@@
 
+Keywords: cuticular conductance, drought tolerance, ecosystem modelling, plant water relations, stomatal conductance model
+
 # I. Introduction
@@
 \
-Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories: conductance of the cuticle only, conductance of detached leaves, and gas-exchange measurements during conditions leading to presumed stomatal closure.
+Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories: conductance of the cuticle only, conductance of detached leaves, and gas-exchange measurements during conditions leading to presumed stomatal closure (Table 1).
 The absolute minimum conductance attained by leaves is that through the cuticle only.
@@
 Although this method resembles conditions that plants may experience during a dry-down or periods of extreme stress, some uncertainties regarding the methods for measurement remain.
-In our own work, we have found that in some leaves, water loss rate increases (rather than decreases) some time after leaf detachment (see Methods S2).
+In our own work, we have found that in some leaves, water loss rate increases (rather than decreases) some time after leaf detachment (see Methods S1).
 It is likely that the relatively rapid dry-down sometimes causes artefacts, and must be carefully avoided [see also @heinsoo1998].
@@
 Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but these were not used since they included no metadata).
-Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see Methods S1 for description and references).
-For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Methods S1 for full details of the compilation).
+Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see Methods S2 for description and references).
+For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Methods S2 for full details of the compilation).
 The database includes a total of 221 species (136 Angiosperm, 49 Gymnosperm, 1 Pteridophyte), from 57 taxonomic families, with woody species making up roughly 2/3 of the data.
@@
 g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~low\ PAR~ leaf conductance during low light (photosynthetically active radiation, PAR); g~low\ A~ leaf conductance during periods of very low photosynthesis.
-See text for data sources and methods.
+See Section II for data sources and methods.
 
@@
 There is a general tendency for a decreased g~min~ in plants acclimated to drought stress [@james2008].
-The magnitude of the decrease of g~min~ with acclimation to drought stress varied from -4 to -70% (across 10 studies, see Table S2), with a typical decrease in the order of 30-40%.
+The magnitude of the decrease of g~min~ with acclimation to drought stress varied from -4 to -70% (across 10 studies, see Table S1), with a typical decrease in the order of 30-40%.
 In each of the studies summarized, plants were grown in well-watered or droughted conditions, and in one case a difference in g~min~ was demonstrated after just four days of drought exposure [@bengtson1978].
@@
 We have summarized five studies in Fig.
-S3, all of which reported an increased g~min~ for older leaves for the majority of species studied.
+S1, all of which reported an increased g~min~ for older leaves for the majority of species studied.
 The data reported by @jordan2007 are particularly impressive, as g~min~ in the woody shrub *Agastachys odorata* gradually increased with leaf age up to ca.
@@
 There was no consistent effect of measurement temperature on g~min~, but leaves of trees grown in elevated temperature showed on average a 45% reduction in g~min~ (p < 0.01, linear mixed-effects model).
+Error bars denote 95% confidence intervals.
 
@@
 1 as published by @ballberry1987 did not include an intercept term.
-It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model.
+It was first introduced in the unpublished manuscript of Ball & Berry (1988, available on request), as an 'intercept' without further discussion of the role it plays in the model.
 In practice, g~0~ is usually estimated from a regression of Eq.
@@
 **Figure 7**.
-Example linear regression of stomatal conductance (g~s~) against a combination of terms, in this case the linearized version of the model of Medlyn et al.
+(a) Example linear regression of stomatal conductance (g~s~) against a combination of terms, in this case the linearized version of the model of Medlyn et al.
 (2011) applied to a leaf gas exchange dataset of @martinstpaul2012.
@@
 The solid line is regression line, and the shaded area a 95% confidence interval for the mean.
-The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression.
+(b) The correlation between estimated slope (g~1~) and intercept (g~0~) of the regression shown in panel (a).
 The dotted ellipse is a bivariate 95% confidence interval for slope and intercept.
@@
 
-**Methods S1** Literature compilation of g~min~ estimates
+**Methods S1** Methodological issues with measurement of g~min~
 
-**Methods S2** Methodological issues with measurement of g~min~
+**Methods S2** Literature compilation of g~min~ estimates
 
